### 2021-08-23 Frontiers/Frontiers.Rmd
as

---
title: "Palmer amaranth (*Amaranthus palmeri*) adaptation to US Midwest agroecosystems"
runtitle: "Palmer amaranth adaptation"
documentclass: frontiersSCNS  # or frontiersHLTH, or frontiersFPHY
author:
  - name: Maxwel C Oliveira
    affiliation: '1'
    etal: Oliveira # First author's last name. 
  - name: Amit J Jhala
    affiliation: '2'
  - name: Mark Bernards
    affiliation: '3'
  - name: Chris Proctor
    affiliation: '2'
  - name: Strahinja Stepanovic
    affiliation: '2'
  - name: Rodrigo Werle
    affiliation: '1'
    email: rwerle@uwisc.edu # Indicates corresponding Author
    institution: Department of Agronomy, University of Wisconsin-Madison, United States
    street: 1575 Linden Dr
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    zip: 53705
    country: United States
affiliation:
  - id: '1'
    department: Department of Agronomy
    institution: University of Wisconsin-Madison
    city: Madison
    state: Wisconsin # only USA, Australia, Canada
    country: United States
  - id: '2' 
    department: Department of Agronomy and Horticulture
    institution: University of Nebraska-Lincoln
    city: Lincoln
    state: Nebraska # only USA, Australia, Canada
    country: United States
  - id: '3' 
    department: Department of Agronomy
    institution: Western Illinois University
    city: Macomb
    state: Illnois # only USA, Australia, Canada
    country: United States
output: rticles::frontiers_article
csl: frontiers.csl
bibliography: adaptation.bib
---

```{r include=FALSE}
library(kableExtra)
library(tidyverse)
```

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}

\end{abstract}

# Introduction {-}
Palmer amaranth (*Amaranthus palmeri* S. Watson) is currently ranked as one of the most economically detrimental weed species to cropping systems in the United States [@vanwychen2020]. Unmanaged Palmer amaranth plants compete for water, light, and nutrients, which can drastically impact crop yields [@berger2015]. For example, Palmer amaranth has been documented to reduce up to 91%, 68%, and 54% corn [@massinga2001], soybean [@klingaman1994], and cotton [@morgan2001] yields, respectively. Moreover, Palmer amaranth has shown a remarkable capacity to evolve resistance to herbicides. To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017]. Thus, Palmer amaranth poses an economical and environmental risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to the Sonoran Desert [@sauer1957]. The species would eventually emerge as a threat to US agriculture in the 1990s. Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology. Farm mechanization, adoption of conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013]. On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992]. With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018]. Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017]. The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.

Light and temperature are the main environment requirements for Palmer amaranth successful grow and development [@jha2010]. Palmer amaranth is reported with an extended germination period [@ward2013]. Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003]. In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010]. Water has not shown to limit Palmer amaranth fitness. Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018]. Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021]. Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019]. Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under selection pressure. Palmer amaranth invasion/adaptation into new habitats is likely to increase with the global temperature warming. Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018]. Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018; @briscoerunquist2019].

Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021]. In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019; @bagavathiannan2016] in the southern United States but its range is expanding northward. For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a]. There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021]. Also, it is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares. In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021]. Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021]. Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021]. Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America. Therefore, its range of expansion into new habitats is increasing. It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future. Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed. Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management. Palmer amaranth management should be built on minimizing the species ability to adapt, grow and develop into agroecossystems.

In the southeastern US, early growing Palmer amaranth is well known to have a higher impact on cotton yields compared to late established plants [@macrae2013]. In the northern states, Palmer amaranth impact on the agroecosystem is recent. Studies investigating Palmer amaranth in those locations are limited due to the plant classification as noxious weed species in some northern states [@yu2021]. Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in such environments. Understanding Palmer amaranth biology and growing strategies under different agroecosystems can enhance our knowledge on species adaptation and management practices. It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and developing resilient and sustainable farming systems [@maclaren2020]. Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.

# Results {-}

<!-- Comments should be HTML-style comments -->

## Plant material and growing conditions {-}

A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with no reported herbicide resistance according to @oliveira2021a was selected for this study. Three weeks prior to the onset of the field experiments, seeds were planted in plastic trays containing potting-mix. Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>). Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors in the Perkins extension office at Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.

## Field study {-}

The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N, 'W)).

Fields were conventionally tilled prior to crop planting. Corn hybrid and soybean varieties were planted in 76-cm row spacing (Table 1). Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)

```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/weather/Figure 1.png")
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
options(knitr.kable.NA = '')
read_csv("../data/table.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  kable(booktabs = TRUE, caption = "Field study attributes ",
        col.names = c("Attributes",
                           "",
                           "Arlington, WI",
                           "Clay Center, NE",
                           "Grant, NE",
                      "Lincoln, NE", "Macomb, IL")) %>% 
#  pack_rows(
 #   index = c("Bareground" = 2, "Corn" = 4, "Soybean" = 8)) %>% 
  kable_styling(font_size = 10, 
                full_width = FALSE, 
                position = "center",
                latex_options = c("scale_down", "hold_position")) %>% 
  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") 
```

The field experimental units were three adjacent 9.1 m wide (12 rows at 76.2 cm row spacing) by 10.7 m long. The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations. Each experimental unit was planted with corn, soybean, or kept under bareground. The two transplanting timings were June 1 (first cohort) and July 1 (second cohort). Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2). After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2). When needed, Palmer amaranth plants were supplied with water during the first week after transplanting.

```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 2.png")
```

After transplanting, Palmer amaranth flowering was monitored until the end of the study. When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant. Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant-1 was recorded).

Plants had to be harvested at flowering because Palmer amaranth is neither endemic in Wisconsin nor in Illinois. In our study, all locations follow the methodology of plant harvest at flowering initiation, except in Grant, NE. In this location, all Palmer amaranth plants were harvest at once on July 06, 2018 and 2019 (first cohort), and on August 17, 2018 and on July 31, 2019 (second cohort).

## Statistical analyses {-}

The statistical analyses were performed using R statistical software version 4.0.1.

Analyses of Palmer amaranth height and biomass were performed with a linear mixed model using *lmer* function from "lme4" package [@bates2015]. Plant height and biomass were log transformed to meet model assumption of normality. In the model, crop (bareground, corn, soybean) was the fixed effect and year nested with location the random effects. Analysis of variance was performed with *anova* function from "car" package [@fox2018]. Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].

The Palmer amaranth flowering timing was estimated as cumulative flowering across all location. Dataset was nested but id did not include Grant, NE. Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].

$$Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))$$

In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year (doy).

The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package. Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crops and cohorts using the *EDcomp* function of drc package. The EDcomp function compares the ratio of cumulative flowering using t-statistics, where P-value < 0.05 indicates that we fail to reject the null hypothesis.

Palmer amaranth sex was fitted to a binary logistic regression [@bangdiwala2018]. Binary logistic regression is used for predicting binary classes, such as the probability of a plant being female in a dioecious species. Prior to the analysis, all missing values were removed from the dataset. Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location. The complete dataset was splitted into 80% train and 20% test data. The 80% train is used for the model training and the 20% test is used for checking how the model generalized on unseen dataset. With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction). The model family was binomial with a logit function. The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package). The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package. The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function. The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package. The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0. F1-score conveys the balance between the precision and the recall [@yacouby2020].

# Results {-}

<!-- Comments should be HTML-style comments -->

## Palmer amaranth height and biomass {-}

Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (figure 3A). Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and 16.3 g plant-1 in bareground, soybean and corn, respectively. Plants from the second cohort produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.

```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 3.png")
```

Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 3B). Palmer amaranth plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74). In addition, no difference in Palmer amaranth height (69.3 cm) was detected from first cohort plants in soybean to first and second cohort plants in bareground (P > 0.75). Palmer amaranth plants from the second cohort were nearly 10 cm lower compared to the first cohort in soybeans (P = 0.04). The tallest (first cohort) and smallest (second cohort) Palmer amaranth plants were found in corn. Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.

## Palmer amaranth cumulative flowering {-}

Palmer amaranth plants from the first cohort growing in corn resulted in a longer flowering window compared to plants growing in bareground and soybean (Figure 4A). The 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June. Palmer amaranth reached 10% flowering in soybean, bareground and corn at doy 180, 180.9 and 181.7, respectively. The 50% Palmer amaranth cumulative flowering occurred in July. Palmer amaranth reached 50% flowering in bareground, soybean and corn at doy 193.4, 194.8, and 206.6, respectively. Similar trend was observed at 90% Palmer amaranth cumulative flowering. Palmer amaranth reached 90% flowering at doy 252.6 in corn (early September), which was 38 and 32 days after reaching 90% flowering in bareground and soybean, respectively.

```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 4.png")
```

Palmer amaranth cumulative flowering at the second cohort ranged from mid July to mid September (Figure 4B). Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn. Palmer amaranth growing in bareground reached 10%, 50%, and 90% flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer amaranth growing in soybean reached 10% flowering at doy 210.9, which was 6 days prior to corn (*P*-value = 0.00). Similar trend was observed at 50% flowering, whereas Palmer amaranth reached 50% flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; *P* = 0.00). The 90% Palmer amaranth cumulative flowering occurred at same day in corn (260.9) and soybean (260.5; *P* = 0.66).

## Palmer amaranth gender {-}

The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively. The likelihood ratio test showed a p-value of < 0.00. The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% less females plants (Table 2). Moreover, increasing a unit doy  increases the probability of having a female plant by 0.4% (Table 2 and Figure4A). Similar trend is observed for weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.

```{r echo = FALSE, warning = FALSE, message=FALSE}
read_csv("../data/table_a.csv") -> table_a

#names(table_a)[2] <- paste0(names(table_a)[2], 
#                                footnote_marker_symbol(1))

table_a %>% 
  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
  kable_styling(font_size = 10, 
                full_width = FALSE, 
                position = "center",
                latex_options = c("hold_position")) %>% 
  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") %>%
  footnote(
           alphabet = c("Average Marginal Effects.", "Standard Error."),
           footnote_as_chunk = TRUE)
```

```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 5.png")
```

The model accuracy evaluation accuracy in the 20% test dataset was 0.62 with a cutoff value for female and male plants of 0.43. The model classification showed a precision of 0.64, recall of 0.66, and a F means score of 0.65. In addition, the area under the curve was 0.64.

# Discussion {-}

Our study showed that Palmer amaranth biomass, height, flowering pattern and gender varied within agroecossystems and cohort timings.  In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort. At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for the species. High biomass and taller Palmer amaranth plants are likely a weed strategy to compete for light in between crop rows in absence of canopy. In such conditions, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem. This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean. The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 6). These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018]. Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].

```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
knitr::include_graphics("../data analysis/figures/Figure 6.png")
```

Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages. Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness. As a result, Palmer amaranth height and biomass was lower compared to its first cohort. Under crop canopy (e.g., second cohort), Palmer amaranth flowering window was near to similar in corn and soybean. Palmer amaranth growing without crop competition produced the highest amounts of biomass and less extended flowering window. The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height. Nonetheless, Palmer amaranth produced 21% less biomass in second cohort compared to first cohort timing. In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015]. These results suggest that crop competition is not the only factor limiting late Palmer amaranth establishment. The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to day length, light availability and thermal units (e.g, growing degree days). It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period. A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021]. In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.

Seed production was not evaluated due to plant harvest at initiation of flowering. Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018]. I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort. Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the cropping system. Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015]. Still, seed production at second cohort is likely to replenish the soil seedbank. Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987]. Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013]. Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize weed impact into agroecossystem [@davis2015].

An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering window. Our study suggests that Palmer amaranth flowering was slightly influenced by cropping systems and cohort timings. Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering window, respectively. When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing. Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965]. We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window. Early flower initiation is plant strategy when growing in stress conditions. For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021]. A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production and exchange of resistant alleles [@jhala2021]. Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].

The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888]. In our study, the gender model performance was decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930]. However, a slight deviation from 1:1 sex ratio might occur in some dioecious species. For example, the dioecious *Halophila stipulacea* is a female-biased plant in its native habitat, but the naturalized *H. stipulacea* have a 1:1 ratio [@nguyen2018]. Naturalized of *H. stipulacea* reduced female-male ratio to expand into its non-native habitat [@nguyen2018]. Also, biotic and/or abiotic stress can influence plant sex determination. Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019]. Our model estimated that late flowering, heavier and taller Palmer amaranth plants deviated from 1:1 ratio in favor to female plants. It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017]. Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each agroecossystem. Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth. Sexual dimorphism is documented in other dioecious species [@barrett2013]. For example, stronger female plant competition and greater male tolerance to herbivory was reported in *Spinacia oleracea* [@perez-llorca2019]. Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far to complete [@montgomery2021; @montgomery2019]. Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including the plant behavior under climate change.

Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems. Is likely that Palmer amaranth range will continue to expanding to new geographies. Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal. Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development. Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation [@oliveira2021a], plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g. cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems. The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on economy and sustainability.

# Disclosure/Conflict-of-Interest Statement {-}

<!--  
Frontiers follows the recommendations by the International Committee of Medical
Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
that all financial, commercial or other relationships that might be perceived by
the academic community as representing a potential conflict of interest must be
disclosed. If no such relationship exists, authors will be asked to declare that
the research was conducted in the absence of any commercial or financial
relationships that could be construed as a potential conflict of interest. When
disclosing the potential conflict of interest, the authors need to address the
following points:

 - Did you or your institution at any time receive payment or services from a
   third party for any aspect of the submitted work?
 - Please declare financial relationships with entities that could be perceived
   to influence, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 - Please declare patents and copyrights, whether pending, issued, licensed
   and/or receiving royalties relevant to the work.
 - Please state other relationships or activities that readers could perceive to
   have influenced, or that give the appearance of potentially influencing, what
   you wrote in the submitted work.
 -->

The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

# Author Contributions {-}

<!--

When determining authorship the following criteria should be observed:

 - Substantial contributions to the conception or design of the work; or the
   acquisition, analysis, or interpretation of data for the work; AND
 - Drafting the work or revising it critically for important intellectual
   content; AND
 - Final approval of the version to be published ; AND
 - Agreement to be accountable for all aspects of the work in ensuring that
   questions related to the accuracy or integrity of any part of the work are
   appropriately investigated and resolved.

Contributors who meet fewer than all 4 of the above criteria for authorship
should not be listed as authors, but they should be acknowledged.
(http://www.icmje.org/roles_a.html)

-->

RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research. All authors reviewed the manuscript.

# Acknowledgments

Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors

# Supplemental Data

Supplementary Material should be uploaded separately on submission, if there are
Supplementary Figures, please include the caption in the same file as the
figure. LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

# References

A Frontier article expect the reference list to be included in this section.
To make that happens, the below syntax can be used. This [feature is from Pandoc citeproc](https://pandoc.org/MANUAL.html#placement-of-the-bibliography) which is used with `frontier_article()` to handle the bibliography

::: {#refs}
:::

# Figures

```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit', out.width = "150mm", out.height = "70mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B). From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
# You can also refer to code chunks from above to place figures at the bottom.
```

### 2021-08-23 Frontiers/Frontiers.tex
as

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass[utf8]{frontiersSCNS}

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% BELOW TAKEN FROM rticles plos template
%
% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Below is from frontiers
%
\bibliographystyle{frontiersinSCNS}
% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\def\Authors{
  Maxwel C Oliveira\,\textsuperscript{1},
  Amit J Jhala\,\textsuperscript{2},
  Mark Bernards\,\textsuperscript{3},
  Chris Proctor\,\textsuperscript{2},
  Strahinja Stepanovic\,\textsuperscript{2},
  Rodrigo Werle\,\textsuperscript{1*}}

\def\Address{

  \textsuperscript{1} Department of Agronomy, University of
Wisconsin-Madison,  Madison,  Wisconsin,  United States

  \textsuperscript{2} Department of Agronomy and
Horticulture, University of
Nebraska-Lincoln,  Lincoln,  Nebraska,  United States

  \textsuperscript{3} Department of Agronomy, Western Illinois
University,  Macomb,  Illnois,  United States
  }

  \def\firstAuthorLast{Oliveira {et~al.}}

  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Department of Agronomy,
University of Wisconsin-Madison, United States\\1575 Linden
Dr\\Madison, Wisconsin, 53705 United
States}\def\corrEmail{\href{mailto:rwerle@uwisc.edu}{\nolinkurl{rwerle@uwisc.edu}}}

\begin{document}
\onecolumn
\firstpage{1}

\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
palmeri}) adaptation to US Midwest agroecosystems}
\author[\firstAuthorLast]{\Authors}
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

Abstract length and content varies depending on article type. Refer to 
\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
and length according to article type.

%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\tiny
 \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}

\end{abstract}

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
ranked as one of the most economically detrimental weed species to
cropping systems in the United States (Van Wychen, 2020). Unmanaged
Palmer amaranth plants compete for water, light, and nutrients, which
can drastically impact crop yields (Berger et al., 2015). For example,
Palmer amaranth has been documented to reduce up to 91\%, 68\%, and 54\%
corn (Massinga et al., 2001), soybean (Klingaman and Oliver, 1994), and
cotton (Morgan et al., 2001) yields, respectively. Moreover, Palmer
amaranth has shown a remarkable capacity to evolve resistance to
herbicides. To date, Palmer amaranth has evolved resistance to eight
herbicide sites of action (Heap, 2021), increasing the weed management
complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
economical and environmental risk to sustainable agriculture.

Palmer amaranth is a fast growing summer annual forb indigenous to the
Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
threat to US agriculture in the 1990s. Palmer amaranth weediness is
likely a result of human-assisted selection in combination with plant
biology. Farm mechanization, adoption of conservation agriculture (e.g.,
no-till), and reliance on herbicides for weed management are the main
human-mediated selection of Palmer amaranth into cropping systems (Ward
et al., 2013). On the other hand, Palmer amaranth is a prolific seed
producer with a C4 photosynthetic apparatus (Wang et al., 1992). With a
dioecy nature, Palmer amaranth male and female plants are obligate
outcrosser species, increasing the chances of exchanging adaptive traits
among plants (Oliveira et al., 2018). Also, Palmer amaranth small seed
size (e.g, 1 mm) tend to thrive in no-tillage systems (Price et al.,
2011), and spread across locations through farm equipment (Sauer, 1972),
manure (Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
dispersal capacity of Palmer amaranth makes the species one of the most
successful cases of weed adaptation to current cropping systems.

Light and temperature are the main environment requirements for Palmer
amaranth successful grow and development (Jha et al., 2010). Palmer
amaranth is reported with an extended germination period (Ward et al.,
2013). Germination of Palmer amaranth was triggered by 18 C soil
temperature at 5 cm depth (Keeley et al., 1987), and optimal germination
and biomass production occurred at 35/30 C day and night temperatures
(Guo and Al-Khatib, 2003). In addition, Palmer amaranth establishment is
human-mediated by tillage timings and preemergence-applied herbicides
(Chahal et al., 2021), which can result in weed germination shifts
(Sbatella and Wilson, 2010). Water has not shown to limit Palmer
amaranth fitness. Under continuous water stress, Palmer amaranth
survived and produced at least 14000 seeds plant-1 (Chahal et al.,
2018). Seeds from Palmer amaranth growing with limited water conditions
were heavier, less dormant, and prompt for germination (Matzrafi et al.,
2021). Growing conditions and management practices also influence Palmer
amaranth sex dimorphism and flowering pattern (Korres et al., 2017;
Rumpa et al., 2019). Therefore, Palmer amaranth has shown plasticity to
evolve and fast adapt under selection pressure. Palmer amaranth
invasion/adaptation into new habitats is likely to increase with the
global temperature warming. Currently, it is estimated that the greatest
climatic risk of Palmer amaranth establishment are agronomic crops in
Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
Temperature is a key factor limiting Palmer amaranth expansion to cooler
geographies (Briscoe Runquist et al., 2019); however, under future
climate change Palmer amaranth is likely to expand northward into Canada
and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
al., 2019).

Palmer amaranth is already found in agronomic crops of South America
(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
et al., 2021). In the United States, Palmer amaranth is well established
in the Cotton Belt (Garetson et al., 2019; Bagavathiannan and
Norsworthy, 2016) in the southern United States but its range is
expanding northward. For example, herbicide resistant Palmer amaranth is
widespread in Nebraska (Oliveira et al., 2021). There are some reported
cases of Palmer amaranth in Michigan (Kohrt et al., 2017) and
Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
amaranth can cause high damage to soybean fields in Illinois (Davis et
al., 2015), which is concerning as soybean along with corn make most of
US Midwest agronomic hectares. In Iowa, a study showed that Palmer
amaranth is still not well adapted compared to waterhemp
(\emph{Amaranthus tuberculatus}) (Baker, 2021). Invasion and successful
eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
2021). Palmer amaranth infestations have not been detected in Canada;
however, Palmer amaranth seeds were detected in sweet potato slips in
the country (Page et al., 2021). Palmer amaranth is still not as well
adapted and established to Northern as it is in the Southern North
America. Therefore, its range of expansion into new habitats is
increasing. It seems fated the need to manage Palmer amaranth in
agronomic crops throughout multiple environments in the near future.
Strategies on Palmer amaranth management should encompass the
agroecosystem level but not only attempts to eradicate the weed. Most
tactics to manage Palmer amaranth are based on technology fixes (Scott,
2011), which are short-term (e.g., herbicide and/or tillage) rather than
long-term weed management. Palmer amaranth management should be built on
minimizing the species ability to adapt, grow and develop into
agroecossystems.

In the southeastern US, early growing Palmer amaranth is well known to
have a higher impact on cotton yields compared to late established
plants (MacRae et al., 2013). In the northern states, Palmer amaranth
impact on the agroecosystem is recent. Studies investigating Palmer
amaranth in those locations are limited due to the plant classification
as noxious weed species in some northern states (Yu et al., 2021).
Nonetheless, the continuous Palmer amaranth dispersal and potential
establishment across the northern United States is concerning and
warrants investigations on species morphology in such environments.
Understanding Palmer amaranth biology and growing strategies under
different agroecosystems can enhance our knowledge on species adaptation
and management practices. It can also aid in designing proactive and
ecological tactics to limit the species range expansion, reduce its
negative impact, and developing resilient and sustainable farming
systems (MacLaren et al., 2020). Therefore, the objective of this study
was to investigate the flowering pattern, gender, biomass production,
and height of Palmer amaranth cohorts growing under corn, soybean and
bareground across five locations in the United States Midwest.

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

\hypertarget{plant-material-and-growing-conditions}{%
\subsection*{Plant material and growing
conditions}\label{plant-material-and-growing-conditions}}
\addcontentsline{toc}{subsection}{Plant material and growing conditions}

A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with
no reported herbicide resistance according to Oliveira et al. (2021) was
selected for this study. Three weeks prior to the onset of the field
experiments, seeds were planted in plastic trays containing potting-mix.
Emerged seedlings (1 cm) were transplanted into 200 cm-3 plastic pots (a
plant pot-1). Palmer amaranth seedlings were supplied with adequate
water and kept under greenhouse conditions at the University of
Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois
University; and kept outdoors in the Perkins extension office at Grant,
NE until the 2-3 leaf stage (5 to 8 cm height) when they were
transported to the field.

\hypertarget{field-study}{%
\subsection*{Field study}\label{field-study}}
\addcontentsline{toc}{subsection}{Field study}

The experiment was conducted in 2018 and 2019 under field conditions at
five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N,
'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N,
'W)).

Fields were conventionally tilled prior to crop planting. Corn hybrid
and soybean varieties were planted in 76-cm row spacing (Table 1).
Monthly mean air temperature and total precipitation were obtained using
Daymet weather data from June through September across the five
locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Field study attributes }
\centering
\resizebox{\linewidth}{!}{
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lllllll}
\toprule
Attributes &  & Arlington, WI & Clay Center, NE & Grant, NE & Lincoln, NE & Macomb, IL\\
\midrule
Bareground &  &  &  &  &  & \\

 & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

Corn &  &  &  &  &  & \\

 & Hybrid &  & DKC60-67 &  &  & \\

 & Seeding rate &  & 86487 &  &  & \\

 & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\

 & Stage 1 cohort & V2-3 &  &  &  & \\

\multirow{-5}{*}{\raggedright\arraybackslash } & Stage 2 cohort & V6-7 &  &  &  & \\

Soybean &  &  &  &  &  & \\

 & Variety & DSR-1950 & AG21X8 &  &  & \\

 & Seeding rate & 296400 & 321237 &  &  & \\

 & Stage 1 cohort & V1-2 &  &  &  & \\

 & Stage 2 cohort & V5-6 &  &  &  & \\

\multirow{-5}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\

Planting day &  & May 10 to 20 & May 10 to 14 &  &  & \\

Soil &  &  &  &  &  & \\

 & Type &  & Crete Silt Loam &  &  & \\

 & Ratio (sand-clay-silt) &  & 58-25-6.5 &  &  & \\

 & pH & 6.6 & 6.5 &  &  & \\

\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) &  & 3 &  &  & \\
\bottomrule
\end{tabular}}
\end{table}

The field experimental units were three adjacent 9.1 m wide (12 rows at
76.2 cm row spacing) by 10.7 m long. The experimental design were
arranged in factorial design with three crops, two transplanting times
simulating two cohorts, repeated across five locations. Each
experimental unit was planted with corn, soybean, or kept under
bareground. The two transplanting timings were June 1 (first cohort) and
July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
plants were equidistantly placed (0.76 m apart) between rows within each
crop (Figure 2). After a week, one plant was eliminated and one was
kept, resulting in 24 plants per experimental unit and transplanting
time (Figure 2). When needed, Palmer amaranth plants were supplied with
water during the first week after transplanting.

After transplanting, Palmer amaranth flowering was monitored until the
end of the study. When a plant flowered, the day was recorded, plant sex
was identified (male or female), plant height was measured from soil
surface to the top of plant. Also, aboveground plant biomass was harvest
near soil surface and oven dried at 65 C until reaching constant weight
before weighing (g plant-1 was recorded).

Plants had to be harvested at flowering because Palmer amaranth is
neither endemic in Wisconsin nor in Illinois. In our study, all
locations follow the methodology of plant harvest at flowering
initiation, except in Grant, NE. In this location, all Palmer amaranth
plants were harvest at once on July 06, 2018 and 2019 (first cohort),
and on August 17, 2018 and on July 31, 2019 (second cohort).

\hypertarget{statistical-analyses}{%
\subsection*{Statistical analyses}\label{statistical-analyses}}
\addcontentsline{toc}{subsection}{Statistical analyses}

The statistical analyses were performed using R statistical software
version 4.0.1.

Analyses of Palmer amaranth height and biomass were performed with a
linear mixed model using \emph{lmer} function from ``lme4'' package
(Bates et al., 2015). Plant height and biomass were log transformed to
meet model assumption of normality. In the model, crop (bareground,
corn, soybean) was the fixed effect and year nested with location the
random effects. Analysis of variance was performed with \emph{anova}
function from ``car'' package (Fox and Weisberg, 2018). Marginal means
and compact letter display were estimated with \emph{emmeans} and
\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
2008).

The Palmer amaranth flowering timing was estimated as cumulative
flowering across all location. Dataset was nested but id did not include
Grant, NE. Palmer amaranth cumulative flowering estimation was
determined using an asymmetrical three parameter log logistic Weibull
model of the drc package (Ritz et al., 2015).

\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\]

In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
\emph{d} is the upper limit (set to 100), and \emph{e} is the XXX, and
\emph{x} day of year (doy).

The doy for 10, 50, and 90\% Palmer amaranth cumulative flowering were
determined using the \emph{ED} function of drc package. Also, the 10,
50, and 90\% Palmer amaranth cumulative flowering were compared among
crops and cohorts using the \emph{EDcomp} function of drc package. The
EDcomp function compares the ratio of cumulative flowering using
t-statistics, where P-value \textless{} 0.05 indicates that we fail to
reject the null hypothesis.

Palmer amaranth sex was fitted to a binary logistic regression
(Bangdiwala, 2018). Binary logistic regression is used for predicting
binary classes, such as the probability of a plant being female in a
dioecious species. Prior to the analysis, all missing values were
removed from the dataset. Also, data from Grant was not used in this
analysis due to the uniform plant harvesting at that location. The
complete dataset was splitted into 80\% train and 20\% test data. The
80\% train is used for the model training and the 20\% test is used for
checking how the model generalized on unseen dataset. With 80\% dataset,
a binary response variable, male (0) and female (1), was fitted to a
generalized linear model (\emph{glm} function) including day of year
harvest, height, weight, crop and month as independent variables
(without interaction). The model family was binomial with a logit
function. The model fit was assessed through pseudo R-squared values
(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
\emph{nagelkerke} function (``rcompanion'' package). The marginal
effects computation was performed with Average Marginal Effects (AMEs)
at every observed value of x and average across the results (Leeper,
2017) using \emph{margins} function from ``margins'' package. The rest
20\% dataset was predicted using \emph{predict} function with a cutoff
estimation for male or female using \emph{performance} function. The
model quality prediction from the classification algorithm was measured
with precision (\emph{precision} function), recall (\emph{recall}
function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
package. The precision determines the accuracy of positive predictions
(female plants), recall determines the fraction of positives that were
correctly identified, and F1-score is a weighted harmonic mean of
precision and recall with the best score of 1 and the worst score of 0.
F1-score conveys the balance between the precision and the recall
(Yacouby and Axman, 2020).

\hypertarget{results-1}{%
\section*{Results}\label{results-1}}
\addcontentsline{toc}{section}{Results}

\hypertarget{palmer-amaranth-height-and-biomass}{%
\subsection*{Palmer amaranth height and
biomass}\label{palmer-amaranth-height-and-biomass}}
\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}

Palmer amaranth plants accumulated more biomass when growing in
bareground compared to plants growing in soybean and corn (figure 3A).
Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and
16.3 g plant-1 in bareground, soybean and corn, respectively. Plants
from the second cohort produced 62.6 g plant in bareground, followed by
6.3 g plant in soybean, and 1.4 g plant in corn.

Palmer amaranth height was more uniform across cohort timings, except
when growing in corn (figure 3B). Palmer amaranth plants from the first
cohort were on average 69.2 cm tall in bareground, which was not
different from the 70.7 cm tall plants from the second cohort timing (P
= 0.74). In addition, no difference in Palmer amaranth height (69.3 cm)
was detected from first cohort plants in soybean to first and second
cohort plants in bareground (P \textgreater{} 0.75). Palmer amaranth
plants from the second cohort were nearly 10 cm lower compared to the
first cohort in soybeans (P = 0.04). The tallest (first cohort) and
smallest (second cohort) Palmer amaranth plants were found in corn.
Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.

\hypertarget{palmer-amaranth-cumulative-flowering}{%
\subsection*{Palmer amaranth cumulative
flowering}\label{palmer-amaranth-cumulative-flowering}}
\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}

Palmer amaranth plants from the first cohort growing in corn resulted in
a longer flowering window compared to plants growing in bareground and
soybean (Figure 4A). The 10\% cumulative Palmer amaranth flowering in
soybean, bareground and corn occurred at the end of June. Palmer
amaranth reached 10\% flowering in soybean, bareground and corn at doy
180, 180.9 and 181.7, respectively. The 50\% Palmer amaranth cumulative
flowering occurred in July. Palmer amaranth reached 50\% flowering in
bareground, soybean and corn at doy 193.4, 194.8, and 206.6,
respectively. Similar trend was observed at 90\% Palmer amaranth
cumulative flowering. Palmer amaranth reached 90\% flowering at doy
252.6 in corn (early September), which was 38 and 32 days after reaching
90\% flowering in bareground and soybean, respectively.

Palmer amaranth cumulative flowering at the second cohort ranged from
mid July to mid September (Figure 4B). Palmer amaranth growing in the
bareground resulted in earlier flowering time compared to soybean and
corn. Palmer amaranth growing in bareground reached 10\%, 50\%, and 90\%
flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer
amaranth growing in soybean reached 10\% flowering at doy 210.9, which
was 6 days prior to corn (\emph{P}-value = 0.00). Similar trend was
observed at 50\% flowering, whereas Palmer amaranth reached 50\%
flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; \emph{P}
= 0.00). The 90\% Palmer amaranth cumulative flowering occurred at same
day in corn (260.9) and soybean (260.5; \emph{P} = 0.66).

\hypertarget{palmer-amaranth-gender}{%
\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
\addcontentsline{toc}{subsection}{Palmer amaranth gender}

The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from
McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
likelihood ratio test showed a p-value of \textless{} 0.00. The average
marginal effects showed that Palmer amaranth growing in corn resulted in
14.8\% less females plants (Table 2). Moreover, increasing a unit doy
increases the probability of having a female plant by 0.4\% (Table 2 and
Figure4A). Similar trend is observed for weight as well as height,
whereas the probability of being female increase by 0.2\% (Figure 4B)
and 0.1\% (Figure 4C) when a unit of weight (g) and height (cm)
increases, respectively.

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrrr}
\toprule
Term & AME & SE & Lower & Upper & Z-score & P-value\\
\midrule
crop\_bareground & -0.048 & 0.054 & -0.154 & 0.059 & -0.876 & 0.381\\

crop\_corn & -0.148 & 0.052 & -0.250 & -0.046 & -2.842 & 0.004\\

doyh & 0.004 & 0.001 & 0.003 & 0.006 & 4.959 & 0.000\\

height & 0.002 & 0.001 & 0.001 & 0.003 & 2.953 & 0.003\\

weight & 0.001 & 0.000 & 0.000 & 0.001 & 2.179 & 0.029\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} Average Marginal Effects. \textsuperscript{b} Standard Error.}\\
\end{tabular}
\end{table}

The model accuracy evaluation accuracy in the 20\% test dataset was 0.62
with a cutoff value for female and male plants of 0.43. The model
classification showed a precision of 0.64, recall of 0.66, and a F means
score of 0.65. In addition, the area under the curve was 0.64.

\hypertarget{discussion}{%
\section*{Discussion}\label{discussion}}
\addcontentsline{toc}{section}{Discussion}

Our study showed that Palmer amaranth biomass, height, flowering pattern
and gender varied within agroecossystems and cohort timings. In general,
first cohort of Palmer amaranth plants were heavier and taller when
compared to the second cohort. At first cohort, resources (e.g., soil
nutrients) and conditions (e.g., light) were more timely available for
the species. High biomass and taller Palmer amaranth plants are likely a
weed strategy to compete for light in between crop rows in absence of
canopy. In such conditions, Palmer amaranth showed an extraordinary
plasticity to adapt upon the agroecosystem. This is evident when
comparing Palmer amaranth canopy shape, and its extended flowering
pattern when growing into corn compared to soybean. The Palmer amaranth
competition strategy was to mimic the crop grow and development (Figure
6). These results suggests that Palmer amaranth can fast evolve
life-history traits to adapt into cropping systems and cultural
practices, which was also showed in a study varying nitrogen
fertilization (Bravo et al., 2018). Our results highlight the Palmer
amaranth as a threat to field crops as breeding more competitive crop
varieties is likely to select more competitive biotypes (Bravo et al.,
2017).

Palmer amaranth grow and development in second cohort was limited due to
the crop competitive ability at advanced development stages. Palmer
amaranth was transplanted when corn canopy was nearly closed, which
reduced Palmer amaranth competitiveness. As a result, Palmer amaranth
height and biomass was lower compared to its first cohort. Under crop
canopy (e.g., second cohort), Palmer amaranth flowering window was near
to similar in corn and soybean. Palmer amaranth growing without crop
competition produced the highest amounts of biomass and less extended
flowering window. The Palmer amaranth strategy in bareground was to
invest biomass in growing plant width and height. Nonetheless, Palmer
amaranth produced 21\% less biomass in second cohort compared to first
cohort timing. In a bareground study, early emerged Palmer amaranth
without competition was 50\% taller than late emerged plants (Webster
and Grey, 2015). These results suggest that crop competition is not the
only factor limiting late Palmer amaranth establishment. The limited
growth of Palmer amaranth at second cohort is likely a reduced plant
response to day length, light availability and thermal units (e.g,
growing degree days). It is hypothesize that reduced day length
contributed to smaller plants at second cohort as well as shorter
flowering period. A study in North Carolina and Illinois predicted that
less than 10\% Palmer amaranth seedlings emergence occurred after June
(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
yields was higher when plants were established near to crop planting.

Seed production was not evaluated due to plant harvest at initiation of
flowering. Nonetheless, it is well documented a strong positive
correlation between Palmer amaranth biomass and seed production
(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
amaranth growing at first cohort accumulated an overall 36\% more
biomass when compared to second cohort. Therefore, Palmer amaranth
plants growing in the second cohort is likely to produce less seeds
regardless the cropping system. Our observation is consistent with the
findings that first Palmer amaranth cohort produced 50\% more seeds per
plant than Palmer amaranth plants established six weeks later in
bareground (Webster and Grey, 2015). Still, seed production at second
cohort is likely to replenish the soil seedbank. Seed production and
deposition in the seedbank is also a key factor for species perpetuation
(Menges, 1987). Palmer amaranth can produce hundred thousands seeds per
plant (Schwartz et al., 2016; Keeley et al., 1987), and stay viable
buried in the seedbank for at least 36 months (Sosnoskie et al., 2013).
Therefore, preventing Palmer amaranth seed production or/and seed
migration to its non-native habitat is an essential strategy to minimize
weed impact into agroecossystem (Davis et al., 2015).

An ecological approach to reduce seed production in Palmer amaranth is
understanding its flowering window. Our study suggests that Palmer
amaranth flowering was slightly influenced by cropping systems and
cohort timings. Palmer amaranth growing in bareground and corn resulted
in the overall shortest and longest flowering window, respectively. When
growing in soybean, Palmer amaranth flowering window was similar to
bareground at first cohort but similar to corn at second cohort timing.
Plant flowering initiation is complex and depends on biological and
ecological factors (Lang, 1965). We hypothesize that when growing in
high competition (e.g., second cohort), Palmer amaranth plants tend to
initiate flowering early, as well as having an extended flowering
window. Early flower initiation is plant strategy when growing in stress
conditions. For example, when growing under water stress, early
flowering in Palmer amaranth resulted in a mismatch between female and
male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
amaranth male and female flowering period can minimize plant outcross,
and thus reduce plant seed production and exchange of resistant alleles
(Jhala et al., 2021). Sex dimorphism manipulation is considered a
potential ecological pest control (McFarlane et al., 2018; Schliekelman
et al., 2005).

The mechanisms of sex-determination in plant species is intriguing and
arouse the curiosity of many scientists, including Charles Darwin
(Darwin, 1888). In our study, the gender model performance was decent
(AIC 0.64) considering the biology of plant flowering. A 1:1 male and
female sex ratio is a general evolutionary stable strategy for plant
species perpetuation (Fisher, 1930). However, a slight deviation from
1:1 sex ratio might occur in some dioecious species. For example, the
dioecious \emph{Halophila stipulacea} is a female-biased plant in its
native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
female-male ratio to expand into its non-native habitat (Nguyen et al.,
2018). Also, biotic and/or abiotic stress can influence plant sex
determination. Palmer amaranth male-to-female ratio was greater under
high plant densities (Korres and Norsworthy, 2017) and after herbicide
application (Rumpa et al., 2019). Our model estimated that late
flowering, heavier and taller Palmer amaranth plants deviated from 1:1
ratio in favor to female plants. It was reported that female Palmer
amaranth plants invested more in height, stem and biomass while male
invested more in leaf area and leaf dry weight under nutrient deficiency
(Korres et al., 2017). Our model also estimated more female plants in
soybean and bareground compared to corn, which might linked to plant
competition strategy in each agroecossystem. Our results showed the
influence of life-history and ecological traits on sexual dimorphism in
Palmer amaranth. Sexual dimorphism is documented in other dioecious
species (Barrett and Hough, 2013). For example, stronger female plant
competition and greater male tolerance to herbivory was reported in
\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
Research on candidate genes for sex determination in \emph{Amaranthus}
species are currently underway but it is far to complete (Montgomery et
al., 2021, 2019). Further studies are also needed to understand the
ecological basis of Palmer amaranth flowering, including the plant
behavior under climate change.

Our study demonstrated the short-term Palmer amaranth plasticity to grow
and develop into cropping-systems. Is likely that Palmer amaranth range
will continue to expanding to new geographies. Therefore, preventive
management is a priority to minimizing Palmer amaranth dispersal.
Reactive management should focus on early-season management programs,
which would have a large negative effect on Palmer amaranth growth and
development. Tactics that promote early-season crop advantage against
Palmer amaranth, including early crop planting, crop rotation (Oliveira
et al., 2021), plant width, preemergence applied herbicide (Sanctis et
al., 2021), and crop residue (e.g.~cover crops) would minimize the
negative impact of Palmer amaranth in agroecossystems. The
aggressiveness and differential Palmer amaranth adaptation to
agroecosystem is striking and require national efforts to minimize the
species impact on economy and sustainability.

\hypertarget{disclosureconflict-of-interest-statement}{%
\section*{Disclosure/Conflict-of-Interest
Statement}\label{disclosureconflict-of-interest-statement}}
\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
Statement}

The authors declare that the research was conducted in the absence of
any commercial or financial relationships that could be construed as a
potential conflict of interest.

\hypertarget{author-contributions}{%
\section*{Author Contributions}\label{author-contributions}}
\addcontentsline{toc}{section}{Author Contributions}

RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP,
MB, MO, SS, and RW: conceptualized the research. All authors reviewed
the manuscript.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

Funding: This work received no specific grant from any funding agency,
commercial, or not-for-profit sectors

\hypertarget{supplemental-data}{%
\section{Supplemental Data}\label{supplemental-data}}

Supplementary Material should be uploaded separately on submission, if
there are Supplementary Figures, please include the caption in the same
file as the figure. LaTeX Supplementary Material templates can be found
in the Frontiers LaTeX folder

\hypertarget{references}{%
\section{References}\label{references}}

A Frontier article expect the reference list to be included in this
section. To make that happens, the below syntax can be used. This
\href{https://pandoc.org/MANUAL.html\#placement-of-the-bibliography}{feature
is from Pandoc citeproc} which is used with \texttt{frontier\_article()}
to handle the bibliography

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-aulakh2021}{}%
Aulakh, J. S., Chahal, P. S., Kumar, V., Price, A. J., and Guillard, K.
(2021). Multiple herbicide-resistant {Palmer} amaranth ({Amaranthus}
palmeri) in {Connecticut}: Confirmation and response to {POST}
herbicides. \emph{Weed Technology} 35, 457--463.
doi:\href{https://doi.org/10.1017/wet.2021.6}{10.1017/wet.2021.6}.

\leavevmode\hypertarget{ref-bagavathiannan2016}{}%
Bagavathiannan, M. V., and Norsworthy, J. K. (2016). Multiple-{Herbicide
Resistance Is Widespread} in {Roadside Palmer Amaranth Populations}.
\emph{PLOS ONE} 11, e0148748.
doi:\href{https://doi.org/10.1371/journal.pone.0148748}{10.1371/journal.pone.0148748}.

\leavevmode\hypertarget{ref-baker2021}{}%
Baker, R. (2021). Comparative analysis of {Palmer} amaranth
({Amaranthus} palmeri) and waterhemp ({A}. Tuberculatus) in {Iowa}.
doi:\href{https://doi.org/10.31274/etd-20210609-11}{10.31274/etd-20210609-11}.

\leavevmode\hypertarget{ref-bangdiwala2018}{}%
Bangdiwala, S. I. (2018). Regression: Binary logistic.
\emph{International Journal of Injury Control and Safety Promotion} 25,
336--338.
doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.

\leavevmode\hypertarget{ref-barrett2013}{}%
Barrett, S. C. H., and Hough, J. (2013). Sexual dimorphism in flowering
plants. \emph{Journal of Experimental Botany} 64, 67--82.
doi:\href{https://doi.org/10.1093/jxb/ers308}{10.1093/jxb/ers308}.

\leavevmode\hypertarget{ref-bates2015}{}%
Bates, D., Mächler, M., Bolker, B., and Walker, S. (2015). Fitting
{Linear Mixed}-{Effects Models Using} Lme4. \emph{Journal of Statistical
Software} 67, 1--48.
doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.

\leavevmode\hypertarget{ref-berger2015}{}%
Berger, S. T., Ferrell, J. A., Rowland, D. L., and Webster, T. M.
(2015). Palmer {Amaranth} ({Amaranthus} palmeri) {Competition} for
{Water} in {Cotton}. \emph{Weed Science} 63, 928--935.
doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.

\leavevmode\hypertarget{ref-bravo2017}{}%
Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
(2017). Differentiation of {Life}-{History Traits} among {Palmer
Amaranth Populations} ({Amaranthus} palmeri) and {Its Relation} to
{Cropping Systems} and {Glyphosate Sensitivity}. \emph{Weed Science} 65,
339--349.
doi:\href{https://doi.org/10.1017/wsc.2017.14}{10.1017/wsc.2017.14}.

\leavevmode\hypertarget{ref-bravo2018}{}%
Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
(2018). Evolutionary {Adaptations} of {Palmer Amaranth} ({Amaranthus}
palmeri) to {Nitrogen Fertilization} and {Crop Rotation History Affect
Morphology} and {Nutrient}-{Use Efficiency}. \emph{Weed Science} 66,
180--189.
doi:\href{https://doi.org/10.1017/wsc.2017.73}{10.1017/wsc.2017.73}.

\leavevmode\hypertarget{ref-briscoerunquist2019}{}%
Briscoe Runquist, R. D., Lake, T., Tiffin, P., and Moeller, D. A.
(2019). Species distribution models throughout the invasion history of
{Palmer} amaranth predict regions at risk of future invasion and reveal
challenges with modeling rapidly shifting geographic ranges. \emph{Sci
Rep} 9, 2426.
doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.

\leavevmode\hypertarget{ref-chahal2021}{}%
Chahal, P. S., Barnes, E. R., and Jhala, A. J. (2021). Emergence pattern
of {Palmer} amaranth ({Amaranthus} palmeri) influenced by tillage
timings and residual herbicides. \emph{Weed Technology} 35, 433--439.
doi:\href{https://doi.org/10.1017/wet.2020.136}{10.1017/wet.2020.136}.

\leavevmode\hypertarget{ref-chahal2018}{}%
Chahal, P. S., Irmak, S., Jugulam, M., and Jhala, A. J. (2018).
Evaluating {Effect} of {Degree} of {Water Stress} on {Growth} and
{Fecundity} of {Palmer} amaranth ({Amaranthus} palmeri) {Using Soil
Moisture Sensors}. \emph{Weed Science} 66, 738--745.
doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.

\leavevmode\hypertarget{ref-correndo2021}{}%
Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
Retrieving and processing agro-meteorological data from {API}-client
sources using {R} software. \emph{BMC Research Notes} 14, 205.
doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.

\leavevmode\hypertarget{ref-darwin1888}{}%
Darwin, C. (1888). \emph{The {Different Forms} of {Flowers} on {Plants}
of the {Same Species}}. {J. Murray} Available at:
\url{http://books.google.com?id=7uMEAAAAYAAJ}.

\leavevmode\hypertarget{ref-davis2015}{}%
Davis, A. S., Schutte, B. J., Hager, A. G., and Young, B. G. (2015).
Palmer {Amaranth} ({Amaranthus} palmeri) {Damage Niche} in {Illinois
Soybean Is Seed Limited}. \emph{Weed Science} 63, 658--668.
doi:\href{https://doi.org/10.1614/WS-D-14-00177.1}{10.1614/WS-D-14-00177.1}.

\leavevmode\hypertarget{ref-farmer2017}{}%
Farmer, J. A., Webb, E. B., Pierce, R. A., and Bradley, K. W. (2017).
Evaluating the potential for weed seed dispersal based on waterfowl
consumption and seed viability. \emph{Pest Management Science} 73,
2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.

\leavevmode\hypertarget{ref-fisher1930}{}%
Fisher, R. A. (1930). The genetical theory of natural selection.
\emph{Eugen Rev} 22, 127--130. Available at:
\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2984947/} {[}Accessed
August 12, 2021{]}.

\leavevmode\hypertarget{ref-fox2018}{}%
Fox, J., and Weisberg, S. (2018). \emph{An {R Companion} to {Applied
Regression}}. {SAGE Publications} Available at:
\url{http://books.google.com?id=uPNrDwAAQBAJ}.

\leavevmode\hypertarget{ref-garetson2019}{}%
Garetson, R., Singh, V., Singh, S., Dotray, P., and Bagavathiannan, M.
(2019). Distribution of herbicide-resistant {Palmer} amaranth
({Amaranthus} palmeri) in row crop production systems in {Texas}.
\emph{Weed Technology} 33, 355--365.
doi:\href{https://doi.org/10.1017/wet.2019.14}{10.1017/wet.2019.14}.

\leavevmode\hypertarget{ref-guo2003}{}%
Guo, P., and Al-Khatib, K. (2003). Temperature effects on germination
and growth of redroot pigweed ({Amaranthus} retroflexus), {Palmer}
amaranth ({A}. Palmeri), and common waterhemp ({A}. rudis). \emph{Weed
Science} 51, 869--875.
doi:\href{https://doi.org/10.1614/P2002-127}{10.1614/P2002-127}.

\leavevmode\hypertarget{ref-hartzler2016}{}%
Hartzler, B., and Anderson, M. (2016). Palmer amaranth: {It}'s here, now
what? 10.

\leavevmode\hypertarget{ref-heap2021}{}%
Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
July 26, 2021{]}.

\leavevmode\hypertarget{ref-hothorn2008}{}%
Hothorn, T., Bretz, F., and Westfall, P. (2008). Simultaneous
{Inference} in {General Parametric Models}. \emph{Biometrical Journal}
50, 346--363.
doi:\href{https://doi.org/10.1002/bimj.200810425}{10.1002/bimj.200810425}.

\leavevmode\hypertarget{ref-jha2010}{}%
Jha, P., Norsworthy, J. K., Riley, M. B., and Bridges, W. (2010). Annual
{Changes} in {Temperature} and {Light Requirements} for {Germination} of
{Palmer Amaranth} ({Amaranthus} palmeri) {Seeds Retrieved} from {Soil}.
\emph{Weed Science} 58, 426--432.
doi:\href{https://doi.org/10.1614/WS-D-09-00038.1}{10.1614/WS-D-09-00038.1}.

\leavevmode\hypertarget{ref-jhala2021}{}%
Jhala, A. J., Norsworthy, J. K., Ganie, Z. A., Sosnoskie, L. M., Beckie,
H. J., Mallory-Smith, C. A., Liu, J., Wei, W., Wang, J., and
Stoltenberg, D. E. (2021). Pollen-mediated gene flow and transfer of
resistance alleles from herbicide-resistant broadleaf weeds. \emph{Weed
Technology} 35, 173--187.
doi:\href{https://doi.org/10.1017/wet.2020.101}{10.1017/wet.2020.101}.

\leavevmode\hypertarget{ref-keeley1987}{}%
Keeley, P. E., Carter, C. H., and Thullen, R. J. (1987). Influence of
{Planting Date} on {Growth} of {Palmer Amaranth} ({Amaranthus} palmeri).
\emph{Weed Science} 35, 199--204.
doi:\href{https://doi.org/10.1017/S0043174500079054}{10.1017/S0043174500079054}.

\leavevmode\hypertarget{ref-kistner2018}{}%
Kistner, E. J., and Hatfield, J. L. (2018). Potential {Geographic
Distribution} of {Palmer Amaranth} under {Current} and {Future
Climates}. \emph{Agricultural \& Environmental Letters} 3, 170044.
doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.

\leavevmode\hypertarget{ref-klingaman1994}{}%
Klingaman, T. E., and Oliver, L. R. (1994). Palmer {Amaranth}
({Amaranthus} palmeri) {Interference} in {Soybeans} ({Glycine} max).
\emph{Weed Science} 42, 523--527.
doi:\href{https://doi.org/10.1017/S0043174500076888}{10.1017/S0043174500076888}.

\leavevmode\hypertarget{ref-kohrt2017}{}%
Kohrt, J. R., Sprague, C. L., Nadakuduti, S. S., and Douches, D. (2017).
Confirmation of a {Three}-{Way} ({Glyphosate}, {ALS}, and {Atrazine})
{Herbicide}-{Resistant Population} of {Palmer Amaranth} ({Amaranthus}
palmeri) in {Michigan}. \emph{Weed Science} 65, 327--338.
doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.

\leavevmode\hypertarget{ref-korres2017a}{}%
Korres, N. E., and Norsworthy, J. K. (2017). Palmer {Amaranth}
({Amaranthus} palmeri) {Demographic} and {Biological Characteristics} in
{Wide}-{Row Soybean}. \emph{Weed Science} 65, 491--503.
doi:\href{https://doi.org/10.1017/wsc.2017.12}{10.1017/wsc.2017.12}.

\leavevmode\hypertarget{ref-korres2017}{}%
Korres, N. E., Norsworthy, J. K., FitzSimons, T., Roberts, T. L., and
Oosterhuis, D. M. (2017). Differential {Response} of {Palmer Amaranth}
({Amaranthus} palmeri) {Gender} to {Abiotic Stress}. \emph{Weed Science}
65, 213--227.
doi:\href{https://doi.org/10.1017/wsc.2016.34}{10.1017/wsc.2016.34}.

\leavevmode\hypertarget{ref-korres2020}{}%
Korres, N. E., Norsworthy, J. K., Mauromoustakos, A., and Williams, M.
M. (2020). Soybean density and {Palmer} amaranth ({Amaranthus} palmeri)
establishment time: Effects on weed biology, crop yield, and economic
returns. \emph{Weed Science} 68, 467--475.
doi:\href{https://doi.org/10.1017/wsc.2020.41}{10.1017/wsc.2020.41}.

\leavevmode\hypertarget{ref-kupper2017}{}%
Küpper, A., Borgato, E. A., Patterson, E. L., Netto, A. G., Nicolai, M.,
Carvalho, S. J. P. de, Nissen, S. J., Gaines, T. A., and Christoffoleti,
P. J. (2017). Multiple {Resistance} to {Glyphosate} and {Acetolactate
Synthase Inhibitors} in {Palmer Amaranth} ({Amaranthus} palmeri)
{Identified} in {Brazil}. \emph{Weed Science} 65, 317--326.
doi:\href{https://doi.org/10.1017/wsc.2017.1}{10.1017/wsc.2017.1}.

\leavevmode\hypertarget{ref-lang1965}{}%
Lang, A. (1965). {``Physiology of flower initiation,''} in
\emph{Differenzierung und {Entwicklung} / {Differentiation} and
{Development}} Handbuch der {Pflanzenphysiologie} / {Encyclopedia} of
{Plant Physiology}., ed. A. Lang ({Berlin, Heidelberg}: {Springer}),
1380--1536.
doi:\href{https://doi.org/10.1007/978-3-642-50088-6_39}{10.1007/978-3-642-50088-6\_39}.

\leavevmode\hypertarget{ref-larran2017}{}%
Larran, A. S., Palmieri, V. E., Perotti, V. E., Lieber, L., Tuesca, D.,
and Permingeat, H. R. (2017). Target-site resistance to acetolactate
synthase ({ALS})-inhibiting herbicides in {Amaranthus} palmeri from
{Argentina}. \emph{Pest Management Science} 73, 2578--2584.
doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.

\leavevmode\hypertarget{ref-leeper2017}{}%
Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
Marginal Eﬀects} with {R}'s margins. 31.

\leavevmode\hypertarget{ref-lindsay2017}{}%
Lindsay, K., Popp, M., Norsworthy, J., Bagavathiannan, M., Powles, S.,
and Lacoste, M. (2017). {PAM}: {Decision Support} for {Long}-{Term
Palmer Amaranth} ({Amaranthus} palmeri) {Control}. \emph{Weed
Technology} 31, 915--927.
doi:\href{https://doi.org/10.1017/wet.2017.69}{10.1017/wet.2017.69}.

\leavevmode\hypertarget{ref-maclaren2020}{}%
MacLaren, C., Storkey, J., Menegat, A., Metcalfe, H., and
Dehnen-Schmutz, K. (2020). An ecological future for weed science to
sustain crop production and the environment. {A} review. \emph{Agron.
Sustain. Dev.} 40, 24.
doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.

\leavevmode\hypertarget{ref-macrae2013}{}%
MacRae, A. W., Webster, T. M., Sosnoskie, L. M., Culpepper, A. S., and
Kichler, J. M. (2013). Cotton {Yield Loss Potential} in {Response} to
{Length} of {Palmer Amaranth} ({Amaranthus} palmeri) {Interference}. 17,
6.

\leavevmode\hypertarget{ref-massinga2001}{}%
Massinga, R. A., Currie, R. S., Horak, M. J., and Boyer, J. (2001).
Interference of {Palmer} amaranth in corn. \emph{Weed Science} 49,
202--208.
doi:\href{https://doi.org/10.1614/0043-1745(2001)049\%5B0202:IOPAIC\%5D2.0.CO;2}{10.1614/0043-1745(2001)049{[}0202:IOPAIC{]}2.0.CO;2}.

\leavevmode\hypertarget{ref-matzrafi2021}{}%
Matzrafi, M., Osipitan, O. A., Ohadi, S., and Mesgaran, M. B. (2021).
Under pressure: Maternal effects promote drought tolerance in progeny
seed of {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed Science}
69, 31--38.
doi:\href{https://doi.org/10.1017/wsc.2020.75}{10.1017/wsc.2020.75}.

\leavevmode\hypertarget{ref-mcfarlane2018}{}%
McFarlane, G. R., Whitelaw, C. B. A., and Lillico, S. G. (2018).
{CRISPR}-{Based Gene Drives} for {Pest Control}. \emph{Trends in
Biotechnology} 36, 130--133.
doi:\href{https://doi.org/10.1016/j.tibtech.2017.10.001}{10.1016/j.tibtech.2017.10.001}.

\leavevmode\hypertarget{ref-menges1987}{}%
Menges, R. M. (1987). Weed {Seed Population Dynamics} during {Six Years}
of {Weed Management Systems} in {Crop Rotations} on {Irrigated Soil}.
\emph{Weed Science} 35, 328--332. Available at:
\url{http://www.jstor.org/stable/4044593}.

\leavevmode\hypertarget{ref-mesgaran2021}{}%
Mesgaran, M. B., Matzrafi, M., and Ohadi, S. (2021). Sex dimorphism in
dioecious {Palmer} amaranth ({Amaranthus} palmeri) in response to water
stress. \emph{Planta} 254, 17.
doi:\href{https://doi.org/10.1007/s00425-021-03664-7}{10.1007/s00425-021-03664-7}.

\leavevmode\hypertarget{ref-milani2021}{}%
Milani, A., Panozzo, S., Farinati, S., Iamonico, D., Sattin, M., Loddo,
D., and Scarabel, L. (2021). Recent {Discovery} of {Amaranthus} palmeri
{S}. {Watson} in {Italy}: {Characterization} of {ALS}-{Resistant
Populations} and {Sensitivity} to {Alternative Herbicides}.
\emph{Sustainability} 13, 7003.
doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.

\leavevmode\hypertarget{ref-montgomery2021}{}%
Montgomery, J. S., Giacomini, D. A., Weigel, D., and Tranel, P. J.
(2021). Male-specific {Y}-chromosomal regions in waterhemp ({Amaranthus}
tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{New
Phytologist} 229, 3522--3533.
doi:\href{https://doi.org/10.1111/nph.17108}{10.1111/nph.17108}.

\leavevmode\hypertarget{ref-montgomery2019}{}%
Montgomery, J. S., Sadeque, A., Giacomini, D. A., Brown, P. J., and
Tranel, P. J. (2019). Sex-specific markers for waterhemp ({Amaranthus}
tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed
Science} 67, 412--418.
doi:\href{https://doi.org/10.1017/wsc.2019.27}{10.1017/wsc.2019.27}.

\leavevmode\hypertarget{ref-morgan2001}{}%
Morgan, G. D., Baumann, P. A., and Chandler, J. M. (2001). Competitive
{Impact} of {Palmer Amaranth} ({Amaranthus} palmeri) on {Cotton}
({Gossypium} hirsutum) {Development} and {Yield}. \emph{Weed Technology}
15, 408--412.
doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.

\leavevmode\hypertarget{ref-nguyen2018}{}%
Nguyen, H. M., Kleitou, P., Kletou, D., Sapir, Y., and Winters, G.
(2018). Differences in flowering sex ratios between native and invasive
populations of the seagrass {Halophila} stipulacea. \emph{Botanica
Marina} 61, 337--342.
doi:\href{https://doi.org/10.1515/bot-2018-0015}{10.1515/bot-2018-0015}.

\leavevmode\hypertarget{ref-oliveira2018}{}%
Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
intraspecific transference of metabolism-based mesotrione resistance in
dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.

\leavevmode\hypertarget{ref-oliveira2021a}{}%
Oliveira, M. C., Giacomini, D. A., Arsenijevic, N., Vieira, G., Tranel,
P. J., and Werle, R. (2021). Distribution and validation of genotypic
and phenotypic glyphosate and {PPO}-inhibitor resistance in {Palmer}
amaranth ({Amaranthus} palmeri) from southwestern {Nebraska}. \emph{Weed
Technology} 35, 65--76.
doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.

\leavevmode\hypertarget{ref-page2021}{}%
Page, E. R., Nurse, R. E., Meloche, S., Bosveld, K., Grainger, C.,
Obeid, K., Filotas, M., Simard, M.-J., and Laforest, M. (2021). Import
of {Palmer} amaranth ({Amaranthus} palmeri {S}. {Wats}.) Seed with sweet
potato ({Ipomea} batatas ({L}.) {Lam}) slips. \emph{Can. J. Plant Sci.},
CJPS-2020-0321.
doi:\href{https://doi.org/10.1139/CJPS-2020-0321}{10.1139/CJPS-2020-0321}.

\leavevmode\hypertarget{ref-perez-llorca2019}{}%
Pérez-Llorca, M., and Sánchez Vilas, J. (2019). Sexual dimorphism in
response to herbivory and competition in the dioecious herb {Spinacia}
oleracea. \emph{Plant Ecol} 220, 57--68.
doi:\href{https://doi.org/10.1007/s11258-018-0902-7}{10.1007/s11258-018-0902-7}.

\leavevmode\hypertarget{ref-piskackova2021}{}%
Piskackova, T. A. R., Reberg-Horton, S. C., Richardson, R. J., Jennings,
K. M., Franca, L., Young, B. G., and Leon, R. G. (2021). Windows of
action for controlling palmer amaranth ({Amaranthus} palmeri) using
emergence and phenology models. \emph{Weed Research} 61, 188--198.
doi:\href{https://doi.org/10.1111/wre.12470}{10.1111/wre.12470}.

\leavevmode\hypertarget{ref-price2011}{}%
Price, A. J., Balkcom, K. S., Culpepper, S. A., Kelton, J. A., Nichols,
R. L., and Schomberg, H. (2011). Glyphosate-resistant {Palmer} amaranth:
{A} threat to conservation tillage. \emph{Journal of Soil and Water
Conservation} 66, 265--275.
doi:\href{https://doi.org/10.2489/jswc.66.4.265}{10.2489/jswc.66.4.265}.

\leavevmode\hypertarget{ref-ritz2015}{}%
Ritz, C., Baty, F., Streibig, J. C., and Gerhard, D. (2015).
Dose-{Response Analysis Using R}. \emph{PLOS ONE} 10, e0146021.
doi:\href{https://doi.org/10.1371/journal.pone.0146021}{10.1371/journal.pone.0146021}.

\leavevmode\hypertarget{ref-rumpa2019}{}%
Rumpa, M. M., Krausz, R. F., Gibson, D. J., and Gage, K. L. (2019).
Effect of {PPO}-{Inhibiting Herbicides} on the {Growth} and {Sex Ratio}
of a {Dioecious Weed Species Amaranthus} palmeri ({Palmer Amaranth}).
\emph{Agronomy} 9, 275.
doi:\href{https://doi.org/10.3390/agronomy9060275}{10.3390/agronomy9060275}.

\leavevmode\hypertarget{ref-sanctis2021}{}%
Sanctis, J. H. S. de, Barnes, E. R., Knezevic, S. Z., Kumar, V., and
Jhala, A. J. (2021). Residual herbicides affect critical time of
{Palmer} amaranth removal in soybean. \emph{Agronomy Journal} 113,
1920--1933.
doi:\href{https://doi.org/10.1002/agj2.20615}{10.1002/agj2.20615}.

\leavevmode\hypertarget{ref-sauer1957}{}%
Sauer, J. (1957). Recent {Migration} and {Evolution} of the {Dioecious
Amaranths}. \emph{Evolution} 11, 11--31.
doi:\href{https://doi.org/10.2307/2405808}{10.2307/2405808}.

\leavevmode\hypertarget{ref-sauer1972}{}%
Sauer, J. D. (1972). The dioecious amaranths: A new species name and
major range extensions. \emph{Madroño} 21, 426--434. Available at:
\url{http://www.jstor.org/stable/41423815}.

\leavevmode\hypertarget{ref-sbatella2010}{}%
Sbatella, G. M., and Wilson, R. G. (2010). Isoxaflutole {Shifts Kochia}
({Kochia} scoparia) {Populations} in {Continuous Corn}. \emph{Weed
Technology} 24, 392--396.
doi:\href{https://doi.org/10.1614/WT-D-09-00023.1}{10.1614/WT-D-09-00023.1}.

\leavevmode\hypertarget{ref-schliekelman2005}{}%
Schliekelman, P., Ellner, S., and Gould, F. (2005). Pest {Control} by
{Genetic Manipulation} of {Sex Ratio}. \emph{Journal of Economic
Entomology} 98, 18--34.
doi:\href{https://doi.org/10.1093/jee/98.1.18}{10.1093/jee/98.1.18}.

\leavevmode\hypertarget{ref-schwartz2016}{}%
Schwartz, L. M., Norsworthy, J. K., Young, B. G., Bradley, K. W.,
Kruger, G. R., Davis, V. M., Steckel, L. E., and Walsh, M. J. (2016).
Tall {Waterhemp} ({Amaranthus} tuberculatus) and {Palmer} amaranth
({Amaranthus} palmeri) {Seed Production} and {Retention} at {Soybean
Maturity}. \emph{Weed Technology} 30, 284--290.
doi:\href{https://doi.org/10.1614/WT-D-15-00130.1}{10.1614/WT-D-15-00130.1}.

\leavevmode\hypertarget{ref-scott2011}{}%
Scott, D. (2011). The {Technological Fix Criticisms} and the
{Agricultural Biotechnology Debate}. \emph{J Agric Environ Ethics} 24,
207--226.
doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.

\leavevmode\hypertarget{ref-sosnoskie2013}{}%
Sosnoskie, L. M., Webster, T. M., and Culpepper, A. S. (2013).
Glyphosate {Resistance Does Not Affect Palmer Amaranth} ({Amaranthus}
palmeri) {Seedbank Longevity}. \emph{Weed Science} 61, 283--288.
doi:\href{https://doi.org/10.1614/WS-D-12-00111.1}{10.1614/WS-D-12-00111.1}.

\leavevmode\hypertarget{ref-spaunhorst2018}{}%
Spaunhorst, D. J., Devkota, P., Johnson, W. G., Smeda, R. J., Meyer, C.
J., and Norsworthy, J. K. (2018). Phenology of {Five Palmer} amaranth
({Amaranthus} palmeri) {Populations Grown} in {Northern Indiana} and
{Arkansas}. \emph{Weed Science} 66, 457--469.
doi:\href{https://doi.org/10.1017/wsc.2018.12}{10.1017/wsc.2018.12}.

\leavevmode\hypertarget{ref-vanwychen2020}{}%
Van Wychen, L. (2020). 2020 {Survey} of the most common and troublesome
weeds in grass crops, pasture, and turf in the {United States} and
{Canada}. Available at:
\url{https://wssa.net/wp-content/uploads/2020-Weed-Survey_grass-crops.xlsx}.

\leavevmode\hypertarget{ref-wang1992}{}%
Wang, J. L., Klessig, D. F., and Berry, J. O. (1992). Regulation of {C4
Gene Expression} in {Developing Amaranth Leaves}. \emph{The Plant Cell}
4, 173--184.
doi:\href{https://doi.org/10.1105/tpc.4.2.173}{10.1105/tpc.4.2.173}.

\leavevmode\hypertarget{ref-ward2013}{}%
Ward, S. M., Webster, T. M., and Steckel, L. E. (2013). Palmer
{Amaranth} ({Amaranthus} palmeri): {A Review}. \emph{Weed Technology}
27, 12--27.
doi:\href{https://doi.org/10.1614/WT-D-12-00113.1}{10.1614/WT-D-12-00113.1}.

\leavevmode\hypertarget{ref-webster2015}{}%
Webster, T. M., and Grey, T. L. (2015). Glyphosate-{Resistant Palmer
Amaranth} ({Amaranthus} palmeri) {Morphology}, {Growth}, and {Seed
Production} in {Georgia}. \emph{Weed Science} 63, 264--272.
doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.

\leavevmode\hypertarget{ref-yacouby2020}{}%
Yacouby, R., and Axman, D. (2020). Probabilistic {Extension} of
{Precision}, {Recall}, and {F1 Score} for {More Thorough Evaluation} of
{Classification Models}. in \emph{Proceedings of the {First Workshop} on
{Evaluation} and {Comparison} of {NLP Systems}} ({Online}: {Association
for Computational Linguistics}), 79--91.
doi:\href{https://doi.org/10.18653/v1/2020.eval4nlp-1.9}{10.18653/v1/2020.eval4nlp-1.9}.

\leavevmode\hypertarget{ref-yu2021}{}%
Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
Gunsolus, J., and Becker, R. (2021). Timeline of {Palmer} amaranth
({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
\emph{Weed Technology}, 1--31.
doi:\href{https://doi.org/10.1017/wet.2021.32}{10.1017/wet.2021.32}.

\end{CSLReferences}

\hypertarget{figures}{%
\section*{Figures}\label{figures}}
\addcontentsline{toc}{section}{Figures}

\begin{figure}

{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}

}

\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}

}

\caption{Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}

}

\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}

}

\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}

}

\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}

}

\caption{MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B). From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
\end{figure}

\end{document}

### 2021-08-23 data analysis/plot_layout.Rmd
as

---
title: "layout"
author: "Maxwel Coura Oliveira"
date: "8/16/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggimage)
library(ggtext)
```

```{r}
read_csv("../data/layout.csv") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  mutate(distance = as.double(distance)) %>% 
  mutate(emoji = case_when(
    crop == "corn" ~ "https://img.icons8.com/color/48/000000/corn.png",
    crop == "soybean" ~ "https://img.icons8.com/color/48/000000/soy.png",
    TRUE ~ NA_character_)) %>% 
  mutate(weed = "https://hotemoji.com/images/emoji/0/1gqjxj81p5mvj0.png") -> layout
```

```{r}
layout %>% 
  ggplot(aes(x = row, y = distance)) +
  facet_grid(~crop) +
#  geom_point() +
#  scale_x_discrete(limits = c(-0.5, 4.5)) +
  geom_image(data = layout,
               aes(image=weed)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "corn"),
               aes(x = 4.5, image=emoji)) +
  labs(x= NULL, y = NULL) +
  theme_test() +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        strip.text = element_markdown(size = 18),
        panel.background = element_rect(fill = "#f7e4d7")) -> fig1
```

```{r}
library(ggimage)
fig1 +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 0.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 1.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 2.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 3.5, image=emoji)) +
  geom_image(data = layout %>% filter(crop == "soybean"),
               aes(x = 4.5, image=emoji)) -> fig2
```

```{r}
palmer <- tibble(crop = "bareground")
corn <- tibble(crop = "corn")
soybean <- tibble(crop = "soybean")
```

```{r}
fig2 +
  geom_text(data = palmer, aes(x = 2.7, y = 270), size = 2, hjust = 0, 
            fontface = "italic",
            label = "24 Palmer amaranth plants \nwere placed 76.2 cm \nat the same distance") +
  geom_text(data = palmer, aes(x = 2.3, y = 228), size = 2,
            label = "76.2 cm", fontface = "bold") +
  geom_text(data = palmer, aes(x = 2.5, y = 390), size = 2,
            label = "Palmer amaranth \nplants", fontface = "bold") +
  geom_text(data = corn, aes(x = 2.5, y = 400), size = 2,
            label = "Corn plants", fontface = "bold") +
  geom_text(data = soybean, aes(x = 2.5, y = 400), size = 2,
            label = "Soybean plants", fontface = "bold") -> fig3

#ggsave("figures/layout.png", height = 4, width = 9)
```

```{r}
fig3 +
  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 264, 
                                            xend = 2.85, yend = 230),
               linetype = "dotted") +
  geom_segment(data = palmer, mapping = aes(x = 2.1, y = 188, 
                                            xend = 2.85, yend = 228),
               linetype = "dotted") +
  geom_segment(data = palmer, mapping = aes(x = 2, y = 250, 
                                            xend = 2, yend = 198),
               linetype = "dotted") -> fig4

#ggsave("figures/layout.png", height = 4, width = 9)
```

```{r}
fig4 +
  geom_curve(data = palmer, aes(x = 2.5, y = 405, 
                               xend = 2.1, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) +
  geom_curve(data = palmer, aes(x = 2.5, y = 375, 
                               xend = 2.85, yend = 375),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = 0.3) +
  geom_curve(data = corn, aes(x = 2.3, y = 405, 
                               xend = 2.4, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) +
  geom_curve(data = corn, aes(x = 2.7, y = 395, 
                               xend = 2.6, yend = 380),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) + 
  geom_curve(data = soybean, aes(x = 2.3, y = 405, 
                               xend = 2.4, yend = 420),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3) +
  geom_curve(data = soybean, aes(x = 2.7, y = 395, 
                               xend = 2.6, yend = 380),
             arrow = arrow(length = unit(0.07, "inch")), 
               size = 0.4, curvature = -0.3)

ggsave("figures/layout.png", height = 4, width = 9)
```

```{r}
library(png)
library(ggpubr)
```

```{r}
june <- png::readPNG("figures/image.png")
july <- png::readPNG("figures/july.png")
```

```{r}
# Plot with background image
ggplot() +
 background_image(june) -> junef

ggplot() +
 background_image(july) -> julyf
```

```{r}
library(patchwork)
```

```{r}
junef + julyf +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))

ggsave("figures/Figure 6.png", width = 7)
```

### 2021-08-23 data analysis/sex_analysis.Rmd
as

---
title: "sex"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

```{r}
library(emoGG)
library(tidyverse)
library(lubridate)
library(tidymodels)
library(yardstick)
library(broom)       # for making model summary tidy
library(visreg)      # for potting logodds and probability 
library(margins)     # to calculate Average Marginal Effects
library(rcompanion)  # to calculate pseudo R2
library(ROCR) 
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-tidymodels-library-in-r-part-1-c1bdce0ac055
#https://towardsdatascience.com/modelling-binary-logistic-regression-using-r-research-oriented-modelling-and-interpretation-f67b3a954101
```

```{r}
read_csv("../data/data2.csv") %>% 
#  filter(location != "grant") %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) %>% 
  janitor::clean_names() %>% 
  rename(planting = planting_4,
         planting_1 = planting_9) -> data1
```

```{r}
data1 %>% 
  rename(month = planting,
         planting = planting_1) %>% 
  mutate(planting = mdy(planting),
         harvest = mdy(harvest),
         doyp = yday(planting),
         doyh = yday(harvest)) %>% 
  mutate_if(is_character, as_factor) %>% 
  filter(!is.na(gender)) %>% 
#  filter(!is.na(weight)) %>% 
  mutate(fem = as_factor(if_else(gender == "f", 1, 0))) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) %>% 
  filter(location != "grant") %>% 
  dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
  na.omit(data2) -> data2
```

```{r}
data2 %>% 
  distinct(location)
```

```{r}
# Total number of rows in the data frame
n <- nrow(data2)
# Number of rows for the training set (80% of the dataset)
n_train <- round(0.80 * n)
# Create a vector of indices which is an 80% random sample
set.seed(123)
train_indices <- sample(1:n, n_train)
# Subset the data frame to training indices only
train <- data2[train_indices, ]
# Exclude the training indices to create the test set
test <- data2[-train_indices, ]
```

```{r}
paste("train sample size: ", dim(train)[1])
paste("test sample size: ", dim(test)[1])
```

```{r}
library(recipes)
recipe(fem ~ doyh + crop + height + weight, data = train) %>% 
  recipes::step_dummy(all_nominal(), -all_outcomes()) -> recip1

#preprocessing
prepped_rec <- prep(recip1, verbose = TRUE, retain = TRUE )

# new data set with juice
juice <- juice(prepped_rec)

# bake to see the preprocessed training data
preproc_train <- recipes::bake(prepped_rec, new_data = NULL)
glimpse(preproc_train)
# bake to see the preprocessed testing data
preproc_test <- recipes::bake(prepped_rec, new_data = test)
glimpse(preproc_test)
```

```{r}
#Fitting a binary logistic regression
model_logi <- glm(fem ~ ., data = preproc_train, family = "binomial")
#Model summary
summary(model_logi)
```

```{r}
# Pseudo R_squared values and Likelyhood ratio test
rcompanion::nagelkerke(model_logi)
```

```{r}
# Probabilities of female with doy
visreg(model_logi, "doyh", scale="response", rug=2, gg=TRUE,
       xlab="Day of year",
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> doy_f
```

```{r}
doy_f +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(170, 270), breaks = seq(170, 270, 20)) +
  theme_test() +
  geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") -> doyf

#ggsave("figures/fig2.png")
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "weight", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Weight (g plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> weight_f
```
```{r}
weight_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  scale_x_continuous(limits = c(0, 1100), breaks = seq(0, 1100, 150)) +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> weightf

#ggsave("figures/fig1.png")
```

```{r}
# Probabilities of female with weight
visreg(model_logi, "height", scale="response", rug=2, gg = TRUE,
       xlab=expression(paste("Height (cm plant "^"-1",")")),
       ylab="P(female)",
       line = list(col="#333333"),
       fill=list(fill= "green", alpha = 0.2),
       points=list(size=1, col = "#333333", pch=1),
       ylim = c(0,1)) -> height_f
height_f
```

```{r}
height_f +
  scale_y_continuous(limits = c(0, 1)) +
  theme_test() +
  geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> heightf

#ggsave("figures/fig.png")
```

```{r}
library(patchwork)

doyf + weightf + heightf +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, hjust = 0, vjust = 0))

ggsave("figures/Figure 5.png", width = 9, height = 3)
```

### 2021-08-23 data analysis/weather/weather.Rmd
as

---
title: "weather"
author: "Maxwel Coura Oliveira"
date: "7/8/2021"
output: html_document
---

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(patchwork)
library(ggtext)
library(nasapower)
library(vegan)
```

# Load data

You can load your data with read_csv if you have it in a spreadsheet.

```{r}
# your data
df_time <- tibble(ID = c('1','2','3','4','5'), 
                  Site = c('Arlington','Macomb','Lincoln',
                               'Clay Center', 'Grant'),
                  # Both coordinates in decimal format
                  latitude = c(43.18, 40.45, 40.85, 40.57, 40.8430),
                  longitude = c(-89.20, -90.67, -96.62, -98.13, -101.72),
                  Start = c('1988-01-01','1988-01-01','1988-01-01',
                                '1988-01-01', '1988-01-01'),
                  Early_planting = c("2018-06-01", "2018-06-01",
                                     "2018-06-01",
                                     "2018-06-01", "2018-06-01"),
                  Late_planting = c("2018-06-01", "2018-06-01", "2018-06-01",
                                    "2018-06-01", "2018-06-01"),
                      End = c('2020-12-31','2020-12-31','2020-12-31',
                              '2020-12-31', '2020-12-31'))

df_time <- df_time %>% 
  mutate_at(5:8, ymd) # as date
```

## Daymet

```{r function}
source("weather_daymet.R")
```

```{r get_data}
#dpp default = 0 - dpp = days prior start date

df_daymet <- weather_daymet(input = df_time, dpp = 0) 
```

```{r}
df_daymet %>% 
  janitor::clean_names() %>% 
  mutate(month_lab = month(date, label = TRUE)) -> adaptation_weather
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) -> adaptation_weather1
```

```{r}
adaptation_weather1 %>% 
  filter(year %in% c(2018, 2019)) %>% 
  group_by(site, year, month) %>% 
  summarise(prec = sum(pp), tmax = mean(tmax), tmin = mean(tmin), tmean = mean(tmean)) %>% 
  mutate_if(is.double, ~round(., 1))
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) %>% 
  mutate(site = fct_recode(site, 
                           "Arlington, WI" = "Arlington",
                           "Clay Center, NE" = "Clay Center",
                           "Grant, NE" = "Grant",
                           "Lincoln, NE" = "Lincoln",
                           "Macomb, IL" = "Macomb")) -> adaptation_weather1
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  group_by(month_lab, site) %>% 
  summarise(tmean = mean(tmean))
```

```{r}
text1 <- tibble(site = "Arlington, WI")
text2 <- tibble(site = "Clay Center, NE")
```

```{r}
adaptation_weather1 %>% 
  ggplot(aes(x = month_lab, y = tmean)) +
  geom_jitter(aes(color = variable), alpha = 0.05, shape = 0) +
  scale_color_manual(values = c("red", "blue", "gray90")) +
  facet_grid(~ site) +
  geom_line(adaptation_weather1 %>% 
               filter(year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "red") +
  geom_line(adaptation_weather1 %>% 
               filter(year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean, group = 1), 
             color = "blue") +
  geom_point(adaptation_weather1 %>% 
               filter(year %in% c(2018)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "red", shape = 22, size = 3) +
  geom_point(adaptation_weather1 %>% 
               filter(year %in% c(2019)) %>% 
               group_by(site, month_lab) %>% 
               summarise(tmean = mean(tmean)),
             mapping = aes(x = month_lab, y = tmean), 
             fill = "blue", shape = 22, size = 3) +
  geom_richtext(data = text1, aes(x = 2, y = 7), lineheight = 0.4, 
            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text2, aes(x = 2.2, y = 10), lineheight = 0.4, 
            label = "30 years <br>daily mean <br>air temperature", size = 6, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text2, 
             x = 2, y = 13, 
             xend = 2.5, yend = 16,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  scale_y_continuous(limits = c(0, 32)) +
  labs(x = NULL, y = "Temperature (C)") +
  theme_test() +
  theme(legend.position = "none",
#        strip.text = element_markdown(size = 15, hjust = 0.5),
#        axis.text.y = element_markdown(size = 25),
#        axis.title = element_markdown(size = 27),
#        axis.text.x = element_markdown(size = 25)#,
#        legend.text = element_markdown(size = 25)
        )  -> temp

ggsave("weather.png", height = 3, width = 6)
```

```{r}
text3 <- tibble(site = "Grant, NE")
text4 <- tibble(site = "Lincoln, NE")
```

```{r}
adaptation_weather %>% 
  filter(month %in% c(6, 7, 8, 9)) %>% 
  group_by(month_lab, site, year) %>% 
  summarise(pp = sum(pp)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) %>% 
  mutate(site = fct_recode(site, 
                           "Arlington, WI" = "Arlington",
                           "Clay Center, NE" = "Clay Center",
                           "Grant, NE" = "Grant",
                           "Lincoln, NE" = "Lincoln",
                           "Macomb, IL" = "Macomb")) -> adaptation_pp
```

```{r message = FALSE, warning = FALSE}
adaptation_pp %>% 
ggplot(aes(x = month_lab, y = pp)) +
  geom_bar(data = adaptation_pp %>% filter(year == c(2018, 2019)), 
           mapping = aes(fill = year),
           stat = "identity", position = "dodge", alpha = 1) +
  geom_point(data = adaptation_pp %>% filter(year != c(2018, 2019)), 
              position = position_jitter(seed = 1),
              alpha = 0.2, color = "gray50", 
              show.legend = FALSE) +
  scale_fill_manual(values = c("red", "blue")) +
  facet_grid(~ site) +
  geom_richtext(data = text3, aes(x = 2.6, y = 330), lineheight = 0.4, 
            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
            label.color = NA, fill = NA) +
  geom_richtext(data = text4, aes(x = 3.6, y = 320), lineheight = 0.4, 
            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
             label.color = NA, fill = NA, color = "gray90") +
  geom_curve(data = text4, 
             x = 2.6, y = 310, 
             xend = 2.1, yend = 320,
             arrow = arrow(length = unit(0.07, "inch")), 
             size = 0.4, curvature = -0.3, color = "gray90") + 
  labs(x = NULL, y = "Precipitation (mm)") +
  theme_test() +
  theme(legend.position = "none",
#        strip.text = element_markdown(size = 28),
#        axis.text.y = element_markdown(size = 25),
#        axis.title = element_markdown(size = 27),
#        axis.text.x = element_markdown(size = 25),
#        legend.text = element_markdown(size = 25)
  ) -> ppp

```

```{r}
temp / ppp

ggsave("Figure 1.png")
```

### 2021-08-23 data analysis/weight.Rmd
as

---
title: "weight"
author: "Maxwel Coura Oliveira"
date: "8/11/2021"
output: html_document
---

###------------
```{r}
library(tidyverse)
library(lubridate)
library(lme4)
library(lmerTest)
library(emmeans)
library(tidytext)
library(ggthemes)
library(ggtext)
library(patchwork)
```

```{r}
data <- read_csv("../data/data2.csv")
```

###--------------------------

```{r}
data %>% 
  mutate(location = str_replace(location, "havelock", "lincoln")) %>% 
  mutate(harvest = case_when(
    location == "clay center" & doyh == 169 ~ "8/18/2018",
    TRUE ~ harvest
  )) %>% 
  janitor::clean_names() %>% 
  rename(month = planting_4) -> data1
```

```{r}
data1 %>% 
#  rename(month = planting) %>% 
  mutate_at(c("year", "location"), as_factor) %>% 
  mutate(crop = fct_recode(crop, 
                           "bareground" = "fallow")) -> data2
```

```{r}
model <- lmer(log(weight) ~ crop * month + 
                                 (1|location/year), data = data2)
```

```{r}
anova(model)
```

```{r}
emmeans(model, ~ crop * month, type = "response") -> emmeans1
```

```{r}
multcomp::cld(emmeans1, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> weight
```

```{r}
weight %>% 
  mutate(crop = fct_relevel(crop, 
                           levels = c("bareground", "corn", "soybean"))) %>% 
  ggplot(aes(y = crop, 
             x = response,
             shape = month,
             color = month))
```

```{r}
weight %>% 
  mutate(crop = fct_relevel(crop, 
                           levels = c("bareground", "corn", "soybean"))) %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%  
  ggplot(aes(y = crop, 
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_jitter(data = data2 %>% 
                rename(response = weight) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.01, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2, "#333333")) +
  geom_text(aes(label = round(response,1), x = upper.CL + 130, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_x_continuous(limits = c(0,1201), breaks = seq(0, 1200, 300)) +
  scale_shape_manual(values = c(1,16)) +
  labs(x = expression(paste("Weight (g plant "^"-1",")")), 
       y = NULL, color = NULL) +
  annotate("text", x = 300, y = 1.7, 
           label = "First cohort",
           size = 7, color = "#333333",
           fontface = "bold") + 
  annotate("text", x = 300, y = 2.3, 
           label = "Second cohort",
           size = 7, color = 2,
           fontface = "bold") + 
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figA

#ggsave("figures/biomass.png", width = 4, height = 4)
```

```{r}
model2 = lmer(log(height) ~ crop * month + 
                                 (1|location/year), data = data2)
anova(model2)
```

```{r}
emmeans(model2, ~ crop * month, type = "response", cont="pairwise", adjust="none") -> emmeans2
```

```{r}
multcomp::cld(emmeans2$emmeans, 
              alpha=0.05, 
              Letters=letters, 
              adjust="none", reversed = TRUE) %>% 
  as_tibble() -> height
```

```{r warning = FALSE}
height %>% 
  mutate(crop = fct_relevel(crop, 
                           levels = c("bareground", "corn", "soybean"))) %>% 
  mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")) %>%
  ggplot(aes(y = crop,
             x = response,
             shape = month,
             color = month)) +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  geom_linerange(aes(xmin = lower.CL, xmax = upper.CL),
                 position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_jitter(data = data2 %>% 
                rename(response = height) %>% 
                mutate(crop = fct_recode(crop,
          "<b style='color:#fb832d;'>bareground</b>" = "bareground",
          "<b style='color:#30123bFF;'>corn</b>" = "corn",
          "<b style='color:#28bbecff;'>soybean</b>" = "soybean")),
#                filter(month == "First transplanting"), 
              alpha = 0.02, aes(color = month, shape = month), size = 1,
              position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_color_manual(values = c(2,"#333333")) +
  geom_text(aes(label = round(response,1), x = upper.CL + 20, color = month),
            size = 8, fontface = "bold",
            position = position_dodge2(width = 0.4, reverse = TRUE)) +
  geom_point(size = 2,
             position = position_dodge2(width = 0.4, reverse = TRUE)) +
  scale_shape_manual(values = c(1,16)) +
#  scale_x_continuous(limits = c(0,300), breaks = c(0,25,50,75,100)) +
  labs(x = expression(paste("Height (cm plant "^"-1",")")), 
       y = NULL, color = NULL) +
  theme_test(base_family = "roboto") +
  theme(legend.position = "none",
        axis.text.y = element_markdown(size = 25),
        axis.title = element_markdown(size = 30),
        axis.text.x = element_markdown(size = 25),
        legend.text = element_markdown(size = 25)) -> figB

#ggsave("figures/height.png", width = 4, height = 4)
```

```{r}
(figA | figB) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))

ggsave("figures/Figure 3.png", width = 6, height = 3)
```

### 2021-08-23 data analysis/weather/weather_daymet.R

@@ -1,0 +1,73 @@
+
+library(tidyverse)
+
+# Constants for ET0 (Cobaner et al., 2017)
+# Solar constant
+Gsc = 0.0820 # (MJ m-2 min-1)
+# Radiation adjustment coefficient (Samani, 2004)
+kRs = 0.17
+
+weather_daymet <- function(input, dpp = 0){ 
+  input %>%
+    mutate(Weather = pmap(list(ID = ID,
+                               lat = latitude,
+                               lon = longitude,
+                               sta = Start - dpp,
+                               end = End),
+                          
+                          # Retrieving daymet data
+                          function(ID,lat,lon,sta,end){
+                            download_daymet(site = ID,
+                                            lat = lat, lon = lon,
+                                            # Extracting year
+                                            start = as.numeric(substr(sta,1,4)),
+                                            end = as.numeric(substr(end,1,4)),
+                                            internal = T, simplify = T)})) %>% 
+    
+    mutate(Weather = Weather %>% 
+             
+             map(~mutate(.,
+                         Date = as_date(as.numeric(yday)-1, # Day of the year
+                                        origin = paste0(year,'-01-01')),
+                         Year = year(Date),
+                         Month = month(Date),
+                         Day = mday(Date)))  %>% 
+             
+             map(~dplyr::select(., yday, Year, Month, Day, Date,
+                                measurement,value)) %>% 
+             map(~spread(., 'measurement','value'))  %>% 
+             map(~rename_all(., ~c("DOY", # Date as Day of the year
+                                   "Year", # Year
+                                   "Month", # Month 
+                                   "Day", # Day of the month
+                                   "Date", # Date as normal format
+                                   "DL", # Day length (sec)
+                                   "PP", # Precipitation (mm)
+                                   "Rad", # Radiation (W/m2)
+                                   "SWE", # Snow water (kg/m2)
+                                   "Tmax", # Max. temp. (degC)
+                                   "Tmin", # Min. temp. (degC)
+                                   "VPD"))))  %>% # Vap Pres Def (Pa)
+    
+    mutate(Weather = pmap(list(sta=Start-dpp,
[…]

### 2021-08-23 Frontiers/Frontiers.Rmd
vh

@@
     city: Macomb
-    state: Illnois # only USA, Australia, Canada
+    state: Illinois # only USA, Australia, Canada
     country: United States
@@
 
-Abstract length and content varies depending on article type. Refer to 
-\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
-and length according to article type.
+Palmer amaranth (*Amaranthus palmeri* S. Watson) is the most troublesome weed species in the Southern United States cropping systems.
 
-%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \tiny
@@
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
-With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@oliveira2018].
+With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021].
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017].
@@
 
-# Results {-}
+# Material and Methods {-}
 
@@
 
-The statistical analyses were performed using R statistical software version 4.0.1.
+The statistical analyses were performed using R statistical software version 4.0.1. Data analyses are stored at Zenodo [@oliveira2021b].
 
@@
 
-In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the XXX, and *x* day of year (doy).
+In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the inflection point, and *x* day of year (doy).
 
@@
 Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems.
-Is likely that Palmer amaranth range will continue to expanding to new geographies.
+Is likely that Palmer amaranth range will continue to expanding into new geographies.
 Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal.
@@
 
-# Acknowledgments
+# Acknowledgments {-}
 
@@
 
-# Supplemental Data
-
-Supplementary Material should be uploaded separately on submission, if there are
-Supplementary Figures, please include the caption in the same file as the
-figure. LaTeX Supplementary Material templates can be found in the Frontiers
-LaTeX folder
-
-# References
+# References {-}
 
@@
 
-# Figures
+# Figures {-}
 
@@
 
-```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B).
 From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}

### 2021-08-23 Frontiers/Frontiers.tex
vh

@@
   \textsuperscript{3} Department of Agronomy, Western Illinois
-University,  Macomb,  Illnois,  United States
+University,  Macomb,  Illinois,  United States
   }
@@
 outcrosser species, increasing the chances of exchanging adaptive traits
-among plants (Oliveira et al., 2018). Also, Palmer amaranth small seed
-size (e.g, 1 mm) tend to thrive in no-tillage systems (Price et al.,
-2011), and spread across locations through farm equipment (Sauer, 1972),
-manure (Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
+among plants (Jhala et al., 2021). Also, Palmer amaranth small seed size
+(e.g, 1 mm) tend to thrive in no-tillage systems (Price et al., 2011),
+and spread across locations through farm equipment (Sauer, 1972), manure
+(Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
 dispersal capacity of Palmer amaranth makes the species one of the most
@@
 
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
+\hypertarget{material-and-methods}{%
+\section*{Material and Methods}\label{material-and-methods}}
+\addcontentsline{toc}{section}{Material and Methods}
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1.
+version 4.0.1. Data analyses are stored at Zenodo (Oliveira, 2021).
 
@@
 In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
-\emph{d} is the upper limit (set to 100), and \emph{e} is the XXX, and
-\emph{x} day of year (doy).
+\emph{d} is the upper limit (set to 100), and \emph{e} is the inflection
+point, and \emph{x} day of year (doy).
 
@@
 
-\hypertarget{results-1}{%
-\section*{Results}\label{results-1}}
+\hypertarget{results}{%
+\section*{Results}\label{results}}
 \addcontentsline{toc}{section}{Results}
@@
 \hypertarget{supplemental-data}{%
-\section{Supplemental Data}\label{supplemental-data}}
+\section*{Supplemental Data}\label{supplemental-data}}
+\addcontentsline{toc}{section}{Supplemental Data}
 
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 
-\leavevmode\hypertarget{ref-oliveira2018}{}%
-Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
-S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
-intraspecific transference of metabolism-based mesotrione resistance in
-dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
-doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.
+\leavevmode\hypertarget{ref-oliveira2021b}{}%
+Oliveira, M. C. (2021). \emph{Maxwelco/palmer\_adaptation}. {Zenodo}
+Available at: \url{ttps://doi.org/10.5281/zenodo.5236831}.
 
@@
 
-\caption{MCO (180 cm) holding harvested Palmer amaranth plants at 40 days after first transplant (1st cohort, A) and 33 days after second transplant (2nd cohort, B).
+\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B).
 From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}

### 2021-08-23 Frontiers/Frontiers.Rmd
s

@@
 
-Palmer amaranth (*Amaranthus palmeri* S. Watson) is the most troublesome weed species in the Southern United States cropping systems.
+Palmer amaranth (\emph{Amaranthus palmeri} S.
+Watson) is one of the most troublesome weed species in the United States.
+Palmer amaranth is endemic in the Southern United States but its range is expanding northward.
+Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
+A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of United States Midwest.
+In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
+At first cohort, Palmer amaranth produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas Palmer amaranth produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} at second cohort in bareground, soybean and corn, respectively.
+The plant height was more impacted by Palmer amaranth growing in corn.
+Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
+Moreover, Palmer amaranth flowering window shifted according to crop and cohort timings.
+Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window.
+Also, Palmer amaranth sex dimorphism was slightly influenced by day of year, weight and height.
+The model estimated that probability of being a female plant increase as biomass and height increases.
+Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems.
+Palmer amaranth is likely to continue its expansion northward.
+Therefore, preventing plant dispersal into new habitat is the most effective management strategy.
+Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability establish into cropping systems, including crop rotation (not only corn and soybean), early/late crop planting, row spacing, and cover crops.
+Our results also shed a light on sex dimorphism as a future Palmer amaranth ecological management.
 
@@
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
-With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021].
+With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
 Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017].
@@
 Water has not shown to limit Palmer amaranth fitness.
-Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant-1 [@chahal2018].
+Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
 Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under selection pressure.
-Palmer amaranth invasion/adaptation into new habitats is likely to increase with the global temperature warming.
+Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under the current conditions.
+Further scenarios show that global temperature warming can impact agriculture, and promote niches for Palmer amaranth invasion/adaptation into new environments.
 Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
@@
 Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America.
-Therefore, its range of expansion into new habitats is increasing.
+Therefore, Palmer amaranth range of expansion into new habitats is increasing.
 It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
@@
 It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and developing resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, sex dimorphism, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
 
@@
 Three weeks prior to the onset of the field experiments, seeds were planted in plastic trays containing potting-mix.
-Emerged seedlings (1 cm) were transplanted into 200 cm<sup>-3</sup> plastic pots (a plant pot<sup>-1</sup>).
-Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors in the Perkins extension office at Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
+Emerged seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
+Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors at the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
 
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N, 'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N, 'W)).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (('N, 'W)), Lincoln, NE ((41.16'N, 96.42'W)), and Macomb, IL (('N, 'W)).
 
@@
                 latex_options = c("scale_down", "hold_position")) %>% 
-  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") 
+  collapse_rows(columns = 1, latex_hline = "none", valign = "middle") %>%
+  footnote(
+           alphabet = c("glyphoste, 840 g ae ha", "S-metolachor, g ai ha", 
+                        "S-metolachlor + trazine + mesotrione, + bicyclopyrone, g ai ha",
+                        "saflufenacil + imazethapyr + pyroxasulfone, g ai ha"),
+           footnote_as_chunk = TRUE)
 ```
+
+<!--  
+ Acuron applied PRE in corn at 2.5 qt/acre; Zidua PRO applied PRE in soybean at 6 fl oz/acre
+-->
 
@@
 When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant.
-Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant-1 was recorded).
+Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant\textsuperscript{-1}) was recorded.
 
@@
 
-Palmer amaranth sex was fitted to a binary logistic regression [@bangdiwala2018].
+Palmer amaranth sex dimorphism was fitted to a binary logistic regression [@bangdiwala2018].
 Binary logistic regression is used for predicting binary classes, such as the probability of a plant being female in a dioecious species.
@@
 Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (figure 3A).
-Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and 16.3 g plant-1 in bareground, soybean and corn, respectively.
-Plants from the second cohort produced 62.6 g plant in bareground, followed by 6.3 g plant in soybean, and 1.4 g plant in corn.
+Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
+Plants from the second cohort produced 62.6 g plant\textsuperscript{-1} in bareground, followed by 6.3 g plant\textsuperscript{-1} in soybean, and 1.4 g plant\textsuperscript{-1}.
 
@@
 
-## Palmer amaranth gender {-}
+## Palmer amaranth sex dimorphism {-}
 
@@
 
-#names(table_a)[2] <- paste0(names(table_a)[2], 
-#                                footnote_marker_symbol(1))
-
 table_a %>% 
-  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
+  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of sex dimorphism logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
   kable_styling(font_size = 10, 
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering pattern and gender varied within agroecossystems and cohort timings.
+Our study showed that Palmer amaranth biomass, height, flowering pattern and sex varied within crops and cohort timings.
 In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort.
@@
 This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean.
-The Palmer amaranth competition strategy was to mimic the crop grow and development (Figure 6).
+The Palmer amaranth competition (e.g., light) strategy was to mimic the crop grow and development (Figure 6).
 These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018].
@@
 Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
-I our study, Palmer amaranth growing at first cohort accumulated an overall 36% more biomass when compared to second cohort.
+I our study, Palmer amaranth growing at first cohort accumulated an overall 42% more biomass when compared to second cohort.
 Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the cropping system.
@@
 Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
-We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
+We hypothesize that when growing in high competition (e.g., second cohort) and/or under low light/short day length, Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
 Early flower initiation is plant strategy when growing in stress conditions.
@@
 The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
-In our study, the gender model performance was decent (AIC 0.64) considering the biology of plant flowering.
+In our study, the sex dimorphism model performance was decent (AIC 0.64) considering the biology of plant flowering.
 A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
@@
 
-Our study demonstrated the short-term Palmer amaranth plasticity to grow and develop into cropping-systems.
+Our study demonstrated the Palmer amaranth plasticity to grow and develop into cropping-systems.
 Is likely that Palmer amaranth range will continue to expanding into new geographies.
@@
 Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development.
-Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation [@oliveira2021a], plant width, preemergence applied herbicide [@sanctis2021], and crop residue (e.g.
+Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide, and crop residue (e.g.
 cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.

### 2021-08-23 Frontiers/Frontiers.tex
s

@@
 
-Abstract length and content varies depending on article type. Refer to 
-\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
-and length according to article type.
-
-%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+Palmer amaranth (\emph{Amaranthus palmeri} S.
+Watson) is one of the most troublesome weed species in the United States.
+Palmer amaranth is endemic in the Southern United States but its range is expanding northward.
+Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
+A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of United States Midwest.
+In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
+At first cohort, Palmer amaranth produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas Palmer amaranth produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} at second cohort in bareground, soybean and corn, respectively.
+The plant height was more impacted by Palmer amaranth growing in corn.
+Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
+Moreover, Palmer amaranth flowering window shifted according to crop and cohort timings.
+Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window.
+Also, Palmer amaranth sex dimorphism was slightly influenced by day of year, weight and height.
+The model estimated that probability of being a female plant increase as biomass and height increases.
+Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems.
+Palmer amaranth is likely to continue its expansion northward.
+Therefore, preventing plant dispersal into new habitat is the most effective management strategy.
+Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability establish into cropping systems, including crop rotation (not only corn and soybean), early/late crop planting, row spacing, and cover crops.
+Our results also shed a light on sex dimorphism as a future Palmer amaranth ecological management.
+
 \tiny
@@
 outcrosser species, increasing the chances of exchanging adaptive traits
-among plants (Jhala et al., 2021). Also, Palmer amaranth small seed size
-(e.g, 1 mm) tend to thrive in no-tillage systems (Price et al., 2011),
-and spread across locations through farm equipment (Sauer, 1972), manure
-(Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
-dispersal capacity of Palmer amaranth makes the species one of the most
-successful cases of weed adaptation to current cropping systems.
+among plants (Jhala et al., 2021; Oliveira et al., 2018). Also, Palmer
+amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage
+systems (Price et al., 2011), and spread across locations through farm
+equipment (Sauer, 1972), manure (Hartzler and Anderson, 2016), animals
+(Farmer et al., 2017). The dispersal capacity of Palmer amaranth makes
+the species one of the most successful cases of weed adaptation to
+current cropping systems.
 
@@
 amaranth fitness. Under continuous water stress, Palmer amaranth
-survived and produced at least 14000 seeds plant-1 (Chahal et al.,
-2018). Seeds from Palmer amaranth growing with limited water conditions
-were heavier, less dormant, and prompt for germination (Matzrafi et al.,
-2021). Growing conditions and management practices also influence Palmer
-amaranth sex dimorphism and flowering pattern (Korres et al., 2017;
-Rumpa et al., 2019). Therefore, Palmer amaranth has shown plasticity to
-evolve and fast adapt under selection pressure. Palmer amaranth
-invasion/adaptation into new habitats is likely to increase with the
-global temperature warming. Currently, it is estimated that the greatest
-climatic risk of Palmer amaranth establishment are agronomic crops in
-Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
-Temperature is a key factor limiting Palmer amaranth expansion to cooler
-geographies (Briscoe Runquist et al., 2019); however, under future
-climate change Palmer amaranth is likely to expand northward into Canada
-and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
-al., 2019).
+survived and produced at least 14000 seeds plant\textsuperscript{-1}
+(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
+water conditions were heavier, less dormant, and prompt for germination
+(Matzrafi et al., 2021). Growing conditions and management practices
+also influence Palmer amaranth sex dimorphism and flowering pattern
+(Korres et al., 2017; Rumpa et al., 2019). Therefore, Palmer amaranth
+has shown plasticity to evolve and fast adapt under the current
+conditions. Further scenarios show that global temperature warming can
+impact agriculture, and promote niches for Palmer amaranth
+invasion/adaptation into new environments. Currently, it is estimated
+that the greatest climatic risk of Palmer amaranth establishment are
+agronomic crops in Australia and Sub-Sahara Africa (Kistner and
+Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
+expansion to cooler geographies (Briscoe Runquist et al., 2019);
+however, under future climate change Palmer amaranth is likely to expand
+northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
+Briscoe Runquist et al., 2019).
 
@@
 adapted and established to Northern as it is in the Southern North
-America. Therefore, its range of expansion into new habitats is
-increasing. It seems fated the need to manage Palmer amaranth in
+America. Therefore, Palmer amaranth range of expansion into new habitats
+is increasing. It seems fated the need to manage Palmer amaranth in
 agronomic crops throughout multiple environments in the near future.
@@
 systems (MacLaren et al., 2020). Therefore, the objective of this study
-was to investigate the flowering pattern, gender, biomass production,
-and height of Palmer amaranth cohorts growing under corn, soybean and
-bareground across five locations in the United States Midwest.
+was to investigate the flowering pattern, sex dimorphism, biomass
+production, and height of Palmer amaranth cohorts growing under corn,
+soybean and bareground across five locations in the United States
+Midwest.
 
@@
 experiments, seeds were planted in plastic trays containing potting-mix.
-Emerged seedlings (1 cm) were transplanted into 200 cm-3 plastic pots (a
-plant pot-1). Palmer amaranth seedlings were supplied with adequate
-water and kept under greenhouse conditions at the University of
-Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois
-University; and kept outdoors in the Perkins extension office at Grant,
-NE until the 2-3 leaf stage (5 to 8 cm height) when they were
-transported to the field.
+Emerged seedlings (1 cm) were transplanted into 200
+cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
+Palmer amaranth seedlings were supplied with adequate water and kept
+under greenhouse conditions at the University of Wisconsin-Madison,
+University of Nebraska-Lincoln, and Western Illinois University; and
+kept outdoors at the Perkins extension office in Grant, NE until the 2-3
+leaf stage (5 to 8 cm height) when they were transported to the field.
 
@@
 The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N,
-'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N,
-'W)).
+five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
+(40.57'N, 9814'W), Grant, NE (('N, 'W)), Lincoln, NE ((41.16'N,
+96.42'W)), and Macomb, IL (('N, 'W)).
 
@@
 \bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, g ai ha \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, g ai ha \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, g ai ha}\\
 \end{tabular}}
@@
 near soil surface and oven dried at 65 C until reaching constant weight
-before weighing (g plant-1 was recorded).
+before weighing (g plant\textsuperscript{-1}) was recorded.
 
@@
 
-Palmer amaranth sex was fitted to a binary logistic regression
-(Bangdiwala, 2018). Binary logistic regression is used for predicting
-binary classes, such as the probability of a plant being female in a
-dioecious species. Prior to the analysis, all missing values were
-removed from the dataset. Also, data from Grant was not used in this
-analysis due to the uniform plant harvesting at that location. The
+Palmer amaranth sex dimorphism was fitted to a binary logistic
+regression (Bangdiwala, 2018). Binary logistic regression is used for
+predicting binary classes, such as the probability of a plant being
+female in a dioecious species. Prior to the analysis, all missing values
+were removed from the dataset. Also, data from Grant was not used in
+this analysis due to the uniform plant harvesting at that location. The
 complete dataset was splitted into 80\% train and 20\% test data. The
@@
 bareground compared to plants growing in soybean and corn (figure 3A).
-Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and
-16.3 g plant-1 in bareground, soybean and corn, respectively. Plants
-from the second cohort produced 62.6 g plant in bareground, followed by
-6.3 g plant in soybean, and 1.4 g plant in corn.
+Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3
+g plant\textsuperscript{-1} in bareground, soybean and corn,
+respectively. Plants from the second cohort produced 62.6 g
+plant\textsuperscript{-1} in bareground, followed by 6.3 g
+plant\textsuperscript{-1} in soybean, and 1.4 g
+plant\textsuperscript{-1}.
 
@@
 
-\hypertarget{palmer-amaranth-gender}{%
-\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
-\addcontentsline{toc}{subsection}{Palmer amaranth gender}
+\hypertarget{palmer-amaranth-sex-dimorphism}{%
+\subsection*{Palmer amaranth sex
+dimorphism}\label{palmer-amaranth-sex-dimorphism}}
+\addcontentsline{toc}{subsection}{Palmer amaranth sex dimorphism}
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of sex dimorphism logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
 \centering
@@
 Our study showed that Palmer amaranth biomass, height, flowering pattern
-and gender varied within agroecossystems and cohort timings. In general,
-first cohort of Palmer amaranth plants were heavier and taller when
-compared to the second cohort. At first cohort, resources (e.g., soil
-nutrients) and conditions (e.g., light) were more timely available for
-the species. High biomass and taller Palmer amaranth plants are likely a
-weed strategy to compete for light in between crop rows in absence of
-canopy. In such conditions, Palmer amaranth showed an extraordinary
-plasticity to adapt upon the agroecosystem. This is evident when
-comparing Palmer amaranth canopy shape, and its extended flowering
-pattern when growing into corn compared to soybean. The Palmer amaranth
-competition strategy was to mimic the crop grow and development (Figure
-6). These results suggests that Palmer amaranth can fast evolve
-life-history traits to adapt into cropping systems and cultural
-practices, which was also showed in a study varying nitrogen
-fertilization (Bravo et al., 2018). Our results highlight the Palmer
-amaranth as a threat to field crops as breeding more competitive crop
-varieties is likely to select more competitive biotypes (Bravo et al.,
-2017).
+and sex varied within crops and cohort timings. In general, first cohort
+of Palmer amaranth plants were heavier and taller when compared to the
+second cohort. At first cohort, resources (e.g., soil nutrients) and
+conditions (e.g., light) were more timely available for the species.
+High biomass and taller Palmer amaranth plants are likely a weed
+strategy to compete for light in between crop rows in absence of canopy.
+In such conditions, Palmer amaranth showed an extraordinary plasticity
+to adapt upon the agroecosystem. This is evident when comparing Palmer
+amaranth canopy shape, and its extended flowering pattern when growing
+into corn compared to soybean. The Palmer amaranth competition (e.g.,
+light) strategy was to mimic the crop grow and development (Figure 6).
+These results suggests that Palmer amaranth can fast evolve life-history
+traits to adapt into cropping systems and cultural practices, which was
+also showed in a study varying nitrogen fertilization (Bravo et al.,
+2018). Our results highlight the Palmer amaranth as a threat to field
+crops as breeding more competitive crop varieties is likely to select
+more competitive biotypes (Bravo et al., 2017).
 
@@
 (Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
-amaranth growing at first cohort accumulated an overall 36\% more
+amaranth growing at first cohort accumulated an overall 42\% more
 biomass when compared to second cohort. Therefore, Palmer amaranth
@@
 ecological factors (Lang, 1965). We hypothesize that when growing in
-high competition (e.g., second cohort), Palmer amaranth plants tend to
-initiate flowering early, as well as having an extended flowering
-window. Early flower initiation is plant strategy when growing in stress
-conditions. For example, when growing under water stress, early
-flowering in Palmer amaranth resulted in a mismatch between female and
-male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant outcross,
-and thus reduce plant seed production and exchange of resistant alleles
-(Jhala et al., 2021). Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
+high competition (e.g., second cohort) and/or under low light/short day
+length, Palmer amaranth plants tend to initiate flowering early, as well
+as having an extended flowering window. Early flower initiation is plant
+strategy when growing in stress conditions. For example, when growing
+under water stress, early flowering in Palmer amaranth resulted in a
+mismatch between female and male plants by seven days (Mesgaran et al.,
+2021). A mismatch in Palmer amaranth male and female flowering period
+can minimize plant outcross, and thus reduce plant seed production and
+exchange of resistant alleles (Jhala et al., 2021). Sex dimorphism
+manipulation is considered a potential ecological pest control
+(McFarlane et al., 2018; Schliekelman et al., 2005).
 
@@
 arouse the curiosity of many scientists, including Charles Darwin
-(Darwin, 1888). In our study, the gender model performance was decent
-(AIC 0.64) considering the biology of plant flowering. A 1:1 male and
-female sex ratio is a general evolutionary stable strategy for plant
+(Darwin, 1888). In our study, the sex dimorphism model performance was
+decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male
+and female sex ratio is a general evolutionary stable strategy for plant
 species perpetuation (Fisher, 1930). However, a slight deviation from
@@
 
-Our study demonstrated the short-term Palmer amaranth plasticity to grow
-and develop into cropping-systems. Is likely that Palmer amaranth range
-will continue to expanding to new geographies. Therefore, preventive
+Our study demonstrated the Palmer amaranth plasticity to grow and
+develop into cropping-systems. Is likely that Palmer amaranth range will
+continue to expanding into new geographies. Therefore, preventive
 management is a priority to minimizing Palmer amaranth dispersal.
@@
 development. Tactics that promote early-season crop advantage against
-Palmer amaranth, including early crop planting, crop rotation (Oliveira
-et al., 2021), plant width, preemergence applied herbicide (Sanctis et
-al., 2021), and crop residue (e.g.~cover crops) would minimize the
-negative impact of Palmer amaranth in agroecossystems. The
-aggressiveness and differential Palmer amaranth adaptation to
-agroecosystem is striking and require national efforts to minimize the
-species impact on economy and sustainability.
+Palmer amaranth, including early crop planting, crop rotation, plant
+width, preemergence applied herbicide, and crop residue (e.g.~cover
+crops) would minimize the negative impact of Palmer amaranth in
+agroecossystems. The aggressiveness and differential Palmer amaranth
+adaptation to agroecosystem is striking and require national efforts to
+minimize the species impact on economy and sustainability.
 
@@
 commercial, or not-for-profit sectors
-
-\hypertarget{supplemental-data}{%
-\section*{Supplemental Data}\label{supplemental-data}}
-\addcontentsline{toc}{section}{Supplemental Data}
-
-Supplementary Material should be uploaded separately on submission, if
-there are Supplementary Figures, please include the caption in the same
-file as the figure. LaTeX Supplementary Material templates can be found
-in the Frontiers LaTeX folder
 
@@
 
+\leavevmode\hypertarget{ref-oliveira2018}{}%
+Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
+S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
+intraspecific transference of metabolism-based mesotrione resistance in
+dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
+doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.
+
 \leavevmode\hypertarget{ref-oliveira2021a}{}%
@@
 
-\leavevmode\hypertarget{ref-sanctis2021}{}%
-Sanctis, J. H. S. de, Barnes, E. R., Knezevic, S. Z., Kumar, V., and
-Jhala, A. J. (2021). Residual herbicides affect critical time of
-{Palmer} amaranth removal in soybean. \emph{Agronomy Journal} 113,
-1920--1933.
-doi:\href{https://doi.org/10.1002/agj2.20615}{10.1002/agj2.20615}.
-
 \leavevmode\hypertarget{ref-sauer1957}{}%
@@
 
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
 \end{figure}

### 2021-08-24 Frontiers/Frontiers.Rmd
as

@@
 Watson) is one of the most troublesome weed species in the United States.
-Palmer amaranth is endemic in the Southern United States but its range is expanding northward.
+Palmer amaranth is endemic to Southern United States but its range is expanding northward.
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of United States Midwest.
+A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
 In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
@@
 The model estimated that probability of being a female plant increase as biomass and height increases.
-Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems.
+Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems from United States Midwest.
 Palmer amaranth is likely to continue its expansion northward.
-Therefore, preventing plant dispersal into new habitat is the most effective management strategy.
+Therefore, preventing plant dispersal into new habitats is the most effective management strategy.
 Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability establish into cropping systems, including crop rotation (not only corn and soybean), early/late crop planting, row spacing, and cover crops.
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology.
-Farm mechanization, adoption of conservation agriculture (e.g., no-till), and reliance on herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
+Farm mechanization, adoption of conservation agriculture (e.g., no-till), and intensive use of herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
@@
 In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
-Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
+Waterhemp, an *Amaranthus* species, is the most troublesome species in that geography [@tranel2011].
+Nonetheless, invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
 Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021].
 Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America.
-Therefore, Palmer amaranth range of expansion into new habitats is increasing.
-It seems fated the need to manage Palmer amaranth in agronomic crops throughout multiple environments in the near future.
+Therefore, Palmer amaranth range of expansion into new habitats can increase.
+It seems certain the need to manage Palmer amaranth in agronomic crops throughout novel environments in the near future.
 Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (('N, 'W)), Lincoln, NE ((41.16'N, 96.42'W)), and Macomb, IL (('N, 'W)).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE ('N, 'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL ('N, 'W).
 
@@
   footnote(
-           alphabet = c("glyphoste, 840 g ae ha", "S-metolachor, g ai ha", 
-                        "S-metolachlor + trazine + mesotrione, + bicyclopyrone, g ai ha",
-                        "saflufenacil + imazethapyr + pyroxasulfone, g ai ha"),
+           alphabet = c("glyphoste, 840 g ae ha", "S-metolachor, 1324 g ai ha;", 
+                        "S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha;",
+                        "saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha"),
            footnote_as_chunk = TRUE)
@@
 Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide).
-Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop (Figure 2).
+Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop.
 After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2).
@@
 
-The statistical analyses were performed using R statistical software version 4.0.1. Data analyses are stored at Zenodo [@oliveira2021b].
+The statistical analyses were performed using R statistical software version 4.0.1 [@rcoreteam2021]. Data analyses are stored at Zenodo [@oliveira2021b].
 
@@
 Plant height and biomass were log transformed to meet model assumption of normality.
-In the model, crop (bareground, corn, soybean) was the fixed effect and year nested with location the random effects.
+In the model, crop (bareground, corn, soybean) and cohort time (first and second) were the fixed effects and year nested with location the random effects.
 Analysis of variance was performed with *anova* function from "car" package [@fox2018].
-Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" and "multcomp" [@hothorn2008].
+Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" [@lenth2021] and "multcomp" [@hothorn2008], respectively.
 
 The Palmer amaranth flowering timing was estimated as cumulative flowering across all location.
-Dataset was nested but id did not include Grant, NE.
+Dataset was nested but it did not include Grant, NE.
 Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
@@
 The complete dataset was splitted into 80% train and 20% test data.
-The 80% train is used for the model training and the 20% test is used for checking how the model generalized on unseen dataset.
-With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (*glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction).
+The 80% train is used for the model training and the 20% test is used for checking model performance on unseen dataset.
+With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (base R *glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction).
 The model family was binomial with a logit function.
-The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function ("rcompanion" package).
-The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package.
+The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function from "rcompanion" package [@mangiafico2021].
+The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package [@leeper2021].
 The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
-The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package.
+The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package [@kuhn2021].
 The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
@@
 Moreover, increasing a unit doy increases the probability of having a female plant by 0.4% (Table 2 and Figure4A).
-Similar trend is observed for weight as well as height, whereas the probability of being female increase by 0.2% (Figure 4B) and 0.1% (Figure 4C) when a unit of weight (g) and height (cm) increases, respectively.
+Similar trend is observed for weight (Figure 4B) as well as height (Figure 4C), whereas the probability of being female increase by 0.2% and 0.1% when a unit of weight (g) and height (cm) increases, respectively.
 
@@
 table_a %>% 
-  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal means of sex dimorphism logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.") %>% 
+  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth sex dimorphism logistic model.
+Factor pararemter values (crop and bareground) is shown compared to soybean.") %>%
   kable_styling(font_size = 10, 
@@
 The model accuracy evaluation accuracy in the 20% test dataset was 0.62 with a cutoff value for female and male plants of 0.43.
-The model classification showed a precision of 0.64, recall of 0.66, and a F means score of 0.65.
+The model classification showed a precision of 0.64, recall of 0.66, and a F1-score of 0.65.
 In addition, the area under the curve was 0.64.
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering pattern and sex varied within crops and cohort timings.
+Our study showed that Palmer amaranth biomass, height, flowering window and sex varied within crops and cohort timings.
 In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort.
@@
 In such conditions, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
-This is evident when comparing Palmer amaranth canopy shape, and its extended flowering pattern when growing into corn compared to soybean.
+This is evident when comparing Palmer amaranth canopy shape.
 The Palmer amaranth competition (e.g., light) strategy was to mimic the crop grow and development (Figure 6).
 These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018].
-Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive biotypes [@bravo2017].
+Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
 
@@
 Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages.
-Palmer amaranth was transplanted when corn canopy was nearly closed, which reduced Palmer amaranth competitiveness.
+Plants were transplanted at greater crop height, which reduced Palmer amaranth competitiveness.
 As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
-Under crop canopy (e.g., second cohort), Palmer amaranth flowering window was near to similar in corn and soybean.
-Palmer amaranth growing without crop competition produced the highest amounts of biomass and less extended flowering window.
+Palmer amaranth growing without crop competition produced the highest amounts of biomass.
 The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
-Nonetheless, Palmer amaranth produced 21% less biomass in second cohort compared to first cohort timing.
+Nonetheless, Palmer amaranth produced 17% less biomass in second cohort compared to first cohort timing.
 In a bareground study, early emerged Palmer amaranth without competition was 50% taller than late emerged plants [@webster2015].
@@
 The limited growth of Palmer amaranth at second cohort is likely a reduced plant response to day length, light availability and thermal units (e.g, growing degree days).
-It is hypothesize that reduced day length contributed to smaller plants at second cohort as well as shorter flowering period.
+The *Amaranthus* species are sensitive to photoperiod [@wu2014].
+It is hypothesize that reduced day length contributed to smaller plants at second cohort regardless the crop.
 A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021].
@@
 Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
-I our study, Palmer amaranth growing at first cohort accumulated an overall 42% more biomass when compared to second cohort.
-Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the cropping system.
+I our study, plants growing at first cohort accumulated an overall 42% more biomass when compared to second cohort.
+Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the crop.
 Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
@@
 Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
-Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize weed impact into agroecossystem [@davis2015].
+Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize the species impact into agroecossystem [@davis2015].
 
 An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering window.
-Our study suggests that Palmer amaranth flowering was slightly influenced by cropping systems and cohort timings.
+Our study suggests that Palmer amaranth flowering was slightly influenced by crops and cohort timings.
 Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering window, respectively.
+The shortest flowering window of second cohorts of Palmer amaranth growing in bareground highlighted the impact of photoperiod on flowering.
 When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing.
 Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
-We hypothesize that when growing in high competition (e.g., second cohort) and/or under low light/short day length, Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
-Early flower initiation is plant strategy when growing in stress conditions.
+We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
+A study has shown that non-native Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
+Early flower initiation could be also a plant strategy when growing in stress conditions.
 For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
-A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, and thus reduce plant seed production and exchange of resistant alleles [@jhala2021].
+A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, reducing plant seed production and/or exchange of resistant alleles [@jhala2021].
 Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].
@@
 It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
-Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each agroecossystem.
+Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each crop.
 Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth.
@@
 
-Our study demonstrated the Palmer amaranth plasticity to grow and develop into cropping-systems.
+Our study demonstrated the Palmer amaranth plasticity to grow and develop into arable land of US Midwest.
 Is likely that Palmer amaranth range will continue to expanding into new geographies.
+The migration of Palmer amaranth into the US Midwest will reshape the landscape as waterhemp and Palmer amaranth will share the same habitat.
+The presence of Palmer amaranth and waterhemp will increase the weed management complexity.
 Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal.
 Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development.
-Tactics that promote early-season crop advantage against Palmer amaranth, including early crop planting, crop rotation, plant width, preemergence applied herbicide, and crop residue (e.g.
-cover crops) would minimize the negative impact of Palmer amaranth in agroecossystems.
-The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on economy and sustainability.
+Long-term tactics that promote early-season crop advantage against Palmer amaranth, including diversity of crops in rotation, early/late crop planting, plant width, and crop residue (e.g.
+cover crops) would minimize the negative impact of Palmer amaranth to cropping systems.
+The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on sustainability and economy.
 
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing.
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
 Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit', out.width = "150mm", out.height = "70mm"}
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "150mm", out.height = "150mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B).
+```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
 From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}

### 2021-08-24 Frontiers/Frontiers.tex
as

@@
 Watson) is one of the most troublesome weed species in the United States.
-Palmer amaranth is endemic in the Southern United States but its range is expanding northward.
+Palmer amaranth is endemic to Southern United States but its range is expanding northward.
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
@@
 biology. Farm mechanization, adoption of conservation agriculture (e.g.,
-no-till), and reliance on herbicides for weed management are the main
-human-mediated selection of Palmer amaranth into cropping systems (Ward
-et al., 2013). On the other hand, Palmer amaranth is a prolific seed
-producer with a C4 photosynthetic apparatus (Wang et al., 1992). With a
-dioecy nature, Palmer amaranth male and female plants are obligate
-outcrosser species, increasing the chances of exchanging adaptive traits
-among plants (Jhala et al., 2021; Oliveira et al., 2018). Also, Palmer
-amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage
-systems (Price et al., 2011), and spread across locations through farm
-equipment (Sauer, 1972), manure (Hartzler and Anderson, 2016), animals
-(Farmer et al., 2017). The dispersal capacity of Palmer amaranth makes
-the species one of the most successful cases of weed adaptation to
-current cropping systems.
+no-till), and intensive use of herbicides for weed management are the
+main human-mediated selection of Palmer amaranth into cropping systems
+(Ward et al., 2013). On the other hand, Palmer amaranth is a prolific
+seed producer with a C4 photosynthetic apparatus (Wang et al., 1992).
+With a dioecy nature, Palmer amaranth male and female plants are
+obligate outcrosser species, increasing the chances of exchanging
+adaptive traits among plants (Jhala et al., 2021; Oliveira et al.,
+2018). Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive
+in no-tillage systems (Price et al., 2011), and spread across locations
+through farm equipment (Sauer, 1972), manure (Hartzler and Anderson,
+2016), animals (Farmer et al., 2017). The dispersal capacity of Palmer
+amaranth makes the species one of the most successful cases of weed
+adaptation to current cropping systems.
 
@@
 amaranth is still not well adapted compared to waterhemp
-(\emph{Amaranthus tuberculatus}) (Baker, 2021). Invasion and successful
+(\emph{Amaranthus tuberculatus}) (Baker, 2021). Waterhemp, an
+\emph{Amaranthus} species, is the most troublesome species in that
+geography (Tranel et al., 2011). Nonetheless, invasion and successful
 eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
@@
 America. Therefore, Palmer amaranth range of expansion into new habitats
-is increasing. It seems fated the need to manage Palmer amaranth in
-agronomic crops throughout multiple environments in the near future.
+can increase. It seems certain the need to manage Palmer amaranth in
+agronomic crops throughout novel environments in the near future.
 Strategies on Palmer amaranth management should encompass the
@@
 
-A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska. with
 no reported herbicide resistance according to Oliveira et al. (2021) was
@@
 five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE (('N, 'W)), Lincoln, NE ((41.16'N,
-96.42'W)), and Macomb, IL (('N, 'W)).
+(40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N,
+96.42'W), and Macomb, IL (XXX'N, XXX'W).
 
@@
 
- & Hybrid &  & DKC60-67 &  &  & \\
-
- & Seeding rate &  & 86487 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
- & Stage 1 cohort & V2-3 &  &  &  & \\
-
-\multirow{-5}{*}{\raggedright\arraybackslash } & Stage 2 cohort & V6-7 &  &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
+
+ & Seeding rate & 88956 & 86487 &  &  & \\
+
+ & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{c} &  &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\
 
@@
 
- & Stage 1 cohort & V1-2 &  &  &  & \\
-
- & Stage 2 cohort & V5-6 &  &  &  & \\
-
-\multirow{-5}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
-Planting day &  & May 10 to 20 & May 10 to 14 &  &  & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr\textsuperscript{d} + pyroxasulfone &  &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
+
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
 
@@
 
- & Type &  & Crete Silt Loam &  &  & \\
-
- & Ratio (sand-clay-silt) &  & 58-25-6.5 &  &  & \\
+ & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\
+
+ & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\
 
@@
 
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) &  & 3 &  &  & \\
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, g ai ha \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, g ai ha \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, g ai ha}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
 \end{tabular}}
@@
 plants were equidistantly placed (0.76 m apart) between rows within each
-crop (Figure 2). After a week, one plant was eliminated and one was
-kept, resulting in 24 plants per experimental unit and transplanting
-time (Figure 2). When needed, Palmer amaranth plants were supplied with
-water during the first week after transplanting.
+crop. After a week, one plant was eliminated and one was kept, resulting
+in 24 plants per experimental unit and transplanting time (Figure 2).
+When needed, Palmer amaranth plants were supplied with water during the
+first week after transplanting.
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1. Data analyses are stored at Zenodo (Oliveira, 2021).
+version 4.0.1 (Team, 2021). Data analyses are stored at Zenodo
+(Oliveira, 2021).
 
@@
 meet model assumption of normality. In the model, crop (bareground,
-corn, soybean) was the fixed effect and year nested with location the
-random effects. Analysis of variance was performed with \emph{anova}
-function from ``car'' package (Fox and Weisberg, 2018). Marginal means
-and compact letter display were estimated with \emph{emmeans} and
-\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
-2008).
+corn, soybean) and cohort time (first and second) were the fixed effects
+and year nested with location the random effects. Analysis of variance
+was performed with \emph{anova} function from ``car'' package (Fox and
+Weisberg, 2018). Marginal means and compact letter display were
+estimated with \emph{emmeans} and \emph{cld} from packages ``emmeans''
+(Lenth et al., 2021) and ``multcomp'' (Hothorn et al., 2008),
+respectively.
 
 The Palmer amaranth flowering timing was estimated as cumulative
-flowering across all location. Dataset was nested but id did not include
+flowering across all location. Dataset was nested but it did not include
 Grant, NE. Palmer amaranth cumulative flowering estimation was
@@
 80\% train is used for the model training and the 20\% test is used for
-checking how the model generalized on unseen dataset. With 80\% dataset,
-a binary response variable, male (0) and female (1), was fitted to a
-generalized linear model (\emph{glm} function) including day of year
-harvest, height, weight, crop and month as independent variables
+checking model performance on unseen dataset. With 80\% dataset, a
+binary response variable, male (0) and female (1), was fitted to a
+generalized linear model (base R \emph{glm} function) including day of
+year harvest, height, weight, crop and month as independent variables
 (without interaction). The model family was binomial with a logit
@@
 (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
-\emph{nagelkerke} function (``rcompanion'' package). The marginal
-effects computation was performed with Average Marginal Effects (AMEs)
-at every observed value of x and average across the results (Leeper,
-2017) using \emph{margins} function from ``margins'' package. The rest
-20\% dataset was predicted using \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function. The
-model quality prediction from the classification algorithm was measured
-with precision (\emph{precision} function), recall (\emph{recall}
-function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
-package. The precision determines the accuracy of positive predictions
-(female plants), recall determines the fraction of positives that were
-correctly identified, and F1-score is a weighted harmonic mean of
-precision and recall with the best score of 1 and the worst score of 0.
-F1-score conveys the balance between the precision and the recall
+\emph{nagelkerke} function from ``rcompanion'' package (Mangiafico,
+2021). The marginal effects computation was performed with Average
+Marginal Effects (AMEs) at every observed value of x and average across
+the results (Leeper, 2017) using \emph{margins} function from
+``margins'' package (Leeper et al., 2021). The rest 20\% dataset was
+predicted using \emph{predict} function with a cutoff estimation for
+male or female using \emph{performance} function. The model quality
+prediction from the classification algorithm was measured with precision
+(\emph{precision} function), recall (\emph{recall} function) and
+F1-score (\emph{f\_meas} function) using the ``yardstick'' package (Kuhn
+et al., 2021). The precision determines the accuracy of positive
+predictions (female plants), recall determines the fraction of positives
+that were correctly identified, and F1-score is a weighted harmonic mean
+of precision and recall with the best score of 1 and the worst score of
+0. F1-score conveys the balance between the precision and the recall
 (Yacouby and Axman, 2020).
@@
 increases the probability of having a female plant by 0.4\% (Table 2 and
-Figure4A). Similar trend is observed for weight as well as height,
-whereas the probability of being female increase by 0.2\% (Figure 4B)
-and 0.1\% (Figure 4C) when a unit of weight (g) and height (cm)
-increases, respectively.
+Figure4A). Similar trend is observed for weight (Figure 4B) as well as
+height (Figure 4C), whereas the probability of being female increase by
+0.2\% and 0.1\% when a unit of weight (g) and height (cm) increases,
+respectively.
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal means of sex dimorphism logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
 \centering
@@
 \toprule
-Term & AME & SE & Lower & Upper & Z-score & P-value\\
+Term & AME\textsuperscript{a} & SE\textsuperscript{b} & Lower & Upper & Z-score & P-value\\
 \midrule
@@
 with a cutoff value for female and male plants of 0.43. The model
-classification showed a precision of 0.64, recall of 0.66, and a F means
-score of 0.65. In addition, the area under the curve was 0.64.
+classification showed a precision of 0.64, recall of 0.66, and a
+F1-score of 0.65. In addition, the area under the curve was 0.64.
 
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering pattern
+Our study showed that Palmer amaranth biomass, height, flowering window
 and sex varied within crops and cohort timings. In general, first cohort
@@
 to adapt upon the agroecosystem. This is evident when comparing Palmer
-amaranth canopy shape, and its extended flowering pattern when growing
-into corn compared to soybean. The Palmer amaranth competition (e.g.,
-light) strategy was to mimic the crop grow and development (Figure 6).
-These results suggests that Palmer amaranth can fast evolve life-history
+amaranth canopy shape. The Palmer amaranth competition (e.g., light)
+strategy was to mimic the crop grow and development (Figure 6). These
+results suggests that Palmer amaranth can fast evolve life-history
 traits to adapt into cropping systems and cultural practices, which was
@@
 crops as breeding more competitive crop varieties is likely to select
-more competitive biotypes (Bravo et al., 2017).
+more competitive weed biotypes (Bravo et al., 2017).
 
 Palmer amaranth grow and development in second cohort was limited due to
-the crop competitive ability at advanced development stages. Palmer
-amaranth was transplanted when corn canopy was nearly closed, which
-reduced Palmer amaranth competitiveness. As a result, Palmer amaranth
-height and biomass was lower compared to its first cohort. Under crop
-canopy (e.g., second cohort), Palmer amaranth flowering window was near
-to similar in corn and soybean. Palmer amaranth growing without crop
-competition produced the highest amounts of biomass and less extended
-flowering window. The Palmer amaranth strategy in bareground was to
-invest biomass in growing plant width and height. Nonetheless, Palmer
-amaranth produced 21\% less biomass in second cohort compared to first
-cohort timing. In a bareground study, early emerged Palmer amaranth
-without competition was 50\% taller than late emerged plants (Webster
-and Grey, 2015). These results suggest that crop competition is not the
-only factor limiting late Palmer amaranth establishment. The limited
-growth of Palmer amaranth at second cohort is likely a reduced plant
-response to day length, light availability and thermal units (e.g,
-growing degree days). It is hypothesize that reduced day length
-contributed to smaller plants at second cohort as well as shorter
-flowering period. A study in North Carolina and Illinois predicted that
-less than 10\% Palmer amaranth seedlings emergence occurred after June
-(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
-on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
-yields was higher when plants were established near to crop planting.
+the crop competitive ability at advanced development stages. Plants were
+transplanted at greater crop height, which reduced Palmer amaranth
+competitiveness. As a result, Palmer amaranth height and biomass was
+lower compared to its first cohort. Palmer amaranth growing without crop
+competition produced the highest amounts of biomass. The Palmer amaranth
+strategy in bareground was to invest biomass in growing plant width and
+height. Nonetheless, Palmer amaranth produced 17\% less biomass in
+second cohort compared to first cohort timing. In a bareground study,
+early emerged Palmer amaranth without competition was 50\% taller than
+late emerged plants (Webster and Grey, 2015). These results suggest that
+crop competition is not the only factor limiting late Palmer amaranth
+establishment. The limited growth of Palmer amaranth at second cohort is
+likely a reduced plant response to day length, light availability and
+thermal units (e.g, growing degree days). The \emph{Amaranthus} species
+are sensitive to photoperiod (Wu and Owen, 2014). It is hypothesize that
+reduced day length contributed to smaller plants at second cohort
+regardless the crop. A study in North Carolina and Illinois predicted
+that less than 10\% Palmer amaranth seedlings emergence occurred after
+June (Piskackova et al., 2021). In addition, Palmer amaranth negative
+impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
+2015) yields was higher when plants were established near to crop
+planting.
 
@@
 correlation between Palmer amaranth biomass and seed production
-(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
-amaranth growing at first cohort accumulated an overall 42\% more
-biomass when compared to second cohort. Therefore, Palmer amaranth
-plants growing in the second cohort is likely to produce less seeds
-regardless the cropping system. Our observation is consistent with the
-findings that first Palmer amaranth cohort produced 50\% more seeds per
-plant than Palmer amaranth plants established six weeks later in
-bareground (Webster and Grey, 2015). Still, seed production at second
-cohort is likely to replenish the soil seedbank. Seed production and
-deposition in the seedbank is also a key factor for species perpetuation
-(Menges, 1987). Palmer amaranth can produce hundred thousands seeds per
-plant (Schwartz et al., 2016; Keeley et al., 1987), and stay viable
-buried in the seedbank for at least 36 months (Sosnoskie et al., 2013).
-Therefore, preventing Palmer amaranth seed production or/and seed
-migration to its non-native habitat is an essential strategy to minimize
-weed impact into agroecossystem (Davis et al., 2015).
+(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, plants
+growing at first cohort accumulated an overall 42\% more biomass when
+compared to second cohort. Therefore, Palmer amaranth plants growing in
+the second cohort is likely to produce less seeds regardless the crop.
+Our observation is consistent with the findings that first Palmer
+amaranth cohort produced 50\% more seeds per plant than Palmer amaranth
+plants established six weeks later in bareground (Webster and Grey,
+2015). Still, seed production at second cohort is likely to replenish
+the soil seedbank. Seed production and deposition in the seedbank is
+also a key factor for species perpetuation (Menges, 1987). Palmer
+amaranth can produce hundred thousands seeds per plant (Schwartz et al.,
+2016; Keeley et al., 1987), and stay viable buried in the seedbank for
+at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
+Palmer amaranth seed production or/and seed migration to its non-native
+habitat is an essential strategy to minimize the species impact into
+agroecossystem (Davis et al., 2015).
 
@@
 understanding its flowering window. Our study suggests that Palmer
-amaranth flowering was slightly influenced by cropping systems and
-cohort timings. Palmer amaranth growing in bareground and corn resulted
-in the overall shortest and longest flowering window, respectively. When
+amaranth flowering was slightly influenced by crops and cohort timings.
+Palmer amaranth growing in bareground and corn resulted in the overall
+shortest and longest flowering window, respectively. The shortest
+flowering window of second cohorts of Palmer amaranth growing in
+bareground highlighted the impact of photoperiod on flowering. When
 growing in soybean, Palmer amaranth flowering window was similar to
@@
 ecological factors (Lang, 1965). We hypothesize that when growing in
-high competition (e.g., second cohort) and/or under low light/short day
-length, Palmer amaranth plants tend to initiate flowering early, as well
-as having an extended flowering window. Early flower initiation is plant
-strategy when growing in stress conditions. For example, when growing
-under water stress, early flowering in Palmer amaranth resulted in a
-mismatch between female and male plants by seven days (Mesgaran et al.,
-2021). A mismatch in Palmer amaranth male and female flowering period
-can minimize plant outcross, and thus reduce plant seed production and
-exchange of resistant alleles (Jhala et al., 2021). Sex dimorphism
-manipulation is considered a potential ecological pest control
-(McFarlane et al., 2018; Schliekelman et al., 2005).
+high competition (e.g., second cohort), Palmer amaranth plants tend to
+initiate flowering early, as well as having an extended flowering
+window. A study has shown that non-native Palmer amaranth initiated
+flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
+Early flower initiation could be also a plant strategy when growing in
+stress conditions. For example, when growing under water stress, early
+flowering in Palmer amaranth resulted in a mismatch between female and
+male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
+amaranth male and female flowering period can minimize plant outcross,
+reducing plant seed production and/or exchange of resistant alleles
+(Jhala et al., 2021). Sex dimorphism manipulation is considered a
+potential ecological pest control (McFarlane et al., 2018; Schliekelman
+et al., 2005).
 
@@
 soybean and bareground compared to corn, which might linked to plant
-competition strategy in each agroecossystem. Our results showed the
-influence of life-history and ecological traits on sexual dimorphism in
-Palmer amaranth. Sexual dimorphism is documented in other dioecious
-species (Barrett and Hough, 2013). For example, stronger female plant
+competition strategy in each crop. Our results showed the influence of
+life-history and ecological traits on sexual dimorphism in Palmer
+amaranth. Sexual dimorphism is documented in other dioecious species
+(Barrett and Hough, 2013). For example, stronger female plant
 competition and greater male tolerance to herbivory was reported in
@@
 Our study demonstrated the Palmer amaranth plasticity to grow and
-develop into cropping-systems. Is likely that Palmer amaranth range will
-continue to expanding into new geographies. Therefore, preventive
-management is a priority to minimizing Palmer amaranth dispersal.
-Reactive management should focus on early-season management programs,
-which would have a large negative effect on Palmer amaranth growth and
-development. Tactics that promote early-season crop advantage against
-Palmer amaranth, including early crop planting, crop rotation, plant
-width, preemergence applied herbicide, and crop residue (e.g.~cover
-crops) would minimize the negative impact of Palmer amaranth in
-agroecossystems. The aggressiveness and differential Palmer amaranth
-adaptation to agroecosystem is striking and require national efforts to
-minimize the species impact on economy and sustainability.
+develop into arable land of US Midwest. Is likely that Palmer amaranth
+range will continue to expanding into new geographies. The migration of
+Palmer amaranth into the US Midwest will reshape the landscape as
+waterhemp and Palmer amaranth will share the same habitat. The presence
+of Palmer amaranth and waterhemp will increase the weed management
+complexity. Therefore, preventive management is a priority to minimizing
+Palmer amaranth dispersal. Reactive management should focus on
+early-season management programs, which would have a large negative
+effect on Palmer amaranth growth and development. Long-term tactics that
+promote early-season crop advantage against Palmer amaranth, including
+diversity of crops in rotation, early/late crop planting, plant width,
+and crop residue (e.g.~cover crops) would minimize the negative impact
+of Palmer amaranth to cropping systems. The aggressiveness and
+differential Palmer amaranth adaptation to agroecosystem is striking and
+require national efforts to minimize the species impact on
+sustainability and economy.
 
@@
 
+\leavevmode\hypertarget{ref-kuhn2021}{}%
+Kuhn, M., Vaughan, D., and RStudio (2021). \emph{Yardstick: {Tidy
+Characterizations} of {Model Performance}}. Available at:
+\url{https://CRAN.R-project.org/package=yardstick} {[}Accessed August
+24, 2021{]}.
+
 \leavevmode\hypertarget{ref-kupper2017}{}%
@@
 
+\leavevmode\hypertarget{ref-leeper2021}{}%
+Leeper, T. J., Arnold, J., Arel-Bundock, V., and Long, J. A. (2021).
+\emph{Margins: {Marginal Effects} for {Model Objects}}. Available at:
+\url{https://CRAN.R-project.org/package=margins} {[}Accessed August 24,
+2021{]}.
+
+\leavevmode\hypertarget{ref-lenth2021}{}%
+Lenth, R. V., Buerkner, P., Herve, M., Love, J., Riebl, H., and
+Singmann, H. (2021). \emph{Emmeans: {Estimated Marginal Means}, aka
+{Least}-{Squares Means}}. Available at:
+\url{https://CRAN.R-project.org/package=emmeans} {[}Accessed August 24,
+2021{]}.
+
 \leavevmode\hypertarget{ref-lindsay2017}{}%
@@
 6.
+
+\leavevmode\hypertarget{ref-mangiafico2021}{}%
+Mangiafico, S. (2021). \emph{Rcompanion: {Functions} to {Support
+Extension Education Program Evaluation}}. Available at:
+\url{https://CRAN.R-project.org/package=rcompanion} {[}Accessed August
+24, 2021{]}.
 
@@
 
+\leavevmode\hypertarget{ref-rcoreteam2021}{}%
+Team, R. C. (2021). \emph{R: {The R Project} for {Statistical
+Computing}}. {Vienna, Austria}: {R Foundation for Statistical Computing}
+Available at: \url{https://www.r-project.org/} {[}Accessed August 24,
+2021{]}.
+
+\leavevmode\hypertarget{ref-tranel2011}{}%
+Tranel, P. J., Riggins, C. W., Bell, M. S., and Hager, A. G. (2011).
+Herbicide {Resistances} in {Amaranthus} tuberculatus: {A Call} for {New
+Options}. \emph{J. Agric. Food Chem.} 59, 5808--5812.
+doi:\href{https://doi.org/10.1021/jf103797n}{10.1021/jf103797n}.
+
 \leavevmode\hypertarget{ref-vanwychen2020}{}%
@@
 doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.
+
+\leavevmode\hypertarget{ref-wu2014}{}%
+Wu, C., and Owen, M. D. K. (2014). When {Is} the {Best Time} to
+{Emerge}: {Reproductive Phenology} and {Success} of {Natural Common
+Waterhemp} ({Amaranthus} rudis) {Cohorts} in the {Midwest United
+States}? \emph{Weed Science} 62, 107--117. Available at:
+\url{http://www.jstor.org/stable/43700638}.
 
@@
 
-\caption{Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
+\caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
 \end{figure}
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
 \end{figure}
@@
 
-\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B).
+\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
 From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}

### 2021-08-24 Frontiers/table.tex
as

--- 2021-08-23 Frontiers/Frontiers.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
-%                                              %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Version 3.4 Generated 2018/06/15 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-
-\documentclass[utf8]{frontiersSCNS}
-
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-
-\linenumbers
-
-% BELOW TAKEN FROM rticles plos template
-%
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% hyperref package, useful for hyperlinks
-\usepackage{hyperref}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-% Sweave(-like)
-\usepackage{fancyvrb}
-\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
-\newenvironment{Schunk}{}{}
-\DefineVerbatimEnvironment{Code}{Verbatim}{}
-\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-\newenvironment{CodeChunk}{}{}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% Below is from frontiers
-%
-\bibliographystyle{frontiersinSCNS}
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace}
-%\doublespacing
-
-% Leave a blank line between paragraphs instead of using \\
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
-
-% Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\def\Authors{
-  Maxwel C Oliveira\,\textsuperscript{1},
-  Amit J Jhala\,\textsuperscript{2},
-  Mark Bernards\,\textsuperscript{3},
-  Chris Proctor\,\textsuperscript{2},
-  Strahinja Stepanovic\,\textsuperscript{2},
-  Rodrigo Werle\,\textsuperscript{1*}}
-
-\def\Address{
-
-  \textsuperscript{1} Department of Agronomy, University of
-Wisconsin-Madison,  Madison,  Wisconsin,  United States
-
-  \textsuperscript{2} Department of Agronomy and
-Horticulture, University of
-Nebraska-Lincoln,  Lincoln,  Nebraska,  United States
-
-  \textsuperscript{3} Department of Agronomy, Western Illinois
-University,  Macomb,  Illinois,  United States
-  }
-
-  \def\firstAuthorLast{Oliveira {et~al.}}
-
-  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Department of Agronomy,
-University of Wisconsin-Madison, United States\\1575 Linden
-Dr\\Madison, Wisconsin, 53705 United
-States}\def\corrEmail{\href{mailto:rwerle@uwisc.edu}{\nolinkurl{rwerle@uwisc.edu}}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
-palmeri}) adaptation to US Midwest agroecosystems}
-\author[\firstAuthorLast]{\Authors}
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
-
-\maketitle
-
-\begin{abstract}
-
-Abstract length and content varies depending on article type. Refer to 
-\url{http://www.frontiersin.org/about/AuthorGuidelines} for abstract requirement
-and length according to article type.
-
-%All article types: you may provide up to 8 keywords; at least 5 are mandatory.
-\tiny
- \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}
-
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
-ranked as one of the most economically detrimental weed species to
-cropping systems in the United States (Van Wychen, 2020). Unmanaged
-Palmer amaranth plants compete for water, light, and nutrients, which
-can drastically impact crop yields (Berger et al., 2015). For example,
-Palmer amaranth has been documented to reduce up to 91\%, 68\%, and 54\%
-corn (Massinga et al., 2001), soybean (Klingaman and Oliver, 1994), and
-cotton (Morgan et al., 2001) yields, respectively. Moreover, Palmer
-amaranth has shown a remarkable capacity to evolve resistance to
-herbicides. To date, Palmer amaranth has evolved resistance to eight
-herbicide sites of action (Heap, 2021), increasing the weed management
-complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
-economical and environmental risk to sustainable agriculture.
-
-Palmer amaranth is a fast growing summer annual forb indigenous to the
-Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
-threat to US agriculture in the 1990s. Palmer amaranth weediness is
-likely a result of human-assisted selection in combination with plant
-biology. Farm mechanization, adoption of conservation agriculture (e.g.,
-no-till), and reliance on herbicides for weed management are the main
-human-mediated selection of Palmer amaranth into cropping systems (Ward
-et al., 2013). On the other hand, Palmer amaranth is a prolific seed
-producer with a C4 photosynthetic apparatus (Wang et al., 1992). With a
-dioecy nature, Palmer amaranth male and female plants are obligate
-outcrosser species, increasing the chances of exchanging adaptive traits
-among plants (Jhala et al., 2021). Also, Palmer amaranth small seed size
-(e.g, 1 mm) tend to thrive in no-tillage systems (Price et al., 2011),
-and spread across locations through farm equipment (Sauer, 1972), manure
-(Hartzler and Anderson, 2016), animals (Farmer et al., 2017). The
-dispersal capacity of Palmer amaranth makes the species one of the most
-successful cases of weed adaptation to current cropping systems.
-
-Light and temperature are the main environment requirements for Palmer
-amaranth successful grow and development (Jha et al., 2010). Palmer
-amaranth is reported with an extended germination period (Ward et al.,
-2013). Germination of Palmer amaranth was triggered by 18 C soil
-temperature at 5 cm depth (Keeley et al., 1987), and optimal germination
-and biomass production occurred at 35/30 C day and night temperatures
-(Guo and Al-Khatib, 2003). In addition, Palmer amaranth establishment is
-human-mediated by tillage timings and preemergence-applied herbicides
-(Chahal et al., 2021), which can result in weed germination shifts
-(Sbatella and Wilson, 2010). Water has not shown to limit Palmer
-amaranth fitness. Under continuous water stress, Palmer amaranth
-survived and produced at least 14000 seeds plant-1 (Chahal et al.,
-2018). Seeds from Palmer amaranth growing with limited water conditions
-were heavier, less dormant, and prompt for germination (Matzrafi et al.,
-2021). Growing conditions and management practices also influence Palmer
-amaranth sex dimorphism and flowering pattern (Korres et al., 2017;
-Rumpa et al., 2019). Therefore, Palmer amaranth has shown plasticity to
-evolve and fast adapt under selection pressure. Palmer amaranth
-invasion/adaptation into new habitats is likely to increase with the
-global temperature warming. Currently, it is estimated that the greatest
-climatic risk of Palmer amaranth establishment are agronomic crops in
-Australia and Sub-Sahara Africa (Kistner and Hatfield, 2018).
-Temperature is a key factor limiting Palmer amaranth expansion to cooler
-geographies (Briscoe Runquist et al., 2019); however, under future
-climate change Palmer amaranth is likely to expand northward into Canada
-and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
-al., 2019).
-
-Palmer amaranth is already found in agronomic crops of South America
-(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
-et al., 2021). In the United States, Palmer amaranth is well established
-in the Cotton Belt (Garetson et al., 2019; Bagavathiannan and
-Norsworthy, 2016) in the southern United States but its range is
-expanding northward. For example, herbicide resistant Palmer amaranth is
-widespread in Nebraska (Oliveira et al., 2021). There are some reported
-cases of Palmer amaranth in Michigan (Kohrt et al., 2017) and
-Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
-amaranth can cause high damage to soybean fields in Illinois (Davis et
-al., 2015), which is concerning as soybean along with corn make most of
-US Midwest agronomic hectares. In Iowa, a study showed that Palmer
-amaranth is still not well adapted compared to waterhemp
-(\emph{Amaranthus tuberculatus}) (Baker, 2021). Invasion and successful
-eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
-2021). Palmer amaranth infestations have not been detected in Canada;
-however, Palmer amaranth seeds were detected in sweet potato slips in
-the country (Page et al., 2021). Palmer amaranth is still not as well
-adapted and established to Northern as it is in the Southern North
-America. Therefore, its range of expansion into new habitats is
-increasing. It seems fated the need to manage Palmer amaranth in
-agronomic crops throughout multiple environments in the near future.
-Strategies on Palmer amaranth management should encompass the
-agroecosystem level but not only attempts to eradicate the weed. Most
-tactics to manage Palmer amaranth are based on technology fixes (Scott,
-2011), which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management. Palmer amaranth management should be built on
-minimizing the species ability to adapt, grow and develop into
-agroecossystems.
-
-In the southeastern US, early growing Palmer amaranth is well known to
-have a higher impact on cotton yields compared to late established
-plants (MacRae et al., 2013). In the northern states, Palmer amaranth
-impact on the agroecosystem is recent. Studies investigating Palmer
-amaranth in those locations are limited due to the plant classification
-as noxious weed species in some northern states (Yu et al., 2021).
-Nonetheless, the continuous Palmer amaranth dispersal and potential
-establishment across the northern United States is concerning and
-warrants investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under
-different agroecosystems can enhance our knowledge on species adaptation
-and management practices. It can also aid in designing proactive and
-ecological tactics to limit the species range expansion, reduce its
-negative impact, and developing resilient and sustainable farming
-systems (MacLaren et al., 2020). Therefore, the objective of this study
-was to investigate the flowering pattern, gender, biomass production,
-and height of Palmer amaranth cohorts growing under corn, soybean and
-bareground across five locations in the United States Midwest.
-
-\hypertarget{material-and-methods}{%
-\section*{Material and Methods}\label{material-and-methods}}
-\addcontentsline{toc}{section}{Material and Methods}
-
-\hypertarget{plant-material-and-growing-conditions}{%
-\subsection*{Plant material and growing
-conditions}\label{plant-material-and-growing-conditions}}
-\addcontentsline{toc}{subsection}{Plant material and growing conditions}
-
-A Palmer amaranth accession (Per1) from Perkins County, Nebraska. with
-no reported herbicide resistance according to Oliveira et al. (2021) was
-selected for this study. Three weeks prior to the onset of the field
-experiments, seeds were planted in plastic trays containing potting-mix.
-Emerged seedlings (1 cm) were transplanted into 200 cm-3 plastic pots (a
-plant pot-1). Palmer amaranth seedlings were supplied with adequate
-water and kept under greenhouse conditions at the University of
-Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois
-University; and kept outdoors in the Perkins extension office at Grant,
-NE until the 2-3 leaf stage (5 to 8 cm height) when they were
-transported to the field.
-
-\hypertarget{field-study}{%
-\subsection*{Field study}\label{field-study}}
-\addcontentsline{toc}{subsection}{Field study}
-
-The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE ('N,
-'W), Grant, NE (('N, 'W)), Lincoln, NE (('N, 'W)), and Macomb, IL (('N,
-'W)).
-
-Fields were conventionally tilled prior to crop planting. Corn hybrid
-and soybean varieties were planted in 76-cm row spacing (Table 1).
-Monthly mean air temperature and total precipitation were obtained using
-Daymet weather data from June through September across the five
-locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)
-
 \begin{table}[!h]
@@
 
- & Hybrid &  & DKC60-67 &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
 
- & Seeding rate &  & 86487 &  &  & \\
+ & Seeding rate & 88956 & 86487 &  &  & \\
 
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+ & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{c} &  &  & \\
 
- & Stage 1 cohort & V2-3 &  &  &  & \\
+ & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
 
-\multirow{-5}{*}{\raggedright\arraybackslash } & Stage 2 cohort & V6-7 &  &  &  & \\
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\
 
@@
 
- & Stage 1 cohort & V1-2 &  &  &  & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr\textsuperscript{d} + pyroxasulfone &  &  & \\
 
- & Stage 2 cohort & V5-6 &  &  &  & \\
+ & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
 
-\multirow{-5}{*}{\raggedright\arraybackslash } & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
 
-Planting day &  & May 10 to 20 & May 10 to 14 &  &  & \\
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
 
@@
 
- & Type &  & Crete Silt Loam &  &  & \\
+ & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\
 
- & Ratio (sand-clay-silt) &  & 58-25-6.5 &  &  & \\
+ & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\
 
@@
 
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) &  & 3 &  &  & \\
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
 \bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
 \end{tabular}}
@@
 
-The field experimental units were three adjacent 9.1 m wide (12 rows at
-76.2 cm row spacing) by 10.7 m long. The experimental design were
-arranged in factorial design with three crops, two transplanting times
-simulating two cohorts, repeated across five locations. Each
-experimental unit was planted with corn, soybean, or kept under
-bareground. The two transplanting timings were June 1 (first cohort) and
-July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
-seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
-plants were equidistantly placed (0.76 m apart) between rows within each
-crop (Figure 2). After a week, one plant was eliminated and one was
-kept, resulting in 24 plants per experimental unit and transplanting
-time (Figure 2). When needed, Palmer amaranth plants were supplied with
-water during the first week after transplanting.
-
-After transplanting, Palmer amaranth flowering was monitored until the
-end of the study. When a plant flowered, the day was recorded, plant sex
-was identified (male or female), plant height was measured from soil
-surface to the top of plant. Also, aboveground plant biomass was harvest
-near soil surface and oven dried at 65 C until reaching constant weight
-before weighing (g plant-1 was recorded).
-
-Plants had to be harvested at flowering because Palmer amaranth is
-neither endemic in Wisconsin nor in Illinois. In our study, all
-locations follow the methodology of plant harvest at flowering
-initiation, except in Grant, NE. In this location, all Palmer amaranth
-plants were harvest at once on July 06, 2018 and 2019 (first cohort),
-and on August 17, 2018 and on July 31, 2019 (second cohort).
-
-\hypertarget{statistical-analyses}{%
-\subsection*{Statistical analyses}\label{statistical-analyses}}
-\addcontentsline{toc}{subsection}{Statistical analyses}
-
-The statistical analyses were performed using R statistical software
-version 4.0.1. Data analyses are stored at Zenodo (Oliveira, 2021).
-
-Analyses of Palmer amaranth height and biomass were performed with a
-linear mixed model using \emph{lmer} function from ``lme4'' package
-(Bates et al., 2015). Plant height and biomass were log transformed to
-meet model assumption of normality. In the model, crop (bareground,
-corn, soybean) was the fixed effect and year nested with location the
-random effects. Analysis of variance was performed with \emph{anova}
-function from ``car'' package (Fox and Weisberg, 2018). Marginal means
-and compact letter display were estimated with \emph{emmeans} and
-\emph{cld} from packages ``emmeans'' and ``multcomp'' (Hothorn et al.,
-2008).
-
-The Palmer amaranth flowering timing was estimated as cumulative
-flowering across all location. Dataset was nested but id did not include
-Grant, NE. Palmer amaranth cumulative flowering estimation was
-determined using an asymmetrical three parameter log logistic Weibull
-model of the drc package (Ritz et al., 2015).
-
-\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\]
-
-In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
-\emph{d} is the upper limit (set to 100), and \emph{e} is the inflection
-point, and \emph{x} day of year (doy).
-
-The doy for 10, 50, and 90\% Palmer amaranth cumulative flowering were
-determined using the \emph{ED} function of drc package. Also, the 10,
-50, and 90\% Palmer amaranth cumulative flowering were compared among
-crops and cohorts using the \emph{EDcomp} function of drc package. The
-EDcomp function compares the ratio of cumulative flowering using
-t-statistics, where P-value \textless{} 0.05 indicates that we fail to
-reject the null hypothesis.
-
-Palmer amaranth sex was fitted to a binary logistic regression
-(Bangdiwala, 2018). Binary logistic regression is used for predicting
-binary classes, such as the probability of a plant being female in a
-dioecious species. Prior to the analysis, all missing values were
-removed from the dataset. Also, data from Grant was not used in this
-analysis due to the uniform plant harvesting at that location. The
-complete dataset was splitted into 80\% train and 20\% test data. The
-80\% train is used for the model training and the 20\% test is used for
-checking how the model generalized on unseen dataset. With 80\% dataset,
-a binary response variable, male (0) and female (1), was fitted to a
-generalized linear model (\emph{glm} function) including day of year
-harvest, height, weight, crop and month as independent variables
-(without interaction). The model family was binomial with a logit
-function. The model fit was assessed through pseudo R-squared values
-(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
-\emph{nagelkerke} function (``rcompanion'' package). The marginal
-effects computation was performed with Average Marginal Effects (AMEs)
-at every observed value of x and average across the results (Leeper,
-2017) using \emph{margins} function from ``margins'' package. The rest
-20\% dataset was predicted using \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function. The
-model quality prediction from the classification algorithm was measured
-with precision (\emph{precision} function), recall (\emph{recall}
-function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
-package. The precision determines the accuracy of positive predictions
-(female plants), recall determines the fraction of positives that were
-correctly identified, and F1-score is a weighted harmonic mean of
-precision and recall with the best score of 1 and the worst score of 0.
-F1-score conveys the balance between the precision and the recall
-(Yacouby and Axman, 2020).
-
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
-
-\hypertarget{palmer-amaranth-height-and-biomass}{%
-\subsection*{Palmer amaranth height and
-biomass}\label{palmer-amaranth-height-and-biomass}}
-\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
-
-Palmer amaranth plants accumulated more biomass when growing in
-bareground compared to plants growing in soybean and corn (figure 3A).
-Palmer amaranth plants in the first cohort produced 75.5, 28.3 g, and
-16.3 g plant-1 in bareground, soybean and corn, respectively. Plants
-from the second cohort produced 62.6 g plant in bareground, followed by
-6.3 g plant in soybean, and 1.4 g plant in corn.
-
-Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (figure 3B). Palmer amaranth plants from the first
-cohort were on average 69.2 cm tall in bareground, which was not
-different from the 70.7 cm tall plants from the second cohort timing (P
-= 0.74). In addition, no difference in Palmer amaranth height (69.3 cm)
-was detected from first cohort plants in soybean to first and second
-cohort plants in bareground (P \textgreater{} 0.75). Palmer amaranth
-plants from the second cohort were nearly 10 cm lower compared to the
-first cohort in soybeans (P = 0.04). The tallest (first cohort) and
-smallest (second cohort) Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
-
-\hypertarget{palmer-amaranth-cumulative-flowering}{%
-\subsection*{Palmer amaranth cumulative
-flowering}\label{palmer-amaranth-cumulative-flowering}}
-\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}
-
-Palmer amaranth plants from the first cohort growing in corn resulted in
-a longer flowering window compared to plants growing in bareground and
-soybean (Figure 4A). The 10\% cumulative Palmer amaranth flowering in
-soybean, bareground and corn occurred at the end of June. Palmer
-amaranth reached 10\% flowering in soybean, bareground and corn at doy
-180, 180.9 and 181.7, respectively. The 50\% Palmer amaranth cumulative
-flowering occurred in July. Palmer amaranth reached 50\% flowering in
-bareground, soybean and corn at doy 193.4, 194.8, and 206.6,
-respectively. Similar trend was observed at 90\% Palmer amaranth
-cumulative flowering. Palmer amaranth reached 90\% flowering at doy
-252.6 in corn (early September), which was 38 and 32 days after reaching
-90\% flowering in bareground and soybean, respectively.
-
-Palmer amaranth cumulative flowering at the second cohort ranged from
-mid July to mid September (Figure 4B). Palmer amaranth growing in the
-bareground resulted in earlier flowering time compared to soybean and
-corn. Palmer amaranth growing in bareground reached 10\%, 50\%, and 90\%
-flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer
-amaranth growing in soybean reached 10\% flowering at doy 210.9, which
-was 6 days prior to corn (\emph{P}-value = 0.00). Similar trend was
-observed at 50\% flowering, whereas Palmer amaranth reached 50\%
-flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; \emph{P}
-= 0.00). The 90\% Palmer amaranth cumulative flowering occurred at same
-day in corn (260.9) and soybean (260.5; \emph{P} = 0.66).
-
-\hypertarget{palmer-amaranth-gender}{%
-\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
-\addcontentsline{toc}{subsection}{Palmer amaranth gender}
-
-The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from
-McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
-likelihood ratio test showed a p-value of \textless{} 0.00. The average
-marginal effects showed that Palmer amaranth growing in corn resulted in
-14.8\% less females plants (Table 2). Moreover, increasing a unit doy
-increases the probability of having a female plant by 0.4\% (Table 2 and
-Figure4A). Similar trend is observed for weight as well as height,
-whereas the probability of being female increase by 0.2\% (Figure 4B)
-and 0.1\% (Figure 4C) when a unit of weight (g) and height (cm)
-increases, respectively.
-
 \begin{table}[!h]
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal means of gender logistic model. Factor pararemter values (e.g. crop) is shown related to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
 \centering
@@
 \end{table}
-
-The model accuracy evaluation accuracy in the 20\% test dataset was 0.62
-with a cutoff value for female and male plants of 0.43. The model
-classification showed a precision of 0.64, recall of 0.66, and a F means
-score of 0.65. In addition, the area under the curve was 0.64.
-
-\hypertarget{discussion}{%
-\section*{Discussion}\label{discussion}}
-\addcontentsline{toc}{section}{Discussion}
-
-Our study showed that Palmer amaranth biomass, height, flowering pattern
-and gender varied within agroecossystems and cohort timings. In general,
-first cohort of Palmer amaranth plants were heavier and taller when
-compared to the second cohort. At first cohort, resources (e.g., soil
-nutrients) and conditions (e.g., light) were more timely available for
-the species. High biomass and taller Palmer amaranth plants are likely a
-weed strategy to compete for light in between crop rows in absence of
-canopy. In such conditions, Palmer amaranth showed an extraordinary
-plasticity to adapt upon the agroecosystem. This is evident when
-comparing Palmer amaranth canopy shape, and its extended flowering
-pattern when growing into corn compared to soybean. The Palmer amaranth
-competition strategy was to mimic the crop grow and development (Figure
-6). These results suggests that Palmer amaranth can fast evolve
-life-history traits to adapt into cropping systems and cultural
-practices, which was also showed in a study varying nitrogen
-fertilization (Bravo et al., 2018). Our results highlight the Palmer
-amaranth as a threat to field crops as breeding more competitive crop
-varieties is likely to select more competitive biotypes (Bravo et al.,
-2017).
-
-Palmer amaranth grow and development in second cohort was limited due to
-the crop competitive ability at advanced development stages. Palmer
-amaranth was transplanted when corn canopy was nearly closed, which
-reduced Palmer amaranth competitiveness. As a result, Palmer amaranth
-height and biomass was lower compared to its first cohort. Under crop
-canopy (e.g., second cohort), Palmer amaranth flowering window was near
-to similar in corn and soybean. Palmer amaranth growing without crop
-competition produced the highest amounts of biomass and less extended
-flowering window. The Palmer amaranth strategy in bareground was to
-invest biomass in growing plant width and height. Nonetheless, Palmer
-amaranth produced 21\% less biomass in second cohort compared to first
-cohort timing. In a bareground study, early emerged Palmer amaranth
-without competition was 50\% taller than late emerged plants (Webster
-and Grey, 2015). These results suggest that crop competition is not the
-only factor limiting late Palmer amaranth establishment. The limited
-growth of Palmer amaranth at second cohort is likely a reduced plant
-response to day length, light availability and thermal units (e.g,
-growing degree days). It is hypothesize that reduced day length
-contributed to smaller plants at second cohort as well as shorter
-flowering period. A study in North Carolina and Illinois predicted that
-less than 10\% Palmer amaranth seedlings emergence occurred after June
-(Piskackova et al., 2021). In addition, Palmer amaranth negative impact
-on soybean (Korres et al., 2020) and cotton (Webster and Grey, 2015)
-yields was higher when plants were established near to crop planting.
-
-Seed production was not evaluated due to plant harvest at initiation of
-flowering. Nonetheless, it is well documented a strong positive
-correlation between Palmer amaranth biomass and seed production
-(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, Palmer
-amaranth growing at first cohort accumulated an overall 36\% more
-biomass when compared to second cohort. Therefore, Palmer amaranth
-plants growing in the second cohort is likely to produce less seeds
-regardless the cropping system. Our observation is consistent with the
-findings that first Palmer amaranth cohort produced 50\% more seeds per
-plant than Palmer amaranth plants established six weeks later in
-bareground (Webster and Grey, 2015). Still, seed production at second
-cohort is likely to replenish the soil seedbank. Seed production and
-deposition in the seedbank is also a key factor for species perpetuation
-(Menges, 1987). Palmer amaranth can produce hundred thousands seeds per
-plant (Schwartz et al., 2016; Keeley et al., 1987), and stay viable
-buried in the seedbank for at least 36 months (Sosnoskie et al., 2013).
-Therefore, preventing Palmer amaranth seed production or/and seed
-migration to its non-native habitat is an essential strategy to minimize
-weed impact into agroecossystem (Davis et al., 2015).
-
-An ecological approach to reduce seed production in Palmer amaranth is
-understanding its flowering window. Our study suggests that Palmer
-amaranth flowering was slightly influenced by cropping systems and
-cohort timings. Palmer amaranth growing in bareground and corn resulted
-in the overall shortest and longest flowering window, respectively. When
-growing in soybean, Palmer amaranth flowering window was similar to
-bareground at first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and depends on biological and
-ecological factors (Lang, 1965). We hypothesize that when growing in
-high competition (e.g., second cohort), Palmer amaranth plants tend to
-initiate flowering early, as well as having an extended flowering
-window. Early flower initiation is plant strategy when growing in stress
-conditions. For example, when growing under water stress, early
-flowering in Palmer amaranth resulted in a mismatch between female and
-male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant outcross,
-and thus reduce plant seed production and exchange of resistant alleles
-(Jhala et al., 2021). Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
-
-The mechanisms of sex-determination in plant species is intriguing and
-arouse the curiosity of many scientists, including Charles Darwin
-(Darwin, 1888). In our study, the gender model performance was decent
-(AIC 0.64) considering the biology of plant flowering. A 1:1 male and
-female sex ratio is a general evolutionary stable strategy for plant
-species perpetuation (Fisher, 1930). However, a slight deviation from
-1:1 sex ratio might occur in some dioecious species. For example, the
-dioecious \emph{Halophila stipulacea} is a female-biased plant in its
-native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
-ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
-female-male ratio to expand into its non-native habitat (Nguyen et al.,
-2018). Also, biotic and/or abiotic stress can influence plant sex
-determination. Palmer amaranth male-to-female ratio was greater under
-high plant densities (Korres and Norsworthy, 2017) and after herbicide
-application (Rumpa et al., 2019). Our model estimated that late
-flowering, heavier and taller Palmer amaranth plants deviated from 1:1
-ratio in favor to female plants. It was reported that female Palmer
-amaranth plants invested more in height, stem and biomass while male
-invested more in leaf area and leaf dry weight under nutrient deficiency
-(Korres et al., 2017). Our model also estimated more female plants in
-soybean and bareground compared to corn, which might linked to plant
-competition strategy in each agroecossystem. Our results showed the
-influence of life-history and ecological traits on sexual dimorphism in
-Palmer amaranth. Sexual dimorphism is documented in other dioecious
-species (Barrett and Hough, 2013). For example, stronger female plant
-competition and greater male tolerance to herbivory was reported in
-\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
-Research on candidate genes for sex determination in \emph{Amaranthus}
-species are currently underway but it is far to complete (Montgomery et
-al., 2021, 2019). Further studies are also needed to understand the
-ecological basis of Palmer amaranth flowering, including the plant
-behavior under climate change.
-
-Our study demonstrated the short-term Palmer amaranth plasticity to grow
-and develop into cropping-systems. Is likely that Palmer amaranth range
-will continue to expanding to new geographies. Therefore, preventive
-management is a priority to minimizing Palmer amaranth dispersal.
-Reactive management should focus on early-season management programs,
-which would have a large negative effect on Palmer amaranth growth and
-development. Tactics that promote early-season crop advantage against
-Palmer amaranth, including early crop planting, crop rotation (Oliveira
-et al., 2021), plant width, preemergence applied herbicide (Sanctis et
-al., 2021), and crop residue (e.g.~cover crops) would minimize the
-negative impact of Palmer amaranth in agroecossystems. The
-aggressiveness and differential Palmer amaranth adaptation to
-agroecosystem is striking and require national efforts to minimize the
-species impact on economy and sustainability.
-
-\hypertarget{disclosureconflict-of-interest-statement}{%
-\section*{Disclosure/Conflict-of-Interest
-Statement}\label{disclosureconflict-of-interest-statement}}
-\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
-Statement}
-
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
-
-\hypertarget{author-contributions}{%
-\section*{Author Contributions}\label{author-contributions}}
-\addcontentsline{toc}{section}{Author Contributions}
-
-RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
-the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP,
-MB, MO, SS, and RW: conceptualized the research. All authors reviewed
-the manuscript.
-
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-Funding: This work received no specific grant from any funding agency,
-commercial, or not-for-profit sectors
-
-\hypertarget{supplemental-data}{%
-\section*{Supplemental Data}\label{supplemental-data}}
-\addcontentsline{toc}{section}{Supplemental Data}
-
-Supplementary Material should be uploaded separately on submission, if
-there are Supplementary Figures, please include the caption in the same
-file as the figure. LaTeX Supplementary Material templates can be found
-in the Frontiers LaTeX folder
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-A Frontier article expect the reference list to be included in this
-section. To make that happens, the below syntax can be used. This
-\href{https://pandoc.org/MANUAL.html\#placement-of-the-bibliography}{feature
-is from Pandoc citeproc} which is used with \texttt{frontier\_article()}
-to handle the bibliography
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-aulakh2021}{}%
-Aulakh, J. S., Chahal, P. S., Kumar, V., Price, A. J., and Guillard, K.
-(2021). Multiple herbicide-resistant {Palmer} amaranth ({Amaranthus}
-palmeri) in {Connecticut}: Confirmation and response to {POST}
-herbicides. \emph{Weed Technology} 35, 457--463.
-doi:\href{https://doi.org/10.1017/wet.2021.6}{10.1017/wet.2021.6}.
-
-\leavevmode\hypertarget{ref-bagavathiannan2016}{}%
-Bagavathiannan, M. V., and Norsworthy, J. K. (2016). Multiple-{Herbicide
-Resistance Is Widespread} in {Roadside Palmer Amaranth Populations}.
-\emph{PLOS ONE} 11, e0148748.
-doi:\href{https://doi.org/10.1371/journal.pone.0148748}{10.1371/journal.pone.0148748}.
-
-\leavevmode\hypertarget{ref-baker2021}{}%
-Baker, R. (2021). Comparative analysis of {Palmer} amaranth
-({Amaranthus} palmeri) and waterhemp ({A}. Tuberculatus) in {Iowa}.
-doi:\href{https://doi.org/10.31274/etd-20210609-11}{10.31274/etd-20210609-11}.
-
-\leavevmode\hypertarget{ref-bangdiwala2018}{}%
-Bangdiwala, S. I. (2018). Regression: Binary logistic.
-\emph{International Journal of Injury Control and Safety Promotion} 25,
-336--338.
-doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.
-
-\leavevmode\hypertarget{ref-barrett2013}{}%
-Barrett, S. C. H., and Hough, J. (2013). Sexual dimorphism in flowering
-plants. \emph{Journal of Experimental Botany} 64, 67--82.
-doi:\href{https://doi.org/10.1093/jxb/ers308}{10.1093/jxb/ers308}.
-
-\leavevmode\hypertarget{ref-bates2015}{}%
-Bates, D., Mächler, M., Bolker, B., and Walker, S. (2015). Fitting
-{Linear Mixed}-{Effects Models Using} Lme4. \emph{Journal of Statistical
-Software} 67, 1--48.
-doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.
-
-\leavevmode\hypertarget{ref-berger2015}{}%
-Berger, S. T., Ferrell, J. A., Rowland, D. L., and Webster, T. M.
-(2015). Palmer {Amaranth} ({Amaranthus} palmeri) {Competition} for
-{Water} in {Cotton}. \emph{Weed Science} 63, 928--935.
-doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.
-
-\leavevmode\hypertarget{ref-bravo2017}{}%
-Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
-(2017). Differentiation of {Life}-{History Traits} among {Palmer
-Amaranth Populations} ({Amaranthus} palmeri) and {Its Relation} to
-{Cropping Systems} and {Glyphosate Sensitivity}. \emph{Weed Science} 65,
-339--349.
-doi:\href{https://doi.org/10.1017/wsc.2017.14}{10.1017/wsc.2017.14}.
-
-\leavevmode\hypertarget{ref-bravo2018}{}%
-Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
-(2018). Evolutionary {Adaptations} of {Palmer Amaranth} ({Amaranthus}
-palmeri) to {Nitrogen Fertilization} and {Crop Rotation History Affect
-Morphology} and {Nutrient}-{Use Efficiency}. \emph{Weed Science} 66,
-180--189.
-doi:\href{https://doi.org/10.1017/wsc.2017.73}{10.1017/wsc.2017.73}.
-
-\leavevmode\hypertarget{ref-briscoerunquist2019}{}%
-Briscoe Runquist, R. D., Lake, T., Tiffin, P., and Moeller, D. A.
-(2019). Species distribution models throughout the invasion history of
-{Palmer} amaranth predict regions at risk of future invasion and reveal
-challenges with modeling rapidly shifting geographic ranges. \emph{Sci
-Rep} 9, 2426.
-doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.
-
-\leavevmode\hypertarget{ref-chahal2021}{}%
-Chahal, P. S., Barnes, E. R., and Jhala, A. J. (2021). Emergence pattern
-of {Palmer} amaranth ({Amaranthus} palmeri) influenced by tillage
-timings and residual herbicides. \emph{Weed Technology} 35, 433--439.
-doi:\href{https://doi.org/10.1017/wet.2020.136}{10.1017/wet.2020.136}.
-
-\leavevmode\hypertarget{ref-chahal2018}{}%
-Chahal, P. S., Irmak, S., Jugulam, M., and Jhala, A. J. (2018).
-Evaluating {Effect} of {Degree} of {Water Stress} on {Growth} and
-{Fecundity} of {Palmer} amaranth ({Amaranthus} palmeri) {Using Soil
-Moisture Sensors}. \emph{Weed Science} 66, 738--745.
-doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.
-
-\leavevmode\hypertarget{ref-correndo2021}{}%
-Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
-Retrieving and processing agro-meteorological data from {API}-client
-sources using {R} software. \emph{BMC Research Notes} 14, 205.
-doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.
-
-\leavevmode\hypertarget{ref-darwin1888}{}%
-Darwin, C. (1888). \emph{The {Different Forms} of {Flowers} on {Plants}
-of the {Same Species}}. {J. Murray} Available at:
-\url{http://books.google.com?id=7uMEAAAAYAAJ}.
-
-\leavevmode\hypertarget{ref-davis2015}{}%
-Davis, A. S., Schutte, B. J., Hager, A. G., and Young, B. G. (2015).
-Palmer {Amaranth} ({Amaranthus} palmeri) {Damage Niche} in {Illinois
-Soybean Is Seed Limited}. \emph{Weed Science} 63, 658--668.
-doi:\href{https://doi.org/10.1614/WS-D-14-00177.1}{10.1614/WS-D-14-00177.1}.
-
-\leavevmode\hypertarget{ref-farmer2017}{}%
-Farmer, J. A., Webb, E. B., Pierce, R. A., and Bradley, K. W. (2017).
-Evaluating the potential for weed seed dispersal based on waterfowl
-consumption and seed viability. \emph{Pest Management Science} 73,
-2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.
-
-\leavevmode\hypertarget{ref-fisher1930}{}%
-Fisher, R. A. (1930). The genetical theory of natural selection.
-\emph{Eugen Rev} 22, 127--130. Available at:
-\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2984947/} {[}Accessed
-August 12, 2021{]}.
-
-\leavevmode\hypertarget{ref-fox2018}{}%
-Fox, J., and Weisberg, S. (2018). \emph{An {R Companion} to {Applied
-Regression}}. {SAGE Publications} Available at:
-\url{http://books.google.com?id=uPNrDwAAQBAJ}.
-
-\leavevmode\hypertarget{ref-garetson2019}{}%
-Garetson, R., Singh, V., Singh, S., Dotray, P., and Bagavathiannan, M.
-(2019). Distribution of herbicide-resistant {Palmer} amaranth
-({Amaranthus} palmeri) in row crop production systems in {Texas}.
-\emph{Weed Technology} 33, 355--365.
-doi:\href{https://doi.org/10.1017/wet.2019.14}{10.1017/wet.2019.14}.
-
-\leavevmode\hypertarget{ref-guo2003}{}%
-Guo, P., and Al-Khatib, K. (2003). Temperature effects on germination
-and growth of redroot pigweed ({Amaranthus} retroflexus), {Palmer}
-amaranth ({A}. Palmeri), and common waterhemp ({A}. rudis). \emph{Weed
-Science} 51, 869--875.
-doi:\href{https://doi.org/10.1614/P2002-127}{10.1614/P2002-127}.
-
-\leavevmode\hypertarget{ref-hartzler2016}{}%
-Hartzler, B., and Anderson, M. (2016). Palmer amaranth: {It}'s here, now
-what? 10.
-
-\leavevmode\hypertarget{ref-heap2021}{}%
-Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
-Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
-July 26, 2021{]}.
-
-\leavevmode\hypertarget{ref-hothorn2008}{}%
-Hothorn, T., Bretz, F., and Westfall, P. (2008). Simultaneous
-{Inference} in {General Parametric Models}. \emph{Biometrical Journal}
-50, 346--363.
-doi:\href{https://doi.org/10.1002/bimj.200810425}{10.1002/bimj.200810425}.
-
-\leavevmode\hypertarget{ref-jha2010}{}%
-Jha, P., Norsworthy, J. K., Riley, M. B., and Bridges, W. (2010). Annual
-{Changes} in {Temperature} and {Light Requirements} for {Germination} of
-{Palmer Amaranth} ({Amaranthus} palmeri) {Seeds Retrieved} from {Soil}.
-\emph{Weed Science} 58, 426--432.
-doi:\href{https://doi.org/10.1614/WS-D-09-00038.1}{10.1614/WS-D-09-00038.1}.
-
-\leavevmode\hypertarget{ref-jhala2021}{}%
-Jhala, A. J., Norsworthy, J. K., Ganie, Z. A., Sosnoskie, L. M., Beckie,
-H. J., Mallory-Smith, C. A., Liu, J., Wei, W., Wang, J., and
-Stoltenberg, D. E. (2021). Pollen-mediated gene flow and transfer of
-resistance alleles from herbicide-resistant broadleaf weeds. \emph{Weed
-Technology} 35, 173--187.
-doi:\href{https://doi.org/10.1017/wet.2020.101}{10.1017/wet.2020.101}.
-
-\leavevmode\hypertarget{ref-keeley1987}{}%
-Keeley, P. E., Carter, C. H., and Thullen, R. J. (1987). Influence of
-{Planting Date} on {Growth} of {Palmer Amaranth} ({Amaranthus} palmeri).
-\emph{Weed Science} 35, 199--204.
-doi:\href{https://doi.org/10.1017/S0043174500079054}{10.1017/S0043174500079054}.
-
-\leavevmode\hypertarget{ref-kistner2018}{}%
-Kistner, E. J., and Hatfield, J. L. (2018). Potential {Geographic
-Distribution} of {Palmer Amaranth} under {Current} and {Future
-Climates}. \emph{Agricultural \& Environmental Letters} 3, 170044.
-doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.
-
-\leavevmode\hypertarget{ref-klingaman1994}{}%
-Klingaman, T. E., and Oliver, L. R. (1994). Palmer {Amaranth}
-({Amaranthus} palmeri) {Interference} in {Soybeans} ({Glycine} max).
-\emph{Weed Science} 42, 523--527.
-doi:\href{https://doi.org/10.1017/S0043174500076888}{10.1017/S0043174500076888}.
-
-\leavevmode\hypertarget{ref-kohrt2017}{}%
-Kohrt, J. R., Sprague, C. L., Nadakuduti, S. S., and Douches, D. (2017).
-Confirmation of a {Three}-{Way} ({Glyphosate}, {ALS}, and {Atrazine})
-{Herbicide}-{Resistant Population} of {Palmer Amaranth} ({Amaranthus}
-palmeri) in {Michigan}. \emph{Weed Science} 65, 327--338.
-doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.
-
-\leavevmode\hypertarget{ref-korres2017a}{}%
-Korres, N. E., and Norsworthy, J. K. (2017). Palmer {Amaranth}
-({Amaranthus} palmeri) {Demographic} and {Biological Characteristics} in
-{Wide}-{Row Soybean}. \emph{Weed Science} 65, 491--503.
-doi:\href{https://doi.org/10.1017/wsc.2017.12}{10.1017/wsc.2017.12}.
-
-\leavevmode\hypertarget{ref-korres2017}{}%
-Korres, N. E., Norsworthy, J. K., FitzSimons, T., Roberts, T. L., and
-Oosterhuis, D. M. (2017). Differential {Response} of {Palmer Amaranth}
-({Amaranthus} palmeri) {Gender} to {Abiotic Stress}. \emph{Weed Science}
-65, 213--227.
-doi:\href{https://doi.org/10.1017/wsc.2016.34}{10.1017/wsc.2016.34}.
-
-\leavevmode\hypertarget{ref-korres2020}{}%
-Korres, N. E., Norsworthy, J. K., Mauromoustakos, A., and Williams, M.
-M. (2020). Soybean density and {Palmer} amaranth ({Amaranthus} palmeri)
-establishment time: Effects on weed biology, crop yield, and economic
-returns. \emph{Weed Science} 68, 467--475.
-doi:\href{https://doi.org/10.1017/wsc.2020.41}{10.1017/wsc.2020.41}.
-
-\leavevmode\hypertarget{ref-kupper2017}{}%
-Küpper, A., Borgato, E. A., Patterson, E. L., Netto, A. G., Nicolai, M.,
-Carvalho, S. J. P. de, Nissen, S. J., Gaines, T. A., and Christoffoleti,
-P. J. (2017). Multiple {Resistance} to {Glyphosate} and {Acetolactate
-Synthase Inhibitors} in {Palmer Amaranth} ({Amaranthus} palmeri)
-{Identified} in {Brazil}. \emph{Weed Science} 65, 317--326.
-doi:\href{https://doi.org/10.1017/wsc.2017.1}{10.1017/wsc.2017.1}.
-
-\leavevmode\hypertarget{ref-lang1965}{}%
-Lang, A. (1965). {``Physiology of flower initiation,''} in
-\emph{Differenzierung und {Entwicklung} / {Differentiation} and
-{Development}} Handbuch der {Pflanzenphysiologie} / {Encyclopedia} of
-{Plant Physiology}., ed. A. Lang ({Berlin, Heidelberg}: {Springer}),
-1380--1536.
-doi:\href{https://doi.org/10.1007/978-3-642-50088-6_39}{10.1007/978-3-642-50088-6\_39}.
-
-\leavevmode\hypertarget{ref-larran2017}{}%
-Larran, A. S., Palmieri, V. E., Perotti, V. E., Lieber, L., Tuesca, D.,
-and Permingeat, H. R. (2017). Target-site resistance to acetolactate
-synthase ({ALS})-inhibiting herbicides in {Amaranthus} palmeri from
-{Argentina}. \emph{Pest Management Science} 73, 2578--2584.
-doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.
-
-\leavevmode\hypertarget{ref-leeper2017}{}%
-Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
-Marginal Eﬀects} with {R}'s margins. 31.
-
-\leavevmode\hypertarget{ref-lindsay2017}{}%
-Lindsay, K., Popp, M., Norsworthy, J., Bagavathiannan, M., Powles, S.,
-and Lacoste, M. (2017). {PAM}: {Decision Support} for {Long}-{Term
-Palmer Amaranth} ({Amaranthus} palmeri) {Control}. \emph{Weed
-Technology} 31, 915--927.
-doi:\href{https://doi.org/10.1017/wet.2017.69}{10.1017/wet.2017.69}.
-
-\leavevmode\hypertarget{ref-maclaren2020}{}%
-MacLaren, C., Storkey, J., Menegat, A., Metcalfe, H., and
-Dehnen-Schmutz, K. (2020). An ecological future for weed science to
-sustain crop production and the environment. {A} review. \emph{Agron.
-Sustain. Dev.} 40, 24.
-doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.
-
-\leavevmode\hypertarget{ref-macrae2013}{}%
-MacRae, A. W., Webster, T. M., Sosnoskie, L. M., Culpepper, A. S., and
-Kichler, J. M. (2013). Cotton {Yield Loss Potential} in {Response} to
-{Length} of {Palmer Amaranth} ({Amaranthus} palmeri) {Interference}. 17,
-6.
-
-\leavevmode\hypertarget{ref-massinga2001}{}%
-Massinga, R. A., Currie, R. S., Horak, M. J., and Boyer, J. (2001).
-Interference of {Palmer} amaranth in corn. \emph{Weed Science} 49,
-202--208.
-doi:\href{https://doi.org/10.1614/0043-1745(2001)049\%5B0202:IOPAIC\%5D2.0.CO;2}{10.1614/0043-1745(2001)049{[}0202:IOPAIC{]}2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-matzrafi2021}{}%
-Matzrafi, M., Osipitan, O. A., Ohadi, S., and Mesgaran, M. B. (2021).
-Under pressure: Maternal effects promote drought tolerance in progeny
-seed of {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed Science}
-69, 31--38.
-doi:\href{https://doi.org/10.1017/wsc.2020.75}{10.1017/wsc.2020.75}.
-
-\leavevmode\hypertarget{ref-mcfarlane2018}{}%
-McFarlane, G. R., Whitelaw, C. B. A., and Lillico, S. G. (2018).
-{CRISPR}-{Based Gene Drives} for {Pest Control}. \emph{Trends in
-Biotechnology} 36, 130--133.
-doi:\href{https://doi.org/10.1016/j.tibtech.2017.10.001}{10.1016/j.tibtech.2017.10.001}.
-
-\leavevmode\hypertarget{ref-menges1987}{}%
-Menges, R. M. (1987). Weed {Seed Population Dynamics} during {Six Years}
-of {Weed Management Systems} in {Crop Rotations} on {Irrigated Soil}.
-\emph{Weed Science} 35, 328--332. Available at:
-\url{http://www.jstor.org/stable/4044593}.
-
-\leavevmode\hypertarget{ref-mesgaran2021}{}%
-Mesgaran, M. B., Matzrafi, M., and Ohadi, S. (2021). Sex dimorphism in
-dioecious {Palmer} amaranth ({Amaranthus} palmeri) in response to water
-stress. \emph{Planta} 254, 17.
-doi:\href{https://doi.org/10.1007/s00425-021-03664-7}{10.1007/s00425-021-03664-7}.
-
-\leavevmode\hypertarget{ref-milani2021}{}%
-Milani, A., Panozzo, S., Farinati, S., Iamonico, D., Sattin, M., Loddo,
-D., and Scarabel, L. (2021). Recent {Discovery} of {Amaranthus} palmeri
-{S}. {Watson} in {Italy}: {Characterization} of {ALS}-{Resistant
-Populations} and {Sensitivity} to {Alternative Herbicides}.
-\emph{Sustainability} 13, 7003.
-doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.
-
-\leavevmode\hypertarget{ref-montgomery2021}{}%
-Montgomery, J. S., Giacomini, D. A., Weigel, D., and Tranel, P. J.
-(2021). Male-specific {Y}-chromosomal regions in waterhemp ({Amaranthus}
-tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{New
-Phytologist} 229, 3522--3533.
-doi:\href{https://doi.org/10.1111/nph.17108}{10.1111/nph.17108}.
-
-\leavevmode\hypertarget{ref-montgomery2019}{}%
-Montgomery, J. S., Sadeque, A., Giacomini, D. A., Brown, P. J., and
-Tranel, P. J. (2019). Sex-specific markers for waterhemp ({Amaranthus}
-tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed
-Science} 67, 412--418.
-doi:\href{https://doi.org/10.1017/wsc.2019.27}{10.1017/wsc.2019.27}.
-
-\leavevmode\hypertarget{ref-morgan2001}{}%
-Morgan, G. D., Baumann, P. A., and Chandler, J. M. (2001). Competitive
-{Impact} of {Palmer Amaranth} ({Amaranthus} palmeri) on {Cotton}
-({Gossypium} hirsutum) {Development} and {Yield}. \emph{Weed Technology}
-15, 408--412.
-doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-nguyen2018}{}%
-Nguyen, H. M., Kleitou, P., Kletou, D., Sapir, Y., and Winters, G.
-(2018). Differences in flowering sex ratios between native and invasive
-populations of the seagrass {Halophila} stipulacea. \emph{Botanica
-Marina} 61, 337--342.
-doi:\href{https://doi.org/10.1515/bot-2018-0015}{10.1515/bot-2018-0015}.
-
-\leavevmode\hypertarget{ref-oliveira2021b}{}%
-Oliveira, M. C. (2021). \emph{Maxwelco/palmer\_adaptation}. {Zenodo}
-Available at: \url{ttps://doi.org/10.5281/zenodo.5236831}.
-
-\leavevmode\hypertarget{ref-oliveira2021a}{}%
-Oliveira, M. C., Giacomini, D. A., Arsenijevic, N., Vieira, G., Tranel,
-P. J., and Werle, R. (2021). Distribution and validation of genotypic
-and phenotypic glyphosate and {PPO}-inhibitor resistance in {Palmer}
-amaranth ({Amaranthus} palmeri) from southwestern {Nebraska}. \emph{Weed
-Technology} 35, 65--76.
-doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.
-
-\leavevmode\hypertarget{ref-page2021}{}%
-Page, E. R., Nurse, R. E., Meloche, S., Bosveld, K., Grainger, C.,
-Obeid, K., Filotas, M., Simard, M.-J., and Laforest, M. (2021). Import
-of {Palmer} amaranth ({Amaranthus} palmeri {S}. {Wats}.) Seed with sweet
-potato ({Ipomea} batatas ({L}.) {Lam}) slips. \emph{Can. J. Plant Sci.},
-CJPS-2020-0321.
-doi:\href{https://doi.org/10.1139/CJPS-2020-0321}{10.1139/CJPS-2020-0321}.
-
-\leavevmode\hypertarget{ref-perez-llorca2019}{}%
-Pérez-Llorca, M., and Sánchez Vilas, J. (2019). Sexual dimorphism in
-response to herbivory and competition in the dioecious herb {Spinacia}
-oleracea. \emph{Plant Ecol} 220, 57--68.
-doi:\href{https://doi.org/10.1007/s11258-018-0902-7}{10.1007/s11258-018-0902-7}.
-
-\leavevmode\hypertarget{ref-piskackova2021}{}%
-Piskackova, T. A. R., Reberg-Horton, S. C., Richardson, R. J., Jennings,
-K. M., Franca, L., Young, B. G., and Leon, R. G. (2021). Windows of
-action for controlling palmer amaranth ({Amaranthus} palmeri) using
-emergence and phenology models. \emph{Weed Research} 61, 188--198.
-doi:\href{https://doi.org/10.1111/wre.12470}{10.1111/wre.12470}.
-
-\leavevmode\hypertarget{ref-price2011}{}%
-Price, A. J., Balkcom, K. S., Culpepper, S. A., Kelton, J. A., Nichols,
-R. L., and Schomberg, H. (2011). Glyphosate-resistant {Palmer} amaranth:
-{A} threat to conservation tillage. \emph{Journal of Soil and Water
-Conservation} 66, 265--275.
-doi:\href{https://doi.org/10.2489/jswc.66.4.265}{10.2489/jswc.66.4.265}.
-
-\leavevmode\hypertarget{ref-ritz2015}{}%
-Ritz, C., Baty, F., Streibig, J. C., and Gerhard, D. (2015).
-Dose-{Response Analysis Using R}. \emph{PLOS ONE} 10, e0146021.
-doi:\href{https://doi.org/10.1371/journal.pone.0146021}{10.1371/journal.pone.0146021}.
-
-\leavevmode\hypertarget{ref-rumpa2019}{}%
-Rumpa, M. M., Krausz, R. F., Gibson, D. J., and Gage, K. L. (2019).
-Effect of {PPO}-{Inhibiting Herbicides} on the {Growth} and {Sex Ratio}
-of a {Dioecious Weed Species Amaranthus} palmeri ({Palmer Amaranth}).
-\emph{Agronomy} 9, 275.
-doi:\href{https://doi.org/10.3390/agronomy9060275}{10.3390/agronomy9060275}.
-
-\leavevmode\hypertarget{ref-sanctis2021}{}%
-Sanctis, J. H. S. de, Barnes, E. R., Knezevic, S. Z., Kumar, V., and
-Jhala, A. J. (2021). Residual herbicides affect critical time of
-{Palmer} amaranth removal in soybean. \emph{Agronomy Journal} 113,
-1920--1933.
-doi:\href{https://doi.org/10.1002/agj2.20615}{10.1002/agj2.20615}.
-
-\leavevmode\hypertarget{ref-sauer1957}{}%
-Sauer, J. (1957). Recent {Migration} and {Evolution} of the {Dioecious
-Amaranths}. \emph{Evolution} 11, 11--31.
-doi:\href{https://doi.org/10.2307/2405808}{10.2307/2405808}.
-
-\leavevmode\hypertarget{ref-sauer1972}{}%
-Sauer, J. D. (1972). The dioecious amaranths: A new species name and
-major range extensions. \emph{Madroño} 21, 426--434. Available at:
-\url{http://www.jstor.org/stable/41423815}.
-
-\leavevmode\hypertarget{ref-sbatella2010}{}%
-Sbatella, G. M., and Wilson, R. G. (2010). Isoxaflutole {Shifts Kochia}
-({Kochia} scoparia) {Populations} in {Continuous Corn}. \emph{Weed
-Technology} 24, 392--396.
-doi:\href{https://doi.org/10.1614/WT-D-09-00023.1}{10.1614/WT-D-09-00023.1}.
-
-\leavevmode\hypertarget{ref-schliekelman2005}{}%
-Schliekelman, P., Ellner, S., and Gould, F. (2005). Pest {Control} by
-{Genetic Manipulation} of {Sex Ratio}. \emph{Journal of Economic
-Entomology} 98, 18--34.
-doi:\href{https://doi.org/10.1093/jee/98.1.18}{10.1093/jee/98.1.18}.
-
-\leavevmode\hypertarget{ref-schwartz2016}{}%
-Schwartz, L. M., Norsworthy, J. K., Young, B. G., Bradley, K. W.,
-Kruger, G. R., Davis, V. M., Steckel, L. E., and Walsh, M. J. (2016).
-Tall {Waterhemp} ({Amaranthus} tuberculatus) and {Palmer} amaranth
-({Amaranthus} palmeri) {Seed Production} and {Retention} at {Soybean
-Maturity}. \emph{Weed Technology} 30, 284--290.
-doi:\href{https://doi.org/10.1614/WT-D-15-00130.1}{10.1614/WT-D-15-00130.1}.
-
-\leavevmode\hypertarget{ref-scott2011}{}%
-Scott, D. (2011). The {Technological Fix Criticisms} and the
-{Agricultural Biotechnology Debate}. \emph{J Agric Environ Ethics} 24,
-207--226.
-doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.
-
-\leavevmode\hypertarget{ref-sosnoskie2013}{}%
-Sosnoskie, L. M., Webster, T. M., and Culpepper, A. S. (2013).
-Glyphosate {Resistance Does Not Affect Palmer Amaranth} ({Amaranthus}
-palmeri) {Seedbank Longevity}. \emph{Weed Science} 61, 283--288.
-doi:\href{https://doi.org/10.1614/WS-D-12-00111.1}{10.1614/WS-D-12-00111.1}.
-
-\leavevmode\hypertarget{ref-spaunhorst2018}{}%
-Spaunhorst, D. J., Devkota, P., Johnson, W. G., Smeda, R. J., Meyer, C.
-J., and Norsworthy, J. K. (2018). Phenology of {Five Palmer} amaranth
-({Amaranthus} palmeri) {Populations Grown} in {Northern Indiana} and
-{Arkansas}. \emph{Weed Science} 66, 457--469.
-doi:\href{https://doi.org/10.1017/wsc.2018.12}{10.1017/wsc.2018.12}.
-
-\leavevmode\hypertarget{ref-vanwychen2020}{}%
-Van Wychen, L. (2020). 2020 {Survey} of the most common and troublesome
-weeds in grass crops, pasture, and turf in the {United States} and
-{Canada}. Available at:
-\url{https://wssa.net/wp-content/uploads/2020-Weed-Survey_grass-crops.xlsx}.
-
-\leavevmode\hypertarget{ref-wang1992}{}%
-Wang, J. L., Klessig, D. F., and Berry, J. O. (1992). Regulation of {C4
-Gene Expression} in {Developing Amaranth Leaves}. \emph{The Plant Cell}
-4, 173--184.
-doi:\href{https://doi.org/10.1105/tpc.4.2.173}{10.1105/tpc.4.2.173}.
-
-\leavevmode\hypertarget{ref-ward2013}{}%
-Ward, S. M., Webster, T. M., and Steckel, L. E. (2013). Palmer
-{Amaranth} ({Amaranthus} palmeri): {A Review}. \emph{Weed Technology}
-27, 12--27.
-doi:\href{https://doi.org/10.1614/WT-D-12-00113.1}{10.1614/WT-D-12-00113.1}.
-
-\leavevmode\hypertarget{ref-webster2015}{}%
-Webster, T. M., and Grey, T. L. (2015). Glyphosate-{Resistant Palmer
-Amaranth} ({Amaranthus} palmeri) {Morphology}, {Growth}, and {Seed
-Production} in {Georgia}. \emph{Weed Science} 63, 264--272.
-doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.
-
-\leavevmode\hypertarget{ref-yacouby2020}{}%
-Yacouby, R., and Axman, D. (2020). Probabilistic {Extension} of
-{Precision}, {Recall}, and {F1 Score} for {More Thorough Evaluation} of
-{Classification Models}. in \emph{Proceedings of the {First Workshop} on
-{Evaluation} and {Comparison} of {NLP Systems}} ({Online}: {Association
-for Computational Linguistics}), 79--91.
-doi:\href{https://doi.org/10.18653/v1/2020.eval4nlp-1.9}{10.18653/v1/2020.eval4nlp-1.9}.
-
-\leavevmode\hypertarget{ref-yu2021}{}%
-Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
-Gunsolus, J., and Becker, R. (2021). Timeline of {Palmer} amaranth
-({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
-\emph{Weed Technology}, 1--31.
-doi:\href{https://doi.org/10.1017/wet.2021.32}{10.1017/wet.2021.32}.
-
-\end{CSLReferences}
-
-\hypertarget{figures}{%
-\section*{Figures}\label{figures}}
-\addcontentsline{toc}{section}{Figures}
-
-\begin{figure}
-
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
-
-}
-
-\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}
-
-}
-
-\caption{Palmer amaranth adaptation study layout of a plant cohort timing. Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}
-
-}
-
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}
-
-}
-
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-4}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
-
-}
-
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
-
-}
-
-\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B).
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
-\end{figure}
-
-\end{document}

### 2021-08-24 data analysis/sex_analysis.Rmd
as

@@
   geom_emoji(aes(x = 220, y = 0.95), emoji = "2640") +
-  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") -> doyf
+  geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") + 
+    theme(legend.position = "none",
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 30),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25)) -> doyf
 
@@
   geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
-  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> weightf
+  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
+    theme(legend.position = "none",
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 30),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25)) -> weightf
 
@@
   geom_emoji(aes(x = 105, y = 0.95), emoji = "2640") +
-  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") -> heightf
+  geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
+    theme(legend.position = "none",
+        axis.text.y = element_markdown(size = 25),
+        axis.title = element_markdown(size = 30),
+        axis.text.x = element_markdown(size = 25),
+        legend.text = element_markdown(size = 25)) -> heightf
 
@@
   plot_annotation(tag_levels = 'A') & 
-  theme(plot.tag = element_text(size = 10, hjust = 0, vjust = 0))
+  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))
 

### 2021-08-26 Frontiers/Frontiers.Rmd
as

@@
 In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
-At first cohort, Palmer amaranth produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas Palmer amaranth produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} at second cohort in bareground, soybean and corn, respectively.
-The plant height was more impacted by Palmer amaranth growing in corn.
+The first Palmer amaranth first cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
+Palmer amaranth height was more impacted when growing in corn.
 Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
@@
 Also, Palmer amaranth sex dimorphism was slightly influenced by day of year, weight and height.
-The model estimated that probability of being a female plant increase as biomass and height increases.
-Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems from United States Midwest.
+The model estimated that probability of being a female plant increased as biomass and height increased.
+Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems from the US Midwest.
 Palmer amaranth is likely to continue its expansion northward.
 Therefore, preventing plant dispersal into new habitats is the most effective management strategy.
-Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability establish into cropping systems, including crop rotation (not only corn and soybean), early/late crop planting, row spacing, and cover crops.
-Our results also shed a light on sex dimorphism as a future Palmer amaranth ecological management.
+Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability to establish into cropping systems, including crop rotation (beyond corn and soybean), early/late crop planting, row spacing, cover crops, and effective chemical control programs.
 
@@
 With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
-Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], animals [@farmer2017].
+Also, Palmer amaranth small seeds (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], wildlife [@farmer2017], etc.
 The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.
 
-Light and temperature are the main environment requirements for Palmer amaranth successful grow and development [@jha2010].
+Light and temperature are the main environmental requirements for Palmer amaranth growth and development [@jha2010].
 Palmer amaranth is reported with an extended germination period [@ward2013].
@@
 In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
-Water has not shown to limit Palmer amaranth fitness.
+Soil water content has not shown to limit Palmer amaranth fitness.
 Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
@@
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under the current conditions.
+Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under the current agroecosystem conditions.
 Further scenarios show that global temperature warming can impact agriculture, and promote niches for Palmer amaranth invasion/adaptation into new environments.
@@
 In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
-Waterhemp, an *Amaranthus* species, is the most troublesome species in that geography [@tranel2011].
+Waterhemp, an *Amaranthus* species, is the most troublesome species in the US Midwest [@tranel2011].
 Nonetheless, invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
@@
 Therefore, Palmer amaranth range of expansion into new habitats can increase.
-It seems certain the need to manage Palmer amaranth in agronomic crops throughout novel environments in the near future.
+It seems certain the need to manage new Palmer amaranth infestations in agronomic crops throughout northern United States in the near future.
 Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
@@
 
-A Palmer amaranth accession (Per1) from Perkins County, Nebraska.
-with no reported herbicide resistance according to @oliveira2021a was selected for this study.
-Three weeks prior to the onset of the field experiments, seeds were planted in plastic trays containing potting-mix.
+A Palmer amaranth accession (Per1) from Perkins County, Nebraska with no reported herbicide resistance according to @oliveira2021a was selected for this study.
+Three weeks prior to the establishment of each cohort, seeds were planted in plastic trays containing potting-mix.
 Emerged seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
@@
 Fields were conventionally tilled prior to crop planting.
-Corn hybrid and soybean varieties were planted in 76-cm row spacing (Table 1).
+Corn and soybean were planted in 76-cm row spacing (Table 1).
 Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
@@
 The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations.
-Each experimental unit was planted with corn, soybean, or kept under bareground.
+Each filed experimental unit was planted with corn, soybean, or kept under bareground.
 The two transplanting timings were June 1 (first cohort) and July 1 (second cohort).
@@
 After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2).
-When needed, Palmer amaranth plants were supplied with water during the first week after transplanting.
+When needed, Palmer amaranth plants were supplied with water during the first week after transplanting to assure seedling survival.
 
@@
 When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant.
-Also, aboveground plant biomass was harvest near soil surface and oven dried at 65 C until reaching constant weight before weighing (g plant\textsuperscript{-1}) was recorded.
+Also, aboveground plant biomass was harvested near soil surface and oven dried at 65 C until reaching constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
 
 Plants had to be harvested at flowering because Palmer amaranth is neither endemic in Wisconsin nor in Illinois.
-In our study, all locations follow the methodology of plant harvest at flowering initiation, except in Grant, NE.
+In our study, all locations followed the methodology of plant harvest at flowering initiation, except in Grant, NE.
 In this location, all Palmer amaranth plants were harvest at once on July 06, 2018 and 2019 (first cohort), and on August 17, 2018 and on July 31, 2019 (second cohort).
@@
 
-The Palmer amaranth flowering timing was estimated as cumulative flowering across all location.
-Dataset was nested but it did not include Grant, NE.
+The Palmer amaranth flowering timing was estimated as cumulative flowering across all locations, except Grant, NE.
 Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
@@
 
-Palmer amaranth sex dimorphism was fitted to a binary logistic regression [@bangdiwala2018].
-Binary logistic regression is used for predicting binary classes, such as the probability of a plant being female in a dioecious species.
-Prior to the analysis, all missing values were removed from the dataset.
+A binary logistic regression was fitted to Palmer amaranth sex dimorphism.
+Binary logistic regression is used for predicting binary classes [@bangdiwala2018], such as the probability of a plant being female in a dioecious species.
+Prior to the analysis, missing values were removed from the dataset.
 Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location.
@@
 The 80% train is used for the model training and the 20% test is used for checking model performance on unseen dataset.
-With 80% dataset, a binary response variable, male (0) and female (1), was fitted to a generalized linear model (base R *glm* function) including day of year harvest, height, weight, crop and month as independent variables (without interaction).
+With 80% dataset, a generalized linear model (base R *glm* function) was fitted to binary response variable, the probability of being female (0 to male and 1 to female).
+The independent variables were day of year harvest, height, weight, and crop (without interaction).
 The model family was binomial with a logit function.
@@
 The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package [@leeper2021].
-The rest 20% dataset was predicted using *predict* function with a cutoff estimation for male or female using *performance* function.
+The 20% test data was predicted using the *predict* function with a cutoff estimation for male or female using *performance* function.
 The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package [@kuhn2021].
@@
 F1-score conveys the balance between the precision and the recall [@yacouby2020].
+The area under the curve was also estimated with XXXXXXXXXXXXXXXXX.
 
@@
 
-Palmer amaranth cumulative flowering at the second cohort ranged from mid July to mid September (Figure 4B).
+Palmer amaranth cumulative flowering in the second cohort ranged from mid July to mid September (Figure 4B).
 Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn.
@@
 
-The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
+The model goodness of fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
 The likelihood ratio test showed a p-value of < 0.00.
@@
 The model classification showed a precision of 0.64, recall of 0.66, and a F1-score of 0.65.
-In addition, the area under the curve was 0.64.
+In addition, the AUC was 0.64.
 
@@
 Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987].
-Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], and stay viable buried in the seedbank for at least 36 months [@sosnoskie2013].
-Therefore, preventing Palmer amaranth seed production or/and seed migration to its non-native habitat is an essential strategy to minimize the species impact into agroecossystem [@davis2015].
+Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
+Therefore, preventing Palmer amaranth seed production or/and seed migration to non-native habitats is an essential strategy to minimize the species impact in agroecosystems [@davis2015].
 
-An ecological approach to reduce seed production in Palmer amaranth is understanding its flowering window.
+An ecological approach to reduce Palmer amaranth seed production is understanding plant biology, including flowering pattern.
 Our study suggests that Palmer amaranth flowering was slightly influenced by crops and cohort timings.
@@
 A study has shown that non-native Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
-Early flower initiation could be also a plant strategy when growing in stress conditions.
+Early flower initiation could also be a plant strategy when growing in stress conditions.
 For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
@@
 It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
-Our model also estimated more female plants in soybean and bareground compared to corn, which might linked to plant competition strategy in each crop.
+Our model also estimated more female plants in soybean and bareground compared to corn, which might be linked to plant competition strategy in each crop.
 Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth.
@@
 For example, stronger female plant competition and greater male tolerance to herbivory was reported in *Spinacia oleracea* [@perez-llorca2019].
-Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far to complete [@montgomery2021; @montgomery2019].
+Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far from complete [@montgomery2021; @montgomery2019].
 Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including the plant behavior under climate change.
@@
 cover crops) would minimize the negative impact of Palmer amaranth to cropping systems.
-The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on sustainability and economy.
+The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on sustainability and profitability of cropping systems.
 
@@
 
-A Frontier article expect the reference list to be included in this section.
-To make that happens, the below syntax can be used.
-This [feature is from Pandoc citeproc](https://pandoc.org/MANUAL.html#placement-of-the-bibliography) which is used with `frontier_article()` to handle the bibliography
-
 ::: {#refs}
@@
 ```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit', out.width = "150mm", out.height = "70mm"}
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit', out.width = "150mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "150mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)', out.width = "170mm", out.height = "70mm"}
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
+Back line represents the model estimation and shade green the 95% confidence intervals', out.width = "170mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 ```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
+From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-08-26 Frontiers/Frontiers.tex
as

@@
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of United States Midwest.
+A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
 In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
-At first cohort, Palmer amaranth produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas Palmer amaranth produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} at second cohort in bareground, soybean and corn, respectively.
-The plant height was more impacted by Palmer amaranth growing in corn.
+The first Palmer amaranth first cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
+Palmer amaranth height was more impacted when growing in corn.
 Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
@@
 adaptive traits among plants (Jhala et al., 2021; Oliveira et al.,
-2018). Also, Palmer amaranth small seed size (e.g, 1 mm) tend to thrive
-in no-tillage systems (Price et al., 2011), and spread across locations
+2018). Also, Palmer amaranth small seeds (e.g, 1 mm) tend to thrive in
+no-tillage systems (Price et al., 2011), and spread across locations
 through farm equipment (Sauer, 1972), manure (Hartzler and Anderson,
-2016), animals (Farmer et al., 2017). The dispersal capacity of Palmer
-amaranth makes the species one of the most successful cases of weed
-adaptation to current cropping systems.
-
-Light and temperature are the main environment requirements for Palmer
-amaranth successful grow and development (Jha et al., 2010). Palmer
-amaranth is reported with an extended germination period (Ward et al.,
-2013). Germination of Palmer amaranth was triggered by 18 C soil
-temperature at 5 cm depth (Keeley et al., 1987), and optimal germination
-and biomass production occurred at 35/30 C day and night temperatures
-(Guo and Al-Khatib, 2003). In addition, Palmer amaranth establishment is
+2016), wildlife (Farmer et al., 2017), etc. The dispersal capacity of
+Palmer amaranth makes the species one of the most successful cases of
+weed adaptation to current cropping systems.
+
+Light and temperature are the main environmental requirements for Palmer
+amaranth growth and development (Jha et al., 2010). Palmer amaranth is
+reported with an extended germination period (Ward et al., 2013).
+Germination of Palmer amaranth was triggered by 18 C soil temperature at
+5 cm depth (Keeley et al., 1987), and optimal germination and biomass
+production occurred at 35/30 C day and night temperatures (Guo and
+Al-Khatib, 2003). In addition, Palmer amaranth establishment is
 human-mediated by tillage timings and preemergence-applied herbicides
 (Chahal et al., 2021), which can result in weed germination shifts
-(Sbatella and Wilson, 2010). Water has not shown to limit Palmer
-amaranth fitness. Under continuous water stress, Palmer amaranth
+(Sbatella and Wilson, 2010). Soil water content has not shown to limit
+Palmer amaranth fitness. Under continuous water stress, Palmer amaranth
 survived and produced at least 14000 seeds plant\textsuperscript{-1}
@@
 has shown plasticity to evolve and fast adapt under the current
-conditions. Further scenarios show that global temperature warming can
-impact agriculture, and promote niches for Palmer amaranth
+agroecosystem conditions. Further scenarios show that global temperature
+warming can impact agriculture, and promote niches for Palmer amaranth
 invasion/adaptation into new environments. Currently, it is estimated
@@
 (\emph{Amaranthus tuberculatus}) (Baker, 2021). Waterhemp, an
-\emph{Amaranthus} species, is the most troublesome species in that
-geography (Tranel et al., 2011). Nonetheless, invasion and successful
+\emph{Amaranthus} species, is the most troublesome species in the US
+Midwest (Tranel et al., 2011). Nonetheless, invasion and successful
 eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
@@
 America. Therefore, Palmer amaranth range of expansion into new habitats
-can increase. It seems certain the need to manage Palmer amaranth in
-agronomic crops throughout novel environments in the near future.
-Strategies on Palmer amaranth management should encompass the
-agroecosystem level but not only attempts to eradicate the weed. Most
-tactics to manage Palmer amaranth are based on technology fixes (Scott,
-2011), which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management. Palmer amaranth management should be built on
-minimizing the species ability to adapt, grow and develop into
-agroecossystems.
+can increase. It seems certain the need to manage new Palmer amaranth
+infestations in agronomic crops throughout northern United States in the
+near future. Strategies on Palmer amaranth management should encompass
+the agroecosystem level but not only attempts to eradicate the weed.
+Most tactics to manage Palmer amaranth are based on technology fixes
+(Scott, 2011), which are short-term (e.g., herbicide and/or tillage)
+rather than long-term weed management. Palmer amaranth management should
+be built on minimizing the species ability to adapt, grow and develop
+into agroecossystems.
 
@@
 
-A Palmer amaranth accession (Kei3) from Perkins County, Nebraska. with
-no reported herbicide resistance according to Oliveira et al. (2021) was
-selected for this study. Three weeks prior to the onset of the field
-experiments, seeds were planted in plastic trays containing potting-mix.
+A Palmer amaranth accession (Per1) from Perkins County, Nebraska with no
+reported herbicide resistance according to Oliveira et al. (2021) was
+selected for this study. Three weeks prior to the establishment of each
+cohort, seeds were planted in plastic trays containing potting-mix.
 Emerged seedlings (1 cm) were transplanted into 200
@@
 five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N,
-96.42'W), and Macomb, IL (XXX'N, XXX'W).
-
-Fields were conventionally tilled prior to crop planting. Corn hybrid
-and soybean varieties were planted in 76-cm row spacing (Table 1).
-Monthly mean air temperature and total precipitation were obtained using
-Daymet weather data from June through September across the five
-locations in 2018 and 2019 (Correndo et al., 2021) (Figure 1)
+(40.57'N, 9814'W), Grant, NE ('N, 'W), Lincoln, NE (41.16'N, 96.42'W),
+and Macomb, IL ('N, 'W).
+
+Fields were conventionally tilled prior to crop planting. Corn and
+soybean were planted in 76-cm row spacing (Table 1). Monthly mean air
+temperature and total precipitation were obtained using Daymet weather
+data from June through September across the five locations in 2018 and
+2019 (Correndo et al., 2021) (Figure 1)
 
@@
 
- & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{c} &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
+ & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+ & Stage at 1 cohort & V2-3 &  &  &  & \\
+
+ & Stage at 2 cohort & V6-7 &  &  &  & \\
 
@@
 
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
 
@@
 
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr\textsuperscript{d} + pyroxasulfone &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
+ & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+ & Stage at 1 cohort & V1-2 &  &  &  & \\
+
+ & Stage at 2 cohort & V5-6 &  &  &  & \\
 
@@
 arranged in factorial design with three crops, two transplanting times
-simulating two cohorts, repeated across five locations. Each
+simulating two cohorts, repeated across five locations. Each filed
 experimental unit was planted with corn, soybean, or kept under
@@
 When needed, Palmer amaranth plants were supplied with water during the
-first week after transplanting.
+first week after transplanting to assure seedling survival.
 
@@
 was identified (male or female), plant height was measured from soil
-surface to the top of plant. Also, aboveground plant biomass was harvest
-near soil surface and oven dried at 65 C until reaching constant weight
-before weighing (g plant\textsuperscript{-1}) was recorded.
+surface to the top of plant. Also, aboveground plant biomass was
+harvested near soil surface and oven dried at 65 C until reaching
+constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
 
@@
 neither endemic in Wisconsin nor in Illinois. In our study, all
-locations follow the methodology of plant harvest at flowering
+locations followed the methodology of plant harvest at flowering
 initiation, except in Grant, NE. In this location, all Palmer amaranth
@@
 The Palmer amaranth flowering timing was estimated as cumulative
-flowering across all location. Dataset was nested but it did not include
-Grant, NE. Palmer amaranth cumulative flowering estimation was
-determined using an asymmetrical three parameter log logistic Weibull
-model of the drc package (Ritz et al., 2015).
+flowering across all locations, except Grant, NE. Palmer amaranth
+cumulative flowering estimation was determined using an asymmetrical
+three parameter log logistic Weibull model of the drc package (Ritz et
+al., 2015).
 
@@
 
-Palmer amaranth sex dimorphism was fitted to a binary logistic
-regression (Bangdiwala, 2018). Binary logistic regression is used for
-predicting binary classes, such as the probability of a plant being
-female in a dioecious species. Prior to the analysis, all missing values
+A binary logistic regression was fitted to Palmer amaranth sex
+dimorphism. Binary logistic regression is used for predicting binary
+classes (Bangdiwala, 2018), such as the probability of a plant being
+female in a dioecious species. Prior to the analysis, missing values
 were removed from the dataset. Also, data from Grant was not used in
@@
 checking model performance on unseen dataset. With 80\% dataset, a
-binary response variable, male (0) and female (1), was fitted to a
-generalized linear model (base R \emph{glm} function) including day of
-year harvest, height, weight, crop and month as independent variables
-(without interaction). The model family was binomial with a logit
-function. The model fit was assessed through pseudo R-squared values
-(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
-\emph{nagelkerke} function from ``rcompanion'' package (Mangiafico,
-2021). The marginal effects computation was performed with Average
-Marginal Effects (AMEs) at every observed value of x and average across
-the results (Leeper, 2017) using \emph{margins} function from
-``margins'' package (Leeper et al., 2021). The rest 20\% dataset was
-predicted using \emph{predict} function with a cutoff estimation for
-male or female using \emph{performance} function. The model quality
-prediction from the classification algorithm was measured with precision
-(\emph{precision} function), recall (\emph{recall} function) and
-F1-score (\emph{f\_meas} function) using the ``yardstick'' package (Kuhn
-et al., 2021). The precision determines the accuracy of positive
-predictions (female plants), recall determines the fraction of positives
-that were correctly identified, and F1-score is a weighted harmonic mean
-of precision and recall with the best score of 1 and the worst score of
-0. F1-score conveys the balance between the precision and the recall
-(Yacouby and Axman, 2020).
+generalized linear model (base R \emph{glm} function) was fitted to
+binary response variable, the probability of being female (0 to male and
+1 to female). The independent variables were day of year harvest,
+height, weight, and crop (without interaction). The model family was
+binomial with a logit function. The model fit was assessed through
+pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
+likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
+package (Mangiafico, 2021). The marginal effects computation was
+performed with Average Marginal Effects (AMEs) at every observed value
+of x and average across the results (Leeper, 2017) using \emph{margins}
+function from ``margins'' package (Leeper et al., 2021). The 20\% test
+data was predicted using the \emph{predict} function with a cutoff
+estimation for male or female using \emph{performance} function. The
+model quality prediction from the classification algorithm was measured
+with precision (\emph{precision} function), recall (\emph{recall}
+function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
+package (Kuhn et al., 2021). The precision determines the accuracy of
+positive predictions (female plants), recall determines the fraction of
+positives that were correctly identified, and F1-score is a weighted
+harmonic mean of precision and recall with the best score of 1 and the
+worst score of 0. F1-score conveys the balance between the precision and
+the recall (Yacouby and Axman, 2020). The area under the curve was also
+estimated with XXXXXXXXXXXXXXXXX.
 
@@
 
-Palmer amaranth cumulative flowering at the second cohort ranged from
+Palmer amaranth cumulative flowering in the second cohort ranged from
 mid July to mid September (Figure 4B). Palmer amaranth growing in the
@@
 
-The model fit was 0.23, 0.32, 0.40 using pseudo R-squared test from
-McFadden, Cox and Snell, and Cragg and Uhler, respectively. The
-likelihood ratio test showed a p-value of \textless{} 0.00. The average
-marginal effects showed that Palmer amaranth growing in corn resulted in
-14.8\% less females plants (Table 2). Moreover, increasing a unit doy
-increases the probability of having a female plant by 0.4\% (Table 2 and
-Figure4A). Similar trend is observed for weight (Figure 4B) as well as
-height (Figure 4C), whereas the probability of being female increase by
-0.2\% and 0.1\% when a unit of weight (g) and height (cm) increases,
-respectively.
+The model goodness of fit was 0.23, 0.32, 0.40 using pseudo R-squared
+test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
+The likelihood ratio test showed a p-value of \textless{} 0.00. The
+average marginal effects showed that Palmer amaranth growing in corn
+resulted in 14.8\% less females plants (Table 2). Moreover, increasing a
+unit doy increases the probability of having a female plant by 0.4\%
+(Table 2 and Figure4A). Similar trend is observed for weight (Figure 4B)
+as well as height (Figure 4C), whereas the probability of being female
+increase by 0.2\% and 0.1\% when a unit of weight (g) and height (cm)
+increases, respectively.
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
 \centering
@@
 \toprule
-Term & AME\textsuperscript{a} & SE\textsuperscript{b} & Lower & Upper & Z-score & P-value\\
+Term & AME & SE & Lower & Upper & Z-score & P-value\\
 \midrule
@@
 classification showed a precision of 0.64, recall of 0.66, and a
-F1-score of 0.65. In addition, the area under the curve was 0.64.
+F1-score of 0.65. In addition, the AUC was 0.64.
 
@@
 amaranth can produce hundred thousands seeds per plant (Schwartz et al.,
-2016; Keeley et al., 1987), and stay viable buried in the seedbank for
-at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
-Palmer amaranth seed production or/and seed migration to its non-native
-habitat is an essential strategy to minimize the species impact into
-agroecossystem (Davis et al., 2015).
-
-An ecological approach to reduce seed production in Palmer amaranth is
-understanding its flowering window. Our study suggests that Palmer
-amaranth flowering was slightly influenced by crops and cohort timings.
-Palmer amaranth growing in bareground and corn resulted in the overall
-shortest and longest flowering window, respectively. The shortest
-flowering window of second cohorts of Palmer amaranth growing in
-bareground highlighted the impact of photoperiod on flowering. When
-growing in soybean, Palmer amaranth flowering window was similar to
-bareground at first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and depends on biological and
-ecological factors (Lang, 1965). We hypothesize that when growing in
-high competition (e.g., second cohort), Palmer amaranth plants tend to
-initiate flowering early, as well as having an extended flowering
-window. A study has shown that non-native Palmer amaranth initiated
-flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
-Early flower initiation could be also a plant strategy when growing in
-stress conditions. For example, when growing under water stress, early
-flowering in Palmer amaranth resulted in a mismatch between female and
-male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant outcross,
-reducing plant seed production and/or exchange of resistant alleles
-(Jhala et al., 2021). Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
+2016; Keeley et al., 1987), which can stay viable in the soil seedbank
+for at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
+Palmer amaranth seed production or/and seed migration to non-native
+habitats is an essential strategy to minimize the species impact in
+agroecosystems (Davis et al., 2015).
+
+An ecological approach to reduce Palmer amaranth seed production is
+understanding plant biology, including flowering pattern. Our study
+suggests that Palmer amaranth flowering was slightly influenced by crops
+and cohort timings. Palmer amaranth growing in bareground and corn
+resulted in the overall shortest and longest flowering window,
+respectively. The shortest flowering window of second cohorts of Palmer
+amaranth growing in bareground highlighted the impact of photoperiod on
+flowering. When growing in soybean, Palmer amaranth flowering window was
+similar to bareground at first cohort but similar to corn at second
+cohort timing. Plant flowering initiation is complex and depends on
+biological and ecological factors (Lang, 1965). We hypothesize that when
+growing in high competition (e.g., second cohort), Palmer amaranth
+plants tend to initiate flowering early, as well as having an extended
+flowering window. A study has shown that non-native Palmer amaranth
+initiated flowering two weeks prior to the native waterhemp in Iowa
+(Baker, 2021). Early flower initiation could also be a plant strategy
+when growing in stress conditions. For example, when growing under water
+stress, early flowering in Palmer amaranth resulted in a mismatch
+between female and male plants by seven days (Mesgaran et al., 2021). A
+mismatch in Palmer amaranth male and female flowering period can
+minimize plant outcross, reducing plant seed production and/or exchange
+of resistant alleles (Jhala et al., 2021). Sex dimorphism manipulation
+is considered a potential ecological pest control (McFarlane et al.,
+2018; Schliekelman et al., 2005).
 
@@
 (Korres et al., 2017). Our model also estimated more female plants in
-soybean and bareground compared to corn, which might linked to plant
+soybean and bareground compared to corn, which might be linked to plant
 competition strategy in each crop. Our results showed the influence of
@@
 Research on candidate genes for sex determination in \emph{Amaranthus}
-species are currently underway but it is far to complete (Montgomery et
-al., 2021, 2019). Further studies are also needed to understand the
+species are currently underway but it is far from complete (Montgomery
+et al., 2021, 2019). Further studies are also needed to understand the
 ecological basis of Palmer amaranth flowering, including the plant
@@
 require national efforts to minimize the species impact on
-sustainability and economy.
+sustainability and profitability of cropping systems.
 
@@
 \addcontentsline{toc}{section}{References}
-
-A Frontier article expect the reference list to be included in this
-section. To make that happens, the below syntax can be used. This
-\href{https://pandoc.org/MANUAL.html\#placement-of-the-bibliography}{feature
-is from Pandoc citeproc} which is used with \texttt{frontier\_article()}
-to handle the bibliography
 
@@
 Oliveira, M. C. (2021). \emph{Maxwelco/palmer\_adaptation}. {Zenodo}
-Available at: \url{ttps://doi.org/10.5281/zenodo.5236831}.
+Available at: \url{https://doi.org/10.5281/zenodo.5236831}.
 
@@
 \caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each experimental unit}\label{fig:Figure-2}
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit}\label{fig:Figure-2}
 \end{figure}
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
 \end{figure}
 
+\textbackslash begin\{figure\}
+
+\{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
+
+\}
+
+\textbackslash caption\{The probability (P) of being female Palmer
+amaranth by day of year (A), weight (B), and height (C). Back line
+represents the model estimation and shade green the 95\% confidence
+intervals\}\label{fig:Figure-5} \textbackslash end\{figure\}
+
 \begin{figure}
 
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second transplant timing (B)}\label{fig:Figure-5}
+\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
+From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
 \end{figure}
 
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
-
-}
-
-\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
-\end{figure}
-
 \end{document}

### 2021-08-26 data analysis/sex_analysis.Rmd
as

@@
 library(tidymodels)
+library(ggtext)
 library(yardstick)
@@
 ```
+
+```{r}
+# predict the preproc_test dataset
+preproc_test <- na.omit(preproc_test)
+pred <- predict(model_logi, preproc_test, type="response") 
+predicted <- round(pred) # round of the value; >0.5 will convert to 
+                         # 1 else 0
+# Creating a contigency table
+tab <- table(Predicted = predicted, Reference = preproc_test$fem)
+tab
+```
+
+```{r}
+#preproc_test <- na.omit(preproc_test)
+pred.rocr <- prediction(pred, preproc_test$fem)
+eval <- performance(pred.rocr, "acc")
+plot(eval)
+```
+
+```{r}
+# Identifying the best cutoff value that maximizes accuracy
+max <- which.max(slot(eval, "y.values")[[1]])
+acc <- slot(eval, "y.values")[[1]][max] #y.values are accuracy 
+                                        #measures
+cut <- slot(eval, "x.values")[[1]][max] #x.values are cutoff 
+                                        #measures
+print(c(Accuracy = acc, Cutoff = cut))
+```
+
+```{r}
+library(yardstick)
+# Creating a actual/observed vs predicted dataframe
+act_pred <- data.frame(observed = preproc_test$fem, predicted =  
+                       factor(predicted))
+# Calculating precision, recall and F1_score
+prec <- precision(act_pred, observed, predicted)
+rec <- recall(act_pred, observed, predicted)
+F1_score <- f_meas(act_pred, observed, predicted) #called f_measure
+print(prec)
+print(rec)
+print(F1_score)
+```
+
+```{r}
+perf_rocr <- performance(pred.rocr, measure = "auc",
+                         x.measure = "cutoff")
+perf_rocr@y.values[[1]] <- round(perf_rocr@y.values[[1]], digits = 
+                                 4)
+perf.tpr.fpr.rocr <- performance(pred.rocr, "tpr", "fpr")
+plot(perf.tpr.fpr.rocr, colorize=T, 
+     main = paste("AUC:", (perf_rocr@y.values)))
+abline(a = 0, b = 1)
+```

### 2021-08-27 Manuscript/Manuscript.Rmd
cx

--- 2021-08-26 Frontiers/Frontiers.Rmd
@@
 A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
-In general, results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
-The first Palmer amaranth first cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
+Results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
+The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
 Palmer amaranth height was more impacted when growing in corn.
@@
 \tiny
- \keyFont{ \section{Keywords:} Evolution Flowering Management Pigweed Weed}
+ \keyFont{ \section{Keywords:} Evolution, Flowering, Management, Pigweed, Weed}
 
@@
 
-A Palmer amaranth accession (Per1) from Perkins County, Nebraska with no reported herbicide resistance according to @oliveira2021a was selected for this study.
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no glyphosate resistance according to @oliveira2021a was selected for this study.
 Three weeks prior to the establishment of each cohort, seeds were planted in plastic trays containing potting-mix.
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE ('N, 'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL ('N, 'W).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (XXX'N, XXX'W).
 
@@
 Each filed experimental unit was planted with corn, soybean, or kept under bareground.
-The two transplanting timings were June 1 (first cohort) and July 1 (second cohort).
+The two transplant timings were June 1 (first cohort) and July 1 (second cohort).
 Palmer amaranth seedlings (potting mix + two seedlings) were transplanted (6 cm deep and 8 cm wide).
@@
 In the model, crop (bareground, corn, soybean) and cohort time (first and second) were the fixed effects and year nested with location the random effects.
-Analysis of variance was performed with *anova* function from "car" package [@fox2018].
+Analysis of variance at $\alpha$ 0.05 was performed with *anova* function from "car" package [@fox2018].
 Marginal means and compact letter display were estimated with *emmeans* and *cld* from packages "emmeans" [@lenth2021] and "multcomp" [@hothorn2008], respectively.
@@
 The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package [@leeper2021].
-The 20% test data was predicted using the *predict* function with a cutoff estimation for male or female using *performance* function.
+The 20% test data was predicted using the *predict* function with a cutoff estimation for male or female using *performance* function from ROCR package [@sing2005].
 The model quality prediction from the classification algorithm was measured with precision (*precision* function), recall (*recall* function) and F1-score (*f_meas* function) using the "yardstick" package [@kuhn2021].
-The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0.
+The precision determines the accuracy of positive predictions (female plants), recall determines the fraction of positives that were correctly identified, and F1-score is a weighted harmonic mean of precision and recall with the best score of 1 and the worst score of 0 [@raoniar2021].
 F1-score conveys the balance between the precision and the recall [@yacouby2020].
-The area under the curve was also estimated with XXXXXXXXXXXXXXXXX.
+The area under the receiver operating curve (AUC-ROC) was also estimated with performance function using the true positive and false positive rates.
+Higher the AUC, better the model is at distinguishing between female and male Palmer amaranth.
 
@@
 Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages.
-Plants were transplanted at greater crop height, which reduced Palmer amaranth competitiveness.
+Plants were transplanted at greater crop height and width, which reduced Palmer amaranth competitiveness.
 As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
-Palmer amaranth growing without crop competition produced the highest amounts of biomass.
+Moreover, Palmer amaranth growing without crop competition produced the highest amounts of biomass.
 The Palmer amaranth strategy in bareground was to invest biomass in growing plant width and height.
@@
 In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.
+Therefore, early management is a key strategy to minimize the damaging impact of Palmer amaranth to US Midwest cropping systems.
 
@@
 We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
-A study has shown that non-native Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
+A study has shown that Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
 Early flower initiation could also be a plant strategy when growing in stress conditions.
@@
 Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019].
-Our model estimated that late flowering, heavier and taller Palmer amaranth plants deviated from 1:1 ratio in favor to female plants.
+Our model estimated that late flowering, heavier and taller Palmer amaranth plants slighly deviated from 1:1 ratio in favor to female plants.
 It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
@@
 Our study demonstrated the Palmer amaranth plasticity to grow and develop into arable land of US Midwest.
-Is likely that Palmer amaranth range will continue to expanding into new geographies.
+It is likely that Palmer amaranth range will continue to expanding into new geographies.
 The migration of Palmer amaranth into the US Midwest will reshape the landscape as waterhemp and Palmer amaranth will share the same habitat.
@@
 
-<!--  
-Frontiers follows the recommendations by the International Committee of Medical
-Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require
-that all financial, commercial or other relationships that might be perceived by
-the academic community as representing a potential conflict of interest must be
-disclosed. If no such relationship exists, authors will be asked to declare that
-the research was conducted in the absence of any commercial or financial
-relationships that could be construed as a potential conflict of interest. When
-disclosing the potential conflict of interest, the authors need to address the
-following points:
-
- - Did you or your institution at any time receive payment or services from a
-   third party for any aspect of the submitted work?
- - Please declare financial relationships with entities that could be perceived
-   to influence, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- - Please declare patents and copyrights, whether pending, issued, licensed
-   and/or receiving royalties relevant to the work.
- - Please state other relationships or activities that readers could perceive to
-   have influenced, or that give the appearance of potentially influencing, what
-   you wrote in the submitted work.
- -->
-
 The authors declare that the research was conducted in the absence of any
@@
 
-<!--
-
-When determining authorship the following criteria should be observed:
-
- - Substantial contributions to the conception or design of the work; or the
-   acquisition, analysis, or interpretation of data for the work; AND
- - Drafting the work or revising it critically for important intellectual
-   content; AND
- - Final approval of the version to be published ; AND
- - Agreement to be accountable for all aspects of the work in ensuring that
-   questions related to the accuracy or integrity of any part of the work are
-   appropriately investigated and resolved.
-
-Contributors who meet fewer than all 4 of the above criteria for authorship
-should not be listed as authors, but they should be acknowledged.
-(http://www.icmje.org/roles_a.html)
-
--->
-
 RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
@@
 
-Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors
+Funding: This work received no specific grant from any funding agency, commercial, or not-for-profit sectors.
 
@@
 :::
+
+\newpage
 
@@
 ```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
-Back line represents the model estimation and shade green the 95% confidence intervals', out.width = "170mm", out.height = "70mm"}
+Black line represents the model estimation and shaded green the confidence intervals', out.width = "170mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2021-08-27 Manuscript/Manuscript.tex
cx

--- 2021-08-26 Frontiers/Frontiers.tex
@@
 
-A Palmer amaranth accession (Per1) from Perkins County, Nebraska with no
-reported herbicide resistance according to Oliveira et al. (2021) was
-selected for this study. Three weeks prior to the establishment of each
-cohort, seeds were planted in plastic trays containing potting-mix.
-Emerged seedlings (1 cm) were transplanted into 200
-cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
-Palmer amaranth seedlings were supplied with adequate water and kept
-under greenhouse conditions at the University of Wisconsin-Madison,
-University of Nebraska-Lincoln, and Western Illinois University; and
-kept outdoors at the Perkins extension office in Grant, NE until the 2-3
-leaf stage (5 to 8 cm height) when they were transported to the field.
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
+glyphosate resistance according to Oliveira et al. (2021) was selected
+for this study. Three weeks prior to the establishment of each cohort,
+seeds were planted in plastic trays containing potting-mix. Emerged
+seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3}
+plastic pots (a plant pot\textsuperscript{-1}). Palmer amaranth
+seedlings were supplied with adequate water and kept under greenhouse
+conditions at the University of Wisconsin-Madison, University of
+Nebraska-Lincoln, and Western Illinois University; and kept outdoors at
+the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to
+8 cm height) when they were transported to the field.
 
@@
 five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE ('N, 'W), Lincoln, NE (41.16'N, 96.42'W),
-and Macomb, IL ('N, 'W).
+(40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N,
+96.42'W), and Macomb, IL (XXX'N, XXX'W).
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}Field study attributes }
+\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}
 \centering
@@
 
- & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{c} &  &  & \\
 
@@
 
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
- & Stage at 1 cohort & V2-3 &  &  &  & \\
-
- & Stage at 2 cohort & V6-7 &  &  &  & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} &  &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
 
@@
 
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\
 
@@
 
- & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
- & Stage at 1 cohort & V1-2 &  &  &  & \\
-
- & Stage at 2 cohort & V5-6 &  &  &  & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
 
@@
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} \emph{S}--metolachor, 1324 g ai ha; \textsuperscript{c} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha; \textsuperscript{d} \emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha}\\
 \end{tabular}}
@@
 experimental unit was planted with corn, soybean, or kept under
-bareground. The two transplanting timings were June 1 (first cohort) and
+bareground. The two transplant timings were June 1 (first cohort) and
 July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
@@
 and year nested with location the random effects. Analysis of variance
-was performed with \emph{anova} function from ``car'' package (Fox and
-Weisberg, 2018). Marginal means and compact letter display were
-estimated with \emph{emmeans} and \emph{cld} from packages ``emmeans''
-(Lenth et al., 2021) and ``multcomp'' (Hothorn et al., 2008),
-respectively.
+at \(\alpha\) 0.05 was performed with \emph{anova} function from ``car''
+package (Fox and Weisberg, 2018). Marginal means and compact letter
+display were estimated with \emph{emmeans} and \emph{cld} from packages
+``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn et al.,
+2008), respectively.
 
@@
 data was predicted using the \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function. The
-model quality prediction from the classification algorithm was measured
-with precision (\emph{precision} function), recall (\emph{recall}
-function) and F1-score (\emph{f\_meas} function) using the ``yardstick''
-package (Kuhn et al., 2021). The precision determines the accuracy of
-positive predictions (female plants), recall determines the fraction of
-positives that were correctly identified, and F1-score is a weighted
-harmonic mean of precision and recall with the best score of 1 and the
-worst score of 0. F1-score conveys the balance between the precision and
-the recall (Yacouby and Axman, 2020). The area under the curve was also
-estimated with XXXXXXXXXXXXXXXXX.
+estimation for male or female using \emph{performance} function from
+ROCR package (Sing et al., 2005). The model quality prediction from the
+classification algorithm was measured with precision (\emph{precision}
+function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
+function) using the ``yardstick'' package (Kuhn et al., 2021). The
+precision determines the accuracy of positive predictions (female
+plants), recall determines the fraction of positives that were correctly
+identified, and F1-score is a weighted harmonic mean of precision and
+recall with the best score of 1 and the worst score of 0 (Raoniar,
+2021). F1-score conveys the balance between the precision and the recall
+(Yacouby and Axman, 2020). The area under the receiver operating curve
+(AUC-ROC) was also estimated with performance function using the true
+positive and false positive rates. Higher the AUC, better the model is
+at distinguishing between female and male Palmer amaranth.
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
 \centering
@@
 the crop competitive ability at advanced development stages. Plants were
-transplanted at greater crop height, which reduced Palmer amaranth
-competitiveness. As a result, Palmer amaranth height and biomass was
-lower compared to its first cohort. Palmer amaranth growing without crop
-competition produced the highest amounts of biomass. The Palmer amaranth
-strategy in bareground was to invest biomass in growing plant width and
-height. Nonetheless, Palmer amaranth produced 17\% less biomass in
-second cohort compared to first cohort timing. In a bareground study,
-early emerged Palmer amaranth without competition was 50\% taller than
-late emerged plants (Webster and Grey, 2015). These results suggest that
-crop competition is not the only factor limiting late Palmer amaranth
-establishment. The limited growth of Palmer amaranth at second cohort is
-likely a reduced plant response to day length, light availability and
-thermal units (e.g, growing degree days). The \emph{Amaranthus} species
-are sensitive to photoperiod (Wu and Owen, 2014). It is hypothesize that
-reduced day length contributed to smaller plants at second cohort
-regardless the crop. A study in North Carolina and Illinois predicted
-that less than 10\% Palmer amaranth seedlings emergence occurred after
-June (Piskackova et al., 2021). In addition, Palmer amaranth negative
-impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
-2015) yields was higher when plants were established near to crop
-planting.
+transplanted at greater crop height and width, which reduced Palmer
+amaranth competitiveness. As a result, Palmer amaranth height and
+biomass was lower compared to its first cohort. Moreover, Palmer
+amaranth growing without crop competition produced the highest amounts
+of biomass. The Palmer amaranth strategy in bareground was to invest
+biomass in growing plant width and height. Nonetheless, Palmer amaranth
+produced 17\% less biomass in second cohort compared to first cohort
+timing. In a bareground study, early emerged Palmer amaranth without
+competition was 50\% taller than late emerged plants (Webster and Grey,
+2015). These results suggest that crop competition is not the only
+factor limiting late Palmer amaranth establishment. The limited growth
+of Palmer amaranth at second cohort is likely a reduced plant response
+to day length, light availability and thermal units (e.g, growing degree
+days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
+and Owen, 2014). It is hypothesize that reduced day length contributed
+to smaller plants at second cohort regardless the crop. A study in North
+Carolina and Illinois predicted that less than 10\% Palmer amaranth
+seedlings emergence occurred after June (Piskackova et al., 2021). In
+addition, Palmer amaranth negative impact on soybean (Korres et al.,
+2020) and cotton (Webster and Grey, 2015) yields was higher when plants
+were established near to crop planting. Therefore, early management is a
+key strategy to minimize the damaging impact of Palmer amaranth to US
+Midwest cropping systems.
 
@@
 plants tend to initiate flowering early, as well as having an extended
-flowering window. A study has shown that non-native Palmer amaranth
-initiated flowering two weeks prior to the native waterhemp in Iowa
-(Baker, 2021). Early flower initiation could also be a plant strategy
-when growing in stress conditions. For example, when growing under water
-stress, early flowering in Palmer amaranth resulted in a mismatch
-between female and male plants by seven days (Mesgaran et al., 2021). A
-mismatch in Palmer amaranth male and female flowering period can
-minimize plant outcross, reducing plant seed production and/or exchange
-of resistant alleles (Jhala et al., 2021). Sex dimorphism manipulation
-is considered a potential ecological pest control (McFarlane et al.,
-2018; Schliekelman et al., 2005).
+flowering window. A study has shown that Palmer amaranth initiated
+flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
+Early flower initiation could also be a plant strategy when growing in
+stress conditions. For example, when growing under water stress, early
+flowering in Palmer amaranth resulted in a mismatch between female and
+male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
+amaranth male and female flowering period can minimize plant outcross,
+reducing plant seed production and/or exchange of resistant alleles
+(Jhala et al., 2021). Sex dimorphism manipulation is considered a
+potential ecological pest control (McFarlane et al., 2018; Schliekelman
+et al., 2005).
 
@@
 application (Rumpa et al., 2019). Our model estimated that late
-flowering, heavier and taller Palmer amaranth plants deviated from 1:1
-ratio in favor to female plants. It was reported that female Palmer
-amaranth plants invested more in height, stem and biomass while male
-invested more in leaf area and leaf dry weight under nutrient deficiency
-(Korres et al., 2017). Our model also estimated more female plants in
-soybean and bareground compared to corn, which might be linked to plant
-competition strategy in each crop. Our results showed the influence of
-life-history and ecological traits on sexual dimorphism in Palmer
-amaranth. Sexual dimorphism is documented in other dioecious species
-(Barrett and Hough, 2013). For example, stronger female plant
+flowering, heavier and taller Palmer amaranth plants slighly deviated
+from 1:1 ratio in favor to female plants. It was reported that female
+Palmer amaranth plants invested more in height, stem and biomass while
+male invested more in leaf area and leaf dry weight under nutrient
+deficiency (Korres et al., 2017). Our model also estimated more female
+plants in soybean and bareground compared to corn, which might be linked
+to plant competition strategy in each crop. Our results showed the
+influence of life-history and ecological traits on sexual dimorphism in
+Palmer amaranth. Sexual dimorphism is documented in other dioecious
+species (Barrett and Hough, 2013). For example, stronger female plant
 competition and greater male tolerance to herbivory was reported in
@@
 Our study demonstrated the Palmer amaranth plasticity to grow and
-develop into arable land of US Midwest. Is likely that Palmer amaranth
-range will continue to expanding into new geographies. The migration of
-Palmer amaranth into the US Midwest will reshape the landscape as
-waterhemp and Palmer amaranth will share the same habitat. The presence
-of Palmer amaranth and waterhemp will increase the weed management
-complexity. Therefore, preventive management is a priority to minimizing
-Palmer amaranth dispersal. Reactive management should focus on
-early-season management programs, which would have a large negative
+develop into arable land of US Midwest. It is likely that Palmer
+amaranth range will continue to expanding into new geographies. The
+migration of Palmer amaranth into the US Midwest will reshape the
+landscape as waterhemp and Palmer amaranth will share the same habitat.
+The presence of Palmer amaranth and waterhemp will increase the weed
+management complexity. Therefore, preventive management is a priority to
+minimizing Palmer amaranth dispersal. Reactive management should focus
+on early-season management programs, which would have a large negative
 effect on Palmer amaranth growth and development. Long-term tactics that
@@
 Funding: This work received no specific grant from any funding agency,
-commercial, or not-for-profit sectors
+commercial, or not-for-profit sectors.
 
@@
 
+\leavevmode\hypertarget{ref-raoniar2021}{}%
+Raoniar, R. (2021). Modelling {Binary Logistic Regression Using R}
+(research-oriented modelling and interpretation). Available at:
+\url{https://bit.ly/3BhucN3} {[}Accessed August 11, 2021{]}.
+
 \leavevmode\hypertarget{ref-ritz2015}{}%
@@
 
+\leavevmode\hypertarget{ref-sing2005}{}%
+Sing, T., Sander, O., Beerenwinkel, N., and Lengauer, T. (2005). {ROCR}:
+Visualizing classifier performance in {R}. \emph{Bioinformatics} 21,
+3940--3941.
+doi:\href{https://doi.org/10.1093/bioinformatics/bti623}{10.1093/bioinformatics/bti623}.
+
 \leavevmode\hypertarget{ref-sosnoskie2013}{}%
@@
 
+\newpage
+
 \hypertarget{figures}{%
@@
 
-\textbackslash begin\{figure\}
-
-\{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
-
-\}
-
-\textbackslash caption\{The probability (P) of being female Palmer
-amaranth by day of year (A), weight (B), and height (C). Back line
-represents the model estimation and shade green the 95\% confidence
-intervals\}\label{fig:Figure-5} \textbackslash end\{figure\}
+\begin{figure}
+
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
+
+}
+
+\caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
+Black line represents the model estimation and shaded green the confidence intervals}\label{fig:Figure-5}
+\end{figure}
 

### 2021-08-27 Manuscript/table.tex
cx

--- 2021-08-27 Manuscript/Manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
-%                                              %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Version 3.4 Generated 2018/06/15 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-
-\documentclass[utf8]{frontiersSCNS}
-
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-
-\linenumbers
-
-% BELOW TAKEN FROM rticles plos template
-%
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% hyperref package, useful for hyperlinks
-\usepackage{hyperref}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-% Sweave(-like)
-\usepackage{fancyvrb}
-\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
-\newenvironment{Schunk}{}{}
-\DefineVerbatimEnvironment{Code}{Verbatim}{}
-\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-\newenvironment{CodeChunk}{}{}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% Below is from frontiers
-%
-\bibliographystyle{frontiersinSCNS}
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace}
-%\doublespacing
-
-% Leave a blank line between paragraphs instead of using \\
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
-
-% Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\def\Authors{
-  Maxwel C Oliveira\,\textsuperscript{1},
-  Amit J Jhala\,\textsuperscript{2},
-  Mark Bernards\,\textsuperscript{3},
-  Chris Proctor\,\textsuperscript{2},
-  Strahinja Stepanovic\,\textsuperscript{2},
-  Rodrigo Werle\,\textsuperscript{1*}}
-
-\def\Address{
-
-  \textsuperscript{1} Department of Agronomy, University of
-Wisconsin-Madison,  Madison,  Wisconsin,  United States
-
-  \textsuperscript{2} Department of Agronomy and
-Horticulture, University of
-Nebraska-Lincoln,  Lincoln,  Nebraska,  United States
-
-  \textsuperscript{3} Department of Agronomy, Western Illinois
-University,  Macomb,  Illinois,  United States
-  }
-
-  \def\firstAuthorLast{Oliveira {et~al.}}
-
-  \def\corrAuthor{Rodrigo Werle}\def\corrAddress{Department of Agronomy,
-University of Wisconsin-Madison, United States\\1575 Linden
-Dr\\Madison, Wisconsin, 53705 United
-States}\def\corrEmail{\href{mailto:rwerle@uwisc.edu}{\nolinkurl{rwerle@uwisc.edu}}}
-
-\begin{document}
-\onecolumn
-\firstpage{1}
-
-\title[Palmer amaranth adaptation]{Palmer amaranth (\emph{Amaranthus
-palmeri}) adaptation to US Midwest agroecosystems}
-\author[\firstAuthorLast]{\Authors}
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
-
-\maketitle
-
-\begin{abstract}
-
-Palmer amaranth (\emph{Amaranthus palmeri} S.
-Watson) is one of the most troublesome weed species in the United States.
-Palmer amaranth is endemic to Southern United States but its range is expanding northward.
-Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
-Results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
-The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
-Palmer amaranth height was more impacted when growing in corn.
-Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
-Moreover, Palmer amaranth flowering window shifted according to crop and cohort timings.
-Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window.
-Also, Palmer amaranth sex dimorphism was slightly influenced by day of year, weight and height.
-The model estimated that probability of being a female plant increased as biomass and height increased.
-Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems from the US Midwest.
-Palmer amaranth is likely to continue its expansion northward.
-Therefore, preventing plant dispersal into new habitats is the most effective management strategy.
-Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability to establish into cropping systems, including crop rotation (beyond corn and soybean), early/late crop planting, row spacing, cover crops, and effective chemical control programs.
-
-\tiny
- \keyFont{ \section{Keywords:} Evolution, Flowering, Management, Pigweed, Weed}
-
-\end{abstract}
-
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-Palmer amaranth (\emph{Amaranthus palmeri} S. Watson) is currently
-ranked as one of the most economically detrimental weed species to
-cropping systems in the United States (Van Wychen, 2020). Unmanaged
-Palmer amaranth plants compete for water, light, and nutrients, which
-can drastically impact crop yields (Berger et al., 2015). For example,
-Palmer amaranth has been documented to reduce up to 91\%, 68\%, and 54\%
-corn (Massinga et al., 2001), soybean (Klingaman and Oliver, 1994), and
-cotton (Morgan et al., 2001) yields, respectively. Moreover, Palmer
-amaranth has shown a remarkable capacity to evolve resistance to
-herbicides. To date, Palmer amaranth has evolved resistance to eight
-herbicide sites of action (Heap, 2021), increasing the weed management
-complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
-economical and environmental risk to sustainable agriculture.
-
-Palmer amaranth is a fast growing summer annual forb indigenous to the
-Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
-threat to US agriculture in the 1990s. Palmer amaranth weediness is
-likely a result of human-assisted selection in combination with plant
-biology. Farm mechanization, adoption of conservation agriculture (e.g.,
-no-till), and intensive use of herbicides for weed management are the
-main human-mediated selection of Palmer amaranth into cropping systems
-(Ward et al., 2013). On the other hand, Palmer amaranth is a prolific
-seed producer with a C4 photosynthetic apparatus (Wang et al., 1992).
-With a dioecy nature, Palmer amaranth male and female plants are
-obligate outcrosser species, increasing the chances of exchanging
-adaptive traits among plants (Jhala et al., 2021; Oliveira et al.,
-2018). Also, Palmer amaranth small seeds (e.g, 1 mm) tend to thrive in
-no-tillage systems (Price et al., 2011), and spread across locations
-through farm equipment (Sauer, 1972), manure (Hartzler and Anderson,
-2016), wildlife (Farmer et al., 2017), etc. The dispersal capacity of
-Palmer amaranth makes the species one of the most successful cases of
-weed adaptation to current cropping systems.
-
-Light and temperature are the main environmental requirements for Palmer
-amaranth growth and development (Jha et al., 2010). Palmer amaranth is
-reported with an extended germination period (Ward et al., 2013).
-Germination of Palmer amaranth was triggered by 18 C soil temperature at
-5 cm depth (Keeley et al., 1987), and optimal germination and biomass
-production occurred at 35/30 C day and night temperatures (Guo and
-Al-Khatib, 2003). In addition, Palmer amaranth establishment is
-human-mediated by tillage timings and preemergence-applied herbicides
-(Chahal et al., 2021), which can result in weed germination shifts
-(Sbatella and Wilson, 2010). Soil water content has not shown to limit
-Palmer amaranth fitness. Under continuous water stress, Palmer amaranth
-survived and produced at least 14000 seeds plant\textsuperscript{-1}
-(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
-water conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). Growing conditions and management practices
-also influence Palmer amaranth sex dimorphism and flowering pattern
-(Korres et al., 2017; Rumpa et al., 2019). Therefore, Palmer amaranth
-has shown plasticity to evolve and fast adapt under the current
-agroecosystem conditions. Further scenarios show that global temperature
-warming can impact agriculture, and promote niches for Palmer amaranth
-invasion/adaptation into new environments. Currently, it is estimated
-that the greatest climatic risk of Palmer amaranth establishment are
-agronomic crops in Australia and Sub-Sahara Africa (Kistner and
-Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
-expansion to cooler geographies (Briscoe Runquist et al., 2019);
-however, under future climate change Palmer amaranth is likely to expand
-northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
-Briscoe Runquist et al., 2019).
-
-Palmer amaranth is already found in agronomic crops of South America
-(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
-et al., 2021). In the United States, Palmer amaranth is well established
-in the Cotton Belt (Garetson et al., 2019; Bagavathiannan and
-Norsworthy, 2016) in the southern United States but its range is
-expanding northward. For example, herbicide resistant Palmer amaranth is
-widespread in Nebraska (Oliveira et al., 2021). There are some reported
-cases of Palmer amaranth in Michigan (Kohrt et al., 2017) and
-Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
-amaranth can cause high damage to soybean fields in Illinois (Davis et
-al., 2015), which is concerning as soybean along with corn make most of
-US Midwest agronomic hectares. In Iowa, a study showed that Palmer
-amaranth is still not well adapted compared to waterhemp
-(\emph{Amaranthus tuberculatus}) (Baker, 2021). Waterhemp, an
-\emph{Amaranthus} species, is the most troublesome species in the US
-Midwest (Tranel et al., 2011). Nonetheless, invasion and successful
-eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
-2021). Palmer amaranth infestations have not been detected in Canada;
-however, Palmer amaranth seeds were detected in sweet potato slips in
-the country (Page et al., 2021). Palmer amaranth is still not as well
-adapted and established to Northern as it is in the Southern North
-America. Therefore, Palmer amaranth range of expansion into new habitats
-can increase. It seems certain the need to manage new Palmer amaranth
-infestations in agronomic crops throughout northern United States in the
-near future. Strategies on Palmer amaranth management should encompass
-the agroecosystem level but not only attempts to eradicate the weed.
-Most tactics to manage Palmer amaranth are based on technology fixes
-(Scott, 2011), which are short-term (e.g., herbicide and/or tillage)
-rather than long-term weed management. Palmer amaranth management should
-be built on minimizing the species ability to adapt, grow and develop
-into agroecossystems.
-
-In the southeastern US, early growing Palmer amaranth is well known to
-have a higher impact on cotton yields compared to late established
-plants (MacRae et al., 2013). In the northern states, Palmer amaranth
-impact on the agroecosystem is recent. Studies investigating Palmer
-amaranth in those locations are limited due to the plant classification
-as noxious weed species in some northern states (Yu et al., 2021).
-Nonetheless, the continuous Palmer amaranth dispersal and potential
-establishment across the northern United States is concerning and
-warrants investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under
-different agroecosystems can enhance our knowledge on species adaptation
-and management practices. It can also aid in designing proactive and
-ecological tactics to limit the species range expansion, reduce its
-negative impact, and developing resilient and sustainable farming
-systems (MacLaren et al., 2020). Therefore, the objective of this study
-was to investigate the flowering pattern, sex dimorphism, biomass
-production, and height of Palmer amaranth cohorts growing under corn,
-soybean and bareground across five locations in the United States
-Midwest.
-
-\hypertarget{material-and-methods}{%
-\section*{Material and Methods}\label{material-and-methods}}
-\addcontentsline{toc}{section}{Material and Methods}
-
-\hypertarget{plant-material-and-growing-conditions}{%
-\subsection*{Plant material and growing
-conditions}\label{plant-material-and-growing-conditions}}
-\addcontentsline{toc}{subsection}{Plant material and growing conditions}
-
-A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
-glyphosate resistance according to Oliveira et al. (2021) was selected
-for this study. Three weeks prior to the establishment of each cohort,
-seeds were planted in plastic trays containing potting-mix. Emerged
-seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3}
-plastic pots (a plant pot\textsuperscript{-1}). Palmer amaranth
-seedlings were supplied with adequate water and kept under greenhouse
-conditions at the University of Wisconsin-Madison, University of
-Nebraska-Lincoln, and Western Illinois University; and kept outdoors at
-the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to
-8 cm height) when they were transported to the field.
-
-\hypertarget{field-study}{%
-\subsection*{Field study}\label{field-study}}
-\addcontentsline{toc}{subsection}{Field study}
-
-The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N,
-96.42'W), and Macomb, IL (XXX'N, XXX'W).
-
-Fields were conventionally tilled prior to crop planting. Corn and
-soybean were planted in 76-cm row spacing (Table 1). Monthly mean air
-temperature and total precipitation were obtained using Daymet weather
-data from June through September across the five locations in 2018 and
-2019 (Correndo et al., 2021) (Figure 1)
-
 \begin{table}[!h]
@@
 
-The field experimental units were three adjacent 9.1 m wide (12 rows at
-76.2 cm row spacing) by 10.7 m long. The experimental design were
-arranged in factorial design with three crops, two transplanting times
-simulating two cohorts, repeated across five locations. Each filed
-experimental unit was planted with corn, soybean, or kept under
-bareground. The two transplant timings were June 1 (first cohort) and
-July 1 (second cohort). Palmer amaranth seedlings (potting mix + two
-seedlings) were transplanted (6 cm deep and 8 cm wide). Forty-eight
-plants were equidistantly placed (0.76 m apart) between rows within each
-crop. After a week, one plant was eliminated and one was kept, resulting
-in 24 plants per experimental unit and transplanting time (Figure 2).
-When needed, Palmer amaranth plants were supplied with water during the
-first week after transplanting to assure seedling survival.
-
-After transplanting, Palmer amaranth flowering was monitored until the
-end of the study. When a plant flowered, the day was recorded, plant sex
-was identified (male or female), plant height was measured from soil
-surface to the top of plant. Also, aboveground plant biomass was
-harvested near soil surface and oven dried at 65 C until reaching
-constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
-
-Plants had to be harvested at flowering because Palmer amaranth is
-neither endemic in Wisconsin nor in Illinois. In our study, all
-locations followed the methodology of plant harvest at flowering
-initiation, except in Grant, NE. In this location, all Palmer amaranth
-plants were harvest at once on July 06, 2018 and 2019 (first cohort),
-and on August 17, 2018 and on July 31, 2019 (second cohort).
-
-\hypertarget{statistical-analyses}{%
-\subsection*{Statistical analyses}\label{statistical-analyses}}
-\addcontentsline{toc}{subsection}{Statistical analyses}
-
-The statistical analyses were performed using R statistical software
-version 4.0.1 (Team, 2021). Data analyses are stored at Zenodo
-(Oliveira, 2021).
-
-Analyses of Palmer amaranth height and biomass were performed with a
-linear mixed model using \emph{lmer} function from ``lme4'' package
-(Bates et al., 2015). Plant height and biomass were log transformed to
-meet model assumption of normality. In the model, crop (bareground,
-corn, soybean) and cohort time (first and second) were the fixed effects
-and year nested with location the random effects. Analysis of variance
-at \(\alpha\) 0.05 was performed with \emph{anova} function from ``car''
-package (Fox and Weisberg, 2018). Marginal means and compact letter
-display were estimated with \emph{emmeans} and \emph{cld} from packages
-``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn et al.,
-2008), respectively.
-
-The Palmer amaranth flowering timing was estimated as cumulative
-flowering across all locations, except Grant, NE. Palmer amaranth
-cumulative flowering estimation was determined using an asymmetrical
-three parameter log logistic Weibull model of the drc package (Ritz et
-al., 2015).
-
-\[Y(x) = 0 + (d-0) exp (-exp(b(log(x)-e)))\]
-
-In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
-\emph{d} is the upper limit (set to 100), and \emph{e} is the inflection
-point, and \emph{x} day of year (doy).
-
-The doy for 10, 50, and 90\% Palmer amaranth cumulative flowering were
-determined using the \emph{ED} function of drc package. Also, the 10,
-50, and 90\% Palmer amaranth cumulative flowering were compared among
-crops and cohorts using the \emph{EDcomp} function of drc package. The
-EDcomp function compares the ratio of cumulative flowering using
-t-statistics, where P-value \textless{} 0.05 indicates that we fail to
-reject the null hypothesis.
-
-A binary logistic regression was fitted to Palmer amaranth sex
-dimorphism. Binary logistic regression is used for predicting binary
-classes (Bangdiwala, 2018), such as the probability of a plant being
-female in a dioecious species. Prior to the analysis, missing values
-were removed from the dataset. Also, data from Grant was not used in
-this analysis due to the uniform plant harvesting at that location. The
-complete dataset was splitted into 80\% train and 20\% test data. The
-80\% train is used for the model training and the 20\% test is used for
-checking model performance on unseen dataset. With 80\% dataset, a
-generalized linear model (base R \emph{glm} function) was fitted to
-binary response variable, the probability of being female (0 to male and
-1 to female). The independent variables were day of year harvest,
-height, weight, and crop (without interaction). The model family was
-binomial with a logit function. The model fit was assessed through
-pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
-likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
-package (Mangiafico, 2021). The marginal effects computation was
-performed with Average Marginal Effects (AMEs) at every observed value
-of x and average across the results (Leeper, 2017) using \emph{margins}
-function from ``margins'' package (Leeper et al., 2021). The 20\% test
-data was predicted using the \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function from
-ROCR package (Sing et al., 2005). The model quality prediction from the
-classification algorithm was measured with precision (\emph{precision}
-function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
-function) using the ``yardstick'' package (Kuhn et al., 2021). The
-precision determines the accuracy of positive predictions (female
-plants), recall determines the fraction of positives that were correctly
-identified, and F1-score is a weighted harmonic mean of precision and
-recall with the best score of 1 and the worst score of 0 (Raoniar,
-2021). F1-score conveys the balance between the precision and the recall
-(Yacouby and Axman, 2020). The area under the receiver operating curve
-(AUC-ROC) was also estimated with performance function using the true
-positive and false positive rates. Higher the AUC, better the model is
-at distinguishing between female and male Palmer amaranth.
-
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
-
-\hypertarget{palmer-amaranth-height-and-biomass}{%
-\subsection*{Palmer amaranth height and
-biomass}\label{palmer-amaranth-height-and-biomass}}
-\addcontentsline{toc}{subsection}{Palmer amaranth height and biomass}
-
-Palmer amaranth plants accumulated more biomass when growing in
-bareground compared to plants growing in soybean and corn (figure 3A).
-Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3
-g plant\textsuperscript{-1} in bareground, soybean and corn,
-respectively. Plants from the second cohort produced 62.6 g
-plant\textsuperscript{-1} in bareground, followed by 6.3 g
-plant\textsuperscript{-1} in soybean, and 1.4 g
-plant\textsuperscript{-1}.
-
-Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (figure 3B). Palmer amaranth plants from the first
-cohort were on average 69.2 cm tall in bareground, which was not
-different from the 70.7 cm tall plants from the second cohort timing (P
-= 0.74). In addition, no difference in Palmer amaranth height (69.3 cm)
-was detected from first cohort plants in soybean to first and second
-cohort plants in bareground (P \textgreater{} 0.75). Palmer amaranth
-plants from the second cohort were nearly 10 cm lower compared to the
-first cohort in soybeans (P = 0.04). The tallest (first cohort) and
-smallest (second cohort) Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
-
-\hypertarget{palmer-amaranth-cumulative-flowering}{%
-\subsection*{Palmer amaranth cumulative
-flowering}\label{palmer-amaranth-cumulative-flowering}}
-\addcontentsline{toc}{subsection}{Palmer amaranth cumulative flowering}
-
-Palmer amaranth plants from the first cohort growing in corn resulted in
-a longer flowering window compared to plants growing in bareground and
-soybean (Figure 4A). The 10\% cumulative Palmer amaranth flowering in
-soybean, bareground and corn occurred at the end of June. Palmer
-amaranth reached 10\% flowering in soybean, bareground and corn at doy
-180, 180.9 and 181.7, respectively. The 50\% Palmer amaranth cumulative
-flowering occurred in July. Palmer amaranth reached 50\% flowering in
-bareground, soybean and corn at doy 193.4, 194.8, and 206.6,
-respectively. Similar trend was observed at 90\% Palmer amaranth
-cumulative flowering. Palmer amaranth reached 90\% flowering at doy
-252.6 in corn (early September), which was 38 and 32 days after reaching
-90\% flowering in bareground and soybean, respectively.
-
-Palmer amaranth cumulative flowering in the second cohort ranged from
-mid July to mid September (Figure 4B). Palmer amaranth growing in the
-bareground resulted in earlier flowering time compared to soybean and
-corn. Palmer amaranth growing in bareground reached 10\%, 50\%, and 90\%
-flowering time at day 203.8, 214.4, and 232.2, respectively. Palmer
-amaranth growing in soybean reached 10\% flowering at doy 210.9, which
-was 6 days prior to corn (\emph{P}-value = 0.00). Similar trend was
-observed at 50\% flowering, whereas Palmer amaranth reached 50\%
-flowering in corn (doy 233.0) 4 days after soybeans (doy 228.9; \emph{P}
-= 0.00). The 90\% Palmer amaranth cumulative flowering occurred at same
-day in corn (260.9) and soybean (260.5; \emph{P} = 0.66).
-
-\hypertarget{palmer-amaranth-sex-dimorphism}{%
-\subsection*{Palmer amaranth sex
-dimorphism}\label{palmer-amaranth-sex-dimorphism}}
-\addcontentsline{toc}{subsection}{Palmer amaranth sex dimorphism}
-
-The model goodness of fit was 0.23, 0.32, 0.40 using pseudo R-squared
-test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
-The likelihood ratio test showed a p-value of \textless{} 0.00. The
-average marginal effects showed that Palmer amaranth growing in corn
-resulted in 14.8\% less females plants (Table 2). Moreover, increasing a
-unit doy increases the probability of having a female plant by 0.4\%
-(Table 2 and Figure4A). Similar trend is observed for weight (Figure 4B)
-as well as height (Figure 4C), whereas the probability of being female
-increase by 0.2\% and 0.1\% when a unit of weight (g) and height (cm)
-increases, respectively.
-
 \begin{table}[!h]
@@
 \end{table}
-
-The model accuracy evaluation accuracy in the 20\% test dataset was 0.62
-with a cutoff value for female and male plants of 0.43. The model
-classification showed a precision of 0.64, recall of 0.66, and a
-F1-score of 0.65. In addition, the AUC was 0.64.
-
-\hypertarget{discussion}{%
-\section*{Discussion}\label{discussion}}
-\addcontentsline{toc}{section}{Discussion}
-
-Our study showed that Palmer amaranth biomass, height, flowering window
-and sex varied within crops and cohort timings. In general, first cohort
-of Palmer amaranth plants were heavier and taller when compared to the
-second cohort. At first cohort, resources (e.g., soil nutrients) and
-conditions (e.g., light) were more timely available for the species.
-High biomass and taller Palmer amaranth plants are likely a weed
-strategy to compete for light in between crop rows in absence of canopy.
-In such conditions, Palmer amaranth showed an extraordinary plasticity
-to adapt upon the agroecosystem. This is evident when comparing Palmer
-amaranth canopy shape. The Palmer amaranth competition (e.g., light)
-strategy was to mimic the crop grow and development (Figure 6). These
-results suggests that Palmer amaranth can fast evolve life-history
-traits to adapt into cropping systems and cultural practices, which was
-also showed in a study varying nitrogen fertilization (Bravo et al.,
-2018). Our results highlight the Palmer amaranth as a threat to field
-crops as breeding more competitive crop varieties is likely to select
-more competitive weed biotypes (Bravo et al., 2017).
-
-Palmer amaranth grow and development in second cohort was limited due to
-the crop competitive ability at advanced development stages. Plants were
-transplanted at greater crop height and width, which reduced Palmer
-amaranth competitiveness. As a result, Palmer amaranth height and
-biomass was lower compared to its first cohort. Moreover, Palmer
-amaranth growing without crop competition produced the highest amounts
-of biomass. The Palmer amaranth strategy in bareground was to invest
-biomass in growing plant width and height. Nonetheless, Palmer amaranth
-produced 17\% less biomass in second cohort compared to first cohort
-timing. In a bareground study, early emerged Palmer amaranth without
-competition was 50\% taller than late emerged plants (Webster and Grey,
-2015). These results suggest that crop competition is not the only
-factor limiting late Palmer amaranth establishment. The limited growth
-of Palmer amaranth at second cohort is likely a reduced plant response
-to day length, light availability and thermal units (e.g, growing degree
-days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
-and Owen, 2014). It is hypothesize that reduced day length contributed
-to smaller plants at second cohort regardless the crop. A study in North
-Carolina and Illinois predicted that less than 10\% Palmer amaranth
-seedlings emergence occurred after June (Piskackova et al., 2021). In
-addition, Palmer amaranth negative impact on soybean (Korres et al.,
-2020) and cotton (Webster and Grey, 2015) yields was higher when plants
-were established near to crop planting. Therefore, early management is a
-key strategy to minimize the damaging impact of Palmer amaranth to US
-Midwest cropping systems.
-
-Seed production was not evaluated due to plant harvest at initiation of
-flowering. Nonetheless, it is well documented a strong positive
-correlation between Palmer amaranth biomass and seed production
-(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, plants
-growing at first cohort accumulated an overall 42\% more biomass when
-compared to second cohort. Therefore, Palmer amaranth plants growing in
-the second cohort is likely to produce less seeds regardless the crop.
-Our observation is consistent with the findings that first Palmer
-amaranth cohort produced 50\% more seeds per plant than Palmer amaranth
-plants established six weeks later in bareground (Webster and Grey,
-2015). Still, seed production at second cohort is likely to replenish
-the soil seedbank. Seed production and deposition in the seedbank is
-also a key factor for species perpetuation (Menges, 1987). Palmer
-amaranth can produce hundred thousands seeds per plant (Schwartz et al.,
-2016; Keeley et al., 1987), which can stay viable in the soil seedbank
-for at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
-Palmer amaranth seed production or/and seed migration to non-native
-habitats is an essential strategy to minimize the species impact in
-agroecosystems (Davis et al., 2015).
-
-An ecological approach to reduce Palmer amaranth seed production is
-understanding plant biology, including flowering pattern. Our study
-suggests that Palmer amaranth flowering was slightly influenced by crops
-and cohort timings. Palmer amaranth growing in bareground and corn
-resulted in the overall shortest and longest flowering window,
-respectively. The shortest flowering window of second cohorts of Palmer
-amaranth growing in bareground highlighted the impact of photoperiod on
-flowering. When growing in soybean, Palmer amaranth flowering window was
-similar to bareground at first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and depends on
-biological and ecological factors (Lang, 1965). We hypothesize that when
-growing in high competition (e.g., second cohort), Palmer amaranth
-plants tend to initiate flowering early, as well as having an extended
-flowering window. A study has shown that Palmer amaranth initiated
-flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
-Early flower initiation could also be a plant strategy when growing in
-stress conditions. For example, when growing under water stress, early
-flowering in Palmer amaranth resulted in a mismatch between female and
-male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant outcross,
-reducing plant seed production and/or exchange of resistant alleles
-(Jhala et al., 2021). Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
-
-The mechanisms of sex-determination in plant species is intriguing and
-arouse the curiosity of many scientists, including Charles Darwin
-(Darwin, 1888). In our study, the sex dimorphism model performance was
-decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male
-and female sex ratio is a general evolutionary stable strategy for plant
-species perpetuation (Fisher, 1930). However, a slight deviation from
-1:1 sex ratio might occur in some dioecious species. For example, the
-dioecious \emph{Halophila stipulacea} is a female-biased plant in its
-native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
-ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
-female-male ratio to expand into its non-native habitat (Nguyen et al.,
-2018). Also, biotic and/or abiotic stress can influence plant sex
-determination. Palmer amaranth male-to-female ratio was greater under
-high plant densities (Korres and Norsworthy, 2017) and after herbicide
-application (Rumpa et al., 2019). Our model estimated that late
-flowering, heavier and taller Palmer amaranth plants slighly deviated
-from 1:1 ratio in favor to female plants. It was reported that female
-Palmer amaranth plants invested more in height, stem and biomass while
-male invested more in leaf area and leaf dry weight under nutrient
-deficiency (Korres et al., 2017). Our model also estimated more female
-plants in soybean and bareground compared to corn, which might be linked
-to plant competition strategy in each crop. Our results showed the
-influence of life-history and ecological traits on sexual dimorphism in
-Palmer amaranth. Sexual dimorphism is documented in other dioecious
-species (Barrett and Hough, 2013). For example, stronger female plant
-competition and greater male tolerance to herbivory was reported in
-\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
-Research on candidate genes for sex determination in \emph{Amaranthus}
-species are currently underway but it is far from complete (Montgomery
-et al., 2021, 2019). Further studies are also needed to understand the
-ecological basis of Palmer amaranth flowering, including the plant
-behavior under climate change.
-
-Our study demonstrated the Palmer amaranth plasticity to grow and
-develop into arable land of US Midwest. It is likely that Palmer
-amaranth range will continue to expanding into new geographies. The
-migration of Palmer amaranth into the US Midwest will reshape the
-landscape as waterhemp and Palmer amaranth will share the same habitat.
-The presence of Palmer amaranth and waterhemp will increase the weed
-management complexity. Therefore, preventive management is a priority to
-minimizing Palmer amaranth dispersal. Reactive management should focus
-on early-season management programs, which would have a large negative
-effect on Palmer amaranth growth and development. Long-term tactics that
-promote early-season crop advantage against Palmer amaranth, including
-diversity of crops in rotation, early/late crop planting, plant width,
-and crop residue (e.g.~cover crops) would minimize the negative impact
-of Palmer amaranth to cropping systems. The aggressiveness and
-differential Palmer amaranth adaptation to agroecosystem is striking and
-require national efforts to minimize the species impact on
-sustainability and profitability of cropping systems.
-
-\hypertarget{disclosureconflict-of-interest-statement}{%
-\section*{Disclosure/Conflict-of-Interest
-Statement}\label{disclosureconflict-of-interest-statement}}
-\addcontentsline{toc}{section}{Disclosure/Conflict-of-Interest
-Statement}
-
-The authors declare that the research was conducted in the absence of
-any commercial or financial relationships that could be construed as a
-potential conflict of interest.
-
-\hypertarget{author-contributions}{%
-\section*{Author Contributions}\label{author-contributions}}
-\addcontentsline{toc}{section}{Author Contributions}
-
-RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
-the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP,
-MB, MO, SS, and RW: conceptualized the research. All authors reviewed
-the manuscript.
-
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-Funding: This work received no specific grant from any funding agency,
-commercial, or not-for-profit sectors.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-aulakh2021}{}%
-Aulakh, J. S., Chahal, P. S., Kumar, V., Price, A. J., and Guillard, K.
-(2021). Multiple herbicide-resistant {Palmer} amaranth ({Amaranthus}
-palmeri) in {Connecticut}: Confirmation and response to {POST}
-herbicides. \emph{Weed Technology} 35, 457--463.
-doi:\href{https://doi.org/10.1017/wet.2021.6}{10.1017/wet.2021.6}.
-
-\leavevmode\hypertarget{ref-bagavathiannan2016}{}%
-Bagavathiannan, M. V., and Norsworthy, J. K. (2016). Multiple-{Herbicide
-Resistance Is Widespread} in {Roadside Palmer Amaranth Populations}.
-\emph{PLOS ONE} 11, e0148748.
-doi:\href{https://doi.org/10.1371/journal.pone.0148748}{10.1371/journal.pone.0148748}.
-
-\leavevmode\hypertarget{ref-baker2021}{}%
-Baker, R. (2021). Comparative analysis of {Palmer} amaranth
-({Amaranthus} palmeri) and waterhemp ({A}. Tuberculatus) in {Iowa}.
-doi:\href{https://doi.org/10.31274/etd-20210609-11}{10.31274/etd-20210609-11}.
-
-\leavevmode\hypertarget{ref-bangdiwala2018}{}%
-Bangdiwala, S. I. (2018). Regression: Binary logistic.
-\emph{International Journal of Injury Control and Safety Promotion} 25,
-336--338.
-doi:\href{https://doi.org/10.1080/17457300.2018.1486503}{10.1080/17457300.2018.1486503}.
-
-\leavevmode\hypertarget{ref-barrett2013}{}%
-Barrett, S. C. H., and Hough, J. (2013). Sexual dimorphism in flowering
-plants. \emph{Journal of Experimental Botany} 64, 67--82.
-doi:\href{https://doi.org/10.1093/jxb/ers308}{10.1093/jxb/ers308}.
-
-\leavevmode\hypertarget{ref-bates2015}{}%
-Bates, D., Mächler, M., Bolker, B., and Walker, S. (2015). Fitting
-{Linear Mixed}-{Effects Models Using} Lme4. \emph{Journal of Statistical
-Software} 67, 1--48.
-doi:\href{https://doi.org/10.18637/jss.v067.i01}{10.18637/jss.v067.i01}.
-
-\leavevmode\hypertarget{ref-berger2015}{}%
-Berger, S. T., Ferrell, J. A., Rowland, D. L., and Webster, T. M.
-(2015). Palmer {Amaranth} ({Amaranthus} palmeri) {Competition} for
-{Water} in {Cotton}. \emph{Weed Science} 63, 928--935.
-doi:\href{https://doi.org/10.1614/WS-D-15-00062.1}{10.1614/WS-D-15-00062.1}.
-
-\leavevmode\hypertarget{ref-bravo2017}{}%
-Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
-(2017). Differentiation of {Life}-{History Traits} among {Palmer
-Amaranth Populations} ({Amaranthus} palmeri) and {Its Relation} to
-{Cropping Systems} and {Glyphosate Sensitivity}. \emph{Weed Science} 65,
-339--349.
-doi:\href{https://doi.org/10.1017/wsc.2017.14}{10.1017/wsc.2017.14}.
-
-\leavevmode\hypertarget{ref-bravo2018}{}%
-Bravo, W., Leon, R. G., Ferrell, J. A., Mulvaney, M. J., and Wood, C. W.
-(2018). Evolutionary {Adaptations} of {Palmer Amaranth} ({Amaranthus}
-palmeri) to {Nitrogen Fertilization} and {Crop Rotation History Affect
-Morphology} and {Nutrient}-{Use Efficiency}. \emph{Weed Science} 66,
-180--189.
-doi:\href{https://doi.org/10.1017/wsc.2017.73}{10.1017/wsc.2017.73}.
-
-\leavevmode\hypertarget{ref-briscoerunquist2019}{}%
-Briscoe Runquist, R. D., Lake, T., Tiffin, P., and Moeller, D. A.
-(2019). Species distribution models throughout the invasion history of
-{Palmer} amaranth predict regions at risk of future invasion and reveal
-challenges with modeling rapidly shifting geographic ranges. \emph{Sci
-Rep} 9, 2426.
-doi:\href{https://doi.org/10.1038/s41598-018-38054-9}{10.1038/s41598-018-38054-9}.
-
-\leavevmode\hypertarget{ref-chahal2021}{}%
-Chahal, P. S., Barnes, E. R., and Jhala, A. J. (2021). Emergence pattern
-of {Palmer} amaranth ({Amaranthus} palmeri) influenced by tillage
-timings and residual herbicides. \emph{Weed Technology} 35, 433--439.
-doi:\href{https://doi.org/10.1017/wet.2020.136}{10.1017/wet.2020.136}.
-
-\leavevmode\hypertarget{ref-chahal2018}{}%
-Chahal, P. S., Irmak, S., Jugulam, M., and Jhala, A. J. (2018).
-Evaluating {Effect} of {Degree} of {Water Stress} on {Growth} and
-{Fecundity} of {Palmer} amaranth ({Amaranthus} palmeri) {Using Soil
-Moisture Sensors}. \emph{Weed Science} 66, 738--745.
-doi:\href{https://doi.org/10.1017/wsc.2018.47}{10.1017/wsc.2018.47}.
-
-\leavevmode\hypertarget{ref-correndo2021}{}%
-Correndo, A. A., Moro Rosso, L. H., and Ciampitti, I. A. (2021).
-Retrieving and processing agro-meteorological data from {API}-client
-sources using {R} software. \emph{BMC Research Notes} 14, 205.
-doi:\href{https://doi.org/10.1186/s13104-021-05622-8}{10.1186/s13104-021-05622-8}.
-
-\leavevmode\hypertarget{ref-darwin1888}{}%
-Darwin, C. (1888). \emph{The {Different Forms} of {Flowers} on {Plants}
-of the {Same Species}}. {J. Murray} Available at:
-\url{http://books.google.com?id=7uMEAAAAYAAJ}.
-
-\leavevmode\hypertarget{ref-davis2015}{}%
-Davis, A. S., Schutte, B. J., Hager, A. G., and Young, B. G. (2015).
-Palmer {Amaranth} ({Amaranthus} palmeri) {Damage Niche} in {Illinois
-Soybean Is Seed Limited}. \emph{Weed Science} 63, 658--668.
-doi:\href{https://doi.org/10.1614/WS-D-14-00177.1}{10.1614/WS-D-14-00177.1}.
-
-\leavevmode\hypertarget{ref-farmer2017}{}%
-Farmer, J. A., Webb, E. B., Pierce, R. A., and Bradley, K. W. (2017).
-Evaluating the potential for weed seed dispersal based on waterfowl
-consumption and seed viability. \emph{Pest Management Science} 73,
-2592--2603. doi:\href{https://doi.org/10.1002/ps.4710}{10.1002/ps.4710}.
-
-\leavevmode\hypertarget{ref-fisher1930}{}%
-Fisher, R. A. (1930). The genetical theory of natural selection.
-\emph{Eugen Rev} 22, 127--130. Available at:
-\url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2984947/} {[}Accessed
-August 12, 2021{]}.
-
-\leavevmode\hypertarget{ref-fox2018}{}%
-Fox, J., and Weisberg, S. (2018). \emph{An {R Companion} to {Applied
-Regression}}. {SAGE Publications} Available at:
-\url{http://books.google.com?id=uPNrDwAAQBAJ}.
-
-\leavevmode\hypertarget{ref-garetson2019}{}%
-Garetson, R., Singh, V., Singh, S., Dotray, P., and Bagavathiannan, M.
-(2019). Distribution of herbicide-resistant {Palmer} amaranth
-({Amaranthus} palmeri) in row crop production systems in {Texas}.
-\emph{Weed Technology} 33, 355--365.
-doi:\href{https://doi.org/10.1017/wet.2019.14}{10.1017/wet.2019.14}.
-
-\leavevmode\hypertarget{ref-guo2003}{}%
-Guo, P., and Al-Khatib, K. (2003). Temperature effects on germination
-and growth of redroot pigweed ({Amaranthus} retroflexus), {Palmer}
-amaranth ({A}. Palmeri), and common waterhemp ({A}. rudis). \emph{Weed
-Science} 51, 869--875.
-doi:\href{https://doi.org/10.1614/P2002-127}{10.1614/P2002-127}.
-
-\leavevmode\hypertarget{ref-hartzler2016}{}%
-Hartzler, B., and Anderson, M. (2016). Palmer amaranth: {It}'s here, now
-what? 10.
-
-\leavevmode\hypertarget{ref-heap2021}{}%
-Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
-Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
-July 26, 2021{]}.
-
-\leavevmode\hypertarget{ref-hothorn2008}{}%
-Hothorn, T., Bretz, F., and Westfall, P. (2008). Simultaneous
-{Inference} in {General Parametric Models}. \emph{Biometrical Journal}
-50, 346--363.
-doi:\href{https://doi.org/10.1002/bimj.200810425}{10.1002/bimj.200810425}.
-
-\leavevmode\hypertarget{ref-jha2010}{}%
-Jha, P., Norsworthy, J. K., Riley, M. B., and Bridges, W. (2010). Annual
-{Changes} in {Temperature} and {Light Requirements} for {Germination} of
-{Palmer Amaranth} ({Amaranthus} palmeri) {Seeds Retrieved} from {Soil}.
-\emph{Weed Science} 58, 426--432.
-doi:\href{https://doi.org/10.1614/WS-D-09-00038.1}{10.1614/WS-D-09-00038.1}.
-
-\leavevmode\hypertarget{ref-jhala2021}{}%
-Jhala, A. J., Norsworthy, J. K., Ganie, Z. A., Sosnoskie, L. M., Beckie,
-H. J., Mallory-Smith, C. A., Liu, J., Wei, W., Wang, J., and
-Stoltenberg, D. E. (2021). Pollen-mediated gene flow and transfer of
-resistance alleles from herbicide-resistant broadleaf weeds. \emph{Weed
-Technology} 35, 173--187.
-doi:\href{https://doi.org/10.1017/wet.2020.101}{10.1017/wet.2020.101}.
-
-\leavevmode\hypertarget{ref-keeley1987}{}%
-Keeley, P. E., Carter, C. H., and Thullen, R. J. (1987). Influence of
-{Planting Date} on {Growth} of {Palmer Amaranth} ({Amaranthus} palmeri).
-\emph{Weed Science} 35, 199--204.
-doi:\href{https://doi.org/10.1017/S0043174500079054}{10.1017/S0043174500079054}.
-
-\leavevmode\hypertarget{ref-kistner2018}{}%
-Kistner, E. J., and Hatfield, J. L. (2018). Potential {Geographic
-Distribution} of {Palmer Amaranth} under {Current} and {Future
-Climates}. \emph{Agricultural \& Environmental Letters} 3, 170044.
-doi:\href{https://doi.org/10.2134/ael2017.12.0044}{10.2134/ael2017.12.0044}.
-
-\leavevmode\hypertarget{ref-klingaman1994}{}%
-Klingaman, T. E., and Oliver, L. R. (1994). Palmer {Amaranth}
-({Amaranthus} palmeri) {Interference} in {Soybeans} ({Glycine} max).
-\emph{Weed Science} 42, 523--527.
-doi:\href{https://doi.org/10.1017/S0043174500076888}{10.1017/S0043174500076888}.
-
-\leavevmode\hypertarget{ref-kohrt2017}{}%
-Kohrt, J. R., Sprague, C. L., Nadakuduti, S. S., and Douches, D. (2017).
-Confirmation of a {Three}-{Way} ({Glyphosate}, {ALS}, and {Atrazine})
-{Herbicide}-{Resistant Population} of {Palmer Amaranth} ({Amaranthus}
-palmeri) in {Michigan}. \emph{Weed Science} 65, 327--338.
-doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.
-
-\leavevmode\hypertarget{ref-korres2017a}{}%
-Korres, N. E., and Norsworthy, J. K. (2017). Palmer {Amaranth}
-({Amaranthus} palmeri) {Demographic} and {Biological Characteristics} in
-{Wide}-{Row Soybean}. \emph{Weed Science} 65, 491--503.
-doi:\href{https://doi.org/10.1017/wsc.2017.12}{10.1017/wsc.2017.12}.
-
-\leavevmode\hypertarget{ref-korres2017}{}%
-Korres, N. E., Norsworthy, J. K., FitzSimons, T., Roberts, T. L., and
-Oosterhuis, D. M. (2017). Differential {Response} of {Palmer Amaranth}
-({Amaranthus} palmeri) {Gender} to {Abiotic Stress}. \emph{Weed Science}
-65, 213--227.
-doi:\href{https://doi.org/10.1017/wsc.2016.34}{10.1017/wsc.2016.34}.
-
-\leavevmode\hypertarget{ref-korres2020}{}%
-Korres, N. E., Norsworthy, J. K., Mauromoustakos, A., and Williams, M.
-M. (2020). Soybean density and {Palmer} amaranth ({Amaranthus} palmeri)
-establishment time: Effects on weed biology, crop yield, and economic
-returns. \emph{Weed Science} 68, 467--475.
-doi:\href{https://doi.org/10.1017/wsc.2020.41}{10.1017/wsc.2020.41}.
-
-\leavevmode\hypertarget{ref-kuhn2021}{}%
-Kuhn, M., Vaughan, D., and RStudio (2021). \emph{Yardstick: {Tidy
-Characterizations} of {Model Performance}}. Available at:
-\url{https://CRAN.R-project.org/package=yardstick} {[}Accessed August
-24, 2021{]}.
-
-\leavevmode\hypertarget{ref-kupper2017}{}%
-Küpper, A., Borgato, E. A., Patterson, E. L., Netto, A. G., Nicolai, M.,
-Carvalho, S. J. P. de, Nissen, S. J., Gaines, T. A., and Christoffoleti,
-P. J. (2017). Multiple {Resistance} to {Glyphosate} and {Acetolactate
-Synthase Inhibitors} in {Palmer Amaranth} ({Amaranthus} palmeri)
-{Identified} in {Brazil}. \emph{Weed Science} 65, 317--326.
-doi:\href{https://doi.org/10.1017/wsc.2017.1}{10.1017/wsc.2017.1}.
-
-\leavevmode\hypertarget{ref-lang1965}{}%
-Lang, A. (1965). {``Physiology of flower initiation,''} in
-\emph{Differenzierung und {Entwicklung} / {Differentiation} and
-{Development}} Handbuch der {Pflanzenphysiologie} / {Encyclopedia} of
-{Plant Physiology}., ed. A. Lang ({Berlin, Heidelberg}: {Springer}),
-1380--1536.
-doi:\href{https://doi.org/10.1007/978-3-642-50088-6_39}{10.1007/978-3-642-50088-6\_39}.
-
-\leavevmode\hypertarget{ref-larran2017}{}%
-Larran, A. S., Palmieri, V. E., Perotti, V. E., Lieber, L., Tuesca, D.,
-and Permingeat, H. R. (2017). Target-site resistance to acetolactate
-synthase ({ALS})-inhibiting herbicides in {Amaranthus} palmeri from
-{Argentina}. \emph{Pest Management Science} 73, 2578--2584.
-doi:\href{https://doi.org/10.1002/ps.4662}{10.1002/ps.4662}.
-
-\leavevmode\hypertarget{ref-leeper2017}{}%
-Leeper, T. J. (2017). Interpreting {Regression Results} using {Average
-Marginal Eﬀects} with {R}'s margins. 31.
-
-\leavevmode\hypertarget{ref-leeper2021}{}%
-Leeper, T. J., Arnold, J., Arel-Bundock, V., and Long, J. A. (2021).
-\emph{Margins: {Marginal Effects} for {Model Objects}}. Available at:
-\url{https://CRAN.R-project.org/package=margins} {[}Accessed August 24,
-2021{]}.
-
-\leavevmode\hypertarget{ref-lenth2021}{}%
-Lenth, R. V., Buerkner, P., Herve, M., Love, J., Riebl, H., and
-Singmann, H. (2021). \emph{Emmeans: {Estimated Marginal Means}, aka
-{Least}-{Squares Means}}. Available at:
-\url{https://CRAN.R-project.org/package=emmeans} {[}Accessed August 24,
-2021{]}.
-
-\leavevmode\hypertarget{ref-lindsay2017}{}%
-Lindsay, K., Popp, M., Norsworthy, J., Bagavathiannan, M., Powles, S.,
-and Lacoste, M. (2017). {PAM}: {Decision Support} for {Long}-{Term
-Palmer Amaranth} ({Amaranthus} palmeri) {Control}. \emph{Weed
-Technology} 31, 915--927.
-doi:\href{https://doi.org/10.1017/wet.2017.69}{10.1017/wet.2017.69}.
-
-\leavevmode\hypertarget{ref-maclaren2020}{}%
-MacLaren, C., Storkey, J., Menegat, A., Metcalfe, H., and
-Dehnen-Schmutz, K. (2020). An ecological future for weed science to
-sustain crop production and the environment. {A} review. \emph{Agron.
-Sustain. Dev.} 40, 24.
-doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.
-
-\leavevmode\hypertarget{ref-macrae2013}{}%
-MacRae, A. W., Webster, T. M., Sosnoskie, L. M., Culpepper, A. S., and
-Kichler, J. M. (2013). Cotton {Yield Loss Potential} in {Response} to
-{Length} of {Palmer Amaranth} ({Amaranthus} palmeri) {Interference}. 17,
-6.
-
-\leavevmode\hypertarget{ref-mangiafico2021}{}%
-Mangiafico, S. (2021). \emph{Rcompanion: {Functions} to {Support
-Extension Education Program Evaluation}}. Available at:
-\url{https://CRAN.R-project.org/package=rcompanion} {[}Accessed August
-24, 2021{]}.
-
-\leavevmode\hypertarget{ref-massinga2001}{}%
-Massinga, R. A., Currie, R. S., Horak, M. J., and Boyer, J. (2001).
-Interference of {Palmer} amaranth in corn. \emph{Weed Science} 49,
-202--208.
-doi:\href{https://doi.org/10.1614/0043-1745(2001)049\%5B0202:IOPAIC\%5D2.0.CO;2}{10.1614/0043-1745(2001)049{[}0202:IOPAIC{]}2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-matzrafi2021}{}%
-Matzrafi, M., Osipitan, O. A., Ohadi, S., and Mesgaran, M. B. (2021).
-Under pressure: Maternal effects promote drought tolerance in progeny
-seed of {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed Science}
-69, 31--38.
-doi:\href{https://doi.org/10.1017/wsc.2020.75}{10.1017/wsc.2020.75}.
-
-\leavevmode\hypertarget{ref-mcfarlane2018}{}%
-McFarlane, G. R., Whitelaw, C. B. A., and Lillico, S. G. (2018).
-{CRISPR}-{Based Gene Drives} for {Pest Control}. \emph{Trends in
-Biotechnology} 36, 130--133.
-doi:\href{https://doi.org/10.1016/j.tibtech.2017.10.001}{10.1016/j.tibtech.2017.10.001}.
-
-\leavevmode\hypertarget{ref-menges1987}{}%
-Menges, R. M. (1987). Weed {Seed Population Dynamics} during {Six Years}
-of {Weed Management Systems} in {Crop Rotations} on {Irrigated Soil}.
-\emph{Weed Science} 35, 328--332. Available at:
-\url{http://www.jstor.org/stable/4044593}.
-
-\leavevmode\hypertarget{ref-mesgaran2021}{}%
-Mesgaran, M. B., Matzrafi, M., and Ohadi, S. (2021). Sex dimorphism in
-dioecious {Palmer} amaranth ({Amaranthus} palmeri) in response to water
-stress. \emph{Planta} 254, 17.
-doi:\href{https://doi.org/10.1007/s00425-021-03664-7}{10.1007/s00425-021-03664-7}.
-
-\leavevmode\hypertarget{ref-milani2021}{}%
-Milani, A., Panozzo, S., Farinati, S., Iamonico, D., Sattin, M., Loddo,
-D., and Scarabel, L. (2021). Recent {Discovery} of {Amaranthus} palmeri
-{S}. {Watson} in {Italy}: {Characterization} of {ALS}-{Resistant
-Populations} and {Sensitivity} to {Alternative Herbicides}.
-\emph{Sustainability} 13, 7003.
-doi:\href{https://doi.org/10.3390/su13137003}{10.3390/su13137003}.
-
-\leavevmode\hypertarget{ref-montgomery2021}{}%
-Montgomery, J. S., Giacomini, D. A., Weigel, D., and Tranel, P. J.
-(2021). Male-specific {Y}-chromosomal regions in waterhemp ({Amaranthus}
-tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{New
-Phytologist} 229, 3522--3533.
-doi:\href{https://doi.org/10.1111/nph.17108}{10.1111/nph.17108}.
-
-\leavevmode\hypertarget{ref-montgomery2019}{}%
-Montgomery, J. S., Sadeque, A., Giacomini, D. A., Brown, P. J., and
-Tranel, P. J. (2019). Sex-specific markers for waterhemp ({Amaranthus}
-tuberculatus) and {Palmer} amaranth ({Amaranthus} palmeri). \emph{Weed
-Science} 67, 412--418.
-doi:\href{https://doi.org/10.1017/wsc.2019.27}{10.1017/wsc.2019.27}.
-
-\leavevmode\hypertarget{ref-morgan2001}{}%
-Morgan, G. D., Baumann, P. A., and Chandler, J. M. (2001). Competitive
-{Impact} of {Palmer Amaranth} ({Amaranthus} palmeri) on {Cotton}
-({Gossypium} hirsutum) {Development} and {Yield}. \emph{Weed Technology}
-15, 408--412.
-doi:\href{https://doi.org/10.1614/0890-037X(2001)015\%5B0408:CIOPAA\%5D2.0.CO;2}{10.1614/0890-037X(2001)015{[}0408:CIOPAA{]}2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-nguyen2018}{}%
-Nguyen, H. M., Kleitou, P., Kletou, D., Sapir, Y., and Winters, G.
-(2018). Differences in flowering sex ratios between native and invasive
-populations of the seagrass {Halophila} stipulacea. \emph{Botanica
-Marina} 61, 337--342.
-doi:\href{https://doi.org/10.1515/bot-2018-0015}{10.1515/bot-2018-0015}.
-
-\leavevmode\hypertarget{ref-oliveira2021b}{}%
-Oliveira, M. C. (2021). \emph{Maxwelco/palmer\_adaptation}. {Zenodo}
-Available at: \url{https://doi.org/10.5281/zenodo.5236831}.
-
-\leavevmode\hypertarget{ref-oliveira2018}{}%
-Oliveira, M. C., Gaines, T. A., Patterson, E. L., Jhala, A. J., Irmak,
-S., Amundsen, K., and Knezevic, S. Z. (2018). Interspecific and
-intraspecific transference of metabolism-based mesotrione resistance in
-dioecious weedy {Amaranthus}. \emph{The Plant Journal} 96, 1051--1063.
-doi:\href{https://doi.org/10.1111/tpj.14089}{10.1111/tpj.14089}.
-
-\leavevmode\hypertarget{ref-oliveira2021a}{}%
-Oliveira, M. C., Giacomini, D. A., Arsenijevic, N., Vieira, G., Tranel,
-P. J., and Werle, R. (2021). Distribution and validation of genotypic
-and phenotypic glyphosate and {PPO}-inhibitor resistance in {Palmer}
-amaranth ({Amaranthus} palmeri) from southwestern {Nebraska}. \emph{Weed
-Technology} 35, 65--76.
-doi:\href{https://doi.org/10.1017/wet.2020.74}{10.1017/wet.2020.74}.
-
-\leavevmode\hypertarget{ref-page2021}{}%
-Page, E. R., Nurse, R. E., Meloche, S., Bosveld, K., Grainger, C.,
-Obeid, K., Filotas, M., Simard, M.-J., and Laforest, M. (2021). Import
-of {Palmer} amaranth ({Amaranthus} palmeri {S}. {Wats}.) Seed with sweet
-potato ({Ipomea} batatas ({L}.) {Lam}) slips. \emph{Can. J. Plant Sci.},
-CJPS-2020-0321.
-doi:\href{https://doi.org/10.1139/CJPS-2020-0321}{10.1139/CJPS-2020-0321}.
-
-\leavevmode\hypertarget{ref-perez-llorca2019}{}%
-Pérez-Llorca, M., and Sánchez Vilas, J. (2019). Sexual dimorphism in
-response to herbivory and competition in the dioecious herb {Spinacia}
-oleracea. \emph{Plant Ecol} 220, 57--68.
-doi:\href{https://doi.org/10.1007/s11258-018-0902-7}{10.1007/s11258-018-0902-7}.
-
-\leavevmode\hypertarget{ref-piskackova2021}{}%
-Piskackova, T. A. R., Reberg-Horton, S. C., Richardson, R. J., Jennings,
-K. M., Franca, L., Young, B. G., and Leon, R. G. (2021). Windows of
-action for controlling palmer amaranth ({Amaranthus} palmeri) using
-emergence and phenology models. \emph{Weed Research} 61, 188--198.
-doi:\href{https://doi.org/10.1111/wre.12470}{10.1111/wre.12470}.
-
-\leavevmode\hypertarget{ref-price2011}{}%
-Price, A. J., Balkcom, K. S., Culpepper, S. A., Kelton, J. A., Nichols,
-R. L., and Schomberg, H. (2011). Glyphosate-resistant {Palmer} amaranth:
-{A} threat to conservation tillage. \emph{Journal of Soil and Water
-Conservation} 66, 265--275.
-doi:\href{https://doi.org/10.2489/jswc.66.4.265}{10.2489/jswc.66.4.265}.
-
-\leavevmode\hypertarget{ref-raoniar2021}{}%
-Raoniar, R. (2021). Modelling {Binary Logistic Regression Using R}
-(research-oriented modelling and interpretation). Available at:
-\url{https://bit.ly/3BhucN3} {[}Accessed August 11, 2021{]}.
-
-\leavevmode\hypertarget{ref-ritz2015}{}%
-Ritz, C., Baty, F., Streibig, J. C., and Gerhard, D. (2015).
-Dose-{Response Analysis Using R}. \emph{PLOS ONE} 10, e0146021.
-doi:\href{https://doi.org/10.1371/journal.pone.0146021}{10.1371/journal.pone.0146021}.
-
-\leavevmode\hypertarget{ref-rumpa2019}{}%
-Rumpa, M. M., Krausz, R. F., Gibson, D. J., and Gage, K. L. (2019).
-Effect of {PPO}-{Inhibiting Herbicides} on the {Growth} and {Sex Ratio}
-of a {Dioecious Weed Species Amaranthus} palmeri ({Palmer Amaranth}).
-\emph{Agronomy} 9, 275.
-doi:\href{https://doi.org/10.3390/agronomy9060275}{10.3390/agronomy9060275}.
-
-\leavevmode\hypertarget{ref-sauer1957}{}%
-Sauer, J. (1957). Recent {Migration} and {Evolution} of the {Dioecious
-Amaranths}. \emph{Evolution} 11, 11--31.
-doi:\href{https://doi.org/10.2307/2405808}{10.2307/2405808}.
-
-\leavevmode\hypertarget{ref-sauer1972}{}%
-Sauer, J. D. (1972). The dioecious amaranths: A new species name and
-major range extensions. \emph{Madroño} 21, 426--434. Available at:
-\url{http://www.jstor.org/stable/41423815}.
-
-\leavevmode\hypertarget{ref-sbatella2010}{}%
-Sbatella, G. M., and Wilson, R. G. (2010). Isoxaflutole {Shifts Kochia}
-({Kochia} scoparia) {Populations} in {Continuous Corn}. \emph{Weed
-Technology} 24, 392--396.
-doi:\href{https://doi.org/10.1614/WT-D-09-00023.1}{10.1614/WT-D-09-00023.1}.
-
-\leavevmode\hypertarget{ref-schliekelman2005}{}%
-Schliekelman, P., Ellner, S., and Gould, F. (2005). Pest {Control} by
-{Genetic Manipulation} of {Sex Ratio}. \emph{Journal of Economic
-Entomology} 98, 18--34.
-doi:\href{https://doi.org/10.1093/jee/98.1.18}{10.1093/jee/98.1.18}.
-
-\leavevmode\hypertarget{ref-schwartz2016}{}%
-Schwartz, L. M., Norsworthy, J. K., Young, B. G., Bradley, K. W.,
-Kruger, G. R., Davis, V. M., Steckel, L. E., and Walsh, M. J. (2016).
-Tall {Waterhemp} ({Amaranthus} tuberculatus) and {Palmer} amaranth
-({Amaranthus} palmeri) {Seed Production} and {Retention} at {Soybean
-Maturity}. \emph{Weed Technology} 30, 284--290.
-doi:\href{https://doi.org/10.1614/WT-D-15-00130.1}{10.1614/WT-D-15-00130.1}.
-
-\leavevmode\hypertarget{ref-scott2011}{}%
-Scott, D. (2011). The {Technological Fix Criticisms} and the
-{Agricultural Biotechnology Debate}. \emph{J Agric Environ Ethics} 24,
-207--226.
-doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.
-
-\leavevmode\hypertarget{ref-sing2005}{}%
-Sing, T., Sander, O., Beerenwinkel, N., and Lengauer, T. (2005). {ROCR}:
-Visualizing classifier performance in {R}. \emph{Bioinformatics} 21,
-3940--3941.
-doi:\href{https://doi.org/10.1093/bioinformatics/bti623}{10.1093/bioinformatics/bti623}.
-
-\leavevmode\hypertarget{ref-sosnoskie2013}{}%
-Sosnoskie, L. M., Webster, T. M., and Culpepper, A. S. (2013).
-Glyphosate {Resistance Does Not Affect Palmer Amaranth} ({Amaranthus}
-palmeri) {Seedbank Longevity}. \emph{Weed Science} 61, 283--288.
-doi:\href{https://doi.org/10.1614/WS-D-12-00111.1}{10.1614/WS-D-12-00111.1}.
-
-\leavevmode\hypertarget{ref-spaunhorst2018}{}%
-Spaunhorst, D. J., Devkota, P., Johnson, W. G., Smeda, R. J., Meyer, C.
-J., and Norsworthy, J. K. (2018). Phenology of {Five Palmer} amaranth
-({Amaranthus} palmeri) {Populations Grown} in {Northern Indiana} and
-{Arkansas}. \emph{Weed Science} 66, 457--469.
-doi:\href{https://doi.org/10.1017/wsc.2018.12}{10.1017/wsc.2018.12}.
-
-\leavevmode\hypertarget{ref-rcoreteam2021}{}%
-Team, R. C. (2021). \emph{R: {The R Project} for {Statistical
-Computing}}. {Vienna, Austria}: {R Foundation for Statistical Computing}
-Available at: \url{https://www.r-project.org/} {[}Accessed August 24,
-2021{]}.
-
-\leavevmode\hypertarget{ref-tranel2011}{}%
-Tranel, P. J., Riggins, C. W., Bell, M. S., and Hager, A. G. (2011).
-Herbicide {Resistances} in {Amaranthus} tuberculatus: {A Call} for {New
-Options}. \emph{J. Agric. Food Chem.} 59, 5808--5812.
-doi:\href{https://doi.org/10.1021/jf103797n}{10.1021/jf103797n}.
-
-\leavevmode\hypertarget{ref-vanwychen2020}{}%
-Van Wychen, L. (2020). 2020 {Survey} of the most common and troublesome
-weeds in grass crops, pasture, and turf in the {United States} and
-{Canada}. Available at:
-\url{https://wssa.net/wp-content/uploads/2020-Weed-Survey_grass-crops.xlsx}.
-
-\leavevmode\hypertarget{ref-wang1992}{}%
-Wang, J. L., Klessig, D. F., and Berry, J. O. (1992). Regulation of {C4
-Gene Expression} in {Developing Amaranth Leaves}. \emph{The Plant Cell}
-4, 173--184.
-doi:\href{https://doi.org/10.1105/tpc.4.2.173}{10.1105/tpc.4.2.173}.
-
-\leavevmode\hypertarget{ref-ward2013}{}%
-Ward, S. M., Webster, T. M., and Steckel, L. E. (2013). Palmer
-{Amaranth} ({Amaranthus} palmeri): {A Review}. \emph{Weed Technology}
-27, 12--27.
-doi:\href{https://doi.org/10.1614/WT-D-12-00113.1}{10.1614/WT-D-12-00113.1}.
-
-\leavevmode\hypertarget{ref-webster2015}{}%
-Webster, T. M., and Grey, T. L. (2015). Glyphosate-{Resistant Palmer
-Amaranth} ({Amaranthus} palmeri) {Morphology}, {Growth}, and {Seed
-Production} in {Georgia}. \emph{Weed Science} 63, 264--272.
-doi:\href{https://doi.org/10.1614/WS-D-14-00051.1}{10.1614/WS-D-14-00051.1}.
-
-\leavevmode\hypertarget{ref-wu2014}{}%
-Wu, C., and Owen, M. D. K. (2014). When {Is} the {Best Time} to
-{Emerge}: {Reproductive Phenology} and {Success} of {Natural Common
-Waterhemp} ({Amaranthus} rudis) {Cohorts} in the {Midwest United
-States}? \emph{Weed Science} 62, 107--117. Available at:
-\url{http://www.jstor.org/stable/43700638}.
-
-\leavevmode\hypertarget{ref-yacouby2020}{}%
-Yacouby, R., and Axman, D. (2020). Probabilistic {Extension} of
-{Precision}, {Recall}, and {F1 Score} for {More Thorough Evaluation} of
-{Classification Models}. in \emph{Proceedings of the {First Workshop} on
-{Evaluation} and {Comparison} of {NLP Systems}} ({Online}: {Association
-for Computational Linguistics}), 79--91.
-doi:\href{https://doi.org/10.18653/v1/2020.eval4nlp-1.9}{10.18653/v1/2020.eval4nlp-1.9}.
-
-\leavevmode\hypertarget{ref-yu2021}{}%
-Yu, E., Blair, S., Hardel, M., Chandler, M., Thiede, D., Cortilet, A.,
-Gunsolus, J., and Becker, R. (2021). Timeline of {Palmer} amaranth
-({Amaranthus} palmeri) invasion and eradication in {Minnesota}.
-\emph{Weed Technology}, 1--31.
-doi:\href{https://doi.org/10.1017/wet.2021.32}{10.1017/wet.2021.32}.
-
-\end{CSLReferences}
-
-\newpage
-
-\hypertarget{figures}{%
-\section*{Figures}\label{figures}}
-\addcontentsline{toc}{section}{Figures}
-
-\begin{figure}
-
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
-
-}
-
-\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}
-
-}
-
-\caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit}\label{fig:Figure-2}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}
-
-}
-
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}
-
-}
-
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
-
-}
-
-\caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
-Black line represents the model estimation and shaded green the confidence intervals}\label{fig:Figure-5}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
-
-}
-
-\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
-\end{figure}
-
-\end{document}

### 2021-08-27 data analysis/sex_analysis.Rmd
cx

@@
   dplyr::select(fem, gender, location, year, month, crop, height, weight, doyh) -> data2
-  na.omit(data2) -> data2
 ```
@@
   geom_emoji(aes(x = 220, y = 0.05), emoji = "2642") + 
-    theme(legend.position = "none",
-        axis.text.y = element_markdown(size = 25),
-        axis.title = element_markdown(size = 30),
-        axis.text.x = element_markdown(size = 25),
-        legend.text = element_markdown(size = 25)) -> doyf
+    theme(legend.position = "none"#,
+#        axis.text.y = element_markdown(size = 25),
+#        axis.title = element_markdown(size = 30),
+#        axis.text.x = element_markdown(size = 25),
+#        legend.text = element_markdown(size = 25)
+        ) -> doyf
 
@@
   geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
-    theme(legend.position = "none",
-        axis.text.y = element_markdown(size = 25),
-        axis.title = element_markdown(size = 30),
-        axis.text.x = element_markdown(size = 25),
-        legend.text = element_markdown(size = 25)) -> weightf
+    theme(legend.position = "none"#,
+#        axis.text.y = element_markdown(size = 25),
+#        axis.title = element_markdown(size = 30),
+#        axis.text.x = element_markdown(size = 25),
+#        legend.text = element_markdown(size = 25)
+        ) -> weightf
 
@@
   geom_emoji(aes(x = 105, y = 0.05), emoji = "2642") +
-    theme(legend.position = "none",
-        axis.text.y = element_markdown(size = 25),
-        axis.title = element_markdown(size = 30),
-        axis.text.x = element_markdown(size = 25),
-        legend.text = element_markdown(size = 25)) -> heightf
+    theme(legend.position = "none"#,
+#        axis.text.y = element_markdown(size = 25),
+#        axis.title = element_markdown(size = 30),
+#        axis.text.x = element_markdown(size = 25),
+#        legend.text = element_markdown(size = 25)
+        ) -> heightf
 
@@
   plot_annotation(tag_levels = 'A') & 
-  theme(plot.tag = element_text(size = 30, hjust = 0, vjust = 0))
+  theme(plot.tag = element_text(size = 12, hjust = 0, vjust = 0))
 
-ggsave("figures/Figure 5.png", width = 9, height = 3)
+ggsave("figures/Figure 5.png", width = 10, height = 4)
 ```

### 2021-08-28 data analysis/weather/weather_2.Rmd
as

--- 2021-08-23 data analysis/weather/weather.Rmd
@@
 ---
-title: "weather"
+title: "weather_2"
 author: "Maxwel Coura Oliveira"
-date: "7/8/2021"
+date: "8/28/2021"
 output: html_document
@@
 library(nasapower)
+library(ggdist)
 library(vegan)
@@
 adaptation_weather1 %>% 
-  ggplot(aes(x = month_lab, y = tmean)) +
-  geom_jitter(aes(color = variable), alpha = 0.05, shape = 0) +
-  scale_color_manual(values = c("red", "blue", "gray90")) +
+  filter(variable != "NA") %>% 
+  ggplot(aes(x = month_lab, y = tmean, fill = variable)) +
+  geom_jitter(data = adaptation_weather1 %>% 
+                filter(variable == "NA"), show.legend = FALSE, 
+              color = "#333333", alpha = 0.02, shape = 16) +
+#   geom_jitter(position = position_dodge2(width = 0.4),
+#              color = "#333333", alpha = 1, shape = 16) +
+  stat_gradientinterval(position ="dodge") +
+  scale_fill_manual(values = c("red", "blue", "#333333")) +
   facet_grid(~ site) +
-  geom_line(adaptation_weather1 %>% 
-               filter(year %in% c(2018)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(tmean = mean(tmean)),
-             mapping = aes(x = month_lab, y = tmean, group = 1), 
-             color = "red") +
-  geom_line(adaptation_weather1 %>% 
-               filter(year %in% c(2019)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(tmean = mean(tmean)),
-             mapping = aes(x = month_lab, y = tmean, group = 1), 
-             color = "blue") +
-  geom_point(adaptation_weather1 %>% 
-               filter(year %in% c(2018)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(tmean = mean(tmean)),
-             mapping = aes(x = month_lab, y = tmean), 
-             fill = "red", shape = 22, size = 3) +
-  geom_point(adaptation_weather1 %>% 
-               filter(year %in% c(2019)) %>% 
-               group_by(site, month_lab) %>% 
-               summarise(tmean = mean(tmean)),
-             mapping = aes(x = month_lab, y = tmean), 
-             fill = "blue", shape = 22, size = 3) +
+  theme_test() +
   geom_richtext(data = text1, aes(x = 2, y = 7), lineheight = 0.4, 
-            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 6, 
+            label = "Mean air temperature <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
             label.color = NA, fill = NA) +
   geom_richtext(data = text2, aes(x = 2.2, y = 10), lineheight = 0.4, 
-            label = "30 years <br>daily mean <br>air temperature", size = 6, 
+            label = "30 years <br>daily mean <br>air temperature", size = 2, 
              label.color = NA, fill = NA, color = "gray90") +
-  geom_curve(data = text2, 
-             x = 2, y = 13, 
-             xend = 2.5, yend = 16,
-             arrow = arrow(length = unit(0.07, "inch")), 
-             size = 0.4, curvature = -0.3, color = "gray90") + 
-  scale_y_continuous(limits = c(0, 32)) +
+#  geom_curve(data = text2, 
+#             x = 2, y = 13, 
+#             xend = 2.5, yend = 16,
+#             arrow = arrow(length = unit(0.07, "inch")), 
+#             size = 0.4, curvature = -0.3, color = "gray90") + 
+  scale_y_continuous(limits = c(0, 33)) +
   labs(x = NULL, y = "Temperature (C)") +
-  theme_test() +
-  theme(legend.position = "none",
-#        strip.text = element_markdown(size = 15, hjust = 0.5),
-#        axis.text.y = element_markdown(size = 25),
-#        axis.title = element_markdown(size = 27),
-#        axis.text.x = element_markdown(size = 25)#,
-#        legend.text = element_markdown(size = 25)
-        )  -> temp
+  theme(legend.position = "none") -> temp
 
-ggsave("weather.png", height = 3, width = 6)
+ggsave("weather.png", height = 3, width = 9)
 ```

### 2021-08-28 weather_@.Rmd
as

---
title: "weather_2"
author: "Maxwel Coura Oliveira"
date: "8/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-09-01 Manuscript/Manuscript.Rmd
jhjh

@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (XXX'N, XXX'W).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (XXX'N, XXX'W).
 
@@
 
-The statistical analyses were performed using R statistical software version 4.0.1 [@rcoreteam2021]. Data analyses are stored at Zenodo [@oliveira2021b].
+The statistical analyses were performed using R statistical software version 4.0.1 [@rcoreteam2021].
 
@@
 
+# Data accessibility {-}
+
+The data and scripts used to analyse the data presented in this work can be found at Zenodo [@oliveira2021b].
+
 # Author Contributions {-}

### 2021-09-01 Manuscript/Manuscript.tex
jhjh

@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
-
- & Seeding rate & 88956 & 86487 &  &  & \\
-
- & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
-
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 &  & \\
+
+ & Seeding rate & 88956 & 86487 & 79000 &  & \\
+
+ & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} & glyphosate &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 &  & V2-3 &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 &  & V6-7 &  & \\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  May 5, 2018 / May 14, 2019 &  & \\
 
@@
 
- & Variety & DSR-1950 & AG21X8 &  &  & \\
-
- & Seeding rate & 296400 & 321237 &  &  & \\
-
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
-
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
+ & Variety & DSR-1950 & AG21X8 & AG28X7 &  & \\
+
+ & Seeding rate & 296400 & 321237 & 220000 &  & \\
+
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  glyphosate &  & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V1-2 &  & V1-2 &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 &  & V5-6 &  & \\
+
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 15, 2018 / May 3, 2019 &  & \\
 

### 2021-09-01 data analysis/weather/weather_2.Rmd
jhjh

@@
   group_by(month_lab, site, year) %>% 
-  summarise(pp = sum(pp)) %>% 
+#  summarise(pp = sum(pp)) %>% 
   mutate(year = as.character(year),
@@
                            "Macomb, IL" = "Macomb")) -> adaptation_pp
+```
+
+```{r}
+adaptation_pp %>% 
+  filter(!year %in% c(2018, 2019)) %>% 
+ggplot(aes(x = month_lab, y = pp)) +
+  stat_eye(position = "dodge")
 ```

### 2021-09-16 Manuscript/Manuscript.Rmd
as

@@
 Palmer amaranth (\emph{Amaranthus palmeri} S.
-Watson) is one of the most troublesome weed species in the United States.
+Watson) is one of the most troublesome agronomic weed species in the United States.
 Palmer amaranth is endemic to Southern United States but its range is expanding northward.
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
+A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
 Results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
@@
 To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Thus, Palmer amaranth poses an economical and environmental risk to sustainable agriculture.
+Thus, Palmer amaranth poses an economical and ecological risk to sustainable agriculture.
 
@@
 Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology.
-Farm mechanization, adoption of conservation agriculture (e.g., no-till), and intensive use of herbicides for weed management are the main human-mediated selection of Palmer amaranth into cropping systems [@ward2013].
+Farm mechanization, adoption of conservation agriculture (e.g., no-till), and intensive use of herbicides for weed management are the main human-mediated selections of Palmer amaranth into cropping systems [@ward2013].
 On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
-With a dioecy nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
-Also, Palmer amaranth small seeds (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], manure [@hartzler2016], wildlife [@farmer2017], etc.
+With a dioecious nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
+Also, Palmer amaranth's small seeds (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], seed mixes [@hartzler2016], wildlife [@farmer2017], etc.
 The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.
@@
 Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003].
-In addition, Palmer amaranth establishment is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
-Soil water content has not shown to limit Palmer amaranth fitness.
-Under continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
+In addition, Palmer amaranth emergence is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
+Also, Palmer amaranth has shown to grow and develop under reduced soil water content.
+In continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
 Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
@@
 Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under the current agroecosystem conditions.
-Further scenarios show that global temperature warming can impact agriculture, and promote niches for Palmer amaranth invasion/adaptation into new environments.
-Currently, it is estimated that the greatest climatic risk of Palmer amaranth establishment are agronomic crops in Australia and Sub-Sahara Africa [@kistner2018].
+Future scenarios suggest that global temperature warming can impact agriculture, and promote niches for Palmer amaranth invasion/adaptation into new environments.
+Agronomic crops in Australia and Sub-Sahara Africa are currently at greatest risk to Palmer invasion as climate warms [@kistner2018].
 Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018; @briscoerunquist2019].
@@
 There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021].
-Also, it is estimated that Palmer amaranth can cause high damage to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn make most of US Midwest agronomic hectares.
-In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021].
-Waterhemp, an *Amaranthus* species, is the most troublesome species in the US Midwest [@tranel2011].
-Nonetheless, invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
+Also, it is estimated that Palmer amaranth can cause yield loss to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn comprise most of US Midwest agronomic hectares.
+In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021], which is considered the most troublesome species in the US Midwest [@tranel2011].
+Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
 Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021].
-Palmer amaranth is still not as well adapted and established to Northern as it is in the Southern North America.
+Palmer amaranth is still not as well adapted and established to Northern as compared to Southern North America.
 Therefore, Palmer amaranth range of expansion into new habitats can increase.
-It seems certain the need to manage new Palmer amaranth infestations in agronomic crops throughout northern United States in the near future.
-Strategies on Palmer amaranth management should encompass the agroecosystem level but not only attempts to eradicate the weed.
-Most tactics to manage Palmer amaranth are based on technology fixes [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
-Palmer amaranth management should be built on minimizing the species ability to adapt, grow and develop into agroecossystems.
+It seems certain the need to manage new Palmer amaranth infestations in agronomic crops throughout northern United States is eminent in the near future.
+Strategies on Palmer amaranth management should encompass the agroecosystem level but not focus only on attempts to eradicate the weed.
+Most tactics to manage Palmer amaranth are based on technological solutions [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
+Palmer amaranth management should be built around minimizing the species ability to adapt, grow and develop successfully in agroecossystems.
 
@@
 In the northern states, Palmer amaranth impact on the agroecosystem is recent.
-Studies investigating Palmer amaranth in those locations are limited due to the plant classification as noxious weed species in some northern states [@yu2021].
-Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under different agroecosystems can enhance our knowledge on species adaptation and management practices.
-It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact, and developing resilient and sustainable farming systems [@maclaren2020].
+Studies investigating Palmer amaranth in northern states are limited due to the plant classification as noxious weed species [@yu2021].
+Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in northern environments.
+Understanding Palmer amaranth biology and morphology under different agroecosystems can enhance our knowledge on species adaptation and management practices.
+It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact on agroecossystems, and to develop resilient and sustainable farming systems [@maclaren2020].
 Therefore, the objective of this study was to investigate the flowering pattern, sex dimorphism, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
@@
 When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and depends on biological and ecological factors [@lang1965].
+Plant flowering initiation is complex and depends on the complex interaction between genetic makeup and the environmental conditions [@lang1965].
 We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
@@
 cover crops) would minimize the negative impact of Palmer amaranth to cropping systems.
-The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require national efforts to minimize the species impact on sustainability and profitability of cropping systems.
+The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require regional collaboration to minimize the species impact on sustainability and profitability of cropping systems.
 
@@
 
-RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
+RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted the experiments; MO: analyzed the data; MO: wrote the manuscript and all authors revised; AJ, CP, MB, MO, SS, and RW: conceptualized the research.
 All authors reviewed the manuscript.

### 2021-09-16 Manuscript/Manuscript.tex
as

--- 2021-08-27 Manuscript/Manuscript.tex
@@
 complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
-economical and environmental risk to sustainable agriculture.
+economical and ecological risk to sustainable agriculture.
 
@@
 no-tillage systems (Price et al., 2011), and spread across locations
-through farm equipment (Sauer, 1972), manure (Hartzler and Anderson,
+through farm equipment (Sauer, 1972), seed mixes (Hartzler and Anderson,
 2016), wildlife (Farmer et al., 2017), etc. The dispersal capacity of
@@
 (Chahal et al., 2021), which can result in weed germination shifts
-(Sbatella and Wilson, 2010). Soil water content has not shown to limit
-Palmer amaranth fitness. Under continuous water stress, Palmer amaranth
-survived and produced at least 14000 seeds plant\textsuperscript{-1}
-(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
-water conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). Growing conditions and management practices
-also influence Palmer amaranth sex dimorphism and flowering pattern
-(Korres et al., 2017; Rumpa et al., 2019). Therefore, Palmer amaranth
-has shown plasticity to evolve and fast adapt under the current
-agroecosystem conditions. Further scenarios show that global temperature
-warming can impact agriculture, and promote niches for Palmer amaranth
-invasion/adaptation into new environments. Currently, it is estimated
-that the greatest climatic risk of Palmer amaranth establishment are
-agronomic crops in Australia and Sub-Sahara Africa (Kistner and
-Hatfield, 2018). Temperature is a key factor limiting Palmer amaranth
-expansion to cooler geographies (Briscoe Runquist et al., 2019);
-however, under future climate change Palmer amaranth is likely to expand
-northward into Canada and Northern Europe (Kistner and Hatfield, 2018;
-Briscoe Runquist et al., 2019).
+(Sbatella and Wilson, 2010). Also, Palmer amaranth has shown to grow and
+develop under reduced soil water content. In continuous water stress,
+Palmer amaranth survived and produced at least 14000 seeds
+plant\textsuperscript{-1} (Chahal et al., 2018). Seeds from Palmer
+amaranth growing with limited water conditions were heavier, less
+dormant, and prompt for germination (Matzrafi et al., 2021). Growing
+conditions and management practices also influence Palmer amaranth sex
+dimorphism and flowering pattern (Korres et al., 2017; Rumpa et al.,
+2019). Therefore, Palmer amaranth has shown plasticity to evolve and
+fast adapt under the current agroecosystem conditions. Further scenarios
+show that global temperature warming can impact agriculture, and promote
+niches for Palmer amaranth invasion/adaptation into new environments.
+Currently, it is estimated that the greatest climatic risk of Palmer
+amaranth establishment are agronomic crops in Australia and Sub-Sahara
+Africa (Kistner and Hatfield, 2018). Temperature is a key factor
+limiting Palmer amaranth expansion to cooler geographies (Briscoe
+Runquist et al., 2019); however, under future climate change Palmer
+amaranth is likely to expand northward into Canada and Northern Europe
+(Kistner and Hatfield, 2018; Briscoe Runquist et al., 2019).
 
@@
 five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE (XXX'N, XXX'W), Lincoln, NE (41.16'N,
+(40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N,
 96.42'W), and Macomb, IL (XXX'N, XXX'W).
@@
 
-\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}
+\caption{\label{tab:unnamed-chunk-2}Field study attributes }
 \centering
@@
 
- & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{c} &  &  & \\
+ & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
 
@@
 
- & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
+ & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+ & Stage at 1 cohort & V2-3 &  &  &  & \\
+
+ & Stage at 2 cohort & V6-7 &  &  &  & \\
 
@@
 
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
- & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
+ & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+ & Stage at 1 cohort & V1-2 &  &  &  & \\
+
+ & Stage at 2 cohort & V5-6 &  &  &  & \\
 
@@
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} \emph{S}--metolachor, 1324 g ai ha; \textsuperscript{c} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha; \textsuperscript{d} \emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
 \end{tabular}}
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1 (Team, 2021). Data analyses are stored at Zenodo
-(Oliveira, 2021).
+version 4.0.1 (Team, 2021).
 
@@
 similar to bareground at first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and depends on
-biological and ecological factors (Lang, 1965). We hypothesize that when
-growing in high competition (e.g., second cohort), Palmer amaranth
-plants tend to initiate flowering early, as well as having an extended
-flowering window. A study has shown that Palmer amaranth initiated
-flowering two weeks prior to the native waterhemp in Iowa (Baker, 2021).
-Early flower initiation could also be a plant strategy when growing in
-stress conditions. For example, when growing under water stress, early
+cohort timing. Plant flowering initiation is complex and depends on the
+complex interaction between genetic makeup and the environmental
+conditions (Lang, 1965). We hypothesize that when growing in high
+competition (e.g., second cohort), Palmer amaranth plants tend to
+initiate flowering early, as well as having an extended flowering
+window. A study has shown that Palmer amaranth initiated flowering two
+weeks prior to the native waterhemp in Iowa (Baker, 2021). Early flower
+initiation could also be a plant strategy when growing in stress
+conditions. For example, when growing under water stress, early
 flowering in Palmer amaranth resulted in a mismatch between female and
@@
 differential Palmer amaranth adaptation to agroecosystem is striking and
-require national efforts to minimize the species impact on
+require regional collaboration to minimize the species impact on
 sustainability and profitability of cropping systems.
@@
 
+\hypertarget{data-accessibility}{%
+\section*{Data accessibility}\label{data-accessibility}}
+\addcontentsline{toc}{section}{Data accessibility}
+
+The data and scripts used to analyse the data presented in this work can
+be found at Zenodo (Oliveira, 2021).
+
 \hypertarget{author-contributions}{%
@@
 RW and MO: designed the experiments; AJ, CP, MB, MO, and SS: conducted
-the experiments; MO: analyzed the data and wrote the manuscript; AJ, CP,
-MB, MO, SS, and RW: conceptualized the research. All authors reviewed
-the manuscript.
+the experiments; MO: analyzed the data; MO: wrote the manuscript and all
+authors revised; AJ, CP, MB, MO, SS, and RW: conceptualized the
+research. All authors reviewed the manuscript.
 

### 2021-09-17 Manuscript/Manuscript.Rmd
as

@@
 A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
-Results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
+Results demonstrated that the first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from the second cohort (established in July).
 The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
@@
 Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window.
-Also, Palmer amaranth sex dimorphism was slightly influenced by day of year, weight and height.
+Also, Palmer amaranth gender was slightly influenced by day of year, weight and height.
 The model estimated that probability of being a female plant increased as biomass and height increased.
@@
 It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact on agroecossystems, and to develop resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, sex dimorphism, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
+Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
 
@@
 The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations.
-Each filed experimental unit was planted with corn, soybean, or kept under bareground.
+Each field experimental unit was planted with corn, soybean, or kept under bareground.
 The two transplant timings were June 1 (first cohort) and July 1 (second cohort).
@@
 After transplanting, Palmer amaranth flowering was monitored until the end of the study.
-When a plant flowered, the day was recorded, plant sex was identified (male or female), plant height was measured from soil surface to the top of plant.
+When a plant flowered, the day was recorded, plant gender was identified (male or female), plant height was measured from soil surface to the top of plant.
 Also, aboveground plant biomass was harvested near soil surface and oven dried at 65 C until reaching constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
@@
 
-A binary logistic regression was fitted to Palmer amaranth sex dimorphism.
+A binary logistic regression was fitted to Palmer amaranth gender.
 Binary logistic regression is used for predicting binary classes [@bangdiwala2018], such as the probability of a plant being female in a dioecious species.
@@
 
-## Palmer amaranth sex dimorphism {-}
+## Palmer amaranth gender {-}
 
@@
 table_a %>% 
-  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth sex dimorphism logistic model.
+  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth gender logistic model.
 Factor pararemter values (crop and bareground) is shown compared to soybean.") %>%
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering window and sex varied within crops and cohort timings.
+Our study showed that Palmer amaranth biomass, height, flowering window and gender varied within crops and cohort timings.
 In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort.
@@
 This is evident when comparing Palmer amaranth canopy shape.
-The Palmer amaranth competition (e.g., light) strategy was to mimic the crop grow and development (Figure 6).
-These results suggests that Palmer amaranth can fast evolve life-history traits to adapt into cropping systems and cultural practices, which was also showed in a study varying nitrogen fertilization [@bravo2018].
-Our results highlight the Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
+The Palmer amaranth competition (e.g., light) strategy was to mimic the crop growth and development (Figure 6).
+These results suggests that Palmer amaranth can quickly evolve life-history traits to adapt to cropping systems and cultural practices, which was also shown in a study varying nitrogen fertilization [@bravo2018].
+Our results highlight Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
 
@@
 
-Palmer amaranth grow and development in second cohort was limited due to the crop competitive ability at advanced development stages.
+Palmer amaranth growth and development in second cohort was limited due to the crop competitive ability at advanced development stages.
 Plants were transplanted at greater crop height and width, which reduced Palmer amaranth competitiveness.
@@
 The *Amaranthus* species are sensitive to photoperiod [@wu2014].
-It is hypothesize that reduced day length contributed to smaller plants at second cohort regardless the crop.
+It is hypothesize that reduced day length or red/far-red effect contributed to smaller plants at second cohort regardless the crop.
 A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021].
@@
 Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
-I our study, plants growing at first cohort accumulated an overall 42% more biomass when compared to second cohort.
+In our study, plants growing from the first cohort accumulated 42% more biomass when compared to the second cohort.
 Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the crop.
-Our observation is consistent with the findings that first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
-Still, seed production at second cohort is likely to replenish the soil seedbank.
+Our observation is consistent with the findings that the first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
+Still, seed production at the second cohort will replenish the soil seedbank.
 Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987].
-Palmer amaranth can produce hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
+Palmer amaranth can produce a hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
 Therefore, preventing Palmer amaranth seed production or/and seed migration to non-native habitats is an essential strategy to minimize the species impact in agroecosystems [@davis2015].
@@
 The shortest flowering window of second cohorts of Palmer amaranth growing in bareground highlighted the impact of photoperiod on flowering.
-When growing in soybean, Palmer amaranth flowering window was similar to bareground at first cohort but similar to corn at second cohort timing.
+When growing in soybean, Palmer amaranth flowering window was similar to bareground for the first cohort but similar to corn at second cohort timing.
 Plant flowering initiation is complex and depends on the complex interaction between genetic makeup and the environmental conditions [@lang1965].
@@
 A study has shown that Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
-Early flower initiation could also be a plant strategy when growing in stress conditions.
+Early flower initiation could also be a plant strategy when growing in stressful conditions.
 For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
-A mismatch in Palmer amaranth male and female flowering period can minimize plant outcross, reducing plant seed production and/or exchange of resistant alleles [@jhala2021].
-Sex dimorphism manipulation is considered a potential ecological pest control [@mcfarlane2018; @schliekelman2005].
+A mismatch in Palmer amaranth male and female flowering period can minimize plant outcrossing, reducing plant seed production and/or exchange of resistant alleles [@jhala2021].
+Sex dimorphism manipulation is considered a potential ecological pest control strategy [@mcfarlane2018; @schliekelman2005].
 
-The mechanisms of sex-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
-In our study, the sex dimorphism model performance was decent (AIC 0.64) considering the biology of plant flowering.
+The mechanisms of gender-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
+In our study, the gender model performance was decent (AUC 0.64) considering the biology of plant flowering.
 A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
@@
 stipulacea* reduced female-male ratio to expand into its non-native habitat [@nguyen2018].
-Also, biotic and/or abiotic stress can influence plant sex determination.
+Also, biotic and/or abiotic stress can influence plant gender determination.
 Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019].

### 2021-09-17 Manuscript/Manuscript.tex
as

@@
 Palmer amaranth (\emph{Amaranthus palmeri} S.
-Watson) is one of the most troublesome weed species in the United States.
+Watson) is one of the most troublesome agronomic weed species in the United States.
 Palmer amaranth is endemic to Southern United States but its range is expanding northward.
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and sex dimorphism from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
-Results demonstrated that first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from second cohort (established in July).
+A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
+Results demonstrated that the first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from the second cohort (established in July).
 The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
@@
 no-till), and intensive use of herbicides for weed management are the
-main human-mediated selection of Palmer amaranth into cropping systems
+main human-mediated selections of Palmer amaranth into cropping systems
 (Ward et al., 2013). On the other hand, Palmer amaranth is a prolific
 seed producer with a C4 photosynthetic apparatus (Wang et al., 1992).
-With a dioecy nature, Palmer amaranth male and female plants are
+With a dioecious nature, Palmer amaranth male and female plants are
 obligate outcrosser species, increasing the chances of exchanging
 adaptive traits among plants (Jhala et al., 2021; Oliveira et al.,
-2018). Also, Palmer amaranth small seeds (e.g, 1 mm) tend to thrive in
+2018). Also, Palmer amaranth's small seeds (e.g, 1 mm) tend to thrive in
 no-tillage systems (Price et al., 2011), and spread across locations
@@
 production occurred at 35/30 C day and night temperatures (Guo and
-Al-Khatib, 2003). In addition, Palmer amaranth establishment is
+Al-Khatib, 2003). In addition, Palmer amaranth emergence is
 human-mediated by tillage timings and preemergence-applied herbicides
@@
 2019). Therefore, Palmer amaranth has shown plasticity to evolve and
-fast adapt under the current agroecosystem conditions. Further scenarios
-show that global temperature warming can impact agriculture, and promote
-niches for Palmer amaranth invasion/adaptation into new environments.
-Currently, it is estimated that the greatest climatic risk of Palmer
-amaranth establishment are agronomic crops in Australia and Sub-Sahara
-Africa (Kistner and Hatfield, 2018). Temperature is a key factor
-limiting Palmer amaranth expansion to cooler geographies (Briscoe
-Runquist et al., 2019); however, under future climate change Palmer
-amaranth is likely to expand northward into Canada and Northern Europe
-(Kistner and Hatfield, 2018; Briscoe Runquist et al., 2019).
+fast adapt under the current agroecosystem conditions. Future scenarios
+suggest that global temperature warming can impact agriculture, and
+promote niches for Palmer amaranth invasion/adaptation into new
+environments. Agronomic crops in Australia and Sub-Sahara Africa are
+currently at greatest risk to Palmer invasion as climate warms (Kistner
+and Hatfield, 2018). Temperature is a key factor limiting Palmer
+amaranth expansion to cooler geographies (Briscoe Runquist et al.,
+2019); however, under future climate change Palmer amaranth is likely to
+expand northward into Canada and Northern Europe (Kistner and Hatfield,
+2018; Briscoe Runquist et al., 2019).
 
@@
 Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
-amaranth can cause high damage to soybean fields in Illinois (Davis et
-al., 2015), which is concerning as soybean along with corn make most of
-US Midwest agronomic hectares. In Iowa, a study showed that Palmer
+amaranth can cause yield loss to soybean fields in Illinois (Davis et
+al., 2015), which is concerning as soybean along with corn comprise most
+of US Midwest agronomic hectares. In Iowa, a study showed that Palmer
 amaranth is still not well adapted compared to waterhemp
-(\emph{Amaranthus tuberculatus}) (Baker, 2021). Waterhemp, an
-\emph{Amaranthus} species, is the most troublesome species in the US
-Midwest (Tranel et al., 2011). Nonetheless, invasion and successful
-eradication of Palmer amaranth is documented in Minnesota (Yu et al.,
-2021). Palmer amaranth infestations have not been detected in Canada;
-however, Palmer amaranth seeds were detected in sweet potato slips in
-the country (Page et al., 2021). Palmer amaranth is still not as well
-adapted and established to Northern as it is in the Southern North
-America. Therefore, Palmer amaranth range of expansion into new habitats
-can increase. It seems certain the need to manage new Palmer amaranth
-infestations in agronomic crops throughout northern United States in the
-near future. Strategies on Palmer amaranth management should encompass
-the agroecosystem level but not only attempts to eradicate the weed.
-Most tactics to manage Palmer amaranth are based on technology fixes
-(Scott, 2011), which are short-term (e.g., herbicide and/or tillage)
-rather than long-term weed management. Palmer amaranth management should
-be built on minimizing the species ability to adapt, grow and develop
-into agroecossystems.
+(\emph{Amaranthus tuberculatus}) (Baker, 2021), which is considered the
+most troublesome species in the US Midwest (Tranel et al., 2011).
+Invasion and successful eradication of Palmer amaranth is documented in
+Minnesota (Yu et al., 2021). Palmer amaranth infestations have not been
+detected in Canada; however, Palmer amaranth seeds were detected in
+sweet potato slips in the country (Page et al., 2021). Palmer amaranth
+is still not as well adapted and established to Northern as compared to
+Southern North America. Therefore, Palmer amaranth range of expansion
+into new habitats can increase. It seems certain the need to manage new
+Palmer amaranth infestations in agronomic crops throughout northern
+United States is eminent in the near future. Strategies on Palmer
+amaranth management should encompass the agroecosystem level but not
+focus only on attempts to eradicate the weed. Most tactics to manage
+Palmer amaranth are based on technological solutions (Scott, 2011),
+which are short-term (e.g., herbicide and/or tillage) rather than
+long-term weed management. Palmer amaranth management should be built
+around minimizing the species ability to adapt, grow and develop
+successfully in agroecossystems.
 
@@
 impact on the agroecosystem is recent. Studies investigating Palmer
-amaranth in those locations are limited due to the plant classification
-as noxious weed species in some northern states (Yu et al., 2021).
-Nonetheless, the continuous Palmer amaranth dispersal and potential
-establishment across the northern United States is concerning and
-warrants investigations on species morphology in such environments.
-Understanding Palmer amaranth biology and growing strategies under
-different agroecosystems can enhance our knowledge on species adaptation
-and management practices. It can also aid in designing proactive and
-ecological tactics to limit the species range expansion, reduce its
-negative impact, and developing resilient and sustainable farming
-systems (MacLaren et al., 2020). Therefore, the objective of this study
-was to investigate the flowering pattern, sex dimorphism, biomass
-production, and height of Palmer amaranth cohorts growing under corn,
-soybean and bareground across five locations in the United States
-Midwest.
+amaranth in northern states are limited due to the plant classification
+as noxious weed species (Yu et al., 2021). Nonetheless, the continuous
+Palmer amaranth dispersal and potential establishment across the
+northern United States is concerning and warrants investigations on
+species morphology in northern environments. Understanding Palmer
+amaranth biology and morphology under different agroecosystems can
+enhance our knowledge on species adaptation and management practices. It
+can also aid in designing proactive and ecological tactics to limit the
+species range expansion, reduce its negative impact on agroecossystems,
+and to develop resilient and sustainable farming systems (MacLaren et
+al., 2020). Therefore, the objective of this study was to investigate
+the flowering pattern, gender, biomass production, and height of Palmer
+amaranth cohorts growing under corn, soybean and bareground across five
+locations in the United States Midwest.
 
@@
 
- & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
-
- & Seeding rate & 88956 & 86487 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
- & Stage at 1 cohort & V2-3 &  &  &  & \\
-
- & Stage at 2 cohort & V6-7 &  &  &  & \\
-
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
-
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & \\
+
+ & Seeding rate & 88956 & 86487 & 79000 & 86485 & \\
+
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & \\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & \\
 
@@
 
- & Variety & DSR-1950 & AG21X8 &  &  & \\
-
- & Seeding rate & 296400 & 321237 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
- & Stage at 1 cohort & V1-2 &  &  &  & \\
-
- & Stage at 2 cohort & V5-6 &  &  &  & \\
-
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
+ & Variety & DSR-1950 & AG21X8 & AG28X7 & 2018: 3017R2X; 2019 P31A22X & \\
+
+ & Seeding rate & 296400 & 321237 & 220000 & 370650 & \\
+
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & \\
+
+ & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 &  & \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 &  & \\
+
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & \\
 
@@
 
- & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\
-
- & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\
-
- & pH & 6.6 & 6.5 &  &  & \\
-
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
+ & Type & Plano-silt-loam & Crete Silt Loam & 2018: Mace silt loam / 2019: Kuma silt loam & Crete silty clay loam & \\
+
+ & pH & 6.6 & 6.5 & 2018: 6.1 / 2019: 5.1 & 5.2 & \\
+
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 & 2018: 2 / 2019: 1.7 & 3.4 & \\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
 \end{tabular}}
@@
 arranged in factorial design with three crops, two transplanting times
-simulating two cohorts, repeated across five locations. Each filed
+simulating two cohorts, repeated across five locations. Each field
 experimental unit was planted with corn, soybean, or kept under
@@
 After transplanting, Palmer amaranth flowering was monitored until the
-end of the study. When a plant flowered, the day was recorded, plant sex
-was identified (male or female), plant height was measured from soil
-surface to the top of plant. Also, aboveground plant biomass was
+end of the study. When a plant flowered, the day was recorded, plant
+gender was identified (male or female), plant height was measured from
+soil surface to the top of plant. Also, aboveground plant biomass was
 harvested near soil surface and oven dried at 65 C until reaching
@@
 
-A binary logistic regression was fitted to Palmer amaranth sex
-dimorphism. Binary logistic regression is used for predicting binary
-classes (Bangdiwala, 2018), such as the probability of a plant being
-female in a dioecious species. Prior to the analysis, missing values
-were removed from the dataset. Also, data from Grant was not used in
-this analysis due to the uniform plant harvesting at that location. The
-complete dataset was splitted into 80\% train and 20\% test data. The
-80\% train is used for the model training and the 20\% test is used for
-checking model performance on unseen dataset. With 80\% dataset, a
-generalized linear model (base R \emph{glm} function) was fitted to
-binary response variable, the probability of being female (0 to male and
-1 to female). The independent variables were day of year harvest,
-height, weight, and crop (without interaction). The model family was
-binomial with a logit function. The model fit was assessed through
-pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
-likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
-package (Mangiafico, 2021). The marginal effects computation was
-performed with Average Marginal Effects (AMEs) at every observed value
-of x and average across the results (Leeper, 2017) using \emph{margins}
-function from ``margins'' package (Leeper et al., 2021). The 20\% test
-data was predicted using the \emph{predict} function with a cutoff
-estimation for male or female using \emph{performance} function from
-ROCR package (Sing et al., 2005). The model quality prediction from the
-classification algorithm was measured with precision (\emph{precision}
-function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
-function) using the ``yardstick'' package (Kuhn et al., 2021). The
-precision determines the accuracy of positive predictions (female
-plants), recall determines the fraction of positives that were correctly
-identified, and F1-score is a weighted harmonic mean of precision and
-recall with the best score of 1 and the worst score of 0 (Raoniar,
-2021). F1-score conveys the balance between the precision and the recall
-(Yacouby and Axman, 2020). The area under the receiver operating curve
-(AUC-ROC) was also estimated with performance function using the true
-positive and false positive rates. Higher the AUC, better the model is
-at distinguishing between female and male Palmer amaranth.
+A binary logistic regression was fitted to Palmer amaranth gender.
+Binary logistic regression is used for predicting binary classes
+(Bangdiwala, 2018), such as the probability of a plant being female in a
+dioecious species. Prior to the analysis, missing values were removed
+from the dataset. Also, data from Grant was not used in this analysis
+due to the uniform plant harvesting at that location. The complete
+dataset was splitted into 80\% train and 20\% test data. The 80\% train
+is used for the model training and the 20\% test is used for checking
+model performance on unseen dataset. With 80\% dataset, a generalized
+linear model (base R \emph{glm} function) was fitted to binary response
+variable, the probability of being female (0 to male and 1 to female).
+The independent variables were day of year harvest, height, weight, and
+crop (without interaction). The model family was binomial with a logit
+function. The model fit was assessed through pseudo R-squared values
+(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
+\emph{nagelkerke} function from ``rcompanion'' package (Mangiafico,
+2021). The marginal effects computation was performed with Average
+Marginal Effects (AMEs) at every observed value of x and average across
+the results (Leeper, 2017) using \emph{margins} function from
+``margins'' package (Leeper et al., 2021). The 20\% test data was
+predicted using the \emph{predict} function with a cutoff estimation for
+male or female using \emph{performance} function from ROCR package (Sing
+et al., 2005). The model quality prediction from the classification
+algorithm was measured with precision (\emph{precision} function),
+recall (\emph{recall} function) and F1-score (\emph{f\_meas} function)
+using the ``yardstick'' package (Kuhn et al., 2021). The precision
+determines the accuracy of positive predictions (female plants), recall
+determines the fraction of positives that were correctly identified, and
+F1-score is a weighted harmonic mean of precision and recall with the
+best score of 1 and the worst score of 0 (Raoniar, 2021). F1-score
+conveys the balance between the precision and the recall (Yacouby and
+Axman, 2020). The area under the receiver operating curve (AUC-ROC) was
+also estimated with performance function using the true positive and
+false positive rates. Higher the AUC, better the model is at
+distinguishing between female and male Palmer amaranth.
 
@@
 
-\hypertarget{palmer-amaranth-sex-dimorphism}{%
-\subsection*{Palmer amaranth sex
-dimorphism}\label{palmer-amaranth-sex-dimorphism}}
-\addcontentsline{toc}{subsection}{Palmer amaranth sex dimorphism}
+\hypertarget{palmer-amaranth-gender}{%
+\subsection*{Palmer amaranth gender}\label{palmer-amaranth-gender}}
+\addcontentsline{toc}{subsection}{Palmer amaranth gender}
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth gender logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
 \centering
@@
 Our study showed that Palmer amaranth biomass, height, flowering window
-and sex varied within crops and cohort timings. In general, first cohort
-of Palmer amaranth plants were heavier and taller when compared to the
-second cohort. At first cohort, resources (e.g., soil nutrients) and
-conditions (e.g., light) were more timely available for the species.
+and gender varied within crops and cohort timings. In general, first
+cohort of Palmer amaranth plants were heavier and taller when compared
+to the second cohort. At first cohort, resources (e.g., soil nutrients)
+and conditions (e.g., light) were more timely available for the species.
 High biomass and taller Palmer amaranth plants are likely a weed
@@
 amaranth canopy shape. The Palmer amaranth competition (e.g., light)
-strategy was to mimic the crop grow and development (Figure 6). These
-results suggests that Palmer amaranth can fast evolve life-history
-traits to adapt into cropping systems and cultural practices, which was
-also showed in a study varying nitrogen fertilization (Bravo et al.,
-2018). Our results highlight the Palmer amaranth as a threat to field
-crops as breeding more competitive crop varieties is likely to select
-more competitive weed biotypes (Bravo et al., 2017).
-
-Palmer amaranth grow and development in second cohort was limited due to
-the crop competitive ability at advanced development stages. Plants were
-transplanted at greater crop height and width, which reduced Palmer
+strategy was to mimic the crop growth and development (Figure 6). These
+results suggests that Palmer amaranth can quickly evolve life-history
+traits to adapt to cropping systems and cultural practices, which was
+also shown in a study varying nitrogen fertilization (Bravo et al.,
+2018). Our results highlight Palmer amaranth as a threat to field crops
+as breeding more competitive crop varieties is likely to select more
+competitive weed biotypes (Bravo et al., 2017).
+
+Palmer amaranth growth and development in second cohort was limited due
+to the crop competitive ability at advanced development stages. Plants
+were transplanted at greater crop height and width, which reduced Palmer
 amaranth competitiveness. As a result, Palmer amaranth height and
@@
 days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
-and Owen, 2014). It is hypothesize that reduced day length contributed
-to smaller plants at second cohort regardless the crop. A study in North
-Carolina and Illinois predicted that less than 10\% Palmer amaranth
-seedlings emergence occurred after June (Piskackova et al., 2021). In
-addition, Palmer amaranth negative impact on soybean (Korres et al.,
-2020) and cotton (Webster and Grey, 2015) yields was higher when plants
-were established near to crop planting. Therefore, early management is a
-key strategy to minimize the damaging impact of Palmer amaranth to US
-Midwest cropping systems.
+and Owen, 2014). It is hypothesize that reduced day length or
+red/far-red effect contributed to smaller plants at second cohort
+regardless the crop. A study in North Carolina and Illinois predicted
+that less than 10\% Palmer amaranth seedlings emergence occurred after
+June (Piskackova et al., 2021). In addition, Palmer amaranth negative
+impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
+2015) yields was higher when plants were established near to crop
+planting. Therefore, early management is a key strategy to minimize the
+damaging impact of Palmer amaranth to US Midwest cropping systems.
 
@@
 correlation between Palmer amaranth biomass and seed production
-(Schwartz et al., 2016; Spaunhorst et al., 2018). I our study, plants
-growing at first cohort accumulated an overall 42\% more biomass when
-compared to second cohort. Therefore, Palmer amaranth plants growing in
-the second cohort is likely to produce less seeds regardless the crop.
-Our observation is consistent with the findings that first Palmer
-amaranth cohort produced 50\% more seeds per plant than Palmer amaranth
-plants established six weeks later in bareground (Webster and Grey,
-2015). Still, seed production at second cohort is likely to replenish
+(Schwartz et al., 2016; Spaunhorst et al., 2018). In our study, plants
+growing from the first cohort accumulated 42\% more biomass when
+compared to the second cohort. Therefore, Palmer amaranth plants growing
+in the second cohort is likely to produce less seeds regardless the
+crop. Our observation is consistent with the findings that the first
+Palmer amaranth cohort produced 50\% more seeds per plant than Palmer
+amaranth plants established six weeks later in bareground (Webster and
+Grey, 2015). Still, seed production at the second cohort will replenish
 the soil seedbank. Seed production and deposition in the seedbank is
 also a key factor for species perpetuation (Menges, 1987). Palmer
-amaranth can produce hundred thousands seeds per plant (Schwartz et al.,
-2016; Keeley et al., 1987), which can stay viable in the soil seedbank
-for at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
-Palmer amaranth seed production or/and seed migration to non-native
-habitats is an essential strategy to minimize the species impact in
-agroecosystems (Davis et al., 2015).
+amaranth can produce a hundred thousands seeds per plant (Schwartz et
+al., 2016; Keeley et al., 1987), which can stay viable in the soil
+seedbank for at least 36 months (Sosnoskie et al., 2013). Therefore,
+preventing Palmer amaranth seed production or/and seed migration to
+non-native habitats is an essential strategy to minimize the species
+impact in agroecosystems (Davis et al., 2015).
 
@@
 flowering. When growing in soybean, Palmer amaranth flowering window was
-similar to bareground at first cohort but similar to corn at second
+similar to bareground for the first cohort but similar to corn at second
 cohort timing. Plant flowering initiation is complex and depends on the
@@
 weeks prior to the native waterhemp in Iowa (Baker, 2021). Early flower
-initiation could also be a plant strategy when growing in stress
+initiation could also be a plant strategy when growing in stressful
 conditions. For example, when growing under water stress, early
@@
 male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant outcross,
-reducing plant seed production and/or exchange of resistant alleles
-(Jhala et al., 2021). Sex dimorphism manipulation is considered a
-potential ecological pest control (McFarlane et al., 2018; Schliekelman
-et al., 2005).
-
-The mechanisms of sex-determination in plant species is intriguing and
-arouse the curiosity of many scientists, including Charles Darwin
-(Darwin, 1888). In our study, the sex dimorphism model performance was
-decent (AIC 0.64) considering the biology of plant flowering. A 1:1 male
-and female sex ratio is a general evolutionary stable strategy for plant
+amaranth male and female flowering period can minimize plant
+outcrossing, reducing plant seed production and/or exchange of resistant
+alleles (Jhala et al., 2021). Sex dimorphism manipulation is considered
+a potential ecological pest control strategy (McFarlane et al., 2018;
+Schliekelman et al., 2005).
+
+The mechanisms of gender-determination in plant species is intriguing
+and arouse the curiosity of many scientists, including Charles Darwin
+(Darwin, 1888). In our study, the gender model performance was decent
+(AUC 0.64) considering the biology of plant flowering. A 1:1 male and
+female sex ratio is a general evolutionary stable strategy for plant
 species perpetuation (Fisher, 1930). However, a slight deviation from
@@
 female-male ratio to expand into its non-native habitat (Nguyen et al.,
-2018). Also, biotic and/or abiotic stress can influence plant sex
+2018). Also, biotic and/or abiotic stress can influence plant gender
 determination. Palmer amaranth male-to-female ratio was greater under

### 2021-09-17 Manuscript/table.tex
as

@@
 
- & Weed control & glyphosate\textsuperscript{a} / \emph{S}-metolachor\textsuperscript{b} & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{c} &  &  & \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & \\
 
- & Seeding rate & 88956 & 86487 &  &  & \\
+ & Seeding rate & 88956 & 86487 & 79000 & 86485 & \\
 
- & Weed control & glyphosate / \emph{S}-metolachor & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{d} &  &  & \\
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
 
- & Stage at 1\textsuperscript{st} cohort & V2-3 &  &  &  & \\
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & \\
 
- & Stage at 2\textsuperscript{nd} cohort & V6-7 &  &  &  & \\
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & \\
 
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & \\
 
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  &  & \\
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & \\
 
@@
 
- & Variety & DSR-1950 & AG21X8 &  &  & \\
+ & Variety & DSR-1950 & AG21X8 & AG28X7 & 2018: 3017R2X; 2019 P31A22X & \\
 
- & Seeding rate & 296400 & 321237 &  &  & \\
+ & Seeding rate & 296400 & 321237 & 220000 & 370650 & \\
 
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & \\
 
- & Stage at 1\textsuperscript{st} cohort & V1-2 &  &  &  & \\
+ & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 &  & \\
 
- & Stage at 2\textsuperscript{nd} cohort & V5-6 &  &  &  & \\
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 &  & \\
 
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & \\
 
@@
 
- & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\
+ & Type & Plano-silt-loam & Crete Silt Loam & 2018: Mace silt loam / 2019: Kuma silt loam & Crete silty clay loam & \\
 
- & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\
+ & pH & 6.6 & 6.5 & 2018: 6.1 / 2019: 5.1 & 5.2 & \\
 
- & pH & 6.6 & 6.5 &  &  & \\
-
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 & 2018: 2 / 2019: 1.7 & 3.4 & \\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} \emph{S}--metolachor, 1324 g ai ha; \textsuperscript{c} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha; \textsuperscript{d} \emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha}\\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
 \end{tabular}}

### 2021-11-04 Manuscript/Manuscript.Rmd
jh

@@
 Corn and soybean were planted in 76-cm row spacing (Table 1).
-Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1)
+Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1).
 

### 2021-11-04 Manuscript/Manuscript.tex
jh

@@
 
- & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
+ & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & \\
-
- & Seeding rate & 88956 & 86487 & 79000 & 86485 & \\
-
- & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
-
- & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & \\
-
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & \\
-
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
+
+ & Seeding rate & 88956 & 86487 &  &  & \\
+
+ & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
+
+ & Stage at 1 cohort & V2-3 &  &  &  & \\
+
+ & Stage at 2 cohort & V6-7 &  &  &  & \\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
 
@@
 
- & Variety & DSR-1950 & AG21X8 & AG28X7 & 2018: 3017R2X; 2019 P31A22X & \\
-
- & Seeding rate & 296400 & 321237 & 220000 & 370650 & \\
-
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & \\
-
- & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 &  & \\
-
- & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 &  & \\
-
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & \\
+ & Variety & DSR-1950 & AG21X8 &  &  & \\
+
+ & Seeding rate & 296400 & 321237 &  &  & \\
+
+ & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+
+ & Stage at 1 cohort & V1-2 &  &  &  & \\
+
+ & Stage at 2 cohort & V5-6 &  &  &  & \\
+
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
 
@@
 
- & Type & Plano-silt-loam & Crete Silt Loam & 2018: Mace silt loam / 2019: Kuma silt loam & Crete silty clay loam & \\
-
- & pH & 6.6 & 6.5 & 2018: 6.1 / 2019: 5.1 & 5.2 & \\
-
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 & 2018: 2 / 2019: 1.7 & 3.4 & \\
+ & Type & Plano-silt-loam & Crete Silt Loam &  &  & \\
+
+ & Ratio (sand-clay-silt) & 10-64-26 & 17-58-25 &  &  & \\
+
+ & pH & 6.6 & 6.5 &  &  & \\
+
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 &  &  & \\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a} glyphoste, 840 g ae ha \textsuperscript{b} S-metolachor, 1324 g ai ha; \textsuperscript{c} S-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha; \textsuperscript{d} saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha}\\
 \end{tabular}}

### 2022-02-18 Manuscript/Manuscript.Rmd
as

@@
 author:
-  - name: Maxwel C Oliveira
+  - name: Maxwel C. Oliveira
     affiliation: '1'
     etal: Oliveira # First author's last name. 
-  - name: Amit J Jhala
+  - name: Amit J. Jhala
     affiliation: '2'
-  - name: Mark Bernards
+  - name: Mark L. Bernards
     affiliation: '3'
@@
   - id: '3' 
-    department: Department of Agronomy
+    department: School of Agriculture
     institution: Western Illinois University
@@
 Watson) is one of the most troublesome agronomic weed species in the United States.
-Palmer amaranth is endemic to Southern United States but its range is expanding northward.
+Palmer amaranth is prevalent in the Southern Great Plains and Southeastern United States, and its range is expanding northward through natural dispersal and human intervention.
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations of US Midwest.
+A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations in the US Midwest.
 Results demonstrated that the first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from the second cohort (established in July).
 The first Palmer amaranth cohort produced 75.5 g plant\textsuperscript{-1} in bareground, 28.3 g plant\textsuperscript{-1} in soybean and 16.3 g plant\textsuperscript{-1} in corn, whereas the second Palmer amaranth cohort produced 62.6, 6.3, and 1.4 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
-Palmer amaranth height was more impacted when growing in corn.
-Palmer amaranth plants averaged 85.2 cm tall in the first cohort but 38.2 cm tall in the second cohort in corn.
+Palmer amaranth height was most impacted when growing in corn, and averaged 85.2 cm tall in the first cohort, and 38.2 cm tall in the second cohort in corn.
 Moreover, Palmer amaranth flowering window shifted according to crop and cohort timings.
 Palmer amaranth growing in intense competition, such as under low light in corn, resulted in the longest flowering window.
-Also, Palmer amaranth gender was slightly influenced by day of year, weight and height.
-The model estimated that probability of being a female plant increased as biomass and height increased.
-Our results showed the fast adaptation and plasticity of Palmer amaranth to grow and adapt to cropping systems from the US Midwest.
-Palmer amaranth is likely to continue its expansion northward.
-Therefore, preventing plant dispersal into new habitats is the most effective management strategy.
-Reactive management to reduce Palmer amaranth impact on cropping systems should encompass diversity of tactics that minimize the species ability to establish into cropping systems, including crop rotation (beyond corn and soybean), early/late crop planting, row spacing, cover crops, and effective chemical control programs.
+Palmer amaranth gender was slightly influenced by day of year, weight and height.
+We documented a high degree of plasticity in Palmer amaranth, which will presumably favor its adaptation and expansion in cropping systems north of its current range.
+Therefore, preventing Palmer amaranth dispersal into new habitats is the most effective management strategy.
 
@@
 Moreover, Palmer amaranth has shown a remarkable capacity to evolve resistance to herbicides.
-To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017].
-Thus, Palmer amaranth poses an economical and ecological risk to sustainable agriculture.
+To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017] and posing an economical and ecological risk to row-crop agriculture.
 
+Palmer amaranth as a problem weed is a function of both inherent adaptations and selected management practices.
 Palmer amaranth is a fast growing summer annual forb indigenous to the Sonoran Desert [@sauer1957].
-The species would eventually emerge as a threat to US agriculture in the 1990s.
-Palmer amaranth weediness is likely a result of human-assisted selection in combination with plant biology.
-Farm mechanization, adoption of conservation agriculture (e.g., no-till), and intensive use of herbicides for weed management are the main human-mediated selections of Palmer amaranth into cropping systems [@ward2013].
-On the other hand, Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
-With a dioecious nature, Palmer amaranth male and female plants are obligate outcrosser species, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
-Also, Palmer amaranth's small seeds (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], seed mixes [@hartzler2016], wildlife [@farmer2017], etc.
-The dispersal capacity of Palmer amaranth makes the species one of the most successful cases of weed adaptation to current cropping systems.
+It became a serious problem weed in US agriculture in the 1990s [@ward2013].
+Palmer amaranth weediness is likely a result of human-assisted selection combined with plant biology.
+Farm mechanization, adoption of conservation agriculture (e.g., no-till), and intensive use of herbicides for weed management are the main human-mediated selections of Palmer amaranth in cropping systems [@ward2013].
+Palmer amaranth is a prolific seed producer with a C4 photosynthetic apparatus [@wang1992].
+With dioecious nature, Palmer amaranth male and female plants are obligate outcrossers, increasing the chances of exchanging adaptive traits among plants [@jhala2021; @oliveira2018].
+Also, Palmer amaranth's small seeds (e.g, 1 mm) tend to thrive in no-tillage systems [@price2011], and spread across locations through farm equipment [@sauer1972], seed mixes [@hartzler2016], wildlife [@farmer2017], etc., making it one of the most successful examples of weed adaptation to current cropping systems.
 
-Light and temperature are the main environmental requirements for Palmer amaranth growth and development [@jha2010].
-Palmer amaranth is reported with an extended germination period [@ward2013].
+Palmer amaranth's plasticity allows it to respond successfully to environmental changes.
+Palmer amaranth demonstrates a high degree of plasticity to light, temperature, water availability, and human management [@jha2010].
+Palmer amaranth has an extended germination period of up to _____ [@ward2013].
 Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003].
-In addition, Palmer amaranth emergence is human-mediated by tillage timings and preemergence-applied herbicides [@chahal2021], which can result in weed germination shifts [@sbatella2010].
-Also, Palmer amaranth has shown to grow and develop under reduced soil water content.
-In continuous water stress, Palmer amaranth survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
+Palmer amaranth emergence can be influenced by tillage and the use of preemergence herbicides [@chahal2021], and may result in weed germination shifts within a population, as documented in *Bassia scoparia* [@sbatella2010].
+In an experiment where Palmer amaranth was subjected to continuous water stress, it survived and produced at least 14000 seeds plant\textsuperscript{-1} [@chahal2018].
 Seeds from Palmer amaranth growing with limited water conditions were heavier, less dormant, and prompt for germination [@matzrafi2021].
 Growing conditions and management practices also influence Palmer amaranth sex dimorphism and flowering pattern [@korres2017; @rumpa2019].
-Therefore, Palmer amaranth has shown plasticity to evolve and fast adapt under the current agroecosystem conditions.
-Future scenarios suggest that global temperature warming can impact agriculture, and promote niches for Palmer amaranth invasion/adaptation into new environments.
-Agronomic crops in Australia and Sub-Sahara Africa are currently at greatest risk to Palmer invasion as climate warms [@kistner2018].
-Temperature is a key factor limiting Palmer amaranth expansion to cooler geographies [@briscoerunquist2019]; however, under future climate change Palmer amaranth is likely to expand northward into Canada and Northern Europe [@kistner2018; @briscoerunquist2019].
 
-Palmer amaranth is already found in agronomic crops of South America [@larran2017; @kupper2017] and Southern Europe [@milani2021].
-In the United States, Palmer amaranth is well established in the Cotton Belt [@garetson2019; @bagavathiannan2016] in the southern United States but its range is expanding northward.
-For example, herbicide resistant Palmer amaranth is widespread in Nebraska  [@oliveira2021a].
-There are some reported cases of Palmer amaranth in Michigan [@kohrt2017] and Connecticut [@aulakh2021].
-Also, it is estimated that Palmer amaranth can cause yield loss to soybean fields in Illinois [@davis2015], which is concerning as soybean along with corn comprise most of US Midwest agronomic hectares.
-In Iowa, a study showed that Palmer amaranth is still not well adapted compared to waterhemp (*Amaranthus tuberculatus*) [@baker2021], which is considered the most troublesome species in the US Midwest [@tranel2011].
-Invasion and successful eradication of Palmer amaranth is documented in Minnesota [@yu2021].
-Palmer amaranth infestations have not been detected in Canada; however, Palmer amaranth seeds were detected in sweet potato slips in the country [@page2021].
-Palmer amaranth is still not as well adapted and established to Northern as compared to Southern North America.
-Therefore, Palmer amaranth range of expansion into new habitats can increase.
-It seems certain the need to manage new Palmer amaranth infestations in agronomic crops throughout northern United States is eminent in the near future.
-Strategies on Palmer amaranth management should encompass the agroecosystem level but not focus only on attempts to eradicate the weed.
-Most tactics to manage Palmer amaranth are based on technological solutions [@scott2011], which are short-term (e.g., herbicide and/or tillage) rather than long-term weed management.
-Palmer amaranth management should be built around minimizing the species ability to adapt, grow and develop successfully in agroecossystems.
+Palmer amaranth's current global range includes agronomic cropland in Italy [@milani2021], soybean producing regions of Brazil and Argentina [@larran2017; @kupper2017], and the Cotton Belt [@garetson2019; @bagavathiannan2016] and southern Great Plains of the United States (@kumar2020, @crespo2016a].
+However, its range appears to be expanding steadily, and in some cases rapidly, in the U.S.
+For example, in states with large Palmer amaranth infestations its range is gradually pushing north as it displaces other weeds in agronomic row crops (R Werle, personal observation), aided by wildlife [@farmer2017] and agricultural machinery.
+But there are other examples where it has suddenly become a problem weed hundreds of miles away from known infestations.
+New infestations in Michigan, Indiana and North Dakota appear to have resulted from Palmer amaranth contaminated livestock feed.
+New infestations in Iowa and Minnesota were associated with planting contaminated pollinator habitat seed [@yu2021].
+Palmer amaranth infestations have not been detected in Canada, but Palmer amaranth seeds were found in sweet potato slips imported into the country [@page2021].
+Global warming will create new opportunities for Palmer amaranth invasion.
+Although agronomic crops in warm environments like Australia and Sub-Saharan Africa are currently at greatest risk for Palmer amaranth invasion, warming temperatures will reduce barriers that may have limited Palmer amaranths spread into cooler climates like Canada and Northern Europe [@kistner2018; @briscoerunquist2019].
 
-In the southeastern US, early growing Palmer amaranth is well known to have a higher impact on cotton yields compared to late established plants [@macrae2013].
-In the northern states, Palmer amaranth impact on the agroecosystem is recent.
-Studies investigating Palmer amaranth in northern states are limited due to the plant classification as noxious weed species [@yu2021].
-Nonetheless, the continuous Palmer amaranth dispersal and potential establishment across the northern United States is concerning and warrants investigations on species morphology in northern environments.
-Understanding Palmer amaranth biology and morphology under different agroecosystems can enhance our knowledge on species adaptation and management practices.
-It can also aid in designing proactive and ecological tactics to limit the species range expansion, reduce its negative impact on agroecossystems, and to develop resilient and sustainable farming systems [@maclaren2020].
-Therefore, the objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing under corn, soybean and bareground across five locations in the United States Midwest.
+There are many areas in the U.S.
+Corn Belt (North Central States) where Palmer amaranth is not yet established, and its potential adaptability is untested.
+Nonetheless, the rapid expansion of Palmer amaranth across the Northern U.S.
+is concerning and warrants investigations on its adaptability.
+We do know that  Palmer amaranth caused yield loss in Illinois soybean fields [@davis2015], but an Iowa study showed that Palmer amaranth was not as well adapted as waterhemp (*Amaranthus tuberculatus*) to conditions in that geography [@baker2021].
+We also know that Palmer amaranth plants that establish shortly after row-crop planting have a much greater impact on crop yield than plants that emerge after the crop has produced several leaves [@macrae2013].
+Understanding Palmer amaranth morphology and development under different agroecosystems and across a wide geography can enhance our knowledge of its adaptability, and may also aid in designing effective tactics to limit is range expansion and minimize its negative effects on row crops.
+The objective of this study was to investigate the flowering pattern, gender, biomass production, and height of Palmer amaranth cohorts growing in corn, soybean and bareground environments across five locations in the North Central United States.
 
@@
 
-A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no glyphosate resistance according to @oliveira2021a was selected for this study.
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no glyphosate resistance was selected for this study [@oliveira2021a].
 Three weeks prior to the establishment of each cohort, seeds were planted in plastic trays containing potting-mix.
@@
 The experimental design were arranged in factorial design with three crops, two transplanting times simulating two cohorts, repeated across five locations.
-Each field experimental unit was planted with corn, soybean, or kept under bareground.
+Each field experimental unit was planted with corn, soybean, or maintained as bareground.
 The two transplant timings were June 1 (first cohort) and July 1 (second cohort).
@@
 When a plant flowered, the day was recorded, plant gender was identified (male or female), plant height was measured from soil surface to the top of plant.
-Also, aboveground plant biomass was harvested near soil surface and oven dried at 65 C until reaching constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
-
-Plants had to be harvested at flowering because Palmer amaranth is neither endemic in Wisconsin nor in Illinois.
+Also, aboveground plant, and all aboveground plant organs were harvested, then oven dried at 65 C until a constant weight was reached, and dry biomass (g plant\textsuperscript{-1}) was recorded.
+Plants were harvested at flowering because Palmer amaranth infestations do not occur at the Wisconsin or Illinois research locations.
 In our study, all locations followed the methodology of plant harvest at flowering initiation, except in Grant, NE.
-In this location, all Palmer amaranth plants were harvest at once on July 06, 2018 and 2019 (first cohort), and on August 17, 2018 and on July 31, 2019 (second cohort).
+At the Grant, NE, location, all plants from the first cohort were harvested on July 6, 2018 or July 6, 2019, and all plants from the second cohort were harvested on August 17, 2018 or July 31, 2019, regardless of flowering status.
 
@@
 Plant height and biomass were log transformed to meet model assumption of normality.
-In the model, crop (bareground, corn, soybean) and cohort time (first and second) were the fixed effects and year nested with location the random effects.
+In the model, crop (bareground, corn, soybean) and cohort time (first and second) were the fixed effects and year nested with location were the random effects.
 Analysis of variance at $\alpha$ 0.05 was performed with *anova* function from "car" package [@fox2018].
@@
 
-The Palmer amaranth flowering timing was estimated as cumulative flowering across all locations, except Grant, NE.
-Palmer amaranth cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015].
+Palmer amaranth cumulative flowering estimation was determined across all locations, except Grant, NE.
+Cumulative flowering estimation was determined using an asymmetrical three parameter log logistic Weibull model of the drc package [@ritz2015]:
 
@@
 
-In this model, *Y* is the Palmer amaranth cumulative flowering, *d* is the upper limit (set to 100), and *e* is the inflection point, and *x* day of year (doy).
+In this model, *Y* is the cumulative flowering, *d* is the upper limit (set to 100), and *e* is the inflection point, and *x* is the day of year (doy).
 
-The doy for 10, 50, and 90% Palmer amaranth cumulative flowering were determined using the *ED* function of drc package.
+The doy for 10, 50, and 90% cumulative flowering were determined using the *ED* function of drc package.
 Also, the 10, 50, and 90% Palmer amaranth cumulative flowering were compared among crops and cohorts using the *EDcomp* function of drc package.
@@
 Binary logistic regression is used for predicting binary classes [@bangdiwala2018], such as the probability of a plant being female in a dioecious species.
-Prior to the analysis, missing values were removed from the dataset.
-Also, data from Grant was not used in this analysis due to the uniform plant harvesting at that location.
-The complete dataset was splitted into 80% train and 20% test data.
-The 80% train is used for the model training and the 20% test is used for checking model performance on unseen dataset.
-With 80% dataset, a generalized linear model (base R *glm* function) was fitted to binary response variable, the probability of being female (0 to male and 1 to female).
-The independent variables were day of year harvest, height, weight, and crop (without interaction).
+Prior to the analysis, missing values were removed from the dataset (including all data from the Grant location).
+The resulting dataset was split into 80% train and 20% test data.
+The 80% train is used for model training and the 20% test is used for checking model performance on unseen dataset.
+Using the 80% train data, a generalized linear model (base R *glm* function) was fitted to binary response variable, the probability of being female (0 to male and 1 to female).
+The independent variables were harvest day of year, height, weight, and crop (without interaction).
 The model family was binomial with a logit function.
 The model fit was assessed through pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using *nagelkerke* function from "rcompanion" package [@mangiafico2021].
-The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and average across the results [@leeper2017] using *margins* function from "margins" package [@leeper2021].
+The marginal effects computation was performed with Average Marginal Effects (AMEs) at every observed value of x and averaged across the results [@leeper2017] using *margins* function from "margins" package [@leeper2021].
 The 20% test data was predicted using the *predict* function with a cutoff estimation for male or female using *performance* function from ROCR package [@sing2005].
@@
 The area under the receiver operating curve (AUC-ROC) was also estimated with performance function using the true positive and false positive rates.
-Higher the AUC, better the model is at distinguishing between female and male Palmer amaranth.
+The higher the AUC, better the model is at distinguishing between female and male Palmer amaranth.
 
@@
 Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
-Plants from the second cohort produced 62.6 g plant\textsuperscript{-1} in bareground, followed by 6.3 g plant\textsuperscript{-1} in soybean, and 1.4 g plant\textsuperscript{-1}.
+Plants from the second cohort produced 62.6 g plant\textsuperscript{-1} in bareground, followed by 6.3 g plant\textsuperscript{-1} in soybean, and 1.4 g plant\textsuperscript{-1} in corn.
 
@@
 
-Palmer amaranth height was more uniform across cohort timings, except when growing in corn (figure 3B).
-Palmer amaranth plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74).
+Palmer amaranth height was less affected by cohort timing than was weight, with the exception of plants growing in corn (Figure 3B).
+Plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74).
 In addition, no difference in Palmer amaranth height (69.3 cm) was detected from first cohort plants in soybean to first and second cohort plants in bareground (P > 0.75).
-Palmer amaranth plants from the second cohort were nearly 10 cm lower compared to the first cohort in soybeans (P = 0.04).
-The tallest (first cohort) and smallest (second cohort) Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
+Palmer amaranth plants from the second cohort were nearly 10 cm shorter compared to the first cohort in soybeans (P = 0.04).
+The tallest (85.2 cm, first cohort) and shortest (38.2 cm, second cohort) plants were observed in corn.
 
@@
 
-Palmer amaranth plants from the first cohort growing in corn resulted in a longer flowering window compared to plants growing in bareground and soybean (Figure 4A).
-The 10% cumulative Palmer amaranth flowering in soybean, bareground and corn occurred at the end of June.
-Palmer amaranth reached 10% flowering in soybean, bareground and corn at doy 180, 180.9 and 181.7, respectively.
-The 50% Palmer amaranth cumulative flowering occurred in July.
-Palmer amaranth reached 50% flowering in bareground, soybean and corn at doy 193.4, 194.8, and 206.6, respectively.
-Similar trend was observed at 90% Palmer amaranth cumulative flowering.
-Palmer amaranth reached 90% flowering at doy 252.6 in corn (early September), which was 38 and 32 days after reaching 90% flowering in bareground and soybean, respectively.
+The initiation and duration of Palmer amaranth flowering was strongly influenced by cohort and surrounding vegetation (Figure 4A, 4B).
+In the first cohort, floral initiation (10% flowering) occurred near the end of June for all three treatments, at doy 180, 180.9, and 181.7 for soybean, bareground, and corn, respectively.
+In the second cohort, floral initiation occurred earlier on plants growing in the bareground than plants growing in soybean or corn (doy 203.8 versus doy 210.9 or 216.8, respectively).
+Palmer amaranth growing in the bareground had the shortest duration of floral initiation (measured as the difference between 10% and 90% cumulative flowering) in both the first (34 days) and second (28 days) cohorts.
+Palmer amaranth growing in soybean had a shorter duration of flowering than corn in the first cohort (40 days vs 71 days), but a longer duration in the second cohort (50 vs 44 days).
 
@@
 The likelihood ratio test showed a p-value of < 0.00.
-The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% less females plants (Table 2).
-Moreover, increasing a unit doy increases the probability of having a female plant by 0.4% (Table 2 and Figure4A).
-Similar trend is observed for weight (Figure 4B) as well as height (Figure 4C), whereas the probability of being female increase by 0.2% and 0.1% when a unit of weight (g) and height (cm) increases, respectively.
+The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% fewer females plants (Table 2).
+Moreover, increasing a cumulative flowering unit doy increases the probability of having a female plant by 0.4% (Table 2 and Figure5A).
+A similar trend was observed for weight (Figure 5B) and height (Figure 5C), where the probability of being female increased 0.1% and 0.2% for each unit increase of weight (g) and height (cm), respectively (Table 2).
 
@@
   kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth gender logistic model.
-Factor pararemter values (crop and bareground) is shown compared to soybean.") %>%
+Factor pararemter values (crop_corn and 'crop_bareground) are calculated relative to soybean.") %>%
   kable_styling(font_size = 10, 
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering window and gender varied within crops and cohort timings.
-In general, first cohort of Palmer amaranth plants were heavier and taller when compared to the second cohort.
-At first cohort, resources (e.g., soil nutrients) and conditions (e.g., light) were more timely available for the species.
-High biomass and taller Palmer amaranth plants are likely a weed strategy to compete for light in between crop rows in absence of canopy.
-In such conditions, Palmer amaranth showed an extraordinary plasticity to adapt upon the agroecosystem.
-This is evident when comparing Palmer amaranth canopy shape.
-The Palmer amaranth competition (e.g., light) strategy was to mimic the crop growth and development (Figure 6).
-These results suggests that Palmer amaranth can quickly evolve life-history traits to adapt to cropping systems and cultural practices, which was also shown in a study varying nitrogen fertilization [@bravo2018].
+Our study confirmed that Palmer amaranth is well adapted to growing throughout the Midwestern United States, and is simply limited in its range by seed dispersal (Davis et al.
+2015).
+We also confirmed Palmer amaranth's extraordinary plasticity to adapt to different agroecosystems.
+For example, Palmer amaranth mimicked crop architecture in competing for light (Figure 6), where plants growing in corn allocated resources to height, while plants growing in bareground allocated resources to numerous branches.
+Plants growing in the absence of a crop produced the greatest biomass, responding to a greater abundance of light, nutrient and water resources (Figure 3A, 3B).
+These results support the argument that Palmer amaranth can quickly evolve life-history traits to adapt to different cultural practices, similar to observations in a Palmer amaranth response to nitrogen study [@bravo2018].
 Our results highlight Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
@@
 
+Floral initiation depends on the complex interaction between a plant's genetic makeup and environmental conditions [@lang1965].
+We observed a significant disruption in flowering caused by surrounding vegetation and cohort timing.
+Although all three treatments began flowering at the same time in the first cohort (about doy 180), it took much less time for the bareground treatment to reach 90% flowering (34 days versus 40 for soybean and 71 for corn).
+The lack of competition (and stress) in the bareground resulted in plants that flowered earlier, and were they not harvested, would have had longer reproductive periods, thereby producing more seed.
+Similarly, in the second cohort, the window for all plants to initiate flowering was much shorter for the bareground (28 days) compared to Palmer amaranth in soybean (50 days) or corn (44 days).
+Significantly, plants in the bareground treatment flowered one week earlier than soybean, and almost two weeks earlier than corn, again allowing for greater seed production per plant had they not been harvested at initial flowering.
+In some cases, early flowering may confer an evolutionary advantage, provided the plant has an indeterminate habit and flowering does not restrict plant growth.
+In an Iowa study, Palmer amaranth initiated flowering two weeks prior to waterhemp [@baker2021], and this characteristic may partially explain why Palmer amaranth can displace waterhemp where both occur in the same agronomic field.
+In other cases, however, early flower initiation may be a response to stressful conditions as a plant attempts to reproduce before running out of resources.
+In a dioecious species like Palmer amaranth, exerting stress on plants to manipulate flowering may be beneficial in limiting seed production [@mcfarlane2018; @schliekelman2005].
+For example, when growing under water stress, there was a seven days flowering mismatch between male and female plants [@mesgaran2021], which can minimize plant outcrossing, reduce seed production and the exchange of resistant alleles [@jhala2021].
+In our study we observed
+
 An ecological approach to reduce Palmer amaranth seed production is understanding plant biology, including flowering pattern.
@@
 
-The mechanisms of gender-determination in plant species is intriguing and arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
+The mechanisms of gender-determination in plant species is intriguing and has arouse the curiosity of many scientists, including Charles Darwin [@darwin1888].
 In our study, the gender model performance was decent (AUC 0.64) considering the biology of plant flowering.
-A 1:1 male and female sex ratio is a general evolutionary stable strategy for plant species perpetuation [@fisher1930].
-However, a slight deviation from 1:1 sex ratio might occur in some dioecious species.
+A 1:1 male and female sex ratio is a general an evolutionarily stable strategy for plant species perpetuation [@fisher1930].
+However, a slight deviation from 1:1 sex ratio occurs in some dioecious species.
 For example, the dioecious *Halophila stipulacea* is a female-biased plant in its native habitat, but the naturalized *H.
-stipulacea* have a 1:1 ratio [@nguyen2018].
-Naturalized of *H.
+stipulacea* has a 1:1 ratio [@nguyen2018].
+Naturalization of *H.
 stipulacea* reduced female-male ratio to expand into its non-native habitat [@nguyen2018].
@@
 Palmer amaranth  male-to-female ratio was greater under high plant densities [@korres2017a] and after herbicide application [@rumpa2019].
-Our model estimated that late flowering, heavier and taller Palmer amaranth plants slighly deviated from 1:1 ratio in favor to female plants.
-It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
-Our model also estimated more female plants in soybean and bareground compared to corn, which might be linked to plant competition strategy in each crop.
-Our results showed the influence of life-history and ecological traits on sexual dimorphism in Palmer amaranth.
+We observed sexual dimorphism in Palmer amaranth in response to surrounding vegetation and plant morphological attributes.
+Our model estimated that late flowering, heavier and taller Palmer amaranth plants slightly deviated from 1:1 ratio in favor of female plants.
+It was reported that female Palmer amaranth plants invested more in height, stem and biomass while male plants invested more in leaf area and leaf dry weight under nutrient deficiency [@korres2017].
+We observed more female plants in soybean and bareground compared to corn.
+Palmer amaranth plants in the corn were more stressed by interspecific competition as evidenced by less biomass and a lower weight:height ratio.
 Sexual dimorphism is documented in other dioecious species [@barrett2013].
 For example, stronger female plant competition and greater male tolerance to herbivory was reported in *Spinacia oleracea* [@perez-llorca2019].
-Research on candidate genes for sex determination in *Amaranthus* species are currently underway but it is far from complete [@montgomery2021; @montgomery2019].
-Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including the plant behavior under climate change.
+Research on candidate genes for sex determination in *Amaranthus* species are currently underway but is far from complete [@montgomery2021; @montgomery2019].
+Further studies are also needed to understand the ecological basis of Palmer amaranth flowering, including plant behavior under climate change.
 
-Our study demonstrated the Palmer amaranth plasticity to grow and develop into arable land of US Midwest.
-It is likely that Palmer amaranth range will continue to expanding into new geographies.
-The migration of Palmer amaranth into the US Midwest will reshape the landscape as waterhemp and Palmer amaranth will share the same habitat.
-The presence of Palmer amaranth and waterhemp will increase the weed management complexity.
-Therefore, preventive management is a priority to minimizing Palmer amaranth dispersal.
-Reactive management should focus on early-season management programs, which would have a large negative effect on Palmer amaranth growth and development.
-Long-term tactics that promote early-season crop advantage against Palmer amaranth, including diversity of crops in rotation, early/late crop planting, plant width, and crop residue (e.g.
-cover crops) would minimize the negative impact of Palmer amaranth to cropping systems.
-The aggressiveness and differential Palmer amaranth adaptation to agroecosystem is striking and require regional collaboration to minimize the species impact on sustainability and profitability of cropping systems.
+Our study demonstrated that Palmer amaranth is adapted to grow on arable land throughout the Midwestern U.S.
+Palmer amaranth's range will continue to expand if current cropping practices are continued.
+As waterhemp and Palmer amaranth begin to share the same habitat, it will increase weed management complexity.
+Preventing Palmer amaranth seed dispersal must be a priority.
+Regional collaboration is necessary to slow the spread of this aggressive and adaptable weed.
+Where Palmer amaranth occurs, management tactics should focus on limiting Palmer amaranth establishment until row crops can shade late-emerging plants.
+Increasing the diversity of crops in rotation, varying row crop planting date, narrowing row width, and increasing residue cover through the use of cover crops are all tactics that can minimize the growth and seed production of Palmer amaranth, and improve the sustainability of Midwestern cropping systems.
 

### 2022-02-18 Manuscript/Manuscript.tex
as

@@
 \def\Authors{
-  Maxwel C Oliveira\,\textsuperscript{1},
-  Amit J Jhala\,\textsuperscript{2},
-  Mark Bernards\,\textsuperscript{3},
+  Maxwel C. Oliveira\,\textsuperscript{1},
+  Amit J. Jhala\,\textsuperscript{2},
+  Mark L. Bernards\,\textsuperscript{3},
   Chris Proctor\,\textsuperscript{2},
@@
 
-  \textsuperscript{3} Department of Agronomy, Western Illinois
+  \textsuperscript{3} School of Agriculture, Western Illinois
 University,  Macomb,  Illinois,  United States
@@
 herbicide sites of action (Heap, 2021), increasing the weed management
-complexity (Lindsay et al., 2017). Thus, Palmer amaranth poses an
-economical and ecological risk to sustainable agriculture.
-
-Palmer amaranth is a fast growing summer annual forb indigenous to the
-Sonoran Desert (Sauer, 1957). The species would eventually emerge as a
-threat to US agriculture in the 1990s. Palmer amaranth weediness is
-likely a result of human-assisted selection in combination with plant
-biology. Farm mechanization, adoption of conservation agriculture (e.g.,
-no-till), and intensive use of herbicides for weed management are the
-main human-mediated selections of Palmer amaranth into cropping systems
-(Ward et al., 2013). On the other hand, Palmer amaranth is a prolific
-seed producer with a C4 photosynthetic apparatus (Wang et al., 1992).
-With a dioecious nature, Palmer amaranth male and female plants are
-obligate outcrosser species, increasing the chances of exchanging
-adaptive traits among plants (Jhala et al., 2021; Oliveira et al.,
-2018). Also, Palmer amaranth's small seeds (e.g, 1 mm) tend to thrive in
-no-tillage systems (Price et al., 2011), and spread across locations
-through farm equipment (Sauer, 1972), seed mixes (Hartzler and Anderson,
-2016), wildlife (Farmer et al., 2017), etc. The dispersal capacity of
-Palmer amaranth makes the species one of the most successful cases of
-weed adaptation to current cropping systems.
-
-Light and temperature are the main environmental requirements for Palmer
-amaranth growth and development (Jha et al., 2010). Palmer amaranth is
-reported with an extended germination period (Ward et al., 2013).
-Germination of Palmer amaranth was triggered by 18 C soil temperature at
-5 cm depth (Keeley et al., 1987), and optimal germination and biomass
-production occurred at 35/30 C day and night temperatures (Guo and
-Al-Khatib, 2003). In addition, Palmer amaranth emergence is
-human-mediated by tillage timings and preemergence-applied herbicides
-(Chahal et al., 2021), which can result in weed germination shifts
-(Sbatella and Wilson, 2010). Also, Palmer amaranth has shown to grow and
-develop under reduced soil water content. In continuous water stress,
-Palmer amaranth survived and produced at least 14000 seeds
-plant\textsuperscript{-1} (Chahal et al., 2018). Seeds from Palmer
-amaranth growing with limited water conditions were heavier, less
-dormant, and prompt for germination (Matzrafi et al., 2021). Growing
-conditions and management practices also influence Palmer amaranth sex
-dimorphism and flowering pattern (Korres et al., 2017; Rumpa et al.,
-2019). Therefore, Palmer amaranth has shown plasticity to evolve and
-fast adapt under the current agroecosystem conditions. Future scenarios
-suggest that global temperature warming can impact agriculture, and
-promote niches for Palmer amaranth invasion/adaptation into new
-environments. Agronomic crops in Australia and Sub-Sahara Africa are
-currently at greatest risk to Palmer invasion as climate warms (Kistner
-and Hatfield, 2018). Temperature is a key factor limiting Palmer
-amaranth expansion to cooler geographies (Briscoe Runquist et al.,
-2019); however, under future climate change Palmer amaranth is likely to
-expand northward into Canada and Northern Europe (Kistner and Hatfield,
-2018; Briscoe Runquist et al., 2019).
-
-Palmer amaranth is already found in agronomic crops of South America
-(Larran et al., 2017; Küpper et al., 2017) and Southern Europe (Milani
-et al., 2021). In the United States, Palmer amaranth is well established
-in the Cotton Belt (Garetson et al., 2019; Bagavathiannan and
-Norsworthy, 2016) in the southern United States but its range is
-expanding northward. For example, herbicide resistant Palmer amaranth is
-widespread in Nebraska (Oliveira et al., 2021). There are some reported
-cases of Palmer amaranth in Michigan (Kohrt et al., 2017) and
-Connecticut (Aulakh et al., 2021). Also, it is estimated that Palmer
-amaranth can cause yield loss to soybean fields in Illinois (Davis et
-al., 2015), which is concerning as soybean along with corn comprise most
-of US Midwest agronomic hectares. In Iowa, a study showed that Palmer
-amaranth is still not well adapted compared to waterhemp
-(\emph{Amaranthus tuberculatus}) (Baker, 2021), which is considered the
-most troublesome species in the US Midwest (Tranel et al., 2011).
-Invasion and successful eradication of Palmer amaranth is documented in
-Minnesota (Yu et al., 2021). Palmer amaranth infestations have not been
-detected in Canada; however, Palmer amaranth seeds were detected in
-sweet potato slips in the country (Page et al., 2021). Palmer amaranth
-is still not as well adapted and established to Northern as compared to
-Southern North America. Therefore, Palmer amaranth range of expansion
-into new habitats can increase. It seems certain the need to manage new
-Palmer amaranth infestations in agronomic crops throughout northern
-United States is eminent in the near future. Strategies on Palmer
-amaranth management should encompass the agroecosystem level but not
-focus only on attempts to eradicate the weed. Most tactics to manage
-Palmer amaranth are based on technological solutions (Scott, 2011),
-which are short-term (e.g., herbicide and/or tillage) rather than
-long-term weed management. Palmer amaranth management should be built
-around minimizing the species ability to adapt, grow and develop
-successfully in agroecossystems.
-
-In the southeastern US, early growing Palmer amaranth is well known to
-have a higher impact on cotton yields compared to late established
-plants (MacRae et al., 2013). In the northern states, Palmer amaranth
-impact on the agroecosystem is recent. Studies investigating Palmer
-amaranth in northern states are limited due to the plant classification
-as noxious weed species (Yu et al., 2021). Nonetheless, the continuous
-Palmer amaranth dispersal and potential establishment across the
-northern United States is concerning and warrants investigations on
-species morphology in northern environments. Understanding Palmer
-amaranth biology and morphology under different agroecosystems can
-enhance our knowledge on species adaptation and management practices. It
-can also aid in designing proactive and ecological tactics to limit the
-species range expansion, reduce its negative impact on agroecossystems,
-and to develop resilient and sustainable farming systems (MacLaren et
-al., 2020). Therefore, the objective of this study was to investigate
-the flowering pattern, gender, biomass production, and height of Palmer
-amaranth cohorts growing under corn, soybean and bareground across five
-locations in the United States Midwest.
+complexity (Lindsay et al., 2017) and posing an economical and
+ecological risk to row-crop agriculture.
+
+Palmer amaranth as a problem weed is a function of both inherent
+adaptations and selected management practices. Palmer amaranth is a fast
+growing summer annual forb indigenous to the Sonoran Desert (Sauer,
+1957). It became a serious problem weed in US agriculture in the 1990s
+(Ward et al., 2013). Palmer amaranth weediness is likely a result of
+human-assisted selection combined with plant biology. Farm
+mechanization, adoption of conservation agriculture (e.g., no-till), and
+intensive use of herbicides for weed management are the main
+human-mediated selections of Palmer amaranth in cropping systems (Ward
+et al., 2013). Palmer amaranth is a prolific seed producer with a C4
+photosynthetic apparatus (Wang et al., 1992). With dioecious nature,
+Palmer amaranth male and female plants are obligate outcrossers,
+increasing the chances of exchanging adaptive traits among plants (Jhala
+et al., 2021; Oliveira et al., 2018). Also, Palmer amaranth's small
+seeds (e.g, 1 mm) tend to thrive in no-tillage systems (Price et al.,
+2011), and spread across locations through farm equipment (Sauer, 1972),
+seed mixes (Hartzler and Anderson, 2016), wildlife (Farmer et al.,
+2017), etc., making it one of the most successful examples of weed
+adaptation to current cropping systems.
+
+Palmer amaranth's plasticity allows it to respond successfully to
+environmental changes. Palmer amaranth demonstrates a high degree of
+plasticity to light, temperature, water availability, and human
+management (Jha et al., 2010). Palmer amaranth has an extended
+germination period of up to \_\_\_\_\_ (Ward et al., 2013). Germination
+of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth
+(Keeley et al., 1987), and optimal germination and biomass production
+occurred at 35/30 C day and night temperatures (Guo and Al-Khatib,
+2003). Palmer amaranth emergence can be influenced by tillage and the
+use of preemergence herbicides (Chahal et al., 2021), and may result in
+weed germination shifts within a population, as documented in
+\emph{Bassia scoparia} (Sbatella and Wilson, 2010). In an experiment
+where Palmer amaranth was subjected to continuous water stress, it
+survived and produced at least 14000 seeds plant\textsuperscript{-1}
+(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
+water conditions were heavier, less dormant, and prompt for germination
+(Matzrafi et al., 2021). Growing conditions and management practices
+also influence Palmer amaranth sex dimorphism and flowering pattern
+(Korres et al., 2017; Rumpa et al., 2019).
+
+Palmer amaranth's current global range includes agronomic cropland in
+Italy (Milani et al., 2021), soybean producing regions of Brazil and
+Argentina (Larran et al., 2017; Küpper et al., 2017), and the Cotton
+Belt (Garetson et al., 2019; Bagavathiannan and Norsworthy, 2016) and
+southern Great Plains of the United States (Kumar et al. (2020), Crespo
+et al. (2016){]}. However, its range appears to be expanding steadily,
+and in some cases rapidly, in the U.S. For example, in states with large
+Palmer amaranth infestations its range is gradually pushing north as it
+displaces other weeds in agronomic row crops (R Werle, personal
+observation), aided by wildlife (Farmer et al., 2017) and agricultural
+machinery. But there are other examples where it has suddenly become a
+problem weed hundreds of miles away from known infestations. New
+infestations in Michigan, Indiana and North Dakota appear to have
+resulted from Palmer amaranth contaminated livestock feed. New
+infestations in Iowa and Minnesota were associated with planting
+contaminated pollinator habitat seed (Yu et al., 2021). Palmer amaranth
+infestations have not been detected in Canada, but Palmer amaranth seeds
+were found in sweet potato slips imported into the country (Page et al.,
+2021). Global warming will create new opportunities for Palmer amaranth
+invasion. Although agronomic crops in warm environments like Australia
+and Sub-Saharan Africa are currently at greatest risk for Palmer
+amaranth invasion, warming temperatures will reduce barriers that may
+have limited Palmer amaranths spread into cooler climates like Canada
+and Northern Europe (Kistner and Hatfield, 2018; Briscoe Runquist et
+al., 2019).
+
+There are many areas in the U.S. Corn Belt (North Central States) where
+Palmer amaranth is not yet established, and its potential adaptability
+is untested. Nonetheless, the rapid expansion of Palmer amaranth across
+the Northern U.S. is concerning and warrants investigations on its
+adaptability. We do know that Palmer amaranth caused yield loss in
+Illinois soybean fields (Davis et al., 2015), but an Iowa study showed
+that Palmer amaranth was not as well adapted as waterhemp
+(\emph{Amaranthus tuberculatus}) to conditions in that geography (Baker,
+2021). We also know that Palmer amaranth plants that establish shortly
+after row-crop planting have a much greater impact on crop yield than
+plants that emerge after the crop has produced several leaves (MacRae et
+al., 2013). Understanding Palmer amaranth morphology and development
+under different agroecosystems and across a wide geography can enhance
+our knowledge of its adaptability, and may also aid in designing
+effective tactics to limit is range expansion and minimize its negative
+effects on row crops. The objective of this study was to investigate the
+flowering pattern, gender, biomass production, and height of Palmer
+amaranth cohorts growing in corn, soybean and bareground environments
+across five locations in the North Central United States.
 
@@
 A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
-glyphosate resistance according to Oliveira et al. (2021) was selected
-for this study. Three weeks prior to the establishment of each cohort,
-seeds were planted in plastic trays containing potting-mix. Emerged
-seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3}
-plastic pots (a plant pot\textsuperscript{-1}). Palmer amaranth
-seedlings were supplied with adequate water and kept under greenhouse
-conditions at the University of Wisconsin-Madison, University of
-Nebraska-Lincoln, and Western Illinois University; and kept outdoors at
-the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to
-8 cm height) when they were transported to the field.
+glyphosate resistance was selected for this study (Oliveira et al.,
+2021). Three weeks prior to the establishment of each cohort, seeds were
+planted in plastic trays containing potting-mix. Emerged seedlings (1
+cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a
+plant pot\textsuperscript{-1}). Palmer amaranth seedlings were supplied
+with adequate water and kept under greenhouse conditions at the
+University of Wisconsin-Madison, University of Nebraska-Lincoln, and
+Western Illinois University; and kept outdoors at the Perkins extension
+office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when
+they were transported to the field.
 
@@
 data from June through September across the five locations in 2018 and
-2019 (Correndo et al., 2021) (Figure 1)
+2019 (Correndo et al., 2021) (Figure 1).
 
@@
 simulating two cohorts, repeated across five locations. Each field
-experimental unit was planted with corn, soybean, or kept under
+experimental unit was planted with corn, soybean, or maintained as
 bareground. The two transplant timings were June 1 (first cohort) and
@@
 gender was identified (male or female), plant height was measured from
-soil surface to the top of plant. Also, aboveground plant biomass was
-harvested near soil surface and oven dried at 65 C until reaching
-constant weight before weigh (g plant\textsuperscript{-1}) was recorded.
-
-Plants had to be harvested at flowering because Palmer amaranth is
-neither endemic in Wisconsin nor in Illinois. In our study, all
-locations followed the methodology of plant harvest at flowering
-initiation, except in Grant, NE. In this location, all Palmer amaranth
-plants were harvest at once on July 06, 2018 and 2019 (first cohort),
-and on August 17, 2018 and on July 31, 2019 (second cohort).
+soil surface to the top of plant. Also, aboveground plant, and all
+aboveground plant organs were harvested, then oven dried at 65 C until a
+constant weight was reached, and dry biomass (g
+plant\textsuperscript{-1}) was recorded. Plants were harvested at
+flowering because Palmer amaranth infestations do not occur at the
+Wisconsin or Illinois research locations. In our study, all locations
+followed the methodology of plant harvest at flowering initiation,
+except in Grant, NE. At the Grant, NE, location, all plants from the
+first cohort were harvested on July 6, 2018 or July 6, 2019, and all
+plants from the second cohort were harvested on August 17, 2018 or July
+31, 2019, regardless of flowering status.
 
@@
 corn, soybean) and cohort time (first and second) were the fixed effects
-and year nested with location the random effects. Analysis of variance
-at \(\alpha\) 0.05 was performed with \emph{anova} function from ``car''
-package (Fox and Weisberg, 2018). Marginal means and compact letter
-display were estimated with \emph{emmeans} and \emph{cld} from packages
-``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn et al.,
-2008), respectively.
-
-The Palmer amaranth flowering timing was estimated as cumulative
-flowering across all locations, except Grant, NE. Palmer amaranth
-cumulative flowering estimation was determined using an asymmetrical
-three parameter log logistic Weibull model of the drc package (Ritz et
-al., 2015).
+and year nested with location were the random effects. Analysis of
+variance at \(\alpha\) 0.05 was performed with \emph{anova} function
+from ``car'' package (Fox and Weisberg, 2018). Marginal means and
+compact letter display were estimated with \emph{emmeans} and \emph{cld}
+from packages ``emmeans'' (Lenth et al., 2021) and ``multcomp'' (Hothorn
+et al., 2008), respectively.
+
+Palmer amaranth cumulative flowering estimation was determined across
+all locations, except Grant, NE. Cumulative flowering estimation was
+determined using an asymmetrical three parameter log logistic Weibull
+model of the drc package (Ritz et al., 2015):
 
@@
 
-In this model, \emph{Y} is the Palmer amaranth cumulative flowering,
-\emph{d} is the upper limit (set to 100), and \emph{e} is the inflection
-point, and \emph{x} day of year (doy).
-
-The doy for 10, 50, and 90\% Palmer amaranth cumulative flowering were
-determined using the \emph{ED} function of drc package. Also, the 10,
-50, and 90\% Palmer amaranth cumulative flowering were compared among
-crops and cohorts using the \emph{EDcomp} function of drc package. The
-EDcomp function compares the ratio of cumulative flowering using
-t-statistics, where P-value \textless{} 0.05 indicates that we fail to
-reject the null hypothesis.
+In this model, \emph{Y} is the cumulative flowering, \emph{d} is the
+upper limit (set to 100), and \emph{e} is the inflection point, and
+\emph{x} is the day of year (doy).
+
+The doy for 10, 50, and 90\% cumulative flowering were determined using
+the \emph{ED} function of drc package. Also, the 10, 50, and 90\% Palmer
+amaranth cumulative flowering were compared among crops and cohorts
+using the \emph{EDcomp} function of drc package. The EDcomp function
+compares the ratio of cumulative flowering using t-statistics, where
+P-value \textless{} 0.05 indicates that we fail to reject the null
+hypothesis.
 
@@
 dioecious species. Prior to the analysis, missing values were removed
-from the dataset. Also, data from Grant was not used in this analysis
-due to the uniform plant harvesting at that location. The complete
-dataset was splitted into 80\% train and 20\% test data. The 80\% train
-is used for the model training and the 20\% test is used for checking
-model performance on unseen dataset. With 80\% dataset, a generalized
-linear model (base R \emph{glm} function) was fitted to binary response
-variable, the probability of being female (0 to male and 1 to female).
-The independent variables were day of year harvest, height, weight, and
-crop (without interaction). The model family was binomial with a logit
-function. The model fit was assessed through pseudo R-squared values
-(McFadden, Cox and Snell, Cragg and Uhler) and likelihood ratio using
-\emph{nagelkerke} function from ``rcompanion'' package (Mangiafico,
-2021). The marginal effects computation was performed with Average
-Marginal Effects (AMEs) at every observed value of x and average across
-the results (Leeper, 2017) using \emph{margins} function from
-``margins'' package (Leeper et al., 2021). The 20\% test data was
-predicted using the \emph{predict} function with a cutoff estimation for
-male or female using \emph{performance} function from ROCR package (Sing
-et al., 2005). The model quality prediction from the classification
-algorithm was measured with precision (\emph{precision} function),
-recall (\emph{recall} function) and F1-score (\emph{f\_meas} function)
-using the ``yardstick'' package (Kuhn et al., 2021). The precision
-determines the accuracy of positive predictions (female plants), recall
-determines the fraction of positives that were correctly identified, and
-F1-score is a weighted harmonic mean of precision and recall with the
-best score of 1 and the worst score of 0 (Raoniar, 2021). F1-score
-conveys the balance between the precision and the recall (Yacouby and
-Axman, 2020). The area under the receiver operating curve (AUC-ROC) was
-also estimated with performance function using the true positive and
-false positive rates. Higher the AUC, better the model is at
-distinguishing between female and male Palmer amaranth.
+from the dataset (including all data from the Grant location). The
+resulting dataset was split into 80\% train and 20\% test data. The 80\%
+train is used for model training and the 20\% test is used for checking
+model performance on unseen dataset. Using the 80\% train data, a
+generalized linear model (base R \emph{glm} function) was fitted to
+binary response variable, the probability of being female (0 to male and
+1 to female). The independent variables were harvest day of year,
+height, weight, and crop (without interaction). The model family was
+binomial with a logit function. The model fit was assessed through
+pseudo R-squared values (McFadden, Cox and Snell, Cragg and Uhler) and
+likelihood ratio using \emph{nagelkerke} function from ``rcompanion''
+package (Mangiafico, 2021). The marginal effects computation was
+performed with Average Marginal Effects (AMEs) at every observed value
+of x and averaged across the results (Leeper, 2017) using \emph{margins}
+function from ``margins'' package (Leeper et al., 2021). The 20\% test
+data was predicted using the \emph{predict} function with a cutoff
+estimation for male or female using \emph{performance} function from
+ROCR package (Sing et al., 2005). The model quality prediction from the
+classification algorithm was measured with precision (\emph{precision}
+function), recall (\emph{recall} function) and F1-score (\emph{f\_meas}
+function) using the ``yardstick'' package (Kuhn et al., 2021). The
+precision determines the accuracy of positive predictions (female
+plants), recall determines the fraction of positives that were correctly
+identified, and F1-score is a weighted harmonic mean of precision and
+recall with the best score of 1 and the worst score of 0 (Raoniar,
+2021). F1-score conveys the balance between the precision and the recall
+(Yacouby and Axman, 2020). The area under the receiver operating curve
+(AUC-ROC) was also estimated with performance function using the true
+positive and false positive rates. The higher the AUC, better the model
+is at distinguishing between female and male Palmer amaranth.
 
@@
 plant\textsuperscript{-1} in soybean, and 1.4 g
-plant\textsuperscript{-1}.
-
-Palmer amaranth height was more uniform across cohort timings, except
-when growing in corn (figure 3B). Palmer amaranth plants from the first
-cohort were on average 69.2 cm tall in bareground, which was not
-different from the 70.7 cm tall plants from the second cohort timing (P
-= 0.74). In addition, no difference in Palmer amaranth height (69.3 cm)
-was detected from first cohort plants in soybean to first and second
-cohort plants in bareground (P \textgreater{} 0.75). Palmer amaranth
-plants from the second cohort were nearly 10 cm lower compared to the
-first cohort in soybeans (P = 0.04). The tallest (first cohort) and
-smallest (second cohort) Palmer amaranth plants were found in corn.
-Palmer amaranth reached 85.2 and 38.2 cm tall, respectively.
+plant\textsuperscript{-1} in corn.
+
+Palmer amaranth height was less affected by cohort timing than was
+weight, with the exception of plants growing in corn (Figure 3B). Plants
+from the first cohort were on average 69.2 cm tall in bareground, which
+was not different from the 70.7 cm tall plants from the second cohort
+timing (P = 0.74). In addition, no difference in Palmer amaranth height
+(69.3 cm) was detected from first cohort plants in soybean to first and
+second cohort plants in bareground (P \textgreater{} 0.75). Palmer
+amaranth plants from the second cohort were nearly 10 cm shorter
+compared to the first cohort in soybeans (P = 0.04). The tallest (85.2
+cm, first cohort) and shortest (38.2 cm, second cohort) plants were
+observed in corn.
 
@@
 
-Palmer amaranth plants from the first cohort growing in corn resulted in
-a longer flowering window compared to plants growing in bareground and
-soybean (Figure 4A). The 10\% cumulative Palmer amaranth flowering in
-soybean, bareground and corn occurred at the end of June. Palmer
-amaranth reached 10\% flowering in soybean, bareground and corn at doy
-180, 180.9 and 181.7, respectively. The 50\% Palmer amaranth cumulative
-flowering occurred in July. Palmer amaranth reached 50\% flowering in
-bareground, soybean and corn at doy 193.4, 194.8, and 206.6,
-respectively. Similar trend was observed at 90\% Palmer amaranth
-cumulative flowering. Palmer amaranth reached 90\% flowering at doy
-252.6 in corn (early September), which was 38 and 32 days after reaching
-90\% flowering in bareground and soybean, respectively.
+The initiation and duration of Palmer amaranth flowering was strongly
+influenced by cohort and surrounding vegetation (Figure 4A, 4B). In the
+first cohort, floral initiation (10\% flowering) occurred near the end
+of June for all three treatments, at doy 180, 180.9, and 181.7 for
+soybean, bareground, and corn, respectively. In the second cohort,
+floral initiation occurred earlier on plants growing in the bareground
+than plants growing in soybean or corn (doy 203.8 versus doy 210.9 or
+216.8, respectively). Palmer amaranth growing in the bareground had the
+shortest duration of floral initiation (measured as the difference
+between 10\% and 90\% cumulative flowering) in both the first (34 days)
+and second (28 days) cohorts. Palmer amaranth growing in soybean had a
+shorter duration of flowering than corn in the first cohort (40 days vs
+71 days), but a longer duration in the second cohort (50 vs 44 days).
 
@@
 average marginal effects showed that Palmer amaranth growing in corn
-resulted in 14.8\% less females plants (Table 2). Moreover, increasing a
-unit doy increases the probability of having a female plant by 0.4\%
-(Table 2 and Figure4A). Similar trend is observed for weight (Figure 4B)
-as well as height (Figure 4C), whereas the probability of being female
-increase by 0.2\% and 0.1\% when a unit of weight (g) and height (cm)
-increases, respectively.
-
-\begin{table}[!h]
-
-\caption{\label{tab:unnamed-chunk-3}Average marginal effects of Palmer amaranth gender logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
-\centering
-\fontsize{10}{12}\selectfont
+resulted in 14.8\% fewer females plants (Table 2). Moreover, increasing
+a cumulative flowering unit doy increases the probability of having a
+female plant by 0.4\% (Table 2 and Figure5A). A similar trend was
+observed for weight (Figure 5B) and height (Figure 5C), where the
+probability of being female increased 0.1\% and 0.2\% for each unit
+increase of weight (g) and height (cm), respectively (Table 2).
+
+\textbackslash begin\{table\}{[}!h{]}
+
+\textbackslash caption\{\label{tab:unnamed-chunk-3}Average marginal
+effects of Palmer amaranth gender logistic model. Factor pararemter
+values (crop\_corn and 'crop\_bareground) are calculated relative to
+soybean.\} \centering \fontsize{10}{12}\selectfont
+
 \begin{tabular}[t]{lrrrrrr}
@@
 \end{tabular}
-\end{table}
+
+\textbackslash end\{table\}
 
@@
 
-Our study showed that Palmer amaranth biomass, height, flowering window
-and gender varied within crops and cohort timings. In general, first
-cohort of Palmer amaranth plants were heavier and taller when compared
-to the second cohort. At first cohort, resources (e.g., soil nutrients)
-and conditions (e.g., light) were more timely available for the species.
-High biomass and taller Palmer amaranth plants are likely a weed
-strategy to compete for light in between crop rows in absence of canopy.
-In such conditions, Palmer amaranth showed an extraordinary plasticity
-to adapt upon the agroecosystem. This is evident when comparing Palmer
-amaranth canopy shape. The Palmer amaranth competition (e.g., light)
-strategy was to mimic the crop growth and development (Figure 6). These
-results suggests that Palmer amaranth can quickly evolve life-history
-traits to adapt to cropping systems and cultural practices, which was
-also shown in a study varying nitrogen fertilization (Bravo et al.,
-2018). Our results highlight Palmer amaranth as a threat to field crops
-as breeding more competitive crop varieties is likely to select more
-competitive weed biotypes (Bravo et al., 2017).
+Our study confirmed that Palmer amaranth is well adapted to growing
+throughout the Midwestern United States, and is simply limited in its
+range by seed dispersal (Davis et al.~2015). We also confirmed Palmer
+amaranth's extraordinary plasticity to adapt to different
+agroecosystems. For example, Palmer amaranth mimicked crop architecture
+in competing for light (Figure 6), where plants growing in corn
+allocated resources to height, while plants growing in bareground
+allocated resources to numerous branches. Plants growing in the absence
+of a crop produced the greatest biomass, responding to a greater
+abundance of light, nutrient and water resources (Figure 3A, 3B). These
+results support the argument that Palmer amaranth can quickly evolve
+life-history traits to adapt to different cultural practices, similar to
+observations in a Palmer amaranth response to nitrogen study (Bravo et
+al., 2018). Our results highlight Palmer amaranth as a threat to field
+crops as breeding more competitive crop varieties is likely to select
+more competitive weed biotypes (Bravo et al., 2017).
 
@@
 
+Floral initiation depends on the complex interaction between a plant's
+genetic makeup and environmental conditions (Lang, 1965). We observed a
+significant disruption in flowering caused by surrounding vegetation and
+cohort timing. Although all three treatments began flowering at the same
+time in the first cohort (about doy 180), it took much less time for the
+bareground treatment to reach 90\% flowering (34 days versus 40 for
+soybean and 71 for corn). The lack of competition (and stress) in the
+bareground resulted in plants that flowered earlier, and were they not
+harvested, would have had longer reproductive periods, thereby producing
+more seed. Similarly, in the second cohort, the window for all plants to
+initiate flowering was much shorter for the bareground (28 days)
+compared to Palmer amaranth in soybean (50 days) or corn (44 days).
+Significantly, plants in the bareground treatment flowered one week
+earlier than soybean, and almost two weeks earlier than corn, again
+allowing for greater seed production per plant had they not been
+harvested at initial flowering. In some cases, early flowering may
+confer an evolutionary advantage, provided the plant has an
+indeterminate habit and flowering does not restrict plant growth. In an
+Iowa study, Palmer amaranth initiated flowering two weeks prior to
+waterhemp (Baker, 2021), and this characteristic may partially explain
+why Palmer amaranth can displace waterhemp where both occur in the same
+agronomic field. In other cases, however, early flower initiation may be
+a response to stressful conditions as a plant attempts to reproduce
+before running out of resources. In a dioecious species like Palmer
+amaranth, exerting stress on plants to manipulate flowering may be
+beneficial in limiting seed production (McFarlane et al., 2018;
+Schliekelman et al., 2005). For example, when growing under water
+stress, there was a seven days flowering mismatch between male and
+female plants (Mesgaran et al., 2021), which can minimize plant
+outcrossing, reduce seed production and the exchange of resistant
+alleles (Jhala et al., 2021). In our study we observed
+
 An ecological approach to reduce Palmer amaranth seed production is
@@
 The mechanisms of gender-determination in plant species is intriguing
-and arouse the curiosity of many scientists, including Charles Darwin
-(Darwin, 1888). In our study, the gender model performance was decent
-(AUC 0.64) considering the biology of plant flowering. A 1:1 male and
-female sex ratio is a general evolutionary stable strategy for plant
-species perpetuation (Fisher, 1930). However, a slight deviation from
-1:1 sex ratio might occur in some dioecious species. For example, the
+and has arouse the curiosity of many scientists, including Charles
+Darwin (Darwin, 1888). In our study, the gender model performance was
+decent (AUC 0.64) considering the biology of plant flowering. A 1:1 male
+and female sex ratio is a general an evolutionarily stable strategy for
+plant species perpetuation (Fisher, 1930). However, a slight deviation
+from 1:1 sex ratio occurs in some dioecious species. For example, the
 dioecious \emph{Halophila stipulacea} is a female-biased plant in its
-native habitat, but the naturalized \emph{H. stipulacea} have a 1:1
-ratio (Nguyen et al., 2018). Naturalized of \emph{H. stipulacea} reduced
+native habitat, but the naturalized \emph{H. stipulacea} has a 1:1 ratio
+(Nguyen et al., 2018). Naturalization of \emph{H. stipulacea} reduced
 female-male ratio to expand into its non-native habitat (Nguyen et al.,
@@
 high plant densities (Korres and Norsworthy, 2017) and after herbicide
-application (Rumpa et al., 2019). Our model estimated that late
-flowering, heavier and taller Palmer amaranth plants slighly deviated
-from 1:1 ratio in favor to female plants. It was reported that female
-Palmer amaranth plants invested more in height, stem and biomass while
-male invested more in leaf area and leaf dry weight under nutrient
-deficiency (Korres et al., 2017). Our model also estimated more female
-plants in soybean and bareground compared to corn, which might be linked
-to plant competition strategy in each crop. Our results showed the
-influence of life-history and ecological traits on sexual dimorphism in
-Palmer amaranth. Sexual dimorphism is documented in other dioecious
-species (Barrett and Hough, 2013). For example, stronger female plant
-competition and greater male tolerance to herbivory was reported in
-\emph{Spinacia oleracea} (Pérez-Llorca and Sánchez Vilas, 2019).
-Research on candidate genes for sex determination in \emph{Amaranthus}
-species are currently underway but it is far from complete (Montgomery
-et al., 2021, 2019). Further studies are also needed to understand the
-ecological basis of Palmer amaranth flowering, including the plant
-behavior under climate change.
-
-Our study demonstrated the Palmer amaranth plasticity to grow and
-develop into arable land of US Midwest. It is likely that Palmer
-amaranth range will continue to expanding into new geographies. The
-migration of Palmer amaranth into the US Midwest will reshape the
-landscape as waterhemp and Palmer amaranth will share the same habitat.
-The presence of Palmer amaranth and waterhemp will increase the weed
-management complexity. Therefore, preventive management is a priority to
-minimizing Palmer amaranth dispersal. Reactive management should focus
-on early-season management programs, which would have a large negative
-effect on Palmer amaranth growth and development. Long-term tactics that
-promote early-season crop advantage against Palmer amaranth, including
-diversity of crops in rotation, early/late crop planting, plant width,
-and crop residue (e.g.~cover crops) would minimize the negative impact
-of Palmer amaranth to cropping systems. The aggressiveness and
-differential Palmer amaranth adaptation to agroecosystem is striking and
-require regional collaboration to minimize the species impact on
-sustainability and profitability of cropping systems.
+application (Rumpa et al., 2019). We observed sexual dimorphism in
+Palmer amaranth in response to surrounding vegetation and plant
+morphological attributes. Our model estimated that late flowering,
+heavier and taller Palmer amaranth plants slightly deviated from 1:1
+ratio in favor of female plants. It was reported that female Palmer
+amaranth plants invested more in height, stem and biomass while male
+plants invested more in leaf area and leaf dry weight under nutrient
+deficiency (Korres et al., 2017). We observed more female plants in
+soybean and bareground compared to corn. Palmer amaranth plants in the
+corn were more stressed by interspecific competition as evidenced by
+less biomass and a lower weight:height ratio. Sexual dimorphism is
+documented in other dioecious species (Barrett and Hough, 2013). For
+example, stronger female plant competition and greater male tolerance to
+herbivory was reported in \emph{Spinacia oleracea} (Pérez-Llorca and
+Sánchez Vilas, 2019). Research on candidate genes for sex determination
+in \emph{Amaranthus} species are currently underway but is far from
+complete (Montgomery et al., 2021, 2019). Further studies are also
+needed to understand the ecological basis of Palmer amaranth flowering,
+including plant behavior under climate change.
+
+Our study demonstrated that Palmer amaranth is adapted to grow on arable
+land throughout the Midwestern U.S. Palmer amaranth's range will
+continue to expand if current cropping practices are continued. As
+waterhemp and Palmer amaranth begin to share the same habitat, it will
+increase weed management complexity. Preventing Palmer amaranth seed
+dispersal must be a priority. Regional collaboration is necessary to
+slow the spread of this aggressive and adaptable weed. Where Palmer
+amaranth occurs, management tactics should focus on limiting Palmer
+amaranth establishment until row crops can shade late-emerging plants.
+Increasing the diversity of crops in rotation, varying row crop planting
+date, narrowing row width, and increasing residue cover through the use
+of cover crops are all tactics that can minimize the growth and seed
+production of Palmer amaranth, and improve the sustainability of
+Midwestern cropping systems.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-aulakh2021}{}%
-Aulakh, J. S., Chahal, P. S., Kumar, V., Price, A. J., and Guillard, K.
-(2021). Multiple herbicide-resistant {Palmer} amaranth ({Amaranthus}
-palmeri) in {Connecticut}: Confirmation and response to {POST}
-herbicides. \emph{Weed Technology} 35, 457--463.
-doi:\href{https://doi.org/10.1017/wet.2021.6}{10.1017/wet.2021.6}.
-
 \leavevmode\hypertarget{ref-bagavathiannan2016}{}%
@@
 
+\leavevmode\hypertarget{ref-crespo2016a}{}%
+Crespo, R. J., Wingeyer, A. B., Borman, C. J., and Bernards, M. L.
+(2016). Baseline {Sensitivity} of {Nebraska Waterhemp} and {Palmer
+Amaranth} to {Dicamba} and 2,4-{D}. \emph{Agronomy Journal} 108,
+1649--1655.
+doi:\href{https://doi.org/10.2134/agronj2015.0465}{10.2134/agronj2015.0465}.
+
 \leavevmode\hypertarget{ref-darwin1888}{}%
@@
 
-\leavevmode\hypertarget{ref-kohrt2017}{}%
-Kohrt, J. R., Sprague, C. L., Nadakuduti, S. S., and Douches, D. (2017).
-Confirmation of a {Three}-{Way} ({Glyphosate}, {ALS}, and {Atrazine})
-{Herbicide}-{Resistant Population} of {Palmer Amaranth} ({Amaranthus}
-palmeri) in {Michigan}. \emph{Weed Science} 65, 327--338.
-doi:\href{https://doi.org/10.1017/wsc.2017.2}{10.1017/wsc.2017.2}.
-
 \leavevmode\hypertarget{ref-korres2017a}{}%
@@
 24, 2021{]}.
+
+\leavevmode\hypertarget{ref-kumar2020}{}%
+Kumar, V., Liu, R., and Stahlman, P. W. (2020). Differential sensitivity
+of {Kansas Palmer} amaranth populations to multiple herbicides.
+\emph{Agronomy Journal} 112, 2152--2163.
+doi:\href{https://doi.org/10.1002/agj2.20178}{10.1002/agj2.20178}.
 
@@
 
-\leavevmode\hypertarget{ref-maclaren2020}{}%
-MacLaren, C., Storkey, J., Menegat, A., Metcalfe, H., and
-Dehnen-Schmutz, K. (2020). An ecological future for weed science to
-sustain crop production and the environment. {A} review. \emph{Agron.
-Sustain. Dev.} 40, 24.
-doi:\href{https://doi.org/10.1007/s13593-020-00631-6}{10.1007/s13593-020-00631-6}.
-
 \leavevmode\hypertarget{ref-macrae2013}{}%
@@
 
-\leavevmode\hypertarget{ref-scott2011}{}%
-Scott, D. (2011). The {Technological Fix Criticisms} and the
-{Agricultural Biotechnology Debate}. \emph{J Agric Environ Ethics} 24,
-207--226.
-doi:\href{https://doi.org/10.1007/s10806-010-9253-7}{10.1007/s10806-010-9253-7}.
-
 \leavevmode\hypertarget{ref-sing2005}{}%
@@
 2021{]}.
-
-\leavevmode\hypertarget{ref-tranel2011}{}%
-Tranel, P. J., Riggins, C. W., Bell, M. S., and Hager, A. G. (2011).
-Herbicide {Resistances} in {Amaranthus} tuberculatus: {A Call} for {New
-Options}. \emph{J. Agric. Food Chem.} 59, 5808--5812.
-doi:\href{https://doi.org/10.1021/jf103797n}{10.1021/jf103797n}.
 

### 2022-02-25 Manuscript/Manuscript.Rmd
as

@@
 \tiny
- \keyFont{ \section{Keywords:} Evolution, Flowering, Management, Pigweed, Weed}
+ \keyFont{ \section{Keywords:} Evolution, Flowering, Weed management, Pigweed}
 
@@
 Palmer amaranth demonstrates a high degree of plasticity to light, temperature, water availability, and human management [@jha2010].
-Palmer amaranth has an extended germination period of up to _____ [@ward2013].
+Palmer amaranth has an extended germination period throughout the growing season [@ward2013].
 Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003].
@@
 
-Palmer amaranth's current global range includes agronomic cropland in Italy [@milani2021], soybean producing regions of Brazil and Argentina [@larran2017; @kupper2017], and the Cotton Belt [@garetson2019; @bagavathiannan2016] and southern Great Plains of the United States (@kumar2020, @crespo2016a].
+Palmer amaranth's current global range includes agronomic cropland in Italy [@milani2021], soybean producing regions of Brazil and Argentina [@larran2017; @kupper2017], and the Cotton Belt [@garetson2019; @bagavathiannan2016] and southern Great Plains of the United States [@kumar2020, @crespo2016a].
 However, its range appears to be expanding steadily, and in some cases rapidly, in the U.S.
-For example, in states with large Palmer amaranth infestations its range is gradually pushing north as it displaces other weeds in agronomic row crops (R Werle, personal observation), aided by wildlife [@farmer2017] and agricultural machinery.
+For example, in states with large Palmer amaranth infestations its range is gradually pushing north as it displaces other weeds in agronomic row crops (R Werle, personal observation), aided by wildlife [@farmer2017] and agricultural machinery [@ward2013].
 But there are other examples where it has suddenly become a problem weed hundreds of miles away from known infestations.
@@
 There are many areas in the U.S.
-Corn Belt (North Central States) where Palmer amaranth is not yet established, and its potential adaptability is untested.
+Corn Belt (North Central states) where Palmer amaranth is not yet established, and its potential adaptability is untested.
 Nonetheless, the rapid expansion of Palmer amaranth across the Northern U.S.
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (XXX'N, XXX'W).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (40.49'N, -90.69'W).
 
@@
 
-Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (figure 3A).
+Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (Figure 3A).
 Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
@@
 Moreover, increasing a cumulative flowering unit doy increases the probability of having a female plant by 0.4% (Table 2 and Figure5A).
-A similar trend was observed for weight (Figure 5B) and height (Figure 5C), where the probability of being female increased 0.1% and 0.2% for each unit increase of weight (g) and height (cm), respectively (Table 2).
+A similar trend was observed for weight (Figure 5B) and height (Figure 5C), where the probability of being female increased 0.1% and 0.2% for each unit increase of weight and height, respectively (Table 2).
 
 ```{r echo = FALSE, warning = FALSE, message=FALSE}
-read_csv("../data/table_a.csv") -> table_a
+readr::read_csv("../data/table_a.csv") -> table_a
 
 table_a %>% 
-  kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth gender logistic model.
+  kableExtra::kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth gender logistic model.
 Factor pararemter values (crop_corn and 'crop_bareground) are calculated relative to soybean.") %>%
@@
 Palmer amaranth can produce a hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
-Therefore, preventing Palmer amaranth seed production or/and seed migration to non-native habitats is an essential strategy to minimize the species impact in agroecosystems [@davis2015].
+Therefore, preventing Palmer amaranth seed production or/and seed migration to non-native habitats is an essential strategy to minimize the species impact in agroecosystems.
 
+An ecological approach to reduce Palmer amaranth seed production is understanding plant biology, including flowering pattern.
 Floral initiation depends on the complex interaction between a plant's genetic makeup and environmental conditions [@lang1965].
@@
 For example, when growing under water stress, there was a seven days flowering mismatch between male and female plants [@mesgaran2021], which can minimize plant outcrossing, reduce seed production and the exchange of resistant alleles [@jhala2021].
-In our study we observed
-
-An ecological approach to reduce Palmer amaranth seed production is understanding plant biology, including flowering pattern.
-Our study suggests that Palmer amaranth flowering was slightly influenced by crops and cohort timings.
-Palmer amaranth growing in bareground and corn resulted in the overall shortest and longest flowering window, respectively.
-The shortest flowering window of second cohorts of Palmer amaranth growing in bareground highlighted the impact of photoperiod on flowering.
-When growing in soybean, Palmer amaranth flowering window was similar to bareground for the first cohort but similar to corn at second cohort timing.
-Plant flowering initiation is complex and depends on the complex interaction between genetic makeup and the environmental conditions [@lang1965].
-We hypothesize that when growing in high competition (e.g., second cohort), Palmer amaranth plants tend to initiate flowering early, as well as having an extended flowering window.
-A study has shown that Palmer amaranth initiated flowering two weeks prior to the native waterhemp in Iowa [@baker2021].
-Early flower initiation could also be a plant strategy when growing in stressful conditions.
-For example, when growing under water stress, early flowering in Palmer amaranth resulted in a mismatch between female and male plants by seven days [@mesgaran2021].
-A mismatch in Palmer amaranth male and female flowering period can minimize plant outcrossing, reducing plant seed production and/or exchange of resistant alleles [@jhala2021].
-Sex dimorphism manipulation is considered a potential ecological pest control strategy [@mcfarlane2018; @schliekelman2005].
 

### 2022-02-25 Manuscript/Manuscript.tex
as

@@
 management (Jha et al., 2010). Palmer amaranth has an extended
-germination period of up to \_\_\_\_\_ (Ward et al., 2013). Germination
-of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth
-(Keeley et al., 1987), and optimal germination and biomass production
-occurred at 35/30 C day and night temperatures (Guo and Al-Khatib,
-2003). Palmer amaranth emergence can be influenced by tillage and the
-use of preemergence herbicides (Chahal et al., 2021), and may result in
-weed germination shifts within a population, as documented in
+germination period throughout the growing season (Ward et al., 2013).
+Germination of Palmer amaranth was triggered by 18 C soil temperature at
+5 cm depth (Keeley et al., 1987), and optimal germination and biomass
+production occurred at 35/30 C day and night temperatures (Guo and
+Al-Khatib, 2003). Palmer amaranth emergence can be influenced by tillage
+and the use of preemergence herbicides (Chahal et al., 2021), and may
+result in weed germination shifts within a population, as documented in
 \emph{Bassia scoparia} (Sbatella and Wilson, 2010). In an experiment
@@
 Belt (Garetson et al., 2019; Bagavathiannan and Norsworthy, 2016) and
-southern Great Plains of the United States (Kumar et al. (2020), Crespo
-et al. (2016){]}. However, its range appears to be expanding steadily,
-and in some cases rapidly, in the U.S. For example, in states with large
-Palmer amaranth infestations its range is gradually pushing north as it
-displaces other weeds in agronomic row crops (R Werle, personal
-observation), aided by wildlife (Farmer et al., 2017) and agricultural
-machinery. But there are other examples where it has suddenly become a
+southern Great Plains of the United States Crespo et al. (2016).
+However, its range appears to be expanding steadily, and in some cases
+rapidly, in the U.S. For example, in states with large Palmer amaranth
+infestations its range is gradually pushing north as it displaces other
+weeds in agronomic row crops (R Werle, personal observation), aided by
+wildlife (Farmer et al., 2017) and agricultural machinery (Ward et al.,
+2013). But there are other examples where it has suddenly become a
 problem weed hundreds of miles away from known infestations. New
@@
 
-There are many areas in the U.S. Corn Belt (North Central States) where
+There are many areas in the U.S. Corn Belt (North Central states) where
 Palmer amaranth is not yet established, and its potential adaptability
@@
 (40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N,
-96.42'W), and Macomb, IL (XXX'N, XXX'W).
+96.42'W), and Macomb, IL (40.49'N, -90.69'W).
 
@@
 
- & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor\textsuperscript{f} \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
-
- & Seeding rate & 88956 & 86487 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
- & Stage at 1 cohort & V2-3 &  &  &  & \\
-
- & Stage at 2 cohort & V6-7 &  &  &  & \\
-
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
-
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & DKC64-34RIB (2018) / G10T63-3000GT (2019) \\
+
+ & Seeding rate & 88956 & 86487 & 79000 & 86485 & 88,000 (2018) / 79,000 (2019)\\
+
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & V4-5\\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & V10-11\\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & May 8, 2018 / April 26, 2019 \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & UAN-32, 202 kg N ha\textsuperscript{-1}\\
 
@@
 
- & Variety & DSR-1950 & AG21X8 &  &  & \\
-
- & Seeding rate & 296400 & 321237 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
-
- & Stage at 1 cohort & V1-2 &  &  &  & \\
-
- & Stage at 2 cohort & V5-6 &  &  &  & \\
-
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 &  &  & \\
+ & Variety & DSR-1950 & AG21X8 & AG28X7 & 3017R2X (2018) / P31A22X (2019) & AG33X8 (2018) / CZ3601 (2019) \\
+
+ & Seeding rate & 296400 & 321237 & 220000 & 370650 & 395000 (2018) / 370000 (2019) \\
+
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & glyphosate + \emph{S}-metolachor \\
+
+ & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 & V1-2 & V1-2 \\
+
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 & V5-6 & V7-8\\
+
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & May 8, 2018 / May 17, 2019 \\
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1 (Team, 2021).
+version 4.0.1 (R Core Team, 2021).
 
@@
 Palmer amaranth plants accumulated more biomass when growing in
-bareground compared to plants growing in soybean and corn (figure 3A).
+bareground compared to plants growing in soybean and corn (Figure 3A).
 Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3
@@
 probability of being female increased 0.1\% and 0.2\% for each unit
-increase of weight (g) and height (cm), respectively (Table 2).
-
-\textbackslash begin\{table\}{[}!h{]}
-
-\textbackslash caption\{\label{tab:unnamed-chunk-3}Average marginal
-effects of Palmer amaranth gender logistic model. Factor pararemter
-values (crop\_corn and 'crop\_bareground) are calculated relative to
-soybean.\} \centering \fontsize{10}{12}\selectfont
-
+increase of weight and height, respectively (Table 2).
+
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
+\centering
+\fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lrrrrrr}
 \toprule
-Term & AME & SE & Lower & Upper & Z-score & P-value\\
+Term & AME\textsuperscript{a} & SE\textsuperscript{b} & Lower & Upper & Z-score & P-value\\
 \midrule
@@
 \end{tabular}
-
-\textbackslash end\{table\}
+\end{table}
 
@@
 non-native habitats is an essential strategy to minimize the species
-impact in agroecosystems (Davis et al., 2015).
-
-Floral initiation depends on the complex interaction between a plant's
-genetic makeup and environmental conditions (Lang, 1965). We observed a
+impact in agroecosystems.
+
+An ecological approach to reduce Palmer amaranth seed production is
+understanding plant biology, including flowering pattern. Floral
+initiation depends on the complex interaction between a plant's genetic
+makeup and environmental conditions (Lang, 1965). We observed a
 significant disruption in flowering caused by surrounding vegetation and
@@
 outcrossing, reduce seed production and the exchange of resistant
-alleles (Jhala et al., 2021). In our study we observed
-
-An ecological approach to reduce Palmer amaranth seed production is
-understanding plant biology, including flowering pattern. Our study
-suggests that Palmer amaranth flowering was slightly influenced by crops
-and cohort timings. Palmer amaranth growing in bareground and corn
-resulted in the overall shortest and longest flowering window,
-respectively. The shortest flowering window of second cohorts of Palmer
-amaranth growing in bareground highlighted the impact of photoperiod on
-flowering. When growing in soybean, Palmer amaranth flowering window was
-similar to bareground for the first cohort but similar to corn at second
-cohort timing. Plant flowering initiation is complex and depends on the
-complex interaction between genetic makeup and the environmental
-conditions (Lang, 1965). We hypothesize that when growing in high
-competition (e.g., second cohort), Palmer amaranth plants tend to
-initiate flowering early, as well as having an extended flowering
-window. A study has shown that Palmer amaranth initiated flowering two
-weeks prior to the native waterhemp in Iowa (Baker, 2021). Early flower
-initiation could also be a plant strategy when growing in stressful
-conditions. For example, when growing under water stress, early
-flowering in Palmer amaranth resulted in a mismatch between female and
-male plants by seven days (Mesgaran et al., 2021). A mismatch in Palmer
-amaranth male and female flowering period can minimize plant
-outcrossing, reducing plant seed production and/or exchange of resistant
-alleles (Jhala et al., 2021). Sex dimorphism manipulation is considered
-a potential ecological pest control strategy (McFarlane et al., 2018;
-Schliekelman et al., 2005).
+alleles (Jhala et al., 2021).
 
@@
 dioecious \emph{Halophila stipulacea} is a female-biased plant in its
-native habitat, but the naturalized \emph{H. stipulacea} has a 1:1 ratio
+native habitat, but naturalized \emph{H. stipulacea} has a 1:1 ratio
 (Nguyen et al., 2018). Naturalization of \emph{H. stipulacea} reduced

### 2022-02-25 Manuscript/table.tex
as

@@
 
- & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor\textsuperscript{f} \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & DKC64-34RIB (2018) / G10T63-3000GT (2019) \\
 
- & Seeding rate & 88956 & 86487 & 79000 & 86485 & \\
+ & Seeding rate & 88956 & 86487 & 79000 & 86485 & 88,000 (2018) / 79,000 (2019)\\
 
- & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & \\
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
 
- & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & \\
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & V4-5\\
 
- & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & \\
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & V10-11\\
 
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & \\
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & May 8, 2018 / April 26, 2019 \\
 
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & \\
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & UAN-32, 202 kg N ha\textsuperscript{-1}\\
 
@@
 
- & Variety & DSR-1950 & AG21X8 & AG28X7 & 2018: 3017R2X; 2019 P31A22X & \\
+ & Variety & DSR-1950 & AG21X8 & AG28X7 & 3017R2X (2018) / P31A22X (2019) & AG33X8 (2018) / CZ3601 (2019) \\
 
- & Seeding rate & 296400 & 321237 & 220000 & 370650 & \\
+ & Seeding rate & 296400 & 321237 & 220000 & 370650 & 395000 (2018) / 370000 (2019) \\
 
- & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & \\
+ & Weed control & glyphosate / \emph{S}-metolachor & saflufenacil + imazethapyr + pyroxasulfone & glyphosate & glyphosate + \emph{S}-metolachor fb glyphosate & glyphosate + \emph{S}-metolachor \\
 
- & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 &  & \\
+ & Stage at 1\textsuperscript{st} cohort & V1-2 & V1-2 & V1-2 & V1-2 & V1-2 \\
 
- & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 &  & \\
+ & Stage at 2\textsuperscript{nd} cohort & V5-6 & V5-6 & V5-6 & V5-6 & V7-8\\
 
-\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & \\
+\multirow{-6}{*}{\raggedright\arraybackslash } & Planting day & May 5, 2018 / May 10, 2009 & May 14, 2018/19 & May 5, 2018 / May 14, 2019 & May 7, 2018 / May 17, 2019 & May 8, 2018 / May 17, 2019 \\
 

### 2022-02-28 Manuscript/Manuscript.Rmd
as

@@
 
-A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no glyphosate resistance was selected for this study [@oliveira2021a].
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska susceptible to glyphosate was selected for this study [@oliveira2021a].
 Three weeks prior to the establishment of each cohort, seeds were planted in plastic trays containing potting-mix.
-Emerged seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a plant pot\textsuperscript{-1}).
-Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors at the Perkins extension office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
+Emerged seedlings (1 cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (one plant pot\textsuperscript{-1}).
+Palmer amaranth seedlings were supplied with adequate water and kept under greenhouse conditions at the University of Wisconsin-Madison, University of Nebraska-Lincoln, and Western Illinois University; and kept outdoors at the Perkins Extension office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when they were transported to the field.
 
@@
 When a plant flowered, the day was recorded, plant gender was identified (male or female), plant height was measured from soil surface to the top of plant.
-Also, aboveground plant, and all aboveground plant organs were harvested, then oven dried at 65 C until a constant weight was reached, and dry biomass (g plant\textsuperscript{-1}) was recorded.
+Also, aboveground plant, and the aboveground plant organs were harvested, then oven dried at 65 C until a constant weight was reached, and dry biomass (g plant\textsuperscript{-1}) was recorded.
 Plants were harvested at flowering because Palmer amaranth infestations do not occur at the Wisconsin or Illinois research locations.
@@
 
-The initiation and duration of Palmer amaranth flowering was strongly influenced by cohort and surrounding vegetation (Figure 4A, 4B).
+The initiation and duration of Palmer amaranth flowering were strongly influenced by cohort and surrounding vegetation (Figure 4A, 4B).
 In the first cohort, floral initiation (10% flowering) occurred near the end of June for all three treatments, at doy 180, 180.9, and 181.7 for soybean, bareground, and corn, respectively.
 In the second cohort, floral initiation occurred earlier on plants growing in the bareground than plants growing in soybean or corn (doy 203.8 versus doy 210.9 or 216.8, respectively).
-Palmer amaranth growing in the bareground had the shortest duration of floral initiation (measured as the difference between 10% and 90% cumulative flowering) in both the first (34 days) and second (28 days) cohorts.
+Palmer amaranth growing in the bareground had the shortest flowering initiation (measured as the difference between 10% and 90% cumulative flowering) in both the first (34 days) and second (28 days) cohorts.
 Palmer amaranth growing in soybean had a shorter duration of flowering than corn in the first cohort (40 days vs 71 days), but a longer duration in the second cohort (50 vs 44 days).
@@
 
-Palmer amaranth cumulative flowering in the second cohort ranged from mid July to mid September (Figure 4B).
+Palmer amaranth cumulative flowering in the second cohort occurred from mid July to mid September (Figure 4B).
 Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn.
@@
 
-Our study confirmed that Palmer amaranth is well adapted to growing throughout the Midwestern United States, and is simply limited in its range by seed dispersal (Davis et al.
+Our study demonstrate that Palmer amaranth is well adapted to growing conditions throughout the Midwestern United States, and is simply limited in its range by seed dispersal (Davis et al.
 2015).
@@
 These results support the argument that Palmer amaranth can quickly evolve life-history traits to adapt to different cultural practices, similar to observations in a Palmer amaranth response to nitrogen study [@bravo2018].
-Our results highlight Palmer amaranth as a threat to field crops as breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
+Our results highlight Palmer amaranth as a threat to field crops and breeding more competitive crop varieties is likely to select more competitive weed biotypes [@bravo2017].
 
@@
 
-Palmer amaranth growth and development in second cohort was limited due to the crop competitive ability at advanced development stages.
-Plants were transplanted at greater crop height and width, which reduced Palmer amaranth competitiveness.
-As a result, Palmer amaranth height and biomass was lower compared to its first cohort.
+Palmer amaranth growth and development in the second cohort was limited due to the crop competitive ability at advanced developmental stages.
+Palmer amaranth seedling were transplanted at greater crop height and width, which reduced Palmer amaranth competitiveness.
+As a result, Palmer amaranth height and biomass were lower compared to its first cohort.
 Moreover, Palmer amaranth growing without crop competition produced the highest amounts of biomass.
@@
 The *Amaranthus* species are sensitive to photoperiod [@wu2014].
-It is hypothesize that reduced day length or red/far-red effect contributed to smaller plants at second cohort regardless the crop.
+We hypothesize that reduced day length or red/far-red effect contributed to smaller plants at second cohort regardless the crop.
 A study in North Carolina and Illinois predicted that less than 10% Palmer amaranth seedlings emergence occurred after June [@piskackova2021].
-In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established near to crop planting.
-Therefore, early management is a key strategy to minimize the damaging impact of Palmer amaranth to US Midwest cropping systems.
+In addition, Palmer amaranth negative impact on soybean [@korres2020] and cotton [@webster2015] yields was higher when plants were established close to crop planting.
+Therefore, early season management is a key strategy to minimize the damaging impact of Palmer amaranth to US Midwest cropping systems.
 
 Seed production was not evaluated due to plant harvest at initiation of flowering.
-Nonetheless, it is well documented a strong positive correlation between Palmer amaranth biomass and seed production [@schwartz2016; @spaunhorst2018].
+Nonetheless, a strong positive correlation between Palmer amaranth biomass and seed production has been documented [@schwartz2016; @spaunhorst2018].
 In our study, plants growing from the first cohort accumulated 42% more biomass when compared to the second cohort.
-Therefore, Palmer amaranth plants growing in the second cohort is likely to produce less seeds regardless the crop.
+Therefore, Palmer amaranth plants growing in the second cohort are likely to produce fewer seeds regardless of crop.
 Our observation is consistent with the findings that the first Palmer amaranth cohort produced 50% more seeds per plant than Palmer amaranth plants established six weeks later in bareground [@webster2015].
@@
 Seed production and deposition in the seedbank is also a key factor for species perpetuation [@menges1987].
-Palmer amaranth can produce a hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
+Palmer amaranth can produce more than hundred thousands seeds per plant [@schwartz2016; @keeley1987], which can stay viable in the soil seedbank for at least 36 months [@sosnoskie2013].
 Therefore, preventing Palmer amaranth seed production or/and seed migration to non-native habitats is an essential strategy to minimize the species impact in agroecosystems.
 
-An ecological approach to reduce Palmer amaranth seed production is understanding plant biology, including flowering pattern.
+An ecological approach to reduce Palmer amaranth seed production encompasses understanding plant biology, including flowering pattern.
 Floral initiation depends on the complex interaction between a plant's genetic makeup and environmental conditions [@lang1965].
@@
 In some cases, early flowering may confer an evolutionary advantage, provided the plant has an indeterminate habit and flowering does not restrict plant growth.
-In an Iowa study, Palmer amaranth initiated flowering two weeks prior to waterhemp [@baker2021], and this characteristic may partially explain why Palmer amaranth can displace waterhemp where both occur in the same agronomic field.
 In other cases, however, early flower initiation may be a response to stressful conditions as a plant attempts to reproduce before running out of resources.
+In an Iowa study, Palmer amaranth initiated flowering two weeks prior to waterhemp [@baker2021], and this characteristic may partially explain why waterhemp can displace Palmer amaranth in Iowa, where Palmer amaranth is not well adapted.
 In a dioecious species like Palmer amaranth, exerting stress on plants to manipulate flowering may be beneficial in limiting seed production [@mcfarlane2018; @schliekelman2005].
@@
 In our study, the gender model performance was decent (AUC 0.64) considering the biology of plant flowering.
-A 1:1 male and female sex ratio is a general an evolutionarily stable strategy for plant species perpetuation [@fisher1930].
+A 1:1 male and female sex ratio is a general and evolutionarily stable strategy for plant species perpetuation [@fisher1930].
 However, a slight deviation from 1:1 sex ratio occurs in some dioecious species.
@@
 
-```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "160mm", out.height = "100mm"}
+```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "160mm", out.height = "100mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 ```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit', out.width = "150mm", out.height = "70mm"}
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit.', out.width = "150mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "90mm"}
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL', out.width = "150mm", out.height = "150mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 ```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
-Black line represents the model estimation and shaded green the confidence intervals', out.width = "170mm", out.height = "70mm"}
+Black line represents the model estimation and shaded green the confidence intervals.', out.width = "170mm", out.height = "70mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 ```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin', out.width = "150mm", out.height = "90mm"}
+From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.

### 2022-02-28 Manuscript/Manuscript.tex
as

--- 2022-02-18 Manuscript/Manuscript.tex
@@
 
-A Palmer amaranth accession (Kei3) from Perkins County, Nebraska with no
-glyphosate resistance was selected for this study (Oliveira et al.,
+A Palmer amaranth accession (Kei3) from Perkins County, Nebraska
+susceptible to glyphosate was selected for this study (Oliveira et al.,
 2021). Three weeks prior to the establishment of each cohort, seeds were
 planted in plastic trays containing potting-mix. Emerged seedlings (1
-cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (a
+cm) were transplanted into 200 cm\textsuperscript{-3} plastic pots (one
 plant pot\textsuperscript{-1}). Palmer amaranth seedlings were supplied
@@
 University of Wisconsin-Madison, University of Nebraska-Lincoln, and
-Western Illinois University; and kept outdoors at the Perkins extension
+Western Illinois University; and kept outdoors at the Perkins Extension
 office in Grant, NE until the 2-3 leaf stage (5 to 8 cm height) when
@@
 gender was identified (male or female), plant height was measured from
-soil surface to the top of plant. Also, aboveground plant, and all
+soil surface to the top of plant. Also, aboveground plant, and the
 aboveground plant organs were harvested, then oven dried at 65 C until a
@@
 
-The initiation and duration of Palmer amaranth flowering was strongly
+The initiation and duration of Palmer amaranth flowering were strongly
 influenced by cohort and surrounding vegetation (Figure 4A, 4B). In the
@@
 216.8, respectively). Palmer amaranth growing in the bareground had the
-shortest duration of floral initiation (measured as the difference
-between 10\% and 90\% cumulative flowering) in both the first (34 days)
-and second (28 days) cohorts. Palmer amaranth growing in soybean had a
-shorter duration of flowering than corn in the first cohort (40 days vs
-71 days), but a longer duration in the second cohort (50 vs 44 days).
-
-Palmer amaranth cumulative flowering in the second cohort ranged from
+shortest flowering initiation (measured as the difference between 10\%
+and 90\% cumulative flowering) in both the first (34 days) and second
+(28 days) cohorts. Palmer amaranth growing in soybean had a shorter
+duration of flowering than corn in the first cohort (40 days vs 71
+days), but a longer duration in the second cohort (50 vs 44 days).
+
+Palmer amaranth cumulative flowering in the second cohort occurred from
 mid July to mid September (Figure 4B). Palmer amaranth growing in the
@@
 probability of being female increased 0.1\% and 0.2\% for each unit
-increase of weight (g) and height (cm), respectively (Table 2).
+increase of weight and height, respectively (Table 2).
 
@@
 
-Our study confirmed that Palmer amaranth is well adapted to growing
-throughout the Midwestern United States, and is simply limited in its
-range by seed dispersal (Davis et al.~2015). We also confirmed Palmer
-amaranth's extraordinary plasticity to adapt to different
-agroecosystems. For example, Palmer amaranth mimicked crop architecture
-in competing for light (Figure 6), where plants growing in corn
-allocated resources to height, while plants growing in bareground
+Our study demonstrate that Palmer amaranth is well adapted to growing
+conditions throughout the Midwestern United States, and is simply
+limited in its range by seed dispersal (Davis et al.~2015). We also
+confirmed Palmer amaranth's extraordinary plasticity to adapt to
+different agroecosystems. For example, Palmer amaranth mimicked crop
+architecture in competing for light (Figure 6), where plants growing in
+corn allocated resources to height, while plants growing in bareground
 allocated resources to numerous branches. Plants growing in the absence
@@
 al., 2018). Our results highlight Palmer amaranth as a threat to field
-crops as breeding more competitive crop varieties is likely to select
+crops and breeding more competitive crop varieties is likely to select
 more competitive weed biotypes (Bravo et al., 2017).
 
-Palmer amaranth growth and development in second cohort was limited due
-to the crop competitive ability at advanced development stages. Plants
-were transplanted at greater crop height and width, which reduced Palmer
-amaranth competitiveness. As a result, Palmer amaranth height and
-biomass was lower compared to its first cohort. Moreover, Palmer
-amaranth growing without crop competition produced the highest amounts
-of biomass. The Palmer amaranth strategy in bareground was to invest
-biomass in growing plant width and height. Nonetheless, Palmer amaranth
-produced 17\% less biomass in second cohort compared to first cohort
-timing. In a bareground study, early emerged Palmer amaranth without
-competition was 50\% taller than late emerged plants (Webster and Grey,
-2015). These results suggest that crop competition is not the only
-factor limiting late Palmer amaranth establishment. The limited growth
-of Palmer amaranth at second cohort is likely a reduced plant response
-to day length, light availability and thermal units (e.g, growing degree
-days). The \emph{Amaranthus} species are sensitive to photoperiod (Wu
-and Owen, 2014). It is hypothesize that reduced day length or
-red/far-red effect contributed to smaller plants at second cohort
-regardless the crop. A study in North Carolina and Illinois predicted
-that less than 10\% Palmer amaranth seedlings emergence occurred after
-June (Piskackova et al., 2021). In addition, Palmer amaranth negative
-impact on soybean (Korres et al., 2020) and cotton (Webster and Grey,
-2015) yields was higher when plants were established near to crop
-planting. Therefore, early management is a key strategy to minimize the
-damaging impact of Palmer amaranth to US Midwest cropping systems.
+Palmer amaranth growth and development in the second cohort was limited
+due to the crop competitive ability at advanced developmental stages.
+Palmer amaranth seedling were transplanted at greater crop height and
+width, which reduced Palmer amaranth competitiveness. As a result,
+Palmer amaranth height and biomass were lower compared to its first
+cohort. Moreover, Palmer amaranth growing without crop competition
+produced the highest amounts of biomass. The Palmer amaranth strategy in
+bareground was to invest biomass in growing plant width and height.
+Nonetheless, Palmer amaranth produced 17\% less biomass in second cohort
+compared to first cohort timing. In a bareground study, early emerged
+Palmer amaranth without competition was 50\% taller than late emerged
+plants (Webster and Grey, 2015). These results suggest that crop
+competition is not the only factor limiting late Palmer amaranth
+establishment. The limited growth of Palmer amaranth at second cohort is
+likely a reduced plant response to day length, light availability and
+thermal units (e.g, growing degree days). The \emph{Amaranthus} species
+are sensitive to photoperiod (Wu and Owen, 2014). We hypothesize that
+reduced day length or red/far-red effect contributed to smaller plants
+at second cohort regardless the crop. A study in North Carolina and
+Illinois predicted that less than 10\% Palmer amaranth seedlings
+emergence occurred after June (Piskackova et al., 2021). In addition,
+Palmer amaranth negative impact on soybean (Korres et al., 2020) and
+cotton (Webster and Grey, 2015) yields was higher when plants were
+established close to crop planting. Therefore, early season management
+is a key strategy to minimize the damaging impact of Palmer amaranth to
+US Midwest cropping systems.
 
 Seed production was not evaluated due to plant harvest at initiation of
-flowering. Nonetheless, it is well documented a strong positive
-correlation between Palmer amaranth biomass and seed production
-(Schwartz et al., 2016; Spaunhorst et al., 2018). In our study, plants
-growing from the first cohort accumulated 42\% more biomass when
-compared to the second cohort. Therefore, Palmer amaranth plants growing
-in the second cohort is likely to produce less seeds regardless the
-crop. Our observation is consistent with the findings that the first
-Palmer amaranth cohort produced 50\% more seeds per plant than Palmer
-amaranth plants established six weeks later in bareground (Webster and
-Grey, 2015). Still, seed production at the second cohort will replenish
-the soil seedbank. Seed production and deposition in the seedbank is
-also a key factor for species perpetuation (Menges, 1987). Palmer
-amaranth can produce a hundred thousands seeds per plant (Schwartz et
-al., 2016; Keeley et al., 1987), which can stay viable in the soil
-seedbank for at least 36 months (Sosnoskie et al., 2013). Therefore,
-preventing Palmer amaranth seed production or/and seed migration to
-non-native habitats is an essential strategy to minimize the species
-impact in agroecosystems (Davis et al., 2015).
-
+flowering. Nonetheless, a strong positive correlation between Palmer
+amaranth biomass and seed production has been documented (Schwartz et
+al., 2016; Spaunhorst et al., 2018). In our study, plants growing from
+the first cohort accumulated 42\% more biomass when compared to the
+second cohort. Therefore, Palmer amaranth plants growing in the second
+cohort are likely to produce fewer seeds regardless of crop. Our
+observation is consistent with the findings that the first Palmer
+amaranth cohort produced 50\% more seeds per plant than Palmer amaranth
+plants established six weeks later in bareground (Webster and Grey,
+2015). Still, seed production at the second cohort will replenish the
+soil seedbank. Seed production and deposition in the seedbank is also a
+key factor for species perpetuation (Menges, 1987). Palmer amaranth can
+produce more than hundred thousands seeds per plant (Schwartz et al.,
+2016; Keeley et al., 1987), which can stay viable in the soil seedbank
+for at least 36 months (Sosnoskie et al., 2013). Therefore, preventing
+Palmer amaranth seed production or/and seed migration to non-native
+habitats is an essential strategy to minimize the species impact in
+agroecosystems.
+
+An ecological approach to reduce Palmer amaranth seed production
+encompasses understanding plant biology, including flowering pattern.
 Floral initiation depends on the complex interaction between a plant's
@@
 confer an evolutionary advantage, provided the plant has an
-indeterminate habit and flowering does not restrict plant growth. In an
-Iowa study, Palmer amaranth initiated flowering two weeks prior to
-waterhemp (Baker, 2021), and this characteristic may partially explain
-why Palmer amaranth can displace waterhemp where both occur in the same
-agronomic field. In other cases, however, early flower initiation may be
-a response to stressful conditions as a plant attempts to reproduce
-before running out of resources. In a dioecious species like Palmer
-amaranth, exerting stress on plants to manipulate flowering may be
-beneficial in limiting seed production (McFarlane et al., 2018;
+indeterminate habit and flowering does not restrict plant growth. In
+other cases, however, early flower initiation may be a response to
+stressful conditions as a plant attempts to reproduce before running out
+of resources. In an Iowa study, Palmer amaranth initiated flowering two
+weeks prior to waterhemp (Baker, 2021), and this characteristic may
+partially explain why waterhemp can displace Palmer amaranth in Iowa,
+where Palmer amaranth is not well adapted. In a dioecious species like
+Palmer amaranth, exerting stress on plants to manipulate flowering may
+be beneficial in limiting seed production (McFarlane et al., 2018;
 Schliekelman et al., 2005). For example, when growing under water
@@
 decent (AUC 0.64) considering the biology of plant flowering. A 1:1 male
-and female sex ratio is a general an evolutionarily stable strategy for
+and female sex ratio is a general and evolutionarily stable strategy for
 plant species perpetuation (Fisher, 1930). However, a slight deviation
@@
 
-\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-1}
+\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-1}
 \end{figure}
@@
 \caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit}\label{fig:Figure-2}
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit.}\label{fig:Figure-2}
 \end{figure}
@@
 
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-3}
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-3}
 \end{figure}
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-4}
 \end{figure}
@@
 \caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
-Black line represents the model estimation and shaded green the confidence intervals}\label{fig:Figure-5}
+Black line represents the model estimation and shaded green the confidence intervals.}\label{fig:Figure-5}
 \end{figure}
@@
 \caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin}\label{fig:Figure-6}
+From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.}\label{fig:Figure-6}
 \end{figure}

### 2022-02-28 data analysis/Untitled.Rmd
SD

---
title: "map"
author: "Maxwel Coura Oliveira"
date: "2/28/2022"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
```

```{r}
library(rnaturalearth)
library(rnaturalearthdata)
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```

```{r}
(sites_1 <- tibble(location = c("Lincoln, NE", "Macomb, IL", "Grant, NE", "Arlington, WI", "Clay Center, NE"),
                 longitude = c(-96.7026, -90.689, -101.70, -89.38, -98.14), 
                 latitude = c(40.8136, 40.489, 40.85, 43.33, 40.57)))
```

```{r}
(sites <- data.frame(location = c("Lincoln, NE", "Macomb, IL", "Grant, NE", "Arlington, WI", "Clay Center, NE"),
                     longitude = c(-96.7026, -90.689, -101.70, -89.38, -98.14), 
                     latitude = c(40.8136, 40.489, 40.85, 43.33, 40.57)))

# #here WGS84, which is the CRS code #4326
(sites <- st_as_sf(sites, coords = c("longitude", "latitude"), 
    crs = 4326, agr = "constant"))
```

```{r}
library(tools)
states$ID <- toTitleCase(states$ID)
head(states)
```

```{r}
#  US National Atlas Equal Area projection (CRS 2163)
usa <- subset(world, admin == "United States of America")
(mainland <- ggplot(data = usa) +
    geom_sf(fill = "cornsilk") +
    geom_sf(data = sites, size = 3, shape = c(15, 16, 17, 18, 19)) +
    coord_sf(crs = st_crs(2163), xlim = c(-2500000, 2500000), ylim = c(-2300000, 
         730000)))
```

```{r}
# Alaska Albers projection (CRS 3467)
(alaska <- ggplot(data = usa) +
     geom_sf(fill = "cornsilk") +
     coord_sf(crs = st_crs(3467), xlim = c(-2400000, 1600000), ylim = c(200000, 
         2500000), expand = FALSE, datum = NA))
```

```{r}
# Old Hawaiian projection (CRS 4135)
(hawaii  <- ggplot(data = usa) +
     geom_sf(fill = "cornsilk") +
     coord_sf(crs = st_crs(4135), xlim = c(-161, -154), ylim = c(18, 
         23), expand = FALSE, datum = NA))
```

```{r}
library(grid)
library(ggrepel)
mainland +
 annotation_custom(
      grob = ggplotGrob(alaska),
      xmin = -2750000,
      xmax = -2750000 + (1600000 - (-2400000))/2.5,
      ymin = -2450000,
      ymax = -2450000 + (2500000 - 200000)/2.5
  ) +
  annotation_custom(
      grob = ggplotGrob(hawaii),
      xmin = -1250000,
      xmax = -1250000 + (-154 - (-161))*120000,
      ymin = -2450000,
      ymax = -2450000 + (23 - 18)*120000
  ) +
  annotation_custom(textGrob(label="Grant, NE", x = 0.40, y = 0.605)) +
  annotation_custom(textGrob(label="Arlington, WI", x = 0.65, y = 0.75)) +
  annotation_custom(textGrob(label="Macomb, IL", x = 0.73, y = 0.60)) +
  annotation_custom(textGrob(label="Lincoln, NE", x = 0.60, y = 0.65)) +
  annotation_custom(textGrob(label="Clay Center, NE", x = 0.58, y = 0.55)) +
  annotation_scale(location = "br", width_hint = 0.25) +
  annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in"),
        style = north_arrow_nautical) 
 # geom_sf_text(data = sites, aes(label = location))

ggsave("figures/map.png", width = 7, heigh = 4)
```

```{r}
library(cowplot)
# Using cowplot
(ratioAlaska <- (2500000 - 200000) / (1600000 - (-2400000)))
(ratioHawaii  <- (23 - 18) / (-154 - (-161)))
ggdraw(mainland) +
    draw_plot(alaska, width = 0.26, height = 0.26 * 10/6 * ratioAlaska, 
        x = 0.05, y = 0.05) +
    draw_plot(hawaii, width = 0.15, height = 0.15 * 10/6 * ratioHawaii, 
        x = 0.3, y = 0.05) +
  annotate("text", label = "Arlington, WI", x = 90, y = 40)
```

### 2022-03-01 Manuscript/Manuscript.Rmd
as

@@
 Moreover, Palmer amaranth has shown a remarkable capacity to evolve resistance to herbicides.
-To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2021], increasing the weed management complexity [@lindsay2017] and posing an economical and ecological risk to row-crop agriculture.
+To date, Palmer amaranth has evolved resistance to eight herbicide sites of action [@heap2022], increasing the weed management complexity [@lindsay2017] and posing an economical and ecological risk to row-crop agriculture.
 
@@
 
-The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE (40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N, 96.42'W), and Macomb, IL (40.49'N, -90.69'W).
+The experiment was conducted in 2018 and 2019 under field conditions at five locations: Arlington, WI, Clay Center, NE,, Grant, NE, Lincoln, NE, and Macomb, IL.
+
+```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 1.png")
+```
 
@@
 Corn and soybean were planted in 76-cm row spacing (Table 1).
-Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 1).
+Monthly mean air temperature and total precipitation were obtained using Daymet weather data from June through September across the five locations in 2018 and 2019 [@correndo2021] (Figure 2).
 
-```{r fig1, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/weather/Figure 1.png")
+```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/weather/Figure 2.png")
 ```
@@
 Forty-eight plants were equidistantly placed (0.76 m apart) between rows within each crop.
-After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 2).
+After a week, one plant was eliminated and one was kept, resulting in 24 plants per experimental unit and transplanting time (Figure 3).
 When needed, Palmer amaranth plants were supplied with water during the first week after transplanting to assure seedling survival.
 
-```{r fig2, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 2.png")
+```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 3.png")
 ```
@@
 
-Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (Figure 3A).
+Palmer amaranth plants accumulated more biomass when growing in bareground compared to plants growing in soybean and corn (Figure 4A).
 Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3 g plant\textsuperscript{-1} in bareground, soybean and corn, respectively.
@@
 
-```{r fig3, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 3.png")
+```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 4.png")
 ```
 
-Palmer amaranth height was less affected by cohort timing than was weight, with the exception of plants growing in corn (Figure 3B).
+Palmer amaranth height was less affected by cohort timing than was weight, with the exception of plants growing in corn (Figure 4B).
 Plants from the first cohort were on average 69.2 cm tall in bareground, which was not different from the 70.7 cm tall plants from the second cohort timing (P = 0.74).
@@
 
-The initiation and duration of Palmer amaranth flowering were strongly influenced by cohort and surrounding vegetation (Figure 4A, 4B).
+The initiation and duration of Palmer amaranth flowering were strongly influenced by cohort and surrounding vegetation (Figure 5A, 4B).
 In the first cohort, floral initiation (10% flowering) occurred near the end of June for all three treatments, at doy 180, 180.9, and 181.7 for soybean, bareground, and corn, respectively.
@@
 
-```{r fig4, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 4.png")
+```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 5.png")
 ```
 
-Palmer amaranth cumulative flowering in the second cohort occurred from mid July to mid September (Figure 4B).
+Palmer amaranth cumulative flowering in the second cohort occurred from mid July to mid September (Figure 5B).
 Palmer amaranth growing in the bareground resulted in earlier flowering time compared to soybean and corn.
@@
 Moreover, increasing a cumulative flowering unit doy increases the probability of having a female plant by 0.4% (Table 2 and Figure5A).
-A similar trend was observed for weight (Figure 5B) and height (Figure 5C), where the probability of being female increased 0.1% and 0.2% for each unit increase of weight and height, respectively (Table 2).
+A similar trend was observed for weight (Figure 6B) and height (Figure 6C), where the probability of being female increased 0.1% and 0.2% for each unit increase of weight and height, respectively (Table 2).
 
@@
   kableExtra::kable(booktabs = TRUE, escape = TRUE, caption = "Average marginal effects of Palmer amaranth gender logistic model.
-Factor pararemter values (crop_corn and 'crop_bareground) are calculated relative to soybean.") %>%
+Factor pararemter values (crop_corn and crop_bareground) are calculated relative to soybean.") %>%
   kable_styling(font_size = 10, 
@@
 
-```{r fig5, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 5.png")
+```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 6.png")
 ```
@@
 We also confirmed Palmer amaranth's extraordinary plasticity to adapt to different agroecosystems.
-For example, Palmer amaranth mimicked crop architecture in competing for light (Figure 6), where plants growing in corn allocated resources to height, while plants growing in bareground allocated resources to numerous branches.
-Plants growing in the absence of a crop produced the greatest biomass, responding to a greater abundance of light, nutrient and water resources (Figure 3A, 3B).
+For example, Palmer amaranth mimicked crop architecture in competing for light (Figure 7), where plants growing in corn allocated resources to height, while plants growing in bareground allocated resources to numerous branches.
+Plants growing in the absence of a crop produced the greatest biomass, responding to a greater abundance of light, nutrient and water resources (Figure 4A, 3B).
 These results support the argument that Palmer amaranth can quickly evolve life-history traits to adapt to different cultural practices, similar to observations in a Palmer amaranth response to nitrogen study [@bravo2018].
@@
 
-```{r fig6, echo = FALSE, message=FALSE, fig.show = "hide"}
-knitr::include_graphics("../data analysis/figures/Figure 6.png")
+```{r fig7, echo = FALSE, message=FALSE, fig.show = "hide"}
+knitr::include_graphics("../data analysis/figures/Figure 7.png")
 ```
@@
 Where Palmer amaranth occurs, management tactics should focus on limiting Palmer amaranth establishment until row crops can shade late-emerging plants.
-Increasing the diversity of crops in rotation, varying row crop planting date, narrowing row width, and increasing residue cover through the use of cover crops are all tactics that can minimize the growth and seed production of Palmer amaranth, and improve the sustainability of Midwestern cropping systems.
+Increasing the diversity of crops in rotation, varying row crop planting date, narrowing row width, and increasing residue cover through the use of cover crops are all tactics that can minimize the growth and seed production of Palmer amaranth, and improve the sustainability of North Central United States cropping systems.
 
@@
 
-```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "160mm", out.height = "100mm"}
+```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Field research locations in the United States north central region.', out.width = "160mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
+```{r, Figure-2, ref.label = "fig2", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "160mm", out.height = "100mm"}
+# You can also refer to code chunks from above to place figures at the bottom.
+```
+
+```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
 Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit.', out.width = "150mm", out.height = "70mm"}
@@
 
-```{r, Figure-3, ref.label = "fig3", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "90mm"}
+```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-4, ref.label = "fig4", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "150mm"}
+```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.', out.width = "150mm", out.height = "150mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-5, ref.label = "fig5", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
+```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
 Black line represents the model estimation and shaded green the confidence intervals.', out.width = "170mm", out.height = "70mm"}
@@
 
-```{r, Figure-6, ref.label = "fig6", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
+```{r, Figure-7, ref.label = "fig7", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (first manuscript author, 180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
 From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.', out.width = "150mm", out.height = "90mm"}

### 2022-03-01 Manuscript/Manuscript.tex
as

@@
 herbicides. To date, Palmer amaranth has evolved resistance to eight
-herbicide sites of action (Heap, 2021), increasing the weed management
+herbicide sites of action (Heap, 2022), increasing the weed management
 complexity (Lindsay et al., 2017) and posing an economical and
@@
 The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington, WI (43°18'N, 89°29'W), Clay Center, NE
-(40.57'N, 9814'W), Grant, NE (40.85'N, -101.70'W), Lincoln, NE (41.16'N,
-96.42'W), and Macomb, IL (40.49'N, -90.69'W).
+five locations: Arlington, WI, Clay Center, NE,, Grant, NE, Lincoln, NE,
+and Macomb, IL.
 
@@
 data from June through September across the five locations in 2018 and
-2019 (Correndo et al., 2021) (Figure 1).
+2019 (Correndo et al., 2021) (Figure 2).
 
@@
 
-\caption{\label{tab:unnamed-chunk-2}Field study attributes }
+\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}
 \centering
@@
 
- & Weed control & glyphosate & saflufenacil + imazethapyr + pyroxasulfone &  &  & \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c}  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate\textsuperscript{c} & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor\textsuperscript{f} \\
 
@@
 
- & Hybrid & NK0142 3120-EZ1 & DKC60-67 &  &  & \\
-
- & Seeding rate & 88956 & 86487 &  &  & \\
-
- & Weed control & glyphosate / S-metolachor & S-metolachlor + trazine + mesotrione, + bicyclopyrone &  &  & \\
-
- & Stage at 1 cohort & V2-3 &  &  &  & \\
-
- & Stage at 2 cohort & V6-7 &  &  &  & \\
-
- & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 &  &  & \\
-
-\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha-1 &  &  &  & \\
+ & Hybrid & NK0142 3120-EZ1 & DKC60-67 & G03C84-5122-EZ1 & P1197AM & DKC64-34RIB (2018) / G10T63-3000GT (2019) \\
+
+ & Seeding rate & 88956 & 86487 & 79000 & 86485 & 88,000 (2018) / 79,000 (2019)\\
+
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
+
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V2-3 & V2-3 & V2-3 & V4-5\\
+
+ & Stage at 2\textsuperscript{nd} cohort & V6-7 & V6-7 & V6-7 & V6-7 & V10-11\\
+
+ & Planting day & April 30, 2018 / May 5, 2019 & May 10, 2018/19 & May 5, 2018 / May 14, 2019 & April 27, 2018 / May 3, 2019 & May 8, 2018 / April 26, 2019 \\
+
+\multirow{-7}{*}{\raggedright\arraybackslash } & Fertilization & N (46-0-0) at 157 kg ha\textsuperscript{-1} &  &  & N (46-0-0) at 336 kg ha\textsuperscript{-1} & UAN-32, 202 kg N ha\textsuperscript{-1}\\
 
@@
 crop. After a week, one plant was eliminated and one was kept, resulting
-in 24 plants per experimental unit and transplanting time (Figure 2).
+in 24 plants per experimental unit and transplanting time (Figure 3).
 When needed, Palmer amaranth plants were supplied with water during the
@@
 Palmer amaranth plants accumulated more biomass when growing in
-bareground compared to plants growing in soybean and corn (Figure 3A).
+bareground compared to plants growing in soybean and corn (Figure 4A).
 Palmer amaranth plants in the first cohort produced 75.5, 28.3, and 16.3
@@
 Palmer amaranth height was less affected by cohort timing than was
-weight, with the exception of plants growing in corn (Figure 3B). Plants
+weight, with the exception of plants growing in corn (Figure 4B). Plants
 from the first cohort were on average 69.2 cm tall in bareground, which
@@
 The initiation and duration of Palmer amaranth flowering were strongly
-influenced by cohort and surrounding vegetation (Figure 4A, 4B). In the
+influenced by cohort and surrounding vegetation (Figure 5A, 4B). In the
 first cohort, floral initiation (10\% flowering) occurred near the end
@@
 Palmer amaranth cumulative flowering in the second cohort occurred from
-mid July to mid September (Figure 4B). Palmer amaranth growing in the
+mid July to mid September (Figure 5B). Palmer amaranth growing in the
 bareground resulted in earlier flowering time compared to soybean and
@@
 female plant by 0.4\% (Table 2 and Figure5A). A similar trend was
-observed for weight (Figure 5B) and height (Figure 5C), where the
+observed for weight (Figure 6B) and height (Figure 6C), where the
 probability of being female increased 0.1\% and 0.2\% for each unit
@@
 
-\textbackslash begin\{table\}{[}!h{]}
-
-\textbackslash caption\{\label{tab:unnamed-chunk-3}Average marginal
-effects of Palmer amaranth gender logistic model. Factor pararemter
-values (crop\_corn and 'crop\_bareground) are calculated relative to
-soybean.\} \centering \fontsize{10}{12}\selectfont
-
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop\_bareground and crop\_corn) is shown compared to soybean.}
+\centering
+\fontsize{10}{12}\selectfont
 \begin{tabular}[t]{lrrrrrr}
 \toprule
-Term & AME & SE & Lower & Upper & Z-score & P-value\\
+Term & AME\textsuperscript{a} & SE\textsuperscript{b} & Lower & Upper & Z-score & P-value\\
 \midrule
@@
 different agroecosystems. For example, Palmer amaranth mimicked crop
-architecture in competing for light (Figure 6), where plants growing in
+architecture in competing for light (Figure 7), where plants growing in
 corn allocated resources to height, while plants growing in bareground
@@
 of a crop produced the greatest biomass, responding to a greater
-abundance of light, nutrient and water resources (Figure 3A, 3B). These
+abundance of light, nutrient and water resources (Figure 4A, 3B). These
 results support the argument that Palmer amaranth can quickly evolve
@@
 production of Palmer amaranth, and improve the sustainability of
-Midwestern cropping systems.
+North Central United States cropping systems.
 
@@
 
-\leavevmode\hypertarget{ref-heap2021}{}%
-Heap, I. (2021). Internation {Herbicide}-{Resistant Weed Database}.
+\leavevmode\hypertarget{ref-heap2022}{}%
+Heap, I. (2022). Internation {Herbicide}-{Resistant Weed Database}.
 Available at: \url{http://www.weedscience.org/Home.aspx} {[}Accessed
-July 26, 2021{]}.
+January 26, 2022{]}.
 
@@
 \leavevmode\hypertarget{ref-rcoreteam2021}{}%
-Team, R. C. (2021). \emph{R: {The R Project} for {Statistical
+R Core Team (2021). \emph{R: {The R Project} for {Statistical
 Computing}}. {Vienna, Austria}: {R Foundation for Statistical Computing}
@@
 
-{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 1}
+{\centering \includegraphics[width=160mm,height=90mm]{../data analysis/figures/Figure 1}
 
@@
 
-\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-1}
+\caption{Field research locations in the United States North Central region.}\label{fig:Figure-1}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 2}
+{\centering \includegraphics[width=160mm,height=100mm]{../data analysis/weather/Figure 2}
 
@@
 
+\caption{Mean average temperature (C) and total montly precipitation (mm) at Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-2}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=150mm,height=70mm]{../data analysis/figures/Figure 3}
+
+}
+
 \caption{Palmer amaranth adaptation study layout of a plant cohort timing in bareground, corn, and soybean.
-Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit.}\label{fig:Figure-2}
+Twenty-four Palmer amaranth plants were place 76.2 cm apart in each field experimental unit.}\label{fig:Figure-3}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 3}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 4}
 
@@
 
-\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-3}
+\caption{Palmer amaranth biomass (A) and height (B) growing in corn, bareground, and soybean nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-4}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 4}
+{\centering \includegraphics[width=150mm,height=150mm]{../data analysis/figures/Figure 5}
 
@@
 
-\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-4}
+\caption{Cumulative flowering of Palmer amaranth at first and second transplant timing (A) and day of year of 10, 50, and 90 cumulative flowering at first and second cohort transplanting time (B) nested across Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}\label{fig:Figure-5}
 \end{figure}
@@
 
-{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 5}
+{\centering \includegraphics[width=170mm,height=70mm]{../data analysis/figures/Figure 6}
 
@@
 \caption{The probability (P) of being female Palmer amaranth by day of year (A), weight (B), and height (C).
-Black line represents the model estimation and shaded green the confidence intervals.}\label{fig:Figure-5}
+Black line represents the model estimation and shaded green the confidence intervals.}\label{fig:Figure-6}
 \end{figure}
@@
 
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 6}
+{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 7}
 
@@
 
-\caption{MCO (180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.}\label{fig:Figure-6}
+\caption{MCO (first manuscript author, 180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
+From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.}\label{fig:Figure-7}
 \end{figure}

### 2022-03-01 data analysis/Untitled.Rmd
as

@@
     geom_sf(fill = "cornsilk") +
-    geom_sf(data = sites, size = 3, shape = c(15, 16, 17, 18, 19)) +
+    geom_sf(data = sites, size = 2, shape = c(15, 16, 17, 18, 19)) +
     coord_sf(crs = st_crs(2163), xlim = c(-2500000, 2500000), ylim = c(-2300000, 
@@
 library(ggrepel)
+library(ggspatial)
+library(ggsflabel)
+library(cowplot)
 mainland +
@@
   ) +
-  annotation_custom(textGrob(label="Grant, NE", x = 0.40, y = 0.605)) +
-  annotation_custom(textGrob(label="Arlington, WI", x = 0.65, y = 0.75)) +
-  annotation_custom(textGrob(label="Macomb, IL", x = 0.73, y = 0.60)) +
-  annotation_custom(textGrob(label="Lincoln, NE", x = 0.60, y = 0.65)) +
-  annotation_custom(textGrob(label="Clay Center, NE", x = 0.58, y = 0.55)) +
+#  annotation_custom(textGrob(label="Grant, NE", x = 0.40, y = 0.605)) +
+#  annotation_custom(textGrob(label="Arlington, WI", x = 0.65, y = 0.75)) +
+#  annotation_custom(textGrob(label="Macomb, IL", x = 0.73, y = 0.60)) +
+#  annotation_custom(textGrob(label="Lincoln, NE", x = 0.60, y = 0.65)) +
+#  annotation_custom(textGrob(label="Clay Center, NE", x = 0.58, y = 0.55)) +
   annotation_scale(location = "br", width_hint = 0.25) +
@@
         pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in"),
-        style = north_arrow_nautical) 
- # geom_sf_text(data = sites, aes(label = location))
+        style = north_arrow_nautical) +
+  geom_sf_text_repel(data = sites, aes(label = location), size = 2.5,
+                     min.segment.length = 1, point.padding = 0.5, 
+                     box.padding = 0.25,
+                     max.overlaps = 10,
+                     seed = 564,
+                     max.iter = 10000,
+                     segment.color = 'grey50') +
+  labs(x= NULL, y = NULL)
 
@@
 ```
-
-```{r}
-library(cowplot)
-# Using cowplot
-(ratioAlaska <- (2500000 - 200000) / (1600000 - (-2400000)))
-(ratioHawaii  <- (23 - 18) / (-154 - (-161)))
-ggdraw(mainland) +
-    draw_plot(alaska, width = 0.26, height = 0.26 * 10/6 * ratioAlaska, 
-        x = 0.05, y = 0.05) +
-    draw_plot(hawaii, width = 0.15, height = 0.15 * 10/6 * ratioHawaii, 
-        x = 0.3, y = 0.05) +
-  annotate("text", label = "Arlington, WI", x = 90, y = 40)
-```

### 2022-03-01 data analysis/map.Rmd
as

--- 2022-03-01 data analysis/Untitled.Rmd
@@
     geom_sf(fill = "cornsilk") +
-    geom_sf(data = sites, size = 2, shape = c(15, 16, 17, 18, 19)) +
+    geom_sf(data = sites, size = 2, shape = 16) +
     coord_sf(crs = st_crs(2163), xlim = c(-2500000, 2500000), ylim = c(-2300000, 
@@
 
-ggsave("figures/map.png", width = 7, heigh = 4)
+ggsave("figures/Figure 1.png", width = 7, heigh = 4)
 ```

### 2022-04-01 Manuscript/Manuscript.Rmd
jhjh

@@
 Palmer amaranth dispersal warrants studies assessing species adaptation into new geographies.
-A study was conducted to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations in the US Midwest.
+A study was conducted in 2018 and 2019 to investigate morphology, flowering and gender from cohorts of Palmer amaranth growing under corn, soybean, and bareground across five locations in the Midwest United States.
 Results demonstrated that the first cohort of Palmer amaranth, established in June, produced 42\% more biomass than plants from the second cohort (established in July).
@@
 Palmer amaranth has an extended germination period throughout the growing season [@ward2013].
-Germination of Palmer amaranth was triggered by 18 C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 C day and night temperatures [@guo2003].
+Germination of Palmer amaranth was triggered by 18 \textdegree C soil temperature at 5 cm depth [@keeley1987], and optimal germination and biomass production occurred at 35/30 \textdegree C day and night temperatures [@guo2003].
 Palmer amaranth emergence can be influenced by tillage and the use of preemergence herbicides [@chahal2021], and may result in weed germination shifts within a population, as documented in *Bassia scoparia* [@sbatella2010].
@@
 
-Palmer amaranth's current global range includes agronomic cropland in Italy [@milani2021], soybean producing regions of Brazil and Argentina [@larran2017; @kupper2017], and the Cotton Belt [@garetson2019; @bagavathiannan2016] and southern Great Plains of the United States [@kumar2020, @crespo2016a].
-However, its range appears to be expanding steadily, and in some cases rapidly, in the U.S.
-For example, in states with large Palmer amaranth infestations its range is gradually pushing north as it displaces other weeds in agronomic row crops (R Werle, personal observation), aided by wildlife [@farmer2017] and agricultural machinery [@ward2013].
+Palmer amaranth's current global range includes agronomic cropland in Italy [@milani2021], soybean producing regions of Brazil and Argentina [@larran2017; @kupper2017], and the Cotton Belt [@garetson2019; @bagavathiannan2016] and southern Great Plains of the United States [@kumar2020; @crespo2016a].
+However, its range appears to be expanding steadily, and in some cases rapidly, in the United States.
+We hypothesize that in states with large Palmer amaranth infestations its range is gradually pushing north as it displaces other weeds in agronomic row crops, aided by wildlife [@farmer2017] and agricultural machinery [@ward2013].
 But there are other examples where it has suddenly become a problem weed hundreds of miles away from known infestations.
@@
 
-There are many areas in the U.S.
-Corn Belt (North Central states) where Palmer amaranth is not yet established, and its potential adaptability is untested.
-Nonetheless, the rapid expansion of Palmer amaranth across the Northern U.S.
-is concerning and warrants investigations on its adaptability.
-We do know that  Palmer amaranth caused yield loss in Illinois soybean fields [@davis2015], but an Iowa study showed that Palmer amaranth was not as well adapted as waterhemp (*Amaranthus tuberculatus*) to conditions in that geography [@baker2021].
-We also know that Palmer amaranth plants that establish shortly after row-crop planting have a much greater impact on crop yield than plants that emerge after the crop has produced several leaves [@macrae2013].
+There are many areas in the US Corn Belt (North Central states) where Palmer amaranth is not yet established, and its potential adaptability is untested.
+Nonetheless, the rapid expansion of Palmer amaranth across the Northern United States is concerning and warrants investigations on its adaptability.
+It has been shown that Palmer amaranth caused yield loss in Illinois soybean fields [@davis2015], but an Iowa study showed that Palmer amaranth was not as well adapted as waterhemp (*Amaranthus tuberculatus*) to conditions in that geography [@baker2021].
+It also known that Palmer amaranth plants that establish shortly after row-crop planting have a much greater impact on crop yield than plants that emerge after the crop has produced several leaves [@macrae2013].
 Understanding Palmer amaranth morphology and development under different agroecosystems and across a wide geography can enhance our knowledge of its adaptability, and may also aid in designing effective tactics to limit is range expansion and minimize its negative effects on row crops.
@@
 When a plant flowered, the day was recorded, plant gender was identified (male or female), plant height was measured from soil surface to the top of plant.
-Also, aboveground plant, and the aboveground plant organs were harvested, then oven dried at 65 C until a constant weight was reached, and dry biomass (g plant\textsuperscript{-1}) was recorded.
-Plants were harvested at flowering because Palmer amaranth infestations do not occur at the Wisconsin or Illinois research locations.
+Also, aboveground plant, and the aboveground plant organs were harvested, then oven dried at 65 \textdegree C until a constant weight was reached, and dry biomass (g plant\textsuperscript{-1}) was recorded.
+Plants were harvested at flowering because Palmer amaranth is not established at the Wisconsin or Illinois research locations.
 In our study, all locations followed the methodology of plant harvest at flowering initiation, except in Grant, NE.
@@
 The model goodness of fit was 0.23, 0.32, 0.40 using pseudo R-squared test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
-The likelihood ratio test showed a p-value of < 0.00.
+The likelihood ratio test showed a p-value of 0.00.
 The average marginal effects showed that Palmer amaranth growing in corn resulted in 14.8% fewer females plants (Table 2).
@@
 
-Our study demonstrate that Palmer amaranth is well adapted to growing conditions throughout the Midwestern United States, and is simply limited in its range by seed dispersal (Davis et al.
+Our study demonstrate that Palmer amaranth is well adapted to growing conditions throughout the Midwestern US, and is simply limited in its range by seed dispersal (Davis et al.
 2015).
 We also confirmed Palmer amaranth's extraordinary plasticity to adapt to different agroecosystems.
-For example, Palmer amaranth mimicked crop architecture in competing for light (Figure 7), where plants growing in corn allocated resources to height, while plants growing in bareground allocated resources to numerous branches.
+For example, Palmer amaranth mimicked crop architecture in competing for light, where plants growing in corn allocated resources to height, while plants growing in bareground allocated resources to numerous branches.
 Plants growing in the absence of a crop produced the greatest biomass, responding to a greater abundance of light, nutrient and water resources (Figure 4A, 3B).
@@
 
-```{r fig7, echo = FALSE, message=FALSE, fig.show = "hide"}
+```{r fig7, echo = FALSE, message=FALSE, fig.show = "hide", include = FALSE}
 knitr::include_graphics("../data analysis/figures/Figure 7.png")
@@
 
-Our study demonstrated that Palmer amaranth is adapted to grow on arable land throughout the Midwestern U.S.
+Our study demonstrated that Palmer amaranth is adapted to grow on arable land throughout the north central United States.
 Palmer amaranth's range will continue to expand if current cropping practices are continued.
@@
 Where Palmer amaranth occurs, management tactics should focus on limiting Palmer amaranth establishment until row crops can shade late-emerging plants.
-Increasing the diversity of crops in rotation, varying row crop planting date, narrowing row width, and increasing residue cover through the use of cover crops are all tactics that can minimize the growth and seed production of Palmer amaranth, and improve the sustainability of North Central United States cropping systems.
+Increasing the diversity of crops in rotation, varying row crop planting date, narrowing row width, and increasing residue cover through the use of cover crops are all tactics that can minimize the growth and seed production of Palmer amaranth, and improve the sustainability of US North Central cropping systems.
 
@@
 
-```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Field research locations in the United States north central region.', out.width = "160mm", out.height = "90mm"}
+```{r, Figure-1, ref.label = "fig1", echo = FALSE, message = FALSE, fig.height=3, fig.width=6, fig.align='center', fig.cap='Field research locations in the north central United States region.', out.width = "160mm", out.height = "90mm"}
 # You can also refer to code chunks from above to place figures at the bottom.
@@
 
-```{r, Figure-7, ref.label = "fig7", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (first manuscript author, 180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
+```{r, Figure-7, include=FALSE, ref.label = "fig7", echo = FALSE, message = FALSE, fig.height=9, fig.width=4, fig.align='center', fig.cap='MCO (first manuscript author, 180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
 From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.', out.width = "150mm", out.height = "90mm"}

### 2022-04-01 Manuscript/Manuscript.tex
jhjh

@@
 germination period throughout the growing season (Ward et al., 2013).
-Germination of Palmer amaranth was triggered by 18 C soil temperature at
-5 cm depth (Keeley et al., 1987), and optimal germination and biomass
-production occurred at 35/30 C day and night temperatures (Guo and
-Al-Khatib, 2003). Palmer amaranth emergence can be influenced by tillage
-and the use of preemergence herbicides (Chahal et al., 2021), and may
-result in weed germination shifts within a population, as documented in
-\emph{Bassia scoparia} (Sbatella and Wilson, 2010). In an experiment
-where Palmer amaranth was subjected to continuous water stress, it
-survived and produced at least 14000 seeds plant\textsuperscript{-1}
-(Chahal et al., 2018). Seeds from Palmer amaranth growing with limited
-water conditions were heavier, less dormant, and prompt for germination
-(Matzrafi et al., 2021). Growing conditions and management practices
-also influence Palmer amaranth sex dimorphism and flowering pattern
-(Korres et al., 2017; Rumpa et al., 2019).
+Germination of Palmer amaranth was triggered by 18 \textdegree C soil
+temperature at 5 cm depth (Keeley et al., 1987), and optimal germination
+and biomass production occurred at 35/30 \textdegree C day and night
+temperatures (Guo and Al-Khatib, 2003). Palmer amaranth emergence can be
+influenced by tillage and the use of preemergence herbicides (Chahal et
+al., 2021), and may result in weed germination shifts within a
+population, as documented in \emph{Bassia scoparia} (Sbatella and
+Wilson, 2010). In an experiment where Palmer amaranth was subjected to
+continuous water stress, it survived and produced at least 14000 seeds
+plant\textsuperscript{-1} (Chahal et al., 2018). Seeds from Palmer
+amaranth growing with limited water conditions were heavier, less
+dormant, and prompt for germination (Matzrafi et al., 2021). Growing
+conditions and management practices also influence Palmer amaranth sex
+dimorphism and flowering pattern (Korres et al., 2017; Rumpa et al.,
+2019).
 
@@
 Belt (Garetson et al., 2019; Bagavathiannan and Norsworthy, 2016) and
-southern Great Plains of the United States Crespo et al. (2016).
-However, its range appears to be expanding steadily, and in some cases
-rapidly, in the U.S. For example, in states with large Palmer amaranth
-infestations its range is gradually pushing north as it displaces other
-weeds in agronomic row crops (R Werle, personal observation), aided by
-wildlife (Farmer et al., 2017) and agricultural machinery (Ward et al.,
-2013). But there are other examples where it has suddenly become a
+southern Great Plains of the United States (Kumar et al., 2020; Crespo
+et al., 2016). However, its range appears to be expanding steadily, and
+in some cases rapidly, in the United States. We hypothesize that in
+states with large Palmer amaranth infestations its range is gradually
+pushing north as it displaces other weeds in agronomic row crops, aided
+by wildlife (Farmer et al., 2017) and agricultural machinery (Ward et
+al., 2013). But there are other examples where it has suddenly become a
 problem weed hundreds of miles away from known infestations. New
@@
 
-There are many areas in the U.S. Corn Belt (North Central states) where
+There are many areas in the US Corn Belt (North Central states) where
 Palmer amaranth is not yet established, and its potential adaptability
 is untested. Nonetheless, the rapid expansion of Palmer amaranth across
-the Northern U.S. is concerning and warrants investigations on its
-adaptability. We do know that Palmer amaranth caused yield loss in
-Illinois soybean fields (Davis et al., 2015), but an Iowa study showed
-that Palmer amaranth was not as well adapted as waterhemp
+the Northern United States is concerning and warrants investigations on
+its adaptability. It has been shown that Palmer amaranth caused yield
+loss in Illinois soybean fields (Davis et al., 2015), but an Iowa study
+showed that Palmer amaranth was not as well adapted as waterhemp
 (\emph{Amaranthus tuberculatus}) to conditions in that geography (Baker,
-2021). We also know that Palmer amaranth plants that establish shortly
+2021). It also known that Palmer amaranth plants that establish shortly
 after row-crop planting have a much greater impact on crop yield than
@@
 
-\hypertarget{field-study}{%
-\subsection*{Field study}\label{field-study}}
-\addcontentsline{toc}{subsection}{Field study}
-
-The experiment was conducted in 2018 and 2019 under field conditions at
-five locations: Arlington, WI, Clay Center, NE,, Grant, NE, Lincoln, NE,
-and Macomb, IL.
-
-Fields were conventionally tilled prior to crop planting. Corn and
-soybean were planted in 76-cm row spacing (Table 1). Monthly mean air
-temperature and total precipitation were obtained using Daymet weather
-data from June through September across the five locations in 2018 and
-2019 (Correndo et al., 2021) (Figure 2).
-
 \begin{table}[!h]
 
-\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL.}
+\caption{\label{tab:unnamed-chunk-2}Field study attributes from Arlington, WI, Clay Center, NE, Grant, NE, Lincoln, NE and Macomb, IL}
 \centering
@@
 
- & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c}  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate\textsuperscript{c} & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor\textsuperscript{f} \\
+ & Weed control\textsuperscript{a} & \emph{S}-metolachor\textsuperscript{b} fb glyphosate\textsuperscript{c} /  & saflufenacil + imazethapyr + pyroxasulfone\textsuperscript{d} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor\textsuperscript{f} \\
 
@@
 
- & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + atrazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
 
@@
 
- & Type & Plano-silt-loam & Crete Silt Loam & 2018: Mace silt loam / 2019: Kuma silt loam & Crete silty clay loam & \\
-
- & pH & 6.6 & 6.5 & 2018: 6.1 / 2019: 5.1 & 5.2 & \\
-
-\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 & 2018: 2 / 2019: 1.7 & 3.4 & \\
+ & Type & Plano-silt-loam & Crete Silt Loam & 2018: Mace silt loam / 2019: Kuma silt loam & Crete silty clay loam &  2018: Osco silt loam / 2019: Keomah silt loam \\
+
+ & pH & 6.6 & 6.5 & 2018: 6.1 / 2019: 5.1 & 5.2 &  2018: 6.8 / 2019: 7.5\\
+
+\multirow{-4}{*}{\raggedright\arraybackslash } & Organic matter (\%) & 3.5 & 3 & 2018: 2 / 2019: 1.7 & 3.4 &  2018: 2.0 / 2019: 1.3\\
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + atrazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
 \end{tabular}}
@@
 soil surface to the top of plant. Also, aboveground plant, and the
-aboveground plant organs were harvested, then oven dried at 65 C until a
-constant weight was reached, and dry biomass (g
+aboveground plant organs were harvested, then oven dried at 65
+\textdegree C until a constant weight was reached, and dry biomass (g
 plant\textsuperscript{-1}) was recorded. Plants were harvested at
-flowering because Palmer amaranth infestations do not occur at the
-Wisconsin or Illinois research locations. In our study, all locations
-followed the methodology of plant harvest at flowering initiation,
-except in Grant, NE. At the Grant, NE, location, all plants from the
-first cohort were harvested on July 6, 2018 or July 6, 2019, and all
-plants from the second cohort were harvested on August 17, 2018 or July
-31, 2019, regardless of flowering status.
+flowering because Palmer amaranth is not established at the Wisconsin or
+Illinois research locations. In our study, all locations followed the
+methodology of plant harvest at flowering initiation, except in Grant,
+NE. At the Grant, NE, location, all plants from the first cohort were
+harvested on July 6, 2018 or July 6, 2019, and all plants from the
+second cohort were harvested on August 17, 2018 or July 31, 2019,
+regardless of flowering status.
 
@@
 The statistical analyses were performed using R statistical software
-version 4.0.1 (R Core Team, 2021).
+version 4.0.1 (Team, 2021).
 
@@
 test from McFadden, Cox and Snell, and Cragg and Uhler, respectively.
-The likelihood ratio test showed a p-value of \textless{} 0.00. The
-average marginal effects showed that Palmer amaranth growing in corn
-resulted in 14.8\% fewer females plants (Table 2). Moreover, increasing
-a cumulative flowering unit doy increases the probability of having a
-female plant by 0.4\% (Table 2 and Figure5A). A similar trend was
-observed for weight (Figure 6B) and height (Figure 6C), where the
-probability of being female increased 0.1\% and 0.2\% for each unit
-increase of weight and height, respectively (Table 2).
+The likelihood ratio test showed a p-value of 0.00. The average marginal
+effects showed that Palmer amaranth growing in corn resulted in 14.8\%
+fewer females plants (Table 2). Moreover, increasing a cumulative
+flowering unit doy increases the probability of having a female plant by
+0.4\% (Table 2 and Figure5A). A similar trend was observed for weight
+(Figure 6B) and height (Figure 6C), where the probability of being
+female increased 0.1\% and 0.2\% for each unit increase of weight and
+height, respectively (Table 2).
 
@@
 Our study demonstrate that Palmer amaranth is well adapted to growing
-conditions throughout the Midwestern United States, and is simply
-limited in its range by seed dispersal (Davis et al.~2015). We also
-confirmed Palmer amaranth's extraordinary plasticity to adapt to
-different agroecosystems. For example, Palmer amaranth mimicked crop
-architecture in competing for light (Figure 7), where plants growing in
-corn allocated resources to height, while plants growing in bareground
-allocated resources to numerous branches. Plants growing in the absence
-of a crop produced the greatest biomass, responding to a greater
-abundance of light, nutrient and water resources (Figure 4A, 3B). These
-results support the argument that Palmer amaranth can quickly evolve
-life-history traits to adapt to different cultural practices, similar to
-observations in a Palmer amaranth response to nitrogen study (Bravo et
-al., 2018). Our results highlight Palmer amaranth as a threat to field
-crops and breeding more competitive crop varieties is likely to select
-more competitive weed biotypes (Bravo et al., 2017).
+conditions throughout the Midwestern US, and is simply limited in its
+range by seed dispersal (Davis et al.~2015). We also confirmed Palmer
+amaranth's extraordinary plasticity to adapt to different
+agroecosystems. For example, Palmer amaranth mimicked crop architecture
+in competing for light, where plants growing in corn allocated resources
+to height, while plants growing in bareground allocated resources to
+numerous branches. Plants growing in the absence of a crop produced the
+greatest biomass, responding to a greater abundance of light, nutrient
+and water resources (Figure 4A, 3B). These results support the argument
+that Palmer amaranth can quickly evolve life-history traits to adapt to
+different cultural practices, similar to observations in a Palmer
+amaranth response to nitrogen study (Bravo et al., 2018). Our results
+highlight Palmer amaranth as a threat to field crops and breeding more
+competitive crop varieties is likely to select more competitive weed
+biotypes (Bravo et al., 2017).
 
@@
 Our study demonstrated that Palmer amaranth is adapted to grow on arable
-land throughout the Midwestern U.S. Palmer amaranth's range will
-continue to expand if current cropping practices are continued. As
+land throughout the north central United States. Palmer amaranth's range
+will continue to expand if current cropping practices are continued. As
 waterhemp and Palmer amaranth begin to share the same habitat, it will
@@
 of cover crops are all tactics that can minimize the growth and seed
-production of Palmer amaranth, and improve the sustainability of
-North Central United States cropping systems.
+production of Palmer amaranth, and improve the sustainability of US
+North Central cropping systems.
 
@@
 \leavevmode\hypertarget{ref-rcoreteam2021}{}%
-R Core Team (2021). \emph{R: {The R Project} for {Statistical
+Team, R. C. (2021). \emph{R: {The R Project} for {Statistical
 Computing}}. {Vienna, Austria}: {R Foundation for Statistical Computing}
@@
 
-\caption{Field research locations in the United States North Central region.}\label{fig:Figure-1}
+\caption{Field research locations in the north central United States region.}\label{fig:Figure-1}
 \end{figure}
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=150mm,height=90mm]{../data analysis/figures/Figure 7}
-
-}
-
-\caption{MCO (first manuscript author, 180 cm) holds harvested Palmer amaranth plants at 40 days after first cohort transplanting (A) and 33 days after second cohort transplanting (B) time.
-From left to right in each image, Palmer amaranth growing in bareground, soybean and corn in Arlington, Wisconsin.}\label{fig:Figure-7}
-\end{figure}
-
 \end{document}

### 2022-04-01 Manuscript/table.tex
as
jhjh

@@
 
- & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + trazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
+ & Weed control & \emph{S}-metolachor fb glyphosate & \emph{S}-metolachlor + atrazine + mesotrione + bicyclopyrone\textsuperscript{e} & glyphosate & glyphosate + \emph{S}-metolachor\textsuperscript{f} fb glyphosate\textsuperscript{g} & glyphosate + \emph{S}-metolachor \\
 
- & Stage at 1\textsuperscript{st} cohort & V2-3 & V6-7 & V6-7 & V2-3 & V4-5\\
+ & Stage at 1\textsuperscript{st} cohort & V2-3 & V2-3 & V2-3 & V2-3 & V4-5\\
 
@@
 \bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + trazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{a}Hoe weeding; \textsuperscript{b}\emph{S}--metolachor, 1324 g ai ha\textsuperscript{-1}; \textsuperscript{c}glyphoste, 1262 g ae ha\textsuperscript{-1}; \textsuperscript{d}saflufenacil + imazethapyr + pyroxasulfone, 215 g ai ha\textsuperscript{-1}; \textsuperscript{e}\emph{S}-metolachlor + atrazine + mesotrione, + bicyclopyrone, 2409 g ai ha\textsuperscript{-1}; \textsuperscript{f}\emph{S}--metolachor, 1070 g ai ha\textsuperscript{-1}; \textsuperscript{g}glyphoste, 870 g ae ha\textsuperscript{-1}} \\
 \end{tabular}}
@@
 
-\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop and bareground) is shown compared to soybean.}
+\caption{\label{tab:unnamed-chunk-3}Average marginal means of Palmer amaranth sex dimorphism logistic model. Factor pararemter values (crop\_bareground and crop\_corn) is shown compared to soybean.}
 \centering
